### 2013-04-28 _posts/2013-04-28-cross-referencing-reviewer-replies-in-latex.md
initial commit after dropbox sync problem

---
layout: post
title: Cross-referencing Revision Line Numbers in LaTeX
published: true
---

One of the more tedious elements of the scientific publishing process is cross-referencing all your revisions in a reply letter to the editor. Typically you make changes in your manuscript and then refer to those changes in the reply letter by page and line number.

For example, you might make a change to your manuscript and then in your reply letter you refer to it:

    We have revised the text. Please see page 3, line 23.

There are many strategies for dealing with this, but most of them require finding all your changes at the end and adding the page numbers and line numbers in the reply letter. This is fine if you only have to do it once, but inevitably something gets changed in the manuscript and all the cross-references need to be redone. Even worse, in working with Word documents that have track changes enabled I've experienced the line numbers changing after all changes are accepted.

I've settled on a fairly simple solution in LaTeX which requires minimal markup and automatically cross-references my changes. While this example deals with the case where the line numbers restart on each page, the solution is even simpler if you have continuous line numbers.

This solution assumes that your reply document is being written at the end of your manuscript. This isn't the only option, but it makes cross-referencing the simplest. At the end, before submitting the manuscript and reply document, I split the PDF in two.

First of all, you'll need to set up page-wise line numbers. Add the `lineno` package to the preamble:

    \usepackage[pagewise]{lineno}

And wherever you want to start the line numbers in your document add:

    \linenumbers

Now we'll set up two new commands in the preamble. First, create a new command to label your changes:

    \newcommand{\r}[1]{\label{#1}\linelabel{#1}}

Then, create a new command to cross-reference your changes:

    \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}

Now, say you wanted to revise a sentence to deal with Reviewer comment A1. You can just add something like this within the manuscript:

    \r{A1} A sentenced we have revised...

In the reply document, you can reference this change like this::

    We have revised the text. Please see \lr{A1}.

Now, when you compile your document, this will get converted to:

    We have revised the text. Please see page 3, line 23.

And as you modify your manuscript the page and line numbers will update.

### 2013-06-20 cv-body.tex
Bankruptcy commit since moving to Linode

\section{Education}

\begin{description}
\item[2011--13]
Ph.D.\ Candidate, Simon Fraser University, Burnaby, Canada Supervisors:
Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper\\Canadian Fulbright
Scholar to the University of Washington in 2012--13
\item[2010]
M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\Thesis title:
Trends, drivers, and ecosystem effects of expanding global invertebrate
fisheries. Supervisor: Dr.\ Heike K. Lotze.
(\href{papers/Anderson_MSc_2010.pdf}{PDF})\\Canadian
\href{http://goo.gl/nA1zE}{Governor General's Academic Gold Medal} for
the top-ranked Master's Natural Sciences and Engineering thesis at
Dalhousie University in 2010
\item[2004--07]
B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
Canada\\Thesis title: How sustainable are emerging low-trophic level
fisheries on the Scotian Shelf? Supervisors: Dr.\ Heike K. Lotze and
Dr.\ Ransom A. Myers. (\href{papers/Anderson_HSc_2006.pdf}{PDF})\\First
Class Honours \& Environmental Programmes Honour Society Medal
\item[2001--03]
B.Sc.\ Candidate (Hons.) Kinesiology, University of Waterloo, Waterloo,
Canada Transferred to Dalhousie University in 2004.
\end{description}

\section{Publications}

\begin{description}
\item[2012]
Phillis, C.C.*, S.M. O'Regan*, S.J. Green*, J.E.B. Bruce*, \textbf{S.C.
Anderson}, J.N. Linton, Earth2Ocean Research Derby, B. Favaro. Multiple
pathways to conservation success. Conservation Letters. Early view
online. \url{http://dx.doi.org/10.1111/j.1755-263X.2012.00294.x}.
(*\emph{Authors contributed equally, listed in reverse alphabetical
order})
\item[2012]
Harnik, P.G., H.K. Lotze, \textbf{S.C. Anderson}, Z.V. Finkel, S.
Finnegan, D.R. Lindberg, L.H. Liow, R. Lockwood, C.R. McClain, J.L.
McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson, D.P. Tittensor.
Extinctions in ancient and modern seas. Trends in Ecology and Evolution.
27(11): 608--617. \url{http://dx.doi.org/10.1016/j.tree.2012.07.010}.
\item[2012]
\textbf{Anderson, S.C.}, T.A. Branch, D. Ricard, H.K. Lotze. Assessing
global marine fishery status with a revised dynamic catch-based method
and stock-assessment reference points. ICES Journal of Marine Science.
69(8): 1491--1500. \url{http://dx.doi.org/10.1093/icesjms/fss105}.
(\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF})
\item[2011]
\textbf{Anderson, S.C.}, B.G. Farmer, F. Ferretti, A.L.S. Houde, and
J.A. Hutchings. Correlates of vertebrate extinction risk in Canada.
BioScience. 61(7): 538--549.
\url{http://dx.doi.org/10.1525/bio.2011.61.7.8}.
(\href{papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF})
\item[2011]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
Serial exploitation of global sea cucumber fisheries. Fish and
Fisheries. 12(3): 317-339.
\url{http://dx.doi.org/10.1111/j.1467-2979.2010.00397.x}.
(\href{papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF})
\href{http://www.sciencemag.org/content/331/6014/129.1.full}{Featured in
Science} on January 14 2011.
\item[2011]
Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down interactions
and temperature control of snow crab abundance in the northwest
Atlantic. Marine Ecology Progress Series. 429: 169--183.
\url{http://dx.doi.org/10.3354/meps09081}.
\item[2011]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze. Rapid
global expansion of invertebrate fisheries: trends, drivers, and
ecosystem effects. PLoS ONE. 6(3): e14735.
\url{http://dx.doi.org/10.1371/journal.pone.0014735}
\item[2008]
\textbf{Anderson, S.C.}, H.K. Lotze, and N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic-level fisheries in Atlantic
Canada. Can. J. Fish. Aquat. Sci. 65(12): 2553--2571.
\url{http://dx.doi.org/10.1139/F08-156}.
(\href{papers/Anderson_etal_2008_knowledge.pdf}{PDF}) Eighth most-read
CJFAS article in 2008.
\end{description}

\section{Submitted manuscripts}

\begin{description}
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Ecological prophets:
Quantifying metapopulation portfolio effects. Submitted with revisions
to Methods in Ecology and Evolution.
\item[2013]
O'Regan, S.M., W.J. Palen, \textbf{S.C. Anderson}. Climate warming
mediates negative impacts of rapid pond drying for three amphibian
species. Submitted to Ecology.
\end{description}

\section{Awards and scholarships}

\begin{description}
\item[2014]
Graduate Fellowship, Simon Fraser University
\item[2012--13]
\href{http://www.fulbright.ca/programs/canadian-students/2012-2013-canadian-fulbright-students.html}{Canadian
Fulbright Scholar award} to the University of Washington in 2012--13
\item[2011]
Canadian \href{http://goo.gl/nA1zE}{Governor General's Academic Gold
Medal} for the top-ranked Master's Natural Sciences and Engineering
thesis at Dalhousie University in 2010
\item[2011--14]
Provost Prize of Distinction, Simon Fraser University
\item[2011--14]
Natural Sciences and Engineering Research Council of Canada Postgraduate
Scholarship (Doctoral)
\item[2007--10]
Faculty Research Grant Scholarship, Dalhousie University
\item[2007--09]
Graduate Studies Scholarship, Dalhousie University
\item[2007]
Environmental Programmes Honour Society Medal, Dalhousie University
\item[2001--03]
Academic All-Canadian, University of Waterloo
\item[2001--02]
Dean's Entrance Scholarship, University of Waterloo
\end{description}

\section{Invited talks and conference presentations}

\begin{description}
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. False prophets: The
challenges of quantifying ecological portfolios
(\href{talks/2013/PE_SAFS_quantsem.pdf}{slides}), Quantitative Seminar,
School of Aquatic and Fishery Sciences, University of Washington,
Seattle, WA, United States.
\item[2012]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. False prophets: The
ecological portfolio effect overestimates the benefit of diversity.
Branch Lab, School of Aquatic and Fishery Sciences, University of
Washington, Seattle, WA, United States.
\item[2012]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Metapopulation
dynamics and the generality of the ecological portfolio effect. Les
Ecologistes Departmental Seminar, Simon Fraser University, Burnaby, BC,
Canada.
\item[2010]
\textbf{Anderson, S.C.} The rise of invertebrates, the fall of sea
cucumbers, and the risk of maturing late. Earth2Ocean Research Group,
Simon Fraser University, Burnaby, BC, Canada.
\item[2009]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
Ecosystem impacts of the global expansion of invertebrate fisheries.
International Oceans Past II Conference \emph{Multidisciplinary
Perspectives on the History and Future of Marine Animal Populations}.
Vancouver, BC, Canada.
\item[2009]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
Global invertebrate fisheries: trends and consequences. NCEAS (National
Center for Ecological Analysis and Synthesis) Working Group
\emph{Finding Common Ground in Marine Conservation and Management}.
Santa Barbara, CA, USA.
\item[2009]
\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic level fisheries in Atlantic
Canada. Harvest Fisheries Seminar Series. Bedford Institute of
Oceanography, Fisheries and Oceans, Dartmouth, NS, Canada.
\item[2008]
\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic level fisheries in Atlantic
Canada. Departmental Seminar, Biology Department, Dalhousie University,
Canada.
\end{description}

\section{Contract employment}

\begin{description}
\item[2011--13]
Database management, data analysis, and visualization for the project
\emph{Global Shark Abundance Baselines} (Hopkins Marine Station,
Stanford University, CA, United States) with F. Micheli, F. Ferretti,
N.K. Dulvy, H.K. Lotze, and B. Worm.
\item[2010]
Assessment of Arctic surfclam on Banquereau Bank using industry
collected data for Clearwater Seafoods, Bedford, NS, Canada.
\item[2010]
Modelling length-frequency distribution changes of Atlantic halibut for
NAFO Divisions 4VWX5Z and 3NOP for Dr.\ K. Trzcinski, Population Ecology
Division, Bedford Institute of Oceanography, Fisheries and Oceans,
Dartmouth, NS, Canada.
\end{description}

\section{Working groups and workshops}

\begin{description}
\item[2011--13]
NESCent (National Evolutionary Synthesis Center, Durham, NC) Working
Group \emph{Determinants of Extinction in Ancient and Modern Seas} led
by Paul Harnik, Seth Finnegan, and Rowan Lockwood.
(\href{http://www.nescent.org/science/awards_summary.php?id=256}{URL})
\item[2010]
Atlantic Halibut Assessment Science Peer Review Meeting, Fisheries and
Oceans, Dartmouth, NS, Canada.
\item[2007--09]
NCEAS (National Center for Ecological Analysis and Synthesis, Santa
Barbara, CA) Distributed Graduate Seminar, in association with the
Working Group \emph{Finding Common Ground in Marine Conservation and
Management} led by Ray Hilborn and Boris Worm.
(\href{http://www.nceas.ucsb.edu/projects/12307}{URL})
\item[2008]
Workshop on Canadian Science and Management Strategies for Sea Cucumber
(\emph{Cucumaria frondosa}), Fisheries and Oceans, Dartmouth, NS, Canada
(work presented \emph{in absentia}).
\item[2007]
Workshop on Canadian Science and Management Strategies for Atlantic
Hagfish, Fisheries and Oceans, Dartmouth, NS, Canada.
(\href{http://www.dfo-mpo.gc.ca/CSAS/Csas/Publications/Pro-CR/2009/2009_009_e.htm}{URL})
\end{description}

\section{Selected courses and training}

\begin{description}
\item[Present]
Proficient in statistical analyses and data management using R, WinBUGS,
JAGS, Sweave, knitr, LaTeX, Perl, PostgreSQL, Git.
\item[2013]
Quantitative Modelling in Ecosystem Based Management, R. Hilborn, School
of Aquatic and Fishery Sciences, University of Washington (auditing).
\item[2013]
Applied Time Series Analysis; E. Holmes, M. Scheuerell, and E. Ward;
School of Aquatic and Fishery Sciences; University of Washington
(audited).
\item[2012]
Numerical Computing for the Natural Resources, Dr.\ A.E. Punt, School of
Aquatic and Fishery Sciences, University of Washington (audited).
\item[2012]
Risk Assessment and Decision Analysis for Management of Natural
Resources, Dr.\ R.M. Peterman, School of Resource and Environmental
Management, Simon Fraser University.
\item[2011]
Presenting Data and Information, Edward Tufte, Seattle, WA, United
States.
\item[2009]
Technical Expertise in Stock Assessment: Model Fitting and
Interpretation by Dr.\ C. Needle, chair of the International Council for
the Exploration of the Sea working group on methods of fish stock
assessment, Dalhousie University.
\item[2008]
Data Analysis, Dr.\ J. Mills Flemming, Department of Mathematics \&
Statistics, Dalhousie University (audited).
\end{description}

\section{Teaching}

\begin{description}
\item[2012]
Introduction to ggplot2.
(\href{courses/12-ggplot2/ggplot2_notes.pdf}{notes},
\href{courses/12-ggplot2/ggplot2_slides_with_examples.pdf}{slides})
Lecture for FISH 507H: Beautiful Graphics in R, School of Aquatic and
Fishery Sciences, University of Washington, Seattle, WA, United States
\item[2012]
Workshop on the R package plyr.
(\href{courses/12-plyr/plyr_2012.pdf}{notes},
\href{courses/12-plyr/plyr_2012_slides.pdf}{slides},
\href{courses/12-plyr/plyr_2012_examples.html}{examples}) Branch Lab,
School of Aquatic and Fishery Sciences, University of Washington,
Seattle, WA, United States
\item[2011]
Multipanel plotting in R with base graphics.
(\href{courses/11-multipanel/multipanel.pdf}{notes},
\href{courses/11-multipanel/multipanel-slides.pdf}{slides}). Earth2Ocean
Research Group, Simon Fraser University, Burnaby, BC, Canada.
\item[2011]
A brief introduction to R. (\href{courses/11-rintro/RIntro.pdf}{notes},
\href{courses/11-rintro/RIntro.R}{workshop code}). Earth2Ocean Research
Group, Simon Fraser University, Burnaby, BC, Canada.
\item[2011]
Workshop on the R package plyr (\href{courses/11-plyr/plyr.pdf}{notes},
\href{courses/11-plyr/plyr-slides.pdf}{slides}). Earth2Ocean Research
Group, Simon Fraser University, Burnaby, BC, Canada.
\item[2007--08]
Teaching Assistant, Organismal Biology and Ecology, Dalhousie
University, two semesters (BIOL 1021).
\item[2007--08]
Teaching Assistant, Marine Mammology, Dalhousie University (BIOL 4060).
\end{description}

\section{Reviews}

Reviewer for Ecology, Conservation Biology, Fish and Fisheries,
International Journal of Tropical Biology and Conservation, Journal of
Environmental Management, Aquatic Conservation

### 2013-06-20 courses/11-multipanel/multipanel.R

@@ -1,0 +1,55 @@
+## same axes
+par(xpd = NA)
+par(mfrow = c(2, 3))
+par(mar = c(0, 0, 0, 0))
+par(oma = c(4, 4, 0.5, 0.5))
+for (i in 1:6) {
+  plot(1, axes = FALSE, xlab = "", ylab = "")
+  box(col = "grey60")
+  if (i %in% c(4, 5, 6)) 
+    axis(1, col = "grey60")
+  if (i %in% c(1, 4)) 
+    axis(2, col = "grey60", las = 1)
+  mtext(letters[i], side = 3, adj = 0.1, line = -2)
+}
+mtext("X axis", side = 1, line = 2.2, outer = TRUE)
+mtext("Y axis", side = 2, line = 2.2, outer = TRUE)
+
+## separate y-axes
+par(xpd = NA)
+par(mfrow = c(2, 3))
+par(mar = c(0, 3, 0, 0))
+par(oma = c(4, 2, 0.5, 0.5))
+for (i in 1:6) {
+  plot(rnorm(40), axes = FALSE, xlab = "", ylab = "", ylim = c(-3, 3.5))
+  box(col = "grey60")
+  if (i %in% c(4, 5, 6)) axis(1, col = "grey60")
+  axis(2, col = "grey60", las = 1)
+  mtext(letters[i], side = 3, adj = 0.1, line = -2)
+}
+mtext("X axis", side = 1, line = 2.2, outer = TRUE)
+mtext("Y axis", side = 2, line = 0.5, outer = TRUE)
+
+## layout()
+pdf("example-with-layout.pdf", width = 5, height = 3.5)
+layout(rbind(c(1, 1), c(2, 3)))
+layout.show(3)
+par(mar = c(2, 4, 1, 1))
+par(oma = c(2, 2, 0, 0))
+par(cex = 0.7)
+plot(1)
+plot(1)
+plot(1)
+dev.off()
+
+## split.screen()
+m <- rbind(c(0.15, 0.2, 0.15, 1), c(0.22, 1, 0.15, 1))
+split.screen(m)
+screen(1)
+par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
+plot(1)
+screen(2)
+par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
+plot(1, axes = TRUE, yaxt = "n")
+close.screen(all = TRUE)
+       

### 2013-06-20 courses/11-plyr/plyr.examples.R

@@ -1,0 +1,51 @@
+library(plyr)
+head(baseball) # see ?baseball
+
+# plot the correlation between stolen bases and homeruns by year:
+plot(ddply(baseball, "year", summarize, correlation = cor(sb, hr)), type = "b")
+
+# capture the output from a linear model of stolen bases and homeruns
+# for each year into a list:
+out <- dlply(subset(baseball, year > 1950), "year", function(x) with(x, 
+  lm(sb ~ hr)))
+head(out)
+summary(out[[1]]) # the first model output
+summary(out[[2]]) # the second...
+
+# browse through the data by year to build up a more complicated
+# function (or debug an existing function)
+# I've commented it out so you don't accidentally run it:
+# out <- dlply(subset(baseball, year > 1950), "year", function(x) {
+#  browser()
+# })
+# hit Return and look at the value of x each time
+# type Q and hit the Return key to quit the browser
+
+# get the slope and intercept from a linear model, fitting a separate
+# linear model for each year:
+out <- ddply(subset(baseball, year > 1950), "year", function(x) {
+  lm.out <- as.numeric(coef(with(x, lm(hr ~ sb))))
+  data.frame(slope = lm.out[2], intercept = lm.out[1])
+})
+head(out)
+
+# create an additional column with the residuals from a linear model
+# fit for each year:
+out.residuals <- ddply(subset(baseball, year > 1950), "year", 
+  transform, residuals = residuals(lm(hr ~ sb)))
+head(out.residuals)
+
+# get the mean stolen bases for each year and team combination:
+out <- ddply(baseball, c("year", "team"), summarise, mean.sb = mean(sb, 
+  na.rm = T))
+head(out)
+
+# Added bonus: plot it
+# first exclude missing data and teams with less than 50 years of data 
+# just for plotting:
+out <- na.omit(out)
+out <- subset(out, team %in% names(table(out$team))[table(out$team) >= 50])
+
+# plot it:
+library(ggplot2)
+ggplot(out, aes(year, mean.sb)) + facet_wrap(~ team) + geom_line()

### 2013-06-20 courses/11-rintro/RIntro.R

@@ -1,0 +1,241 @@
+########################################
+## A brief introduction to R
+## Sean Anderson, sean_anderson@sfu.ca
+## September 2011
+########################################
+
+## This file is meant to accompany the handout RIntro.pdf. I've left
+## space between comments for you to write code. A lot of programming
+## is about figuring out how to do it once (i.e. suffering) and then
+## looking back at what you did when you need to do it again.
+## Hopefully this file will serve as a reference for many of the
+## common tasks you will run into.
+
+## I'll have you read sections of the handout. Then we'll talk about
+## that section briefly and I'll have you try and work out the
+## relevant code in this worksheet.  Then we'll go over the code as a
+## group and discuss any questions you have.
+  
+## First, read sections 1, 2, and 3 in the handout and open up RStudio
+## or your preferred R script editor and console. Poke around the
+## menus and options a bit if you aren't familiar with it. We'll
+## discuss how the software is set up briefly.
+
+###################################
+## A quick introduction to objects
+###################################
+## Read sections 4, 5, 6 and have a go at the following:
+
+# Make a vector of character that's 3 elements long. I'll start:
+character.vector <- c("a", "b", "c")
+# Now print it out in the console:
+character.vector
+# Make a vector of numeric that's 3 elements long:
+numeric.vector <-
+# Multiply your numeric.vector by 2:
+numeric.vector <- 
+# Make a logical vector (TRUE and FALSE) that's 3 elements long:
+logical.vector <-
+# Combine your 3 objects into a data frame with a new object name.
+# You'll need to pick names for your columns.
+my.data.frame <- 
+# Make that data frame the first element of a list with a new object name:
+my.list <- list()
+my.list[[1] <- 
+# Think about when you would want to use these various object types.
+
+# Remove one of the objects:
+
+# Remove all objects in your workspace:
+
+# Now re-run the above code to get your objects back in your
+# workspace. You'll need them for the next steps. Get in the habit
+# of sending over code using the keyboard shortcuts instead of
+# cutting and pasting or clicking on the buttons. In RStudio, you
+# can select the relevant code and hit command-return on a Mac or
+# control-return on a PC. 
+# See Edit -> Run Code for all your options.
+
+#################################################
+## Looking at, manipulating, and subsetting data
+#################################################
+## Read the sections 7, 8, 9 and try the following:
+
+# Let's start with some basic indexing and subsetting of the objects
+# you created above.
[…]

### 2013-06-20 courses/12-ggplot2/ggplot2_exercises_blank.R

@@ -1,0 +1,41 @@
+library(ggplot2)
+
+# Introductory code from slides:
+d <- data.frame(x = c(1:10, 1:10), 
+                y = runif(20), 
+                group1 = rep(gl(2, 5, labels = c("a", "b")), 2), 
+                group2 = gl(2, 10))
+head(d)
+ggplot(d) + geom_point(aes(x = x, y = y, colour = group1)) + facet_grid(~group2)
+
+# Exercises:
+head(diamonds) # from the ggplot2 package
+set.seed(1)
+diamonds2 <- diamonds[sample(nrow(diamonds), 1000), ]
+
+# Here's a basic plot with points to start with
+ggplot(diamonds2, aes(x = carat, y = price)) + geom_point()
+
+# Look up some geoms at http://docs.ggplot2.org/ and try at least 3
+# with the diamonds2 dataset. geoms you might want to try: geom_line,
+# geom_histogram, geom_bar, geom_violin, geom_point, geom_freqpoly ...
+
+
+# By combining the aesthetics of geom_point: x, y, alpha, colour,
+# shape, and size, see how many dimensions of the data you can display
+# simultaneously with aesthetics.
+
+
+# If you have time, try expressing the same data with different
+# aesthetics.
+
+
+# What happens when you specify a continuous vs. a factor variable for
+# an aesthetic like colour?
+
+
+# Try replacing one of the aesthetics with a facet_wrap.
+
+
+# Now use facet_grid with the diamonds dataset.
+

### 2013-06-20 cv-body.tex
Update CV

@@
 Ph.D.\ Candidate, Simon Fraser University, Burnaby, Canada Supervisors:
-Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper\\Canadian Fulbright
+Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper Canadian Fulbright
 Scholar to the University of Washington in 2012--13
 \item[2010]
-M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\Thesis title:
+M.Sc.\ Biology, Dalhousie University, Halifax, Canada Thesis title:
 Trends, drivers, and ecosystem effects of expanding global invertebrate
 fisheries. Supervisor: Dr.\ Heike K. Lotze.
-(\href{papers/Anderson_MSc_2010.pdf}{PDF})\\Canadian
+(\href{papers/Anderson_MSc_2010.pdf}{PDF}) Canadian
 \href{http://goo.gl/nA1zE}{Governor General's Academic Gold Medal} for
@@
 B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
-Canada\\Thesis title: How sustainable are emerging low-trophic level
+Canada Thesis title: How sustainable are emerging low-trophic level
 fisheries on the Scotian Shelf? Supervisors: Dr.\ Heike K. Lotze and
-Dr.\ Ransom A. Myers. (\href{papers/Anderson_HSc_2006.pdf}{PDF})\\First
+Dr.\ Ransom A. Myers. (\href{papers/Anderson_HSc_2006.pdf}{PDF}) First
 Class Honours \& Environmental Programmes Honour Society Medal
@@
 global expansion of invertebrate fisheries: trends, drivers, and
-ecosystem effects. PLoS ONE. 6(3): e14735.
+ecosystem effects. PLOS ONE. 6(3): e14735.
 \url{http://dx.doi.org/10.1371/journal.pone.0014735}
@@
 
-\section{Submitted manuscripts}
+\section{Manuscripts in review}
 
@@
 \textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Ecological prophets:
-Quantifying metapopulation portfolio effects. Submitted with revisions
-to Methods in Ecology and Evolution.
+Quantifying metapopulation portfolio effects. Completing minor revisions
+for Methods in Ecology and Evolution.
 \item[2013]
@@
 mediates negative impacts of rapid pond drying for three amphibian
-species. Submitted to Ecology.
+species. In review at Ecology.
+\item[2013]
+Artelle K.A., \textbf{S.C. Anderson}, A.B. Cooper, Paul C. Paquet, J.D.
+Reynolds, C.T. Darimont. Confronting uncertainty in wildlife management:
+performance of grizzly bear management in British Columbia, Canada. In
+review at PLOS ONE.
 \end{description}

### 2013-10-20 _posts/2013-04-28-cross-referencing-reviewer-replies-in-latex.md
Bankruptcy commit since moving to Linode
Add subtitles to existing blog posts

@@
 layout: post
-title: Cross-referencing Revision Line Numbers in LaTeX
+title: Cross-Referencing Line Numbers in LaTeX
+subtitle: How to make writing reviewer response documents easier.
 published: true
@@
 
-    \newcommand{\r}[1]{\label{#1}\linelabel{#1}}
+    \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
 
@@
 
-    \r{A1} A sentenced we have revised...
+    \R{A1} A sentenced we have revised...
 
@@
 
-And as you modify your manuscript the page and line numbers will update.
+And as you modify your manuscript, the page and line numbers will update.

### 2013-10-20 _knitr-posts/2013-10-20-plyr.Rmd
Change date on plyr post

---
layout: post
title: "plyr: Split-Apply-Combine for Mortals"
subtitle: An introduction to the plyr R package.
published: true
---

`plyr` is an `R` package that makes it simple to split data apart, do
stuff to it, and mash it back together. This is a common
data-manipulation step. Importantly, `plyr` makes it easy to control the
input and output data format with a consistent syntax.

Or, from the documentation:

“`plyr` is a set of tools that solves a common set of problems: you need
to break a big problem down into manageable pieces, operate on each
piece and then put all the pieces back together. It’s already possible
to do this with split and the apply functions, but `plyr` just makes it
all a bit easier…”

This is a very quick introduction to `plyr`. For more details see Hadley
Wickham’s introductory guide [The split-apply-combine strategy for data
analysis](http://www.jstatsoft.org/v40/i01) (2011, Journal of
Statistical Software, Vol 40). There’s quite a bit of discussion online
in general, and especially on
[stackoverflow.com](http://stackoverflow.com/questions/tagged/plyr).

Why use `apply` functions instead of `for` loops?
=================================================

1.  The code is cleaner (once you’re familiar with the concept). The
    code can be easier to code and read, and less error prone because you
    don’t have to deal with subsetting and you don’t have to deal with saving
    your results.

2.  `apply` functions can be faster than `for` loops, sometimes dramatically.

`plyr` basics
=============

`plyr` builds on the built-in `apply` functions by giving you control
over the input and output formats and keeping the syntax consistent
across all variations. It also adds some niceties like error processing,
parallel processing, and progress bars.

The basic format is 2 letters followed by `ply()`. The first letter
refers to the format in and the second to the format out.

The 3 main letters are:

1.  `d` = data frame

2.  `a` = array (includes matrices)

3.  `l` = list

So, `ddply` means: take a data frame, split it up, do something to it,
and return a data frame. I find I use this the majority of the time
since I often work with data frames. `ldply` means: take a list, split
it up, do something to it, and return a data frame. This extends to all
combinations. In the following table, the columns are the input formats
and the rows are the output format:

object type| data frame | list    | array
-----------|------------|---------|--------
data frame | `ddply`    | `ldply` | `adply`
list       | `dlply`    | `llply` | `alply`
array      | `daply`    | `laply` | `aaply`

I’ve ignored some less common format options:

1.  `m` = multi-argument function input
2.  `r` = replicate a function `n` times.
3.  `_` = throw away the output

For plotting, you might find the underscore (`_`) option useful. It will
do something with the data (say add line segments to a plot) and then
throw away the output (e.g., `d_ply()`).

Base R `apply` functions and `plyr`
===================================

`plyr` provides a consistent and easy-to-work-with format for `apply`
functions with control over the input and output formats. Some of the
functionality can be duplicated with base R functions (but with less
consistent syntax). Also, few R `apply` functions work directly with
data frames as input and output and data frames are a common object
class to work with.

Base R `apply` functions (from a presentation given by Hadley):

object type        | array       | data frame  | list        | nothing
-------------------|-------------|-------------|-------------|--------
array              | `apply`     | .           | .           | .
data frame         | .           | `aggregate` | `by`        | .
list               | `sapply`    | .           | `lapply`    | .
n replicates       | `replicate` | .           | `replicate` | .
function arguments | `mapply`    | .           | `mapply`    | .

A general example with `plyr`
=============================

Let’s take a simple example. We’ll take a data frame, split it up by
`year`, calculate the coefficient of variation of the `count`, and
return a data frame. This could easily be done on one line, but I’m
expanding it here to show the format a more complex function could take.

```{r}
set.seed(1)
d <- data.frame(year = rep(2000:2002, each = 3),
  count = round(runif(9, 0, 20)))
print(d)
```

```{r}
library(plyr)
ddply(d, "year", function(x) {
  mean.count <- mean(x$count)
  sd.count <- sd(x$count)
  cv <- sd.count/mean.count
  data.frame(cv.count = cv)
})
```

`transform` and `summarise`
===========================

It is often convenient to use these functions within `plyr`. `transform`
acts as it would normally as the base R function and modifies an
existing data frame. `summarise` creates a new condensed data
frame.

```{r}
ddply(d, "year", summarise, mean.count = mean(count))
```

```{r}
ddply(d, "year", transform, total.count = sum(count))
```

Bonus function: `mutate`. `mutate` works like `transform` but lets you
build on columns you build.

```{r}
ddply(d, "year", mutate, mu = mean(count), sigma = sd(count),
  cv = sigma/mu)
```

Plotting with `plyr`
====================

You can use `plyr` to plot data by throwing away the output with an
underscore (`_`). This is a bit cleaner than a for loop since you don’t
have to subset the data manually.

```{r d_ply_plot, fig.width=6, fig.height=2.5, fig.cap=""}
par(mfrow = c(1, 3), mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0))
d_ply(d, "year", transform, plot(count, main = unique(year), type = "o"))
mtext("count", side = 1, outer = TRUE, line = 1)
mtext("frequency", side = 2, outer = TRUE, line = 1)
```

Nested chunking of the data
===========================

The basic syntax can be easily extended to break apart the data based on
multiple columns:

```{r}
baseball.dat <- subset(baseball, year > 2000) # data from the plyr package
x <- ddply(baseball.dat, c("year", "team"), summarize,
  homeruns = sum(hr))
head(x)
```

Other useful options
====================

Dealing with errors
-------------------

You can use the `failwith` function to control how errors are dealt
with.

```{r}
f <- function(x) if (x == 1) stop("Error!") else 1
safe.f <- failwith(NA, f, quiet = TRUE)
# llply(1:2, f)
llply(1:2, safe.f)
```

Parallel processing
-------------------

In conjunction with `doMC` (or `doSMP` on Windows) you can run your
function separately on each core of your computer. On a dual core
machine this could double your speed in some situations. Set
`.parallel = TRUE`.

```{r}
x <- c(1:10)
wait <- function(i) Sys.sleep(0.1)
system.time(llply(x, wait))
```

```{r}
system.time(sapply(x, wait))
```

```{r}
library(doMC)
registerDoMC(2)
system.time(llply(x, wait, .parallel = TRUE))
```

So, why would I *not* want to use `plyr`?
=========================================

`plyr` can be slow — particularly if you are working with very large
datasets that involve a lot of subsetting. Hadley is working on this and
an in-development version of `plyr`, `dplyr`, can run much faster (<https://github.com/hadley/dplyr>).
However, it’s important to remember that typically the speed that you
can write code and understand it later is the rate-limiting step.

A couple faster options:

Use a base R `apply` function:

```{r}
system.time(ddply(baseball, "id", summarize, length(year)))
```

```{r}
system.time(tapply(baseball$year, baseball$id, 
  function(x) length(x)))
```

Use the `data.table` package:

```{r}
library(data.table)
dt <- data.table(baseball, key = "id")
system.time(dt[, length(year), by = list(id)])
```

### 2013-10-20 _knitr-posts/2013-10-19-reshape.Rmd
Add reshape .Rmd file

---
layout: post
title: An Introduction to reshape2
subtitle: Reshaping data easily with the reshape2 R package.
published: true
---

[`reshape2`](http://cran.r-project.org/web/packages/reshape2/index.html) is an
`R` package written by [Hadley Wickham](http://had.co.nz) that makes it easy
to transform data between wide and long formats.

What makes data wide or long? 
=============================

Wide data has a column for each variable. For example, this is
wide-format data:

```{r, include=FALSE, echo=FALSE}
library(reshape2)
names(airquality) <- tolower(names(airquality))
aqm <- melt(airquality, id=c("month", "day"), na.rm=TRUE)
d <- dcast(aqm, month ~ variable, mean, margins = c("month", "variable"))
d <- d[1:4, -c(1, 3, 6)]
```

```{r, echo=FALSE}
head(d)
```

And this is long-format data:

```{r, echo=FALSE}
melt(d)
```

Long-format data has a column for possible variable types and a column
for the values of those variables. Long-format data isn’t necessarily
only two columns. For example, we might have ozone measurements for each
day of the year. In that case, we could have another column for day. In
other words, there are different levels of “longness”. The ultimate
shape you want to get your data into will depend on what you are doing
with it.

It turns out that you need wide-format data for some types of data analysis
and long-format data for others. In reality, you need long-format data much
more commonly than wide-format data. For example, `ggplot2` requires
long-format data 
(technically [tidy data](http://vita.had.co.nz/papers/tidy-data.html)),
`plyr` requires long-format data, and most modelling functions (such as
`lm()`, `glm()`, and `gam()`) require long-format data. But people often find
it easier to record their data in wide format.

The reshape2 package 
======================

`reshape2` is based around two key functions: `melt` and `cast`:

`melt` takes wide-format data and melts it into long-format data.

`cast` takes long-format data and casts it into wide-format data.

Think of working with metal: if you melt metal, it drips and becomes
long. If you cast it into a mould, it becomes wide.

Wide- to long-format data: the melt function 
==============================================

For this example we’ll work with the `airquality` dataset that is built
into `R`. First we’ll change the column names to lower case to make them
easier to work with. Then we’ll look at the data:

```{r}
names(airquality) <- tolower(names(airquality))
head(airquality)
```

What happens if we run the function `melt` with all the default argument
values?

```{r}
aql <- melt(airquality) # [a]ir [q]uality [l]ong format
head(aql)
```

```{r}
tail(aql)
```

By default, `melt` has assumed that all columns with numeric values are
variables with values. Often this is what you want. Maybe here we want
to know the values of `ozone`, `solar.r`, `wind`, and `temp` for each
`month` and `day`. We can do that with `melt` by telling it that we want
`month` and `day` to be “ID variables”. ID variables are the variables
that identify individual rows of data.

```{r}
aql <- melt(airquality, id.vars = c("month", "day"))
head(aql)
```

What if we wanted to control the column names in our long-format data?
`melt` lets us set those too all in one step:

```{r}
aql <- melt(airquality, id.vars = c("month", "day"),
  variable.name = "climate_variable", 
  value.name = "climate_value")
head(aql)
```

Long- to wide-format data: the cast functions 
===============================================

Whereas going from wide- to long-format data is pretty straightforward,
going from long- to wide-format data can take a bit more thought. It
usually involves some head scratching and some trial and error for all
but the simplest cases. Let’s go through some examples.

In `reshape2` there are multiple `cast` functions. Since you will most
commonly work with `data.frame` objects, we’ll explore the
`dcast` function. (There is also `acast` to return a vector, matrix, or
array.)

Let’s take the long-format `airquality` data and `cast` it into some
different wide formats. To start with, we’ll recover the same format we
started with and compare the two.

`dcast` uses a formula to describe the shape of the data. The arguments
on the left refer to the ID variables and the arguments on the right
refer to the measured variables. Coming up with the right formula can
take some trial and error at first. So, if you’re stuck don’t feel bad
about just experimenting with formulas. There are usually only so many
ways you can write the formula.

Here, we need to tell `dcast` that `month` and `day` are the ID
variables (we want a column for each) and that `variable` describes the
measured variables. Since there is only one remaining column,
`dcast` will figure out that it contains the values themselves. We could
explicitly declare this with `value.var`. (And in some cases it will be
necessary to do so.)

```{r}
aql <- melt(airquality, id.vars = c("month", "day"))
aqw <- dcast(aql, month + day ~ variable)
head(aqw)
```

```{r}
head(airquality) # original data
```

So, besides re-arranging the columns, we’ve recovered our original data.

If it isn’t clear to you what just happened there, then have a look at
this illustration:

![](/images/dcast-illustration.png)

Figure 1: An illustration of the `dcast` function. The blue shading indicates
ID variables that we want to represent individual rows. The red shading
represents variable names that we want to swing into column names. The grey
shading represents the data values that we want to fill in the cells with.

One confusing “mistake” you might make is casting a dataset in which
there is more than one value per data cell. For example, this time we
won’t include `day` as an ID variable:

```{r}
dcast(aql, month ~ variable)
```

When you run this in `R`, you’ll notice the warning message:

    # Aggregation function missing: defaulting to length

And if you look at the output, the cells are filled with the number of
data rows for each month-climate combination. The numbers we’re seeing
are the number of days recorded in each month. When you `cast` your data
and there are multiple values per cell, you also need to tell
`dcast` how to aggregate the data. For example, maybe you want to take
the `mean`, or the `median`, or the `sum`. Let’s try the last example,
but this time we’ll take the mean of the climate values. We’ll also pass
the option `na.rm = TRUE` through the `...` argument to remove `NA`
values. (The `...` let’s you pass on additional arguments to your
`fun.aggregate` function, here `mean`.)

```{r}
dcast(aql, month ~ variable, fun.aggregate = mean, 
  na.rm = TRUE)
```

Unlike `melt`, there are some other fancy things you can do with
`dcast` that I’m not covering here. It’s worth reading the help file
`?dcast`. For example, you can compute summaries for rows and columns,
subset the columns, and fill in missing cells in one call to `dcast`.

Additional help 
===============

Read the package help:
`help(package = "reshape2")`

See the `reshape2` website:
<http://had.co.nz/reshape/>

And read the paper on `reshape`:
Wickham, H. (2007). Reshaping data with the `reshape` package.
21(12):1–20.
<http://www.jstatsoft.org/v21/i12>  
(But note that the paper is written for the `reshape` package not the
`reshape2` package.)

### 2013-10-20 _knitr-posts/2013-10-20-plyr.Rmd
Switch to doParallel

@@
 
-In conjunction with `doMC` (or `doSMP` on Windows) you can run your
-function separately on each core of your computer. On a dual core
-machine this could double your speed in some situations. Set
-`.parallel = TRUE`.
+In conjunction with a package such as `doParallel` you can run your function
+separately on each core of your computer. On a dual core machine this make
+your code up to twice as fast. Simply register the cores and then set
+`.parallel = TRUE`. Look at the `elapsed` time in these examples:
 
@@
 ```{r}
-library(doMC)
-registerDoMC(2)
+library(doParallel)
+registerDoParallel(cores = 2)
 system.time(llply(x, wait, .parallel = TRUE))

### 2013-10-20 _knitr-posts/2013-10-20-plyr.Rmd
Left-align table headers and re-knit

@@
 
-object type| data frame | list    | array
------------|------------|---------|--------
+object type |data frame  |list     |array
+------------|------------|---------|--------
 data frame  | `ddply`    | `ldply` | `adply`

### 2013-10-20 _plugins/lightbox.rb
Update lightbox plugin

@@ -1,20 +1,0 @@
-# A Liquid tag for Jekyll sites that allows embedding Lightbox images.
-# by: kyoendo
-# Source URL: https://gist.github.com/4035604
-#
-# Example usage: {% lightbox 2012/abc.png, Title of Image, Alt Title %}
-module Jekyll
-  class LightboxTag < Liquid::Tag
-    def initialize(tag_name, text, token)
-      super
-      @text = text
-    end
-
-    def render(context)
-      thumb, path, title, alt = @text.split(',').map(&:strip)
-      %{<a href="#{path}" rel="lightbox" title="#{title}"><img src="#{thumb}" alt="#{alt || title}" /></a>}
-    end
-  end
-end
-
-Liquid::Template.register_tag('lightbox', Jekyll::LightboxTag)

### 2013-10-22 _knitr-posts/2013-10-21-panel-letters.Rmd
Add panel letters post draft

---
layout: post
title: Labelling Plot Panels
subtitle: Strategies for placing letters in figure panels.
---

Cole Monnahan writes:

> I've got a multipanel plot and simply want to add "(a)", "(b)", etc. to it.
> I've always cheated and used legend:
>
> ```
> add.label <- function(label, ...) legend("topleft", legend=" ", title=label,
> bty='n', ...)
> ```
>
> which works great for a topleft position. But for bottomright it isn't
> right. I can adapt it to:
>
> ```
> add.label2 <- function(label, position="topleft", ...) legend(x=position,
> legend=label, bty='n', ...)
> ```
>
> and that works great, but then doesn't work for topleft!
>
> There has got to be a non-hack way to do this simple (and common) task.

I'd never thought of using `legend()` for that purpose, but that's a neat
hack.

I tend to use `mtext()` for quick cases where the labels are all the same
width. For example if they're single letters, then you can align the labels
like this:

```{r panel_letters1, fig.width=6, fig.height=3}
par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
for(i in 1:2) {
  plot(1)
  mtext(paste0("(", letters[i], ")"), side = 3, adj = 0.05, 
    line = -1.3)
}
```

But the labels won't align if they are different widths:

```{r panel_letters2, fig.width=6, fig.height=3}
labels <- c("Short label", "A longer unaligned panel label")
par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
for(i in 1:2) {
  plot(1)
  mtext(labels[i], side = 3, adj = 0.05, line = -1.3)
}
```

And we can use `text()` combined with `par("usr")` to align these labels.
`par("usr")` gives us the coordinates of the plotting region in the order `x1,
x2, y1, y2`. Therefore, we can do the following to place the labels 2% over
from the left and 7% down from the top:

```{r panel_letters3, fig.width=6, fig.height=3}
labels <- c("Short label", "A longer panel label aligned")
par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
for(i in 1:2) {
  plot(1)
  u <- par("usr")
  x <- u[1] + 0.02 * (u[2] - u[1])
  y <- u[4] - 0.07 * (u[4] - u[3])
  text(x, y, labels[i], pos = 4)
}
```

And we could turn that into a little function:

```{r}
#' @param xfrac The fraction over from the left side.
#' @param yfrac The fraction down from the top.
#' @param label The text to label with.
#' @param ... Anything extra to pass to text(), e.g. cex, col.
add_label <- function(xfrac, yfrac, label, ...) {
  u <- par("usr")
  x <- u[1] + xfrac * (u[2] - u[1])
  y <- u[4] - yfrac * (u[4] - u[3])
  text(x, y, label, pos = 4, ...)
}
```

And then use it like this:

```{r panel_letters4, fig.width=6, fig.height=3}
labels <- c("Short", "A longer panel label")
par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
for(i in 1:2) {
  plot(1)
  add_label(0.02, 0.07, labels[i])
}
```

### 2013-10-22 _knitr-posts/2013-10-21-panel-letters.Rmd
Update panel letters and plyr posts

@@
 > which works great for a topleft position. But for bottomright it isn't
-> right. I can adapt it to:
->
-> ```
-> add.label2 <- function(label, position="topleft", ...) legend(x=position,
-> legend=label, bty='n', ...)
-> ```
->
-> and that works great, but then doesn't work for topleft!
+> right. I can adapt, and that works great, but then it doesn't
+> work for topleft!
 >
@@
 I'd never thought of using `legend()` for that purpose, but that's a neat
-hack.
+hack. If you wanted to stick with `legend()`, you could just
+pass the position to the function:
+
+```{r panel_letters5, fig.width=6, fig.height=3}
+add_label_legend <- function(pos = "topleft", label, ...) {
+  legend(pos, label, bty = "n", ...)
+}
+
+par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
+for(i in 1:2) {
+  plot(1)
+  add_label_legend("topleft", paste0("(", letters[i], ")"))
+  add_label_legend("topright", paste0("(", letters[i], ")"))
+  add_label_legend("bottomright", paste0("(", letters[i],
+      ")"))
+  add_label_legend("bottomleft", paste0("(", letters[i], ")"))
+}
+```
+
+But, you don't get a lot of control over precisely how the
+label is positioned.
 
@@
 
-And we can use `text()` combined with `par("usr")` to align these labels.
+You can use `text()` combined with `par("usr")` to align these labels.
 `par("usr")` gives us the coordinates of the plotting region in the order `x1,

### 2013-10-22 _knitr-posts/2013-10-21-panel-letters.Rmd
Add "R" to panel post
Add pos as argument to function

@@
 layout: post
-title: Labelling Plot Panels
+title: Labelling Panels in R
 subtitle: Strategies for placing letters in figure panels.
@@
 #' @param label The text to label with.
+#' @param pos Position to pass to text()
 #' @param ... Anything extra to pass to text(), e.g. cex, col.
-add_label <- function(xfrac, yfrac, label, ...) {
+add_label <- function(xfrac, yfrac, label, pos = 4, ...) {
   u <- par("usr")
@@
   y <- u[4] - yfrac * (u[4] - u[3])
-  text(x, y, label, pos = 4, ...)
+  text(x, y, label, pos = pos, ...)
 }

### 2013-12-02 _knitr-posts/2013-12-01-plyr.Rmd
Add revised plyr intro post

--- 2013-10-20 _knitr-posts/2013-10-20-plyr.Rmd
@@
 ---
+
+*Along with my [earlier post](http://localhost:4000/2013/10/19/reshape.html) on the `reshape2` package, I will continue to post my course notes from Data Wrangling and Visualization in R, a graduate-level course I co-taught last semester at Simon Fraser University.*
 
@@
 
-“`plyr` is a set of tools that solves a common set of problems: you need
+"`plyr` is a set of tools that solves a common set of problems: you need
 to break a big problem down into manageable pieces, operate on each
-piece and then put all the pieces back together. It’s already possible
+piece and then put all the pieces back together. It's already possible
 to do this with split and the apply functions, but `plyr` just makes it
-all a bit easier…”
+all a bit easier..."
 
 This is a very quick introduction to `plyr`. For more details see Hadley
-Wickham’s introductory guide [The split-apply-combine strategy for data
-analysis](http://www.jstatsoft.org/v40/i01) (2011, Journal of
-Statistical Software, Vol 40). There’s quite a bit of discussion online
-in general, and especially on
+Wickham's introductory guide [The split-apply-combine strategy for data
+analysis](http://www.jstatsoft.org/v40/i01). There's quite a bit of discussion
+online in general, and especially on
 [stackoverflow.com](http://stackoverflow.com/questions/tagged/plyr).
@@
 
-1.  The code is cleaner (once you’re familiar with the concept). The
+1.  The code is cleaner (once you're familiar with the concept). The
     code can be easier to code and read, and less error prone because you
-    don’t have to deal with subsetting and you don’t have to deal with saving
+    don't have to deal with subsetting and you don't have to deal with saving
     your results.
@@
 
-The basic format is 2 letters followed by `ply()`. The first letter
+The basic format is two letters followed by `ply()`. The first letter
 refers to the format in and the second to the format out.
 
-The 3 main letters are:
+The three main letters are:
 
@@
 
-I’ve ignored some less common format options:
+I've ignored some less common format options:
 
@@
 functions with control over the input and output formats. Some of the
-functionality can be duplicated with base R functions (but with less
-consistent syntax). Also, few R `apply` functions work directly with
+functionality can be duplicated with base `R` functions (but with less
+consistent syntax). Also, few `R` `apply` functions work directly with
 data frames as input and output and data frames are a common object
@@
 
-Base R `apply` functions (from a presentation given by Hadley):
+Base `R` `apply` functions (from a presentation given by Hadley Wickham):
 
@@
 
-Let’s take a simple example. We’ll take a data frame, split it up by
+Let's take a simple example. We'll take a data frame, split it up by
 `year`, calculate the coefficient of variation of the `count`, and
-return a data frame. This could easily be done on one line, but I’m
+return a data frame. This could easily be done on one line, but I'm
 expanding it here to show the format a more complex function could take.
@@
 
-It is often convenient to use these functions within `plyr`. `transform`
-acts as it would normally as the base R function and modifies an
-existing data frame. `summarise` creates a new condensed data
+It is often convenient to use these functions within one of the `**ply`
+functions. `transform` acts as it would normally as the base `R` function and
+modifies an existing data frame. `summarise` creates a new condensed data
 frame.
@@
 Bonus function: `mutate`. `mutate` works like `transform` but lets you
-build on columns you build.
+build on columns.
 
@@
 You can use `plyr` to plot data by throwing away the output with an
-underscore (`_`). This is a bit cleaner than a for loop since you don’t
+underscore (`_`). This is a bit cleaner than a for loop since you don't
 have to subset the data manually.
@@
 
-`plyr` can be slow — particularly if you are working with very large
+`plyr` can be slow --- particularly if you are working with very large
 datasets that involve a lot of subsetting. Hadley is working on this and
 an in-development version of `plyr`, `dplyr`, can run much faster (<https://github.com/hadley/dplyr>).
-However, it’s important to remember that typically the speed that you
+However, it's important to remember that typically the speed that you
 can write code and understand it later is the rate-limiting step.
@@
 
-Use a base R `apply` function:
+Use a base `R` `apply` function:
 

### 2013-12-10 _knitr-posts/2013-12-01-plyr.Rmd
Fix link in plyr post

@@
 
-*Along with my [earlier post](http://localhost:4000/2013/10/19/reshape.html) on the `reshape2` package, I will continue to post my course notes from Data Wrangling and Visualization in R, a graduate-level course I co-taught last semester at Simon Fraser University.*
+*Along with my [earlier post](http://seananderson.ca/2013/10/19/reshape.html) on the `reshape2` package, I will continue to post my course notes from Data Wrangling and Visualization in R, a graduate-level course I co-taught last semester at Simon Fraser University.*
 

### 2014-04-08 _knitr-posts/2014-04-08-gamma-glms.Rmd
Add Gamma GLM post

---
layout: post
title: Fitting Gamma GLMs Multiple Ways
subtitle: Understanding GLMs through simulation
published: true
---

```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, autodep=TRUE, tidy=FALSE)
```

A [Gamma error distribution](http://en.wikipedia.org/wiki/Gamma_distribution) with a log link is a common family to fit [GLMs](http://en.wikipedia.org/wiki/Generalized_linear_model) with in ecology. It works well for positive-only data with positively-skewed errors. The Gamma distribution is flexible and can mimic, among other shapes, a [log-normal](http://en.wikipedia.org/wiki/Log-normal_distribution) shape. The log link can represent an underlying multiplicate process, which is common in ecology.

Here, I'll fit a GLM with Gamma errors and a log link in four different ways. (1) With the built-in `glm()` function in `R`, (2) by optimizing our own likelihood function, (3) by the MCMC Gibbs sampler with [JAGS](http://mcmc-jags.sourceforge.net/), and (4) by the MCMC No U-Turn Sampler in [Stan](http://mc-stan.org/) (the shiny new Bayesian toolbox toy).

I wrote this code for myself to make sure I understood what was going on during the fitting process, but I thought I'd share it here. When I'm learning a new method, I often find it useful to see the same problem solved using a method I know along with a method I'm learning.

There are [multiple ways](http://stat.ethz.ch/R-manual/R-patched/library/stats/html/GammaDist.html) to parameterize the Gamma distribution in `R`. Here I've chosen to use the "rate" and "shape" parameters to match how it is parameterized in JAGS.

We'll start by simulating some data:

```{r}
set.seed(999)
N <- 100
x <- runif(N, -1, 1)
a <- 0.5
b <- 1.2
y_true <- exp(a + b * x)
shape <- 10
y <- rgamma(N, rate = shape / y_true, shape = shape)
```

Let's look at what we just simulated:

```{r plot-gamma-sim-data, fig.width=4, fig.height=4}
plot(x, y)
lines(sort(x), y_true[order(x)])
```

First, we'll fit a model to our data with `glm()` to make sure we can recover the parameters underlying our simulated data:

```{r, message=FALSE}
m_glm <- glm(y ~ x, family = Gamma(link = "log"))
m_glm_ci <- confint(m_glm)
coef(m_glm)
```

That's pretty close to our "true" simulated values. We'll use these values as a reference for our methods below.

To make sure we understand what's going on, we'll write our own likelihood function and optimize the fit to the data ourselves with Ben Bolker's `mle2()` function. You'll need the `bbmle` library.

```{r}
library("bbmle")
nll_gamma <- function(a, b, logshape) {
  linear_predictor <- a + b * x
  # rate = shape / mean:
  rate <- exp(logshape) / exp(linear_predictor)
  # sum of negative log likelihoods:
  -sum(dgamma(y, rate = rate, shape = exp(logshape),
      log = TRUE))
}
(m_mle2 <- bbmle::mle2(nll_gamma,
  start = list(a = rnorm(1), b = rnorm(1),
    logshape = rlnorm(1))))
```

One of the niceties built into `mle2()` is the ability to estimate profile likelihood intervals with a `confint()` function:

```{r, message=FALSE}
(m_mle2_ci <- confint(m_mle2))
```

Now, we'll use the same approach to fit our model, but this time via MCMC with JAGS:

```{r, message=FALSE, results="hide"}
library("R2jags")
m <- function() {
  # Priors:
  a ~ dnorm(0, 0.0001) # mean, precision = N(0, 10^4)
  b ~ dnorm(0, 0.0001)
  shape ~ dunif(0, 100)

  # Likelihood data model:
  for (i in 1:N) {
    linear_predictor[i] <- a + b * x[i]
    # dgamma(shape, rate) in JAGS:
    y[i] ~ dgamma(shape, shape / exp(linear_predictor[i]))
  }
}

m_jags <- R2jags::jags(data = list(N = N, y = y, x = x),
  parameters.to.save = c("a", "b", "shape"), model.file = m)
```

It's important to inspect the output from any MCMC simulation to make sure the chains are mixing well and have converged. Rest assured that I have. Let's extract the mean estimates:

```{r}
m_jags_sims <- m_jags$BUGSoutput$sims.list
lapply(m_jags_sims, mean)[c(1, 2, 4)]
```

And plot, as an example, the density of MCMC samples for the parameter *b*:

```{r jags-gamma-density-b, fig.width=4, fig.height=3.5}
plot(density(m_jags_sims$b), main = "")
```

We'll also calculate 95% credible intervals. We'll use HPD intervals ([highest probability denity](http://en.wikipedia.org/wiki/Credible_interval)):

```{r}
(m_jags_ci <- coda::HPDinterval(coda::as.mcmc(m_jags_sims$b)))
```

Now with Stan:

```{r, message=FALSE, results="hide"}
library(rstan)
stan_code <- "
data {
  int<lower=0> N;
  vector[N] x;
  vector[N] y;
}
parameters {
  real a;
  real b;
  real<lower=0.001,upper=100> shape;
}
model {
  a ~ normal(0,100);
  b ~ normal(0,100);
  for(i in 1:N)
    y[i] ~ gamma(shape, (shape / exp(a + b * x[i])));
}
"
m_stan <- stan(model_code = stan_code,
    data = list(x = x, y = y, N = N))
```

We'll look at the means of each estimated parameter:

```{r}
lapply(extract(m_stan), mean)[c(1, 2, 3)]
```

And we'll extract the 95% HPD interval again:

```{r}

stan_b <- extract(m_stan)$b
(m_stan_ci <- coda::HPDinterval(coda::as.mcmc(as.vector(stan_b))))
```

Finally, we'll plot the four fits to show that they are all qualitatively the same:

```{r compare-gamma-fits, fig.width=7.5, fig.height=3.6}
par(mfrow = c(1, 2), las = 1, mgp = c(2, 0.5, 0),
  tcl = -0.15, mar = c(4, 3, 0, 0.5), oma = c(0, 0, 1, 1))

# left panel:
plot(x, y)
lines(x[order(x)], y_true[order(x)], col = "#00000050",
  lty = 2, lwd = 2)
X <- seq(min(x), max(x), length.out = 100)
lines(X, exp(coef(m_glm)[1] + coef(m_glm)[2] * X),
  col = "red")
lines(X, exp(coef(m_mle2)[1] + coef(m_mle2)[2] * X),
  col = "blue")
lines(X, exp(mean(m_jags_sims$a) + mean(m_jags_sims$b) * X),
  col = "green")
lines(X, exp(get_posterior_mean(m_stan)[1,5] +
    get_posterior_mean(m_stan)[2,5] * X), col = "black")

# right panel:
plot(1, 1, xlim = c(1, 4), ylim = c(0.5, 2), type = "n",
  xaxt = "n", xlab = "", ylab = "Estimate of b")
abline(h = b, col = "#00000050", lwd = 2, lty = 2)
points(1, coef(m_glm)[2])
segments(1, m_glm_ci[2, 1], 1, m_glm_ci[2, 2])
points(2, coef(m_mle2)[2])
segments(2, m_mle2_ci[2, 1], 2, m_mle2_ci[2, 2])
points(3, mean(m_jags_sims$b))
segments(3, m_jags_ci[1, 1], 3, m_jags_ci[1, 2])
points(4, get_posterior_mean(m_stan)["b", "mean-all chains"])
segments(4, m_stan_ci[1, 1], 4, m_stan_ci[1, 2])

axis(1, at = c(1, 2, 3, 4), labels = c("glm()", "mle2()",
    "JAGS", "Stan"))
```

# Take home messages

As expected, all four methods give us a similar result (although some with different interpretations — Bayesian vs. frequentist).

Simulation, as always, is a great way to understand statistical models.

Although we call it a "log link", if we're working with the Gamma distribution directly, we exponentiate the linear predictor instead of logging the data. This ensures that we don't propose negative mean values to the Gamma distribution.

There are multiple ways to parameterize the Gamma distribution, so it's important to pay attention when moving between languages and functions.

As with many optimization exercises, we can force a term (here `shape`) to be positive by fitting in log-space. Hence, using `logspace` and exponentiating within the optimization function for some of the examples. This creates a smooth likelihood or probability surface near the boundaries that optimization algorithms can better navigate.

### 2014-04-09 _knitr-posts/2014-04-08-gamma-glms.Rmd
Fix typos

@@
 
-We'll also calculate 95% credible intervals. We'll use HPD intervals ([highest probability denity](http://en.wikipedia.org/wiki/Credible_interval)):
+We'll also calculate 95% credible intervals. We'll use HPD intervals ([highest probability density](http://en.wikipedia.org/wiki/Credible_interval)):
 
@@
 As with many optimization exercises, we can force a term (here `shape`) to be positive by fitting in log-space.
-Hence, using `logspace` and exponentiating within the optimization function for some of the examples.
+Hence, using `logshape` and exponentiating within the optimization function for some of the examples.
 This creates a smooth likelihood or probability surface near the boundaries that optimization algorithms can better navigate.

### 2014-05-20 _knitr-posts/2014-05-14-centering-interactions.Rmd
Add two posts; centering one is still draft

---
layout: post
title: Centering, Interactions, and Model Averaging
subtitle: Why centering factor predictors can be important
published: false
---

I've encountered a number of people wondering about centering and scaling predictors in the context of model averaging. Here, centering refers to subtracting the mean and scaling refers to dividing by some measure of variability ([usually one or two standard deviations](http://www.stat.columbia.edu/~gelman/research/published/standardizing7.pdf)).

First of all, if you haven't already, go read the paper [Simple means to improve the interpretability of
regression coefficients](http://doi.org/10.1111/j.2041-210X.2010.00012.x) by Holger Schielzeth. This paper is a wealth of useful information. The code I've included below is just a simulated example to demonstrate an aspect of this paper.

Bottom line, if you are going to average coefficients across models with and without interactions you're going to get gibberish unless you center your predictors by subtracting the mean. This is true across continuous, binary, and multi-level factor predictors, although setting up centered factor predictors can take an extra step.

Whether or not averaging coefficient values across models is a good idea is a separate topic that I won't get into here. Just realize that subscribing to the church of coefficient averaging is not the only way and there are also good arguments for three alternatives: maintaining multiple models, model selection (e.g. as often suggested in [Zuur's books](http://www.highstat.com/books.htm)), or, assuming reasonable power, fitting the most complicated model you can justify and interpret and drawing inference from that model (e.g. [Andrew Gelman](http://andrewgelman.com/2004/12/10/against_parsimo/)). After all, removing a predictor is equivalent to assuming the coefficient for that predictor is precisely zero, which is a strong assumption in itself. (This last philosophy often works best in Bayesian context where weakly informative priors are a possibility.)

First, we'll simulate some data. `b0` is an intercept, `b1` a slope on a continuous predictor, `b2` a binary factor coefficient, and `b1.2` an interaction coefficient between `b1` and `b2`.

```{r}
set.seed(999)
b0 <- 1.4 # intercept
b1 <- 0.2 # continuous slope
b2 <- 1.7 # factor level 1 coefficient
b1.2 <- 0.5 # interaction between b1 and b2
sigma <- 2.0 # residual standard deviation
N <- 25 # number of data points

x1 <- runif(N, 0, 20) # continuous predictor data
x2 <- rbinom(N, size = 1, prob = 0.4) # binary predictor data

# generate response data:
y <- rnorm(N, mean = b0 +
  b1 * x1 +
  b2 * x2 +
  x1 * x2 * b1.2,
  sd = sigma)
dat <- data.frame(x1, x2, y)
head(dat)
```

Let's look at the data we created:

```{r cent1, fig.width=6, fig.height=4}
library(ggplot2)
ggplot(dat, aes(x1, y, colour = as.factor(x2))) + geom_point()
```

Now, we'll fit a model with and without an interaction and look at the coefficients:

```{r}
m <- lm(y ~ x1 * x2, data = dat)
m_no_inter <- lm(y ~ x1 + x2, data = dat)
round(coef(m), 2)
round(coef(m_no_inter), 2)
```

Notice how the main effects (everything except the interaction) change dramatically when the interaction is removed. This is because when the interaction is included the main effects are relevant to when the predictors are equal to 0. I.e. `x1 = 0` and the binary predictor `x2` is at its reference (0) level. But, when the interaction is excluded the main effects are relevant across all the values of the other predictors, or equivalently, their mean values. So, if we center, the main effects will represent the same thing in both cases:

```{r}
dat$x2_cent <- dat$x2 - mean(dat$x2)
dat$x1_cent <- dat$x1 - mean(dat$x1)
m_center <- lm(y ~ x1_cent * x2_cent, data = dat)
m_center_no_inter <- lm(y ~ x1_cent + x2_cent,
  data = dat)
round(coef(m_center), 2)
round(coef(m_center_no_inter), 2)
```

Notice that the intercept, `x1`, and `x2` coefficient estimates are now similar regardless of whether the interaction is included. Now, because we've centered the predictors, the predictors equal zero at their mean. So, the main effects are estimating approximately the same thing regardless of whether we include the interaction. In other words, adding the interaction adds more predictive information but doesn't modify the meaning of the main effects.

# A 3-level factor example

Now, let's repeat the above with a three-level factor predictor to make sure it's clear how it works.

```{r}
set.seed(999)
b0 <- 1.4 # intercept
b1 <- 0.2 # continuous slope
b2 <- 1.7 # factor level 0-1 coefficient
b3 <- 2.9 # factor level 0-2 coefficient
b1.2 <- 0.5 # interaction between b1 and b2
b1.3 <- 0.9 # interaction between b1 and b3
sigma <- 2.0 # residual standard deviation
N <- 30 # number of data points

x1 <- runif(N, 0, 20) # continuous predictor data
# 3-factor predictor data:
f <- sample(c(1, 2, 3), N, replace = TRUE)

x2 <- ifelse(f == 2, 1, 0)
x3 <- ifelse(f == 3, 1, 0)

y <- rnorm(N, mean = b0 +
  b1 * x1 +
  b2 * x2 +
  b3 * x3 +
  x1 * x2 * b1.2 +
  x1 * x3 * b1.3,
  sd = sigma)

dat <- data.frame(x1, x2, x3, y,
  f = as.factor(f))
head(dat)
```

```{r cent2, fig.width=6, fig.height=4}
ggplot(dat, aes(x1, y, colour = f)) + geom_point()
```

Now we'll fit the model. First with `factor()` notation:

```{r}
m2 <- lm(y ~ x1 * f, data = dat)
round(coef(m2), 2)
```

And now with "dummy" variable notation. This will make centering in the next step easier:

```{r}
m2.1 <- lm(y ~ x1 * x2 + x1 * x3, data = dat)
round(coef(m2.1), 2)
```

Notice we get the same estimates.

Now, let's compare to a model without the interaction:

```{r}
m2.1_no_inter <- lm(y ~ x1 + x2 + x1 + x3, data = dat)
round(coef(m2.1), 2)
round(coef(m2.1_no_inter), 2)
```

Again, if we model averaged here across models with and without the interaction, we'd be getting gibberish.

Now, we'll fit the same model with centered predictors:

```{r}
dat$x1_cent <- dat$x1 - mean(dat$x1)
dat$x2_cent <- dat$x2 - mean(dat$x2)
dat$x3_cent <- dat$x3 - mean(dat$x3)
m2.2 <- lm(y ~ x1_cent * x2_cent + x1_cent * x3_cent, data = dat)
m2.2_no_inter <- lm(y ~ x1_cent + x2_cent + x3_cent, data = dat)
```

Again, notice that the main effects stay fairly consistent regardless of whether we include the interaction because they are estimated across the centered three-level factor:

```{r}
round(coef(m2.2), 2)
round(coef(m2.2_no_inter), 2)
```

You can see how this works across multiple model comparisons using the `MuMIn::dredge()` function on any one of the above models.

So, *if* you're going to average across models with and without interactions, centering the predictors is important. Even if you're not going to be averaging or comparing coefficient values across models with and without interactions, centering predictors can be useful for computational and interpretation reasons, but that's a separate topic.

# Plotting the predictions

One more thing. Say we wanted to plot the predicted response from our model. This isn't quite as simple as normal once we've centered our predictors and coded the factor levels as centered dummy variable comparisons.

We'll make predictions at each of the 3 levels. Note how we create these levels through the 3 valid combinations of our dummy variables. We can't just use `expand.grid()` and get all combinations of our dummy variables.

```{r}
x1_cent <- seq(-10, 10, length.out = 100)
newdata1 <- data.frame(x1_cent = x1_cent,
  x2_cent = min(dat$x2_cent), x3_cent = min(dat$x3_cent))
newdata2 <- data.frame(x1_cent = x1_cent,
  x2_cent = max(dat$x2_cent), x3_cent = min(dat$x3_cent))
newdata3 <- data.frame(x1_cent = x1_cent,
  x2_cent = min(dat$x2_cent), x3_cent = max(dat$x3_cent))
newdata <- rbind(newdata1, newdata2, newdata3)
```

Now we make our predictions and plot the predictions:

```{r cent3, fig.width=6, fig.height=5.5}
newdata$pred <- predict(m2.2, newdata = newdata)
with(dat, plot(x1_cent + mean(dat$x1), y, col = f))
plyr::d_ply(newdata, "x2_cent", function(x) {
  with(x, lines(x1_cent + mean(dat$x1), pred, col = "grey"))
  })
```

```{r, echo=FALSE}
# # Adjusting factor level references
#
# combine_coefs <- function(model, ref_coef, comp_coef) {
#   if(class(model)[1] == "lme")
#     x <- as.data.frame(summary(model)$tTable)
#   if(class(model)[1] == "glm" | class(model)[1] == "lm") {
#     x <- as.data.frame(summary(model)$coef)
#     names(x)[c(1, 2)] <- c("Value", "Std.Error")
#   }
#   x$coef_name <- row.names(x)
#   row.names(x) <- NULL
#   base_coef <- x[x$coef_name == ref_coef, "Value"]
#   base_se <- x[x$coef_name == ref_coef, "Std.Error"]
#   comparison_coef <- x[x$coef_name == comp_coef, "Value"]
#   comparison_se <- x[x$coef_name == comp_coef, "Std.Error"]
#   combined_coef <- base_coef + comparison_coef
#   combined_se <- sqrt(comparison_se^2 - base_se^2)
#   data.frame(coef = combined_coef, se = combined_se)
# }
#
# x2_coef <- combine_coefs(m2.2, ref_coef = "x1_cent", comp_coef = "x1_cent:x2_cent", unit_value = min(dat$x2_cent))
# x3_coef <- combine_coefs(m2.2, ref_coef = "x1_cent", comp_coef = "x1_cent:x3_cent", unit_value = min(dat$x3_cent))
#
# newdata$pred <- predict(m2.2, newdata = newdata)
# with(dat, plot(x1_cent + mean(dat$x1), y, col = f))
# plyr::d_ply(newdata, "x2_cent", function(x) {
#   with(x, lines(x1_cent + mean(dat$x1), pred, col = "lightgrey"))
#   })
#
# # very roughly:
# # see coef(m2.2)
# abline(a = 3, b = 0.6684 + 0.567*-0.333 + 0.988 * -0.3) # 1st level
# abline(a = 3, b = 0.6684 + 0.567*0.667 + 0.988 * -0.3) # 2nd level
# abline(a = 3, b = 0.6684 + 0.567*-0.333 + 0.988 * 0.7) # 3rd level

```

### 2014-05-20 _knitr-posts/2014-05-18-gamma-hurdle.Rmd
Add two posts; centering one is still draft

---
layout: post
title: Gamma Hurdle Models
subtitle: Fitting and interpreting Gamma hurdle models
published: true
---

```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(fig.align='center', fig.pos="htpb", cache=FALSE, echo=TRUE,
message=FALSE, autodep=TRUE, fig.width=7, fig.height=5)
opts_chunk$set(warning=FALSE, tidy=FALSE, refresh=TRUE)
library(ggplot2)
theme_set(theme_bw())
```

# Zero inflation

Ecologists often run into a scenario where their response data have more zeros than expected if the process generating their data was purely from a standard probability distribution. For example, there might be more zeros than you'd expect from a negative binomial distribution with a given mean. [Two common methods for dealing with zero-inflated data](http://stats.stackexchange.com/questions/81457/what-is-the-difference-between-zero-inflated-and-hurdle-distributions-models) are:

1. Modelling a zero-inflation parameter that represents the probability a given zero comes from the main distribution (say the negative binomial distribution) or is an excess zero.

2. Modelling the zero and non-zero data with one model and then modelling the non-zero data with another. This is often called a "hurdle model".

So, zero-inflation models separate the zeros into "true" and "extra" categories.

Hurdle models model the zeros and non-zeros as two separate processes.

Zero-inflation models may be more elegant and informative if the same predictors are thought to contribute to the extra and real zeros.

Hurdle models can be useful in that they allow you to model the zeros and non-zeros with different predictors or different roles of the same predictors. Maybe one process leads to the zero/non-zero data and another leads to the non-zero magnitude.

For count data (probably the most typical situation where zero inflation is a problem), there are already a variety of R packages to fit either zero-inflated or hurdle models. E.g. [pscl](http://cran.r-project.org/web/packages/pscl/index.html) for models without random effects, [glmmADMB](http://glmmadmb.r-forge.r-project.org/) for models with or without random effects, and [MCMCglmm](http://cran.r-project.org/web/packages/MCMCglmm/index.html) for Bayesian models with or without random effects.

# Zero-inflated continuous data

Sometimes ecologists end up with a response dataset that is positive and continuous but has zero inflation. (Technically, any number of zeros is "zero inflation" for probability distributions that only represent non-zero positive values.) There are probably many scenarios that can result in this kind of data. Sometimes the "zeros" may not really be zeros but just really small values that weren't measured. But there are also plausible scenarios where the zeros are real. For example, if the response variable is biomass at a given site, it's possible there were just no individuals at some sites. If you want to model the biomass with a distribution such as the log-normal or Gamma, these distributions don't allow for zero values. One solution is to model the zeros separately from the non-zeros in a binomial-Gamma hurdle model.

This post illustrates a small simulated example of one of these hurdle models where we estimate an intercept only. I describe how to fit the model, interpret the coefficients, and generate predictions with confidence intervals.

Note that the methods as illustrated here don't apply exactly to count-based hurdle models since count-based models require a truncated count distribution. I.e., they need to account for the fact that the negative binomial or Poisson distribution can represent zeros too, but these zeros were already accounted for in the binomial model. Since the Gamma distribution can not represent zero values we don't need to worry about that. Count hurdle models can be fitted with the packages mentioned above.

# Fitting the models

We'll start by simulating some data and plotting it. First we'll simulate a binomial process (0 or 1) with a probability of a non-zero value (1) of 0.7. Then we'll simulate a Gamma process with a shape (mean) of 0.4.

```{r hurdle1}
set.seed(1)
x <- 1:150
y <- rbinom(length(x), size = 1, prob = 0.7)
y <- y * rgamma(length(x), shape = 0.4)
non_zero <- ifelse(y > 0, 1, 0)
d <- data.frame(x, y, non_zero)
head(d)
library(ggplot2)
p <- ggplot(d, aes(x, y, colour = as.factor(non_zero))) + geom_point()
print(p)
```

Now we'll fit a logistic regression to predict the probability of a non-zero value and a Gamma GLM with a log link to predict the mean of the non-zero data.

```{r}
m1 <- glm(non_zero ~ 1, data = d, family = binomial(link = logit))
m2 <- glm(y ~ 1, data = subset(d, non_zero == 1), family = Gamma(link = log))
```

We'll extract the coefficients and show the 95% confidence intervals (derived from profile likelihoods). Note that the Gamma coefficients come out on a log-scale and we'll exponentiate them as we go. The logistic regression coefficients come out on the logit scale and we'll inverse that link as we go as well.

```{r, message=FALSE}
(bin_coef <- plogis(coef(m1)[[1]])) # close to prob = 0.7 as specified
(gamma_coef <- exp(coef(m2)[[1]])) # close to shape = 0.4 as specified
(plogis(confint(m1)))
(exp(confint(m2)))
```

So, we've managed to recover our simulated means. The logistic model indicates that the mean probability of observing a non-zero value is about 0.7. The Gamma model indicates that given we have observed a non-zero value, the mean of our response value `y` is about 0.4.

# Predictions

We could plot these coefficients along with their confidence intervals. We could also plot predictions for these two separate processes.
For example, we'll plot the predictions along with confidence intervals for these two models here. Note that because we only modelled an intercept, we'll get the same predictions regardless of the data to be predicted on.

```{r hurdle2}
pred1 <- predict(m1, se = TRUE, type = "link")
pred2 <- predict(m2, se = TRUE, type = "link")

# Zero/non-zero binomial model:
ggplot(d, aes(x, non_zero)) + geom_point() +
  geom_hline(yintercept = plogis(pred1$fit)) +
  geom_hline(yintercept = plogis(pred1$fit +
      c(1.96, -1.96) * pred1$se.fit),
    colour = "darkgrey")

# Non-zero Gamma model:
ggplot(subset(d, non_zero == 1),
  aes(x, y)) + geom_point() +
  geom_hline(yintercept = exp(pred2$fit)) +
  geom_hline(yintercept = exp(pred2$fit +
      c(1.96, -1.96) * pred2$se.fit),
    colour = "darkgrey")
```

We can also check the predictions incorporating both the binomial and Gamma models. To do that, we take the two means, add them on a log scale and re-exponentiate them. We can check that the predicted mean (intercept) matches the arithmetic mean.

```{r hurdle4}
(pred <- exp(log(bin_coef) + log(gamma_coef)))
mean(d$y) # same
p + geom_hline(yintercept = pred)
```

# Prediction confidence intervals

Let's derive confidence intervals on our predictions. One way to do that is using a non-parametric bootstrap method. To do that we'll define a little function that carries out our hurdle model as above. `boot` requires that the function's first argument be the data and the second argument be a vector of row indices for each bootstrap replicate. The function should return the statistic we want to bootstrap (here the prediction).

We'll run 500 replicates. In a real analysis you'd want to make sure you'd run enough replicates that the estimates had stabilized. You could test this by increasing the number of replicates until you start getting the same confidence intervals (to whatever level of precision you are interested in) regardless of whether you add more replicates.

We'll use the bias-corrected confidence intervals. Read the documentation for `boot.ci` and the recommended references in the documentation for details.

```{r hurdle3}
hurdle_fn <- function(data, i) {
  dat_boot <- data[i, ]
  m1 <- glm(non_zero ~ 1, data = dat_boot,
    family = binomial(link = logit))
  m2 <- glm(y ~ 1, data = subset(dat_boot, non_zero == 1),
    family = Gamma(link = log))
  bin_coef <- plogis(coef(m1)[[1]])
  gamma_coef <- exp(coef(m2)[[1]])
  exp(log(bin_coef) + log(gamma_coef))
}

library(boot)
b <- boot(d, hurdle_fn, R = 500)
b.ci <- boot.ci(b, type = "bca")
print(b.ci)
p + geom_hline(yintercept = pred) +
  geom_hline(yintercept = b.ci$bca[c(4:5)],
    colour = "darkgrey")
```

### 2014-06-03 _knitr-posts/2014-05-30-centering-interactions.Rmd
Add renamed and updated centering post

--- 2014-05-20 _knitr-posts/2014-05-14-centering-interactions.Rmd
@@
 layout: post
-title: Centering, Interactions, and Model Averaging
+title: Centering factor variables with interactions
 subtitle: Why centering factor predictors can be important
@@
 First of all, if you haven't already, go read the paper [Simple means to improve the interpretability of
-regression coefficients](http://doi.org/10.1111/j.2041-210X.2010.00012.x) by Holger Schielzeth.
-This paper is a wealth of useful information.
-The code I've included below is just a simulated example to demonstrate an aspect of this paper.
+regression coefficients](http://doi.org/10.1111/j.2041-210X.2010.00012.x) by Holger Schielzeth. This paper is a wealth of useful information.
 
@@
 
-Now we make our predictions and plot the predictions:
+Now we make our predictions and plot the predictions. We'll undo the centering of the `x1` variable here by adding the mean of the original data:
 
@@
 
-```{r, echo=FALSE}
-# # Adjusting factor level references
-#
-# combine_coefs <- function(model, ref_coef, comp_coef) {
-#   if(class(model)[1] == "lme")
-#     x <- as.data.frame(summary(model)$tTable)
-#   if(class(model)[1] == "glm" | class(model)[1] == "lm") {
-#     x <- as.data.frame(summary(model)$coef)
-#     names(x)[c(1, 2)] <- c("Value", "Std.Error")
-#   }
-#   x$coef_name <- row.names(x)
-#   row.names(x) <- NULL
-#   base_coef <- x[x$coef_name == ref_coef, "Value"]
-#   base_se <- x[x$coef_name == ref_coef, "Std.Error"]
-#   comparison_coef <- x[x$coef_name == comp_coef, "Value"]
-#   comparison_se <- x[x$coef_name == comp_coef, "Std.Error"]
-#   combined_coef <- base_coef + comparison_coef
-#   combined_se <- sqrt(comparison_se^2 - base_se^2)
-#   data.frame(coef = combined_coef, se = combined_se)
-# }
-#
-# x2_coef <- combine_coefs(m2.2, ref_coef = "x1_cent", comp_coef = "x1_cent:x2_cent", unit_value = min(dat$x2_cent))
-# x3_coef <- combine_coefs(m2.2, ref_coef = "x1_cent", comp_coef = "x1_cent:x3_cent", unit_value = min(dat$x3_cent))
-#
-# newdata$pred <- predict(m2.2, newdata = newdata)
-# with(dat, plot(x1_cent + mean(dat$x1), y, col = f))
-# plyr::d_ply(newdata, "x2_cent", function(x) {
-#   with(x, lines(x1_cent + mean(dat$x1), pred, col = "lightgrey"))
-#   })
-#
-# # very roughly:
-# # see coef(m2.2)
-# abline(a = 3, b = 0.6684 + 0.567*-0.333 + 0.988 * -0.3) # 1st level
-# abline(a = 3, b = 0.6684 + 0.567*0.667 + 0.988 * -0.3) # 2nd level
-# abline(a = 3, b = 0.6684 + 0.567*-0.333 + 0.988 * 0.7) # 3rd level
+# Calculating slopes and standard errors at interaction levels
+
+Currently, our factor variable `f` has three levels.
+Our estimate `b1` represents the slope across the mean of the factors.
+However, we might want to present the slope with confidence intervals at each level of factor `f`.
+
+Calculating the slope coefficients is fairly easy but the standard errors require a bit more math.
+The trick is to correctly combine the coefficient variances (or equivalently the squared standard errors) with the covariances between coefficients.
+
+The way we combine the standard errors is the same way we combine the standard deviation of any normal distributions. For the case:
 
 ```
+N1 ~ N(mu1, sigma1)
+N2 ~ N(mu2, sigma2)
+```
+
+where `mu` represents the mean, `sigma` represents the standard deviation, and `N` represents a normal distribution.
+
+We can calculate the combination of these two distributions multiplied by the constants `a` and `b` (which we'll call `N3`) as:
+
+```
+N3 = a * N1 + b * N2
+N3 ~ N(mu3, sigma3)
+mu3 = a * mu1 + b * mu2
+sigma3 = sqrt(a^2 * sigma1^2 + b^2 * sigma2^2 + 2 * ab * cov(ab))
+```
+
+First, to illustrate combining the coefficient means, here's an example plotting the predictions by specifying the intercept and slopes.
+The first line in each of these `abline` function calls represents the intercepts and each factor level and the second line represents the slopes.
+Notice how we get these slopes by turning the two indicator columns to low or high values.
+
+```{r cent4, fig.width=6, fig.height=5.5}
+with(dat, plot(x1_cent, y, col = f))
+b <- coef(m2.2)
+se <- summary(m2.2)$coef[,2]
+# The intercept and slope for the first level of the factor f:
+abline(a = b[1] + b[3]*-0.333 + b[4] * -0.3, 
+  b = b[2] + b[5]*-0.333 + b[6] * -0.3, col = 1)
+# And the second level:
+abline(a = b[1] + b[3]*0.667 + b[4] * -0.3, 
+  b = b[2] + b[5]*0.667 + b[6] * -0.3, col = 2)
+# And the third:
+abline(a = b[1] + b[3]*-0.333 + b[4] * 0.7, 
+  b = b[2] + b[5]*-0.333 + b[6] * 0.7, col = 3)
+```
+
+Before we calculate the standard errors on the slope at the three factor levels let's return to a more straightforward parameterization so we know what we're aiming for.
+I'll re-write the model here for clarity:
+
+```{r, cent5, fig.width=6, fig.height=5.5}
+m2 <- lm(y ~ x1 * f, data = dat)
+b <- coef(m2)
+se <- summary(m2)$coef[,2]
+
+# Slopes:
+with(dat, plot(x1, y, col = f))
+abline(a = b[1], b = b[2], col = 1)
+abline(a = b[1] + b[3], b = b[2] + b[5], col = 2)
+abline(a = b[1] + b[4], b = b[2] + b[6], col = 3)
+
+# SEs on those slopes:
+# (as in Schielzeth 2010 Methods Ecol. Evol.)
+se[2]
+sqrt(se[5]^2 - se[2]^2)
+sqrt(se[6]^2 - se[2]^2)
+```
+
+Now, let's derive those same standard errors for the centered version. We'll start be extracting the standard errors and covariances of the coefficients. Then we'll combine them:
+
+```{r}
+se <- summary(m2.2)$coef[,2]
+x1x2_low <- min(dat$x2_cent)
+x1x2_high <- max(dat$x2_cent)
+x1x3_low <- min(dat$x3_cent)
+x1x3_high <- max(dat$x3_cent)
+vcov_x1_x1x2 <- vcov(m2.2)[5, 2]
+vcov_x1_x1x3 <- vcov(m2.2)[6, 2]
+vcov_x1x2_x1x3 <- vcov(m2.2)[6, 5]
+se_x1 <- se[2]
+se_x1x2 <- se[5]
+se_x1x3 <- se[6]
+```
+
+Standard error on the slope at the first factor level:
+
+```{r}
+as.numeric(
+sqrt(x1x2_low^2 * se_x1x2^2 + 
+     x1x3_low^2 * se_x1x3^2 +
+     se_x1^2 +
+     2 * x1x2_low * vcov_x1_x1x2 +
+     2 * x1x3_low * vcov_x1_x1x3 +
+     2 * x1x2_low * x1x3_low * vcov_x1x2_x1x3)
+)
+```
+
+This matches the version from the uncentered model:
+
+```{r}
+summary(m2)$coef[2,2] 
+```
+
+And the second level:
+
+```{r}
+as.numeric(
+sqrt(x1x2_high^2 * se_x1x2^2 + 
+     x1x3_low^2 * se_x1x3^2 +
+     se_x1^2 +
+     2 * x1x2_high * vcov_x1_x1x2 +
+     2 * x1x3_low * vcov_x1_x1x3 +
+     2 * x1x2_high * x1x3_low * vcov_x1x2_x1x3)
+)
+```
+
+And the third level:
+
+```{r}
+as.numeric(
+sqrt(x1x2_low^2 * se_x1x2^2 + 
+     x1x3_high^2 * se_x1x3^2 +
+     se_x1^2 +
+     2 * x1x2_low * vcov_x1_x1x2 +
+     2 * x1x3_high * vcov_x1_x1x3 +
+     2 * x1x2_low * x1x3_high * vcov_x1x2_x1x3)
+)
+```
+
+Again, comparing to the uncentered version, they match:
+
+```{r}
+se <- summary(m2)$coef[,2]
+sqrt(se[5]^2 - se[2]^2)
+sqrt(se[6]^2 - se[2]^2)
+```

### 2014-07-08 refs.R
Add very rough script to create bibliography

@@ -1,0 +1,95 @@
+library("bibtex")
+d <- read.bib("refs.bib")
+
+#str(x)
+
+#init <- list()
+
+ref <- list()
+
+for(i in seq_len(length(d))) {
+  name <- list()
+  x <- d[[i]]
+  # authors:
+  for(j in seq_len(length(x$author))) {
+    init <- paste0(paste(substr(x$author$given[[j]], 1, 1), collapse = "."), ".")
+    family <- x$author$family[[j]]
+    if(j == 1) {
+      name[[j]] <- paste0(family, ", ", init)
+      if(name[[j]] == "Anderson, S.C.") name[[j]] <- "**Anderson, S.C.**"
+    } else {
+      name[[j]] <- paste(init, family)
+      if(name[[j]] == "S.C. Anderson") name[[j]] <- "**S.C. Anderson**"
+    }
+  }
+  name0 <- paste0(paste0(name, collase = ""), collapse = ", ")
+
+  vol_num_p <- ""
+  if(!is.null(x$number) & !is.null(x$volume) & !is.null(x$pages)) {
+    vol_num_p <- paste0(x$volume, "(", x$number, "): ", x$pages)
+  }
+  # no number:
+  if(is.null(x$number) & !is.null(x$volume) & !is.null(x$pages)) {
+    vol_num_p <- paste0(x$volume, ": ", x$pages)
+  }
+  # no pages:
+  if(is.null(x$number) & !is.null(x$volume) & is.null(x$pages)) {
+    vol_num_p <- paste0(x$volume, "(", x$number, ")")
+  }
+  # no number or pages:
+  if(is.null(x$number) & !is.null(x$volume) & is.null(x$pages)) {
+    vol_num_p <- x$volume
+  }
+  # no vol, number, or pages:
+  if(is.null(x$number) & is.null(x$volume) & is.null(x$pages)) {
+    vol_num_p <- ""
+  }
+
+  if(vol_num_p != "" & vol_num_p != "In review" & vol_num_p != "In revision") {
+    main <- paste0(paste0(name0, ". ", x$title, ". ", x$journal, ". ", vol_num_p), ".")
+  } else {
+    main <-paste0(name0, ". ", x$title, ". ", vol_num_p, ".")
+  }
+
+  if(!is.null(x$doi)) {
+    main <- paste0(main, " <http://doi.org/", x$doi, ">")
+  }
+
[…]

### 2014-07-08 _knitr-posts/2014-06-12-centering-interactions.Rmd
Add draft version of interaction blog post

--- 2014-06-03 _knitr-posts/2014-05-30-centering-interactions.Rmd
@@
 layout: post
-title: Centering factor variables with interactions
-subtitle: Why centering factor predictors can be important
+title: Centering factor interactions
+subtitle: Why and how you must center interactions for model averaging
 published: false
@@
 Notice how the main effects (everything except the interaction) change dramatically when the interaction is removed.
-This is because when the interaction is included the main effects are relevant to when the predictors are equal to 0.
+This is because when the interaction is included the main effects are relevant to when the other predictors are equal to 0.
 I.e.
-`x1 = 0` and the binary predictor `x2` is at its reference (0) level.
-But, when the interaction is excluded the main effects are relevant across all the values of the other predictors, or equivalently, their mean values.
+`x1 = 0` or the binary predictor `x2` is at its reference `0` level.
+But, when the interaction is excluded the main effects are relevant when the other predictors are at their mean values.
 So, if we center, the main effects will represent the same thing in both cases:
@@
 In other words, adding the interaction adds more predictive information but doesn't modify the meaning of the main effects.
+
+Here's an illustration:
+
+<img src="/images/interactions.png" alt="Illustration of centering interactions" width=580 height=436>
 
@@
 m2 <- lm(y ~ x1 * f, data = dat)
-round(coef(m2), 2)
+t(round(coef(m2), 2))
 ```
@@
 m2.1 <- lm(y ~ x1 * x2 + x1 * x3, data = dat)
-round(coef(m2.1), 2)
+t(round(coef(m2.1), 2))
 ```
@@
 
-Now, let's compare to a model without the interaction:
+Now, let's compare to a model without the interaction before we center the data:
 
@@
 m2.1_no_inter <- lm(y ~ x1 + x2 + x1 + x3, data = dat)
-round(coef(m2.1), 2)
-round(coef(m2.1_no_inter), 2)
+t(round(coef(m2.1_no_inter), 2))
 ```
 
-Again, if we model averaged here across models with and without the interaction, we'd be getting gibberish.
+The main effects look dramatically different. Again, if we model averaged here across models with and without the interaction, we'd be getting gibberish.
 
@@
 dat$x3_cent <- dat$x3 - mean(dat$x3)
-m2.2 <- lm(y ~ x1_cent * x2_cent + x1_cent * x3_cent, data = dat)
-m2.2_no_inter <- lm(y ~ x1_cent + x2_cent + x3_cent, data = dat)
+m2.2 <- lm(y ~ x1_cent * x2_cent + x1_cent * x3_cent, 
+  data = dat)
+m2.2_no_inter <- lm(y ~ x1_cent + x2_cent + x3_cent, 
+  data = dat)
 ```
@@
 ```{r}
-round(coef(m2.2), 2)
-round(coef(m2.2_no_inter), 2)
+t(round(coef(m2.2), 2)) # no interaction 
+t(round(coef(m2.2_no_inter), 2)) # with interaction
 ```
@@
 
-One more thing.
-Say we wanted to plot the predicted response from our model.
+Say we wanted to plot the predicted response from our model on new data.
 This isn't quite as simple as normal once we've centered our predictors and coded the factor levels as centered dummy variable comparisons.
@@
 We'll make predictions at each of the 3 levels.
-Note how we create these levels through the 3 valid combinations of our dummy variables.
+Note how we create these levels through the 3 valid combinations of high and low values of our dummy variables.
 We can't just use `expand.grid()` and get all combinations of our dummy variables.
@@
 
-We can calculate the combination of these two distributions multiplied by the constants `a` and `b` (which we'll call `N3`) as:
+We can calculate the combination of these two distributions multiplied by the constants `a` and `b` as: (we'll this new distribution `N3`)
 
@@
 mu3 = a * mu1 + b * mu2
-sigma3 = sqrt(a^2 * sigma1^2 + b^2 * sigma2^2 + 2 * ab * cov(ab))
+sigma3 = sqrt(a^2 * sigma1^2 + b^2 * sigma2^2 + 
+  2 * ab * cov(ab))
 ```
@@
 The first line in each of these `abline` function calls represents the intercepts and each factor level and the second line represents the slopes.
-Notice how we get these slopes by turning the two indicator columns to low or high values.
+As before, notice how we get these slopes by turning the two indicator columns to low or high values.
 
@@
 ```
+
+Whether or not you center your predictors, transforming your factor coefficient interactions like the above can be a powerful way of displaying data.
+For example see Figure 1 in [O'Regan et al.
+2014, Ecology, 95(4)](http://www.esajournals.org/doi/pdf/10.1890/13-0916.1) where we use this method to illustrate the effect of climate warming and drying on four responses across three amphibian species.
+Each panel represents a single model and we've adjusted the coefficients so that each effect is with respect to zero instead of with respect to another factor level.
+If this kind of presentation of the results is all you're after, then there's no need to center the factor levels.
+If, however, you also want to apply model averaging to coefficient values that include interactions, you will need to center your predictors and combine the coefficients as in the above example.

### 2014-07-08 refs.R
Update ref bib file

@@ -45,7 +45,7 @@ for(i in seq_len(length(d))) {
     vol_num_p <- ""
   }
 
-  if(vol_num_p != "" & vol_num_p != "In review" & vol_num_p != "In revision") {
+  if(vol_num_p != "" & !vol_num_p %in% c("In review", "In revision", "Submitted")) {
     main <- paste0(paste0(name0, ". ", x$title, ". ", x$journal, ". ", vol_num_p), ".")
   } else {
     main <-paste0(name0, ". ", x$title, ". ", vol_num_p, ".")

### 2014-07-30 _knitr-posts/2014-07-30-centering-interactions.Rmd
Publish centering post
Other trivial changes

--- 2014-07-08 _knitr-posts/2014-06-12-centering-interactions.Rmd
@@
 subtitle: Why and how you must center interactions for model averaging
-published: false
+published: true
 ---
@@
 
-Now, let's derive those same standard errors for the centered version. We'll start be extracting the standard errors and covariances of the coefficients. Then we'll combine them:
+Now, let's derive those same standard errors for the centered version.
+We'll start be extracting the standard errors and covariances of the coefficients.
+Then we'll combine them.
+Although we could do this more elegantly, we'll write it out in full to make it obvious what's happening.
 

### 2014-08-07 refs.R
Switch to PDFs linked from Dropbox for speed
GitHub server was much too slow

@@ -60,7 +60,7 @@ for(i in seq_len(length(d))) {
       main <- paste0(main, " ([PDF](", x$pdf, ")).")
     } else {
       main <- paste0(main,
-        " ([PDF](http://seananderson.ca/papers/", x$pdf, ")).")
+        " ([PDF](https://dl.dropboxusercontent.com/u/254940/papers/", x$pdf, ")).")
     }
   }
 

### 2014-08-07 refs.R
Fixes to publication list

@@ -78,7 +78,7 @@ for(i in seq_len(length(d))) {
 
   if(!is.null(x$f1000)) {
     main <- paste0(main,
-      "\\\n     <a href=\"http://f1000.com/prime/",
+      "\n     <a href=\"http://f1000.com/prime/",
       x$f1000,
       "\" target=\"_blank\"><img src=\"http://cdn.f1000.com.s3.amazonaws.com/images/badges/badgef1000.gif\" id=\"bg\"/></a>")
   }

### 2014-08-07 refs.R
Small fixes to CV

@@ -1,9 +1,9 @@
+# TODO J.E.M. Flemming should be J.E. Mills Flemming
+# TODO J.E. Mills Flemming needs to be fixed throughout refs.bib too
+# TODO Earth2Oceans Derby needs to be mannually fixed
+
 library("bibtex")
 d <- read.bib("refs.bib")
-
-#str(x)
-
-#init <- list()
 
 ref <- list()
 
@@ -85,10 +85,6 @@ for(i in seq_len(length(d))) {
 
   main <- paste0(x$year, "\n:    ", main)
   ref[[i]] <- main
-
-
-
-
 }
 
 ref <- paste(ref, collapse = "\n\n")

### 2014-09-15 _knitr-posts/2014-09-13-dplyr-intro.Rmd
Missed adding the dplyr post last time

---
layout: post
title: "dplyr and pipes: the basics"
subtitle: "Revolutionize your data analysis with dplyr and pipes"
published: true
---

The [dplyr](http://cran.r-project.org/package=dplyr) R package is awesome. Pipes from the [magrittr](http://cran.r-project.org/package=magrittr) R package are awesome. Put the two together and you have one of the most exciting things to happen to R in a long time.

dplyr is [Hadley Wickham's](http://had.co.nz/) re-imagined plyr package (with underlying C++ secret sauce co-written by [Romain Francois](http://blog.r-enthusiasts.com/)). plyr 2.0 if you will. It does less than dplyr, but what it does it does more elegantly and much more quickly.

dplyr is built around 5 verbs. These verbs make up the majority of the data manipulation you tend to do. You might need to:

*Select* certain columns of data.

*Filter* your data to select specific rows.

*Arrange* the rows of your data into an order.

*Mutate* your data frame to contain new columns.

*Summarise* chunks of you data in some way.

Let's look at how those work.

# The data

We're going to work with a dataset of mammal life-history, geography, and ecology traits from the PanTHERIA database:

Jones, K.E., *et al*. PanTHERIA: a species-level database of life history, ecology, and geography of extant and recently extinct mammals. Ecology 90:2648. <http://esapubs.org/archive/ecol/E090/184/>

First we'll download the data:

```{r download-data, eval=FALSE}
pantheria <-
  "http://esapubs.org/archive/ecol/E090/184/PanTHERIA_1-0_WR05_Aug2008.txt"
download.file(pantheria, destfile = "mammals.txt")
```

Next we'll read it in and simplify it. This gets a bit ugly, but you can safely just run this code chunk and ignore the details:

```{r}
mammals <- read.table("mammals.txt", sep = "\t", header = TRUE, 
  stringsAsFactors = FALSE)
names(mammals) <- sub("X[0-9._]+", "", names(mammals))
names(mammals) <- sub("MSW05_", "", names(mammals))
mammals <- dplyr::select(mammals, Order, Binomial, AdultBodyMass_g, 
  AdultHeadBodyLen_mm, HomeRange_km2, LitterSize)
names(mammals) <- gsub("([A-Z])", "_\\L\\1", names(mammals), perl = TRUE)
names(mammals) <- gsub("^_", "", names(mammals), perl = TRUE)
mammals[mammals == -999] <- NA
names(mammals)[names(mammals) == "binomial"] <- "species"
mammals <- dplyr::tbl_df(mammals) # for prettier printing
```

Next we'll load the dplyr package:

```{r, message=FALSE}
library(dplyr)
```

# Looking at the data

Data frames look a bit different in dplyr. Above, I called the `tbl_df()` function on our data. This provides more useful printing of data frames in the console. Ever accidentally printed a massive data frame in the console before? Yeah... this avoids that. You don't need to change your data to a data frame tbl first — the dplyr functions will automatically convert your data when you call them. This is what the data look like on the console:

```{r}
mammals
```

dplyr also provides a function `glimpse()` that makes it easy to look at our data in a transposed view. It's similar to the `str()` (structure) function, but has a few advantages (see `?glimpse`).

```{r}
glimpse(mammals)
```

# Selecting columns

`select()` lets you subset by columns. This is similar to `subset()` in base R, but it also allows for some fancy use of helper functions such as `contains()`, `starts_with()` and, `ends_with()`. I think these examples are self explanatory, so I'll just include them here:

```{r}
select(mammals, adult_head_body_len_mm)
select(mammals, adult_head_body_len_mm, litter_size)
select(mammals, adult_head_body_len_mm:litter_size)
select(mammals, -adult_head_body_len_mm)
select(mammals, contains("body"))
select(mammals, starts_with("adult"))
select(mammals, ends_with("g"))
select(mammals, 1:3)
```

# Filtering rows

`filter()` lets you subset by rows. You can use any valid logical statements:

```{r}
filter(mammals, adult_body_mass_g > 1e7)[ , 1:3]
filter(mammals, species == "Balaena mysticetus")
filter(mammals, order == "Carnivora" & adult_body_mass_g < 200)
```

# Arranging rows

`arrange()` lets you order the rows by one or more columns in ascending or descending order. I'm selecting the first three columns only to make the output easier to read:

```{r}
arrange(mammals, adult_body_mass_g)[ , 1:3]
arrange(mammals, desc(adult_body_mass_g))[ , 1:3]
arrange(mammals, order, adult_body_mass_g)[ , 1:3]
```

# Mutating columns

`mutate()` lets you add new columns. Notice that the new columns you create can build on each other. I will wrap these in `glimpse()` to make the new columns easy to see:

```{r}
glimpse(mutate(mammals, adult_body_mass_kg = adult_body_mass_g / 1000))
glimpse(mutate(mammals, 
    g_per_mm = adult_body_mass_g / adult_head_body_len_mm))
glimpse(mutate(mammals, 
    g_per_mm = adult_body_mass_g / adult_head_body_len_mm,
    kg_per_mm = g_per_mm / 1000))
```

# Summarising columns

Finally, `summarise()` lets you calculate summary statistics. On its own `summarise()` isn't that useful, but when combined with `group_by()` you can summarise by chunks of data. This is similar to what you might be familiar with through `ddply()` and `summarise()` from the plyr package:

```{r}
summarise(mammals, mean_mass = mean(adult_body_mass_g, na.rm = TRUE))
# summarise with group_by:
head(summarise(group_by(mammals, order),
  mean_mass = mean(adult_body_mass_g, na.rm = TRUE)))
```

# Piping data

Pipes take the output from one function and feed it to the first argument of the next function. You may have encountered the Unix pipe `|` before.

The magrittr R package contains the pipe function `%>%`. Yes it might look bizarre at first but it makes more sense when you think about it. The R language allows symbols wrapped in `%` to be defined as functions, the `>` helps imply a chain, and you can hit these 2 characters one after the other very quickly on a keyboard by holding down the Shift key. Try it!

Try pronouncing `%>%` "then" whenever you see it. If you want to see the help page, you'll need to wrap it in back ticks like so:

```{r, eval=FALSE}
?magrittr::`%>%`
```

# A trivial pipe example

Pipes can work with nearly any functions. Let's start with a non-dplyr example:

```{r}
x <- rnorm(10)
x %>% max
# is the same thing as:
max(x)
```

So, we took the value of `x` (what would have been printed on the console), captured it, and fed it to the first argument of `max()`. It's probably not clear why this is cool yet, but hang on.

# A silly dplyr example with pipes

Let's try a single-pipe dplyr example. We'll pipe the `mammals` data frame to the arrange function's first argument, and choose to arrange by the `adult_body_mass_g` column:

```{r}
mammals %>% arrange(adult_body_mass_g)
```

# An awesome example

OK, here's where it gets cool. We can chain dplyr functions in succession. This lets us write data manipulation steps in the order we think of them and avoid creating temporary variables in the middle to capture the output. This works because the output from every dplyr function is a data frame and the first argument of every dplyr function is a data frame.

Say we wanted to find the species with the highest body-mass-to-length ratio:

```{r}
mammals %>%
  mutate(mass_to_length = adult_body_mass_g / adult_head_body_len_mm) %>%
  arrange(desc(mass_to_length)) %>%
  select(species, mass_to_length)
```

So, we took `mammals`, fed it to `mutate()` to create a mass-length ratio column, arranged the resulting data frame in descending order by that ratio, and selected the columns we wanted to see. This is just the beginning. If you can imagine it, you can string it together. If you want to debug your code, just pull a pipe off the end and run the code down to that step. Or build your analysis up and add successive pipes.

The above is equivalent to:

```{r}
select(
  arrange(
    mutate(mammals,
      mass_to_length = adult_body_mass_g / adult_head_body_len_mm),
    desc(mass_to_length)),
  species, mass_to_length)
```

But the problem here is that you have to read it inside out, it's easy to miss a bracket, and the arguments get separated from the function (e.g. see `mutate()` and `desc(mass_to_length))`). Plus, this is a rather trivial example. Chain together even more steps and it quickly gets out of hand.

Here's one more example. Let's ask what taxonomic orders have a median litter size greater than 3.

```{r}
mammals %>% group_by(order) %>%
  summarise(median_litter = median(litter_size, na.rm = TRUE)) %>%
  filter(median_litter > 3) %>%
  arrange(desc(median_litter)) %>%
  select(order, median_litter)
```

These examples don't even highlight one of the best things about dplyr. It's *really* fast. The internal C++ code makes quick work of massive data frames that would make plyr slow to a crawl.

dplyr can do much more, but the above are the basics of the 5 verbs and pipes. Try them for a bit. Once they click I think they'll revolutionize your data analysis.

### 2014-09-15 cv-body.tex
Missed adding the dplyr post last time

--- 2013-06-20 cv-body.tex
@@
-\section{Education}
+\section{Education}\label{education}
 
 \begin{description}
-\item[2011--13]
-Ph.D.\ Candidate, Simon Fraser University, Burnaby, Canada Supervisors:
-Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper\\Canadian Fulbright
+\item[2011--14]
+Ph.D.\ Candidate, Simon Fraser University, Burnaby, Canada\\ Supervisors:
+Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper\\ Canadian Fulbright
 Scholar to the University of Washington in 2012--13
-\item[2010]
-M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\Thesis title:
+\item[2008--10]
+M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\ Thesis title:
 Trends, drivers, and ecosystem effects of expanding global invertebrate
 fisheries. Supervisor: Dr.\ Heike K. Lotze.
-(\href{papers/Anderson_MSc_2010.pdf}{PDF})\\Canadian
-\href{http://goo.gl/nA1zE}{Governor General's Academic Gold Medal} for
-the top-ranked Master's Natural Sciences and Engineering thesis at
-Dalhousie University in 2010
+(\href{http://seananderson.ca/papers/Anderson_MSc_2010.pdf}{PDF})\\
+Canadian \href{http://goo.gl/nA1zE}{Governor General's Academic Gold
+Medal} for the top-ranked Master's Natural Sciences and Engineering
+thesis at Dalhousie University in 2010
 \item[2004--07]
 B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
-Canada\\Thesis title: How sustainable are emerging low-trophic level
+Canada\\ Thesis title: How sustainable are emerging low-trophic level
 fisheries on the Scotian Shelf? Supervisors: Dr.\ Heike K. Lotze and
-Dr.\ Ransom A. Myers. (\href{papers/Anderson_HSc_2006.pdf}{PDF})\\First
-Class Honours \& Environmental Programmes Honour Society Medal
+Dr.\ Ransom A. Myers.
+(\href{http://seananderson.ca/papers/Anderson_HSc_2006.pdf}{PDF})\\
+First Class Honours \& Environmental Programmes Honour Society Medal
 \item[2001--03]
@@
 
-\section{Publications}
+\section{Publications}\label{publications}
 
 \begin{description}
-\item[2012]
-Phillis, C.C.*, S.M. O'Regan*, S.J. Green*, J.E.B. Bruce*, \textbf{S.C.
-Anderson}, J.N. Linton, Earth2Ocean Research Derby, B. Favaro. Multiple
-pathways to conservation success. Conservation Letters. Early view
-online. \url{http://dx.doi.org/10.1111/j.1755-263X.2012.00294.x}.
+\item[2014]
+Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down and bottom-up
+forces interact at thermal range extremes on American lobster. Accepted
+with minor revisions at Journal of Animal Ecology.
+\item[2014]
+\textbf{Anderson, S.C.}, J.W. Moore, M.M. McClure, N.K. Dulvy, A.B.
+Cooper. Portfolio conservation of metapopulations under climate change.
+Ecological Applications. In press.
+\url{http://doi.org/10.1890/14-0266.1}.
+\item[2014]
+Farmer, R.G., M.L. Leonard, J.E. Mills Flemming, \textbf{S.C. Anderson}.
+Observer aging and long-term avian survey data quality. Ecology and
+Evolution. 4(12): 2563--2576. \url{http://doi.org/10.1002/ece3.1101}
+(\href{http://onlinelibrary.wiley.com/doi/10.1002/ece3.1101/pdf}{PDF}).
+\href{http://news.nationalgeographic.com/news/2014/08/140805-aging-birders-breeding-bird-survey-volunteers-science/}{National
+Geographic news story}.
+\item[2014]
+\textbf{Anderson, S.C.}, C.C. Monnahan, K.F. Johnson, K. Ono, J.L.
+Valero. ss3sim: An R package for fisheries stock assessment simulation
+with Stock Synthesis. PLOS ONE. 9(4): e92725.
+\url{http://doi.org/10.1371/journal.pone.0092725}
+(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0092725\&representation=PDF}{PDF}).
+(\href{http://cran.r-project.org/web/packages/ss3sim/index.html}{R
+package}, \href{https://github.com/ss3sim/ss3sim}{code},
+\href{https://dl.dropboxusercontent.com/u/254940/ss3sim-vignette.pdf}{vignette}).
+\item[2014]
+Johnson, K.F., C.C. Monnahan, C.R. McGilliard, K.A. Vert-pre,
+\textbf{S.C. Anderson}, C.J. Cunningham, F. Hurtado-Ferro, R.R.
+Licandeo, M.L. Muradian, K. Ono, C.S. Szuwalski, J.L. Valero, A.R.
+Whitten, A.E. Punt. Time-varying natural mortality in fisheries stock
+assessment models: identifying a default approach. ICES Journal of
+Marine Science. In press. \url{http://doi.org/10.1093/icesjms/fsu055}
+(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
+\item[2014]
+Ono, K., R. Licandeo, M.L. Muradian, C.J. Cunningham, \textbf{S.C.
+Anderson}, F. Hurtado-Ferro, K.F. Johnson, C.R. McGilliard, C.C.
+Monnahan, C.S. Szuwalski, J.L. Valero, K.A. Vert-pre, A.R. Whitten, A.E.
+Punt. The importance of length and age composition data in statistical
+catch-at-age models for marine species. ICES Journal of Marine Science.
+In press. \url{http://doi.org/10.1093/icesjms/fsu007}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Ono_etal_2014_importance_of_length_and_age_composition_data.pdf}{PDF}).
+\item[2014]
+O'Regan, S.M., W.J. Palen, \textbf{S.C. Anderson}. Climate warming
+mediates negative impacts of rapid pond drying for three amphibian
+species. Ecology. 95(4): 845--855.
+\url{http://doi.org/10.1890/13-0916.1}
+(\href{http://www.esajournals.org/doi/pdf/10.1890/13-0916.1}{PDF}).
+\href{http://www.esajournals.org/userimages/ContentEditor/1398099670397/i0012-9658-95-4-cov.pdf}{Cover
+Article}.
+\item[2013]
+Favaro, B., D.C. Braun, \textbf{Earth2Ocean Research Derby}. Research
+Derby: A pressure cooker for creative collaborative science. Ideas in
+Ecology and Evolution. Ideas in Ecology and Evolution. 6: 40--46.
+\url{http://doi.org/10.4033/iee.2013.6.9.n}
+(\href{http://library.queensu.ca/ojs/index.php/IEE/article/download/4931/4899}{PDF}).
+\item[2013]
+\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Ecological prophets:
+Quantifying metapopulation portfolio effects. Methods in Ecology and
+Evolution. 4(10): 971--981. \url{http://doi.org/10.1111/2041-210X.12093}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2013_ecological_prophets_with_SOM.pdf}{PDF}).
+\item[2013]
+Artelle, K.A., \textbf{S.C. Anderson}, A.B. Cooper, P.C. Paquet, J.D.
+Reynolds, C.T. Darimont. Confronting uncertainty in wildlife management:
+performance of grizzly bear management. PLOS ONE. 8(11): e78041.
+\url{http://doi.org/10.1371/journal.pone.0078041}
+(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0078041\&representation=PDF}{PDF}).
+\item[2013]
+Phillis*, C.C., S.M. O'Regan*, S.J. Green*, J.E. Bruce*, \textbf{S.C.
+Anderson}, J.N. Linton, E.R. Derby, B. Favaro. Multiple pathways to
+conservation success. Conservation Letters. 6(2): 98--106.
+\url{http://doi.org/10.1111/j.1755-263X.2012.00294.x}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf}{PDF}).
 (*\emph{Authors contributed equally, listed in reverse alphabetical
-order})
+order}).
 \item[2012]
@@
 Extinctions in ancient and modern seas. Trends in Ecology and Evolution.
-27(11): 608--617. \url{http://dx.doi.org/10.1016/j.tree.2012.07.010}.
+27(11): 608--617. \url{http://doi.org/10.1016/j.tree.2012.07.010}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Harnik_etal_2012_Extinctions_in_ancient_and_modern_seas.pdf}{PDF}).
 \item[2012]
@@
 and stock-assessment reference points. ICES Journal of Marine Science.
-69(8): 1491--1500. \url{http://dx.doi.org/10.1093/icesjms/fss105}.
-(\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF})
+69(8): 1491--1500. \url{http://doi.org/10.1093/icesjms/fss105}
+(\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF}).
 \item[2011]
-\textbf{Anderson, S.C.}, B.G. Farmer, F. Ferretti, A.L.S. Houde, and
-J.A. Hutchings. Correlates of vertebrate extinction risk in Canada.
+\textbf{Anderson, S.C.}, R.G. Farmer, F. Ferretti, A.L.S. Houde, J.A.
+Hutchings. Correlates of vertebrate extinction risk in Canada.
 BioScience. 61(7): 538--549.
-\url{http://dx.doi.org/10.1525/bio.2011.61.7.8}.
-(\href{papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF})
+\url{http://doi.org/10.1525/bio.2011.61.7.8}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF}).
 \item[2011]
-\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
+\textbf{Anderson, S.C.}, J.E. Mills Flemming, R. Watson, H.K. Lotze.
 Serial exploitation of global sea cucumber fisheries. Fish and
-Fisheries. 12(3): 317-339.
-\url{http://dx.doi.org/10.1111/j.1467-2979.2010.00397.x}.
-(\href{papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF})
+Fisheries. 12(3): 317--339.
+\url{http://doi.org/10.1111/j.1467-2979.2010.00397.x}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF}).
 \href{http://www.sciencemag.org/content/331/6014/129.1.full}{Featured in
-Science} on January 14 2011.
+Science}.
 \item[2011]
 Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down interactions
-and temperature control of snow crab abundance in the northwest
-Atlantic. Marine Ecology Progress Series. 429: 169--183.
-\url{http://dx.doi.org/10.3354/meps09081}.
+and temperature control of snow crab abundance in the northwest Atlantic
+Ocean. Marine Ecology Progress Series. 429: 169--183.
+\url{http://doi.org/10.3354/meps09081}
+(\href{http://www.int-res.com/articles/meps_oa/m429p169.pdf}{PDF}).
 \item[2011]
-\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze. Rapid
-global expansion of invertebrate fisheries: trends, drivers, and
-ecosystem effects. PLoS ONE. 6(3): e14735.
-\url{http://dx.doi.org/10.1371/journal.pone.0014735}
+\textbf{Anderson, S.C.}, J.E. Mills Flemming, R. Watson, H.K. Lotze.
+Rapid global expansion of invertebrate fisheries: trends, drivers, and
+ecosystem effects. PLOS ONE. 6(3): e14735.
+\url{http://doi.org/10.1371/journal.pone.0014735}
+(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0014735\&representation=PDF}{PDF}).
 \item[2008]
-\textbf{Anderson, S.C.}, H.K. Lotze, and N.L. Shackell. Evaluating the
+\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
 knowledge base for expanding low-trophic-level fisheries in Atlantic
-Canada. Can. J. Fish. Aquat. Sci. 65(12): 2553--2571.
-\url{http://dx.doi.org/10.1139/F08-156}.
-(\href{papers/Anderson_etal_2008_knowledge.pdf}{PDF}) Eighth most-read
-CJFAS article in 2008.
+Canada. Canadian Journal of Fisheries and Aquatic Sciences. 65(12):
+2553--2571. \url{http://doi.org/10.1139/F08-156}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2008_knowledge.pdf}{PDF}).
+Eighth most-read CJFAS article in 2008.
 \end{description}
 
-\section{Submitted manuscripts}
-
-\begin{description}
-\item[2013]
-\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Ecological prophets:
-Quantifying metapopulation portfolio effects. Submitted with revisions
-to Methods in Ecology and Evolution.
-\item[2013]
-O'Regan, S.M., W.J. Palen, \textbf{S.C. Anderson}. Climate warming
-mediates negative impacts of rapid pond drying for three amphibian
-species. Submitted to Ecology.
-\end{description}
-
-\section{Awards and scholarships}
+\section{Manuscripts in review}\label{manuscripts-in-review}
 
@@
 \item[2014]
-Graduate Fellowship, Simon Fraser University
+Finnegan, S., \textbf{S.C. Anderson}, P.G. Harnik, C. Simpson, D.P.
+Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
+Lockwood, H.K. Lotze, C.M. McClain, J.L. McGuire, A. O'Dea, J.M.
+Pandolfi. Paleontological baselines for evaluating extinction risk in
+the modern oceans. In revision.
+\item[2014]
+Trebilco, R., N.K. Dulvy, \textbf{S.C. Anderson}, A.K. Salomon. The
+paradox of inverted biomass pyramids in kelp forest fish communities. In
+review at Proceedings of the Royal Society B.
+\item[2014]
+Hurtado-Ferro, F., C.S. Szuwalski, J.L. Valero, \textbf{S.C. Anderson},
+C.J. Cunningham, K.F. Johnson, R.R. Licandeo, C.R. McGilliard, C.C.
+Monahan, M.L. Muradian, K. Ono, K.A. Vert-Pre, A.R. Whitten, A.E. Punt.
+Looking in the rear-view mirror: bias and retrospective patterns in
+integrated, age-structured stock assessment models. In review at ICES
+Journal of Marine Science.
+\end{description}
+
+\section{Awards and scholarships}\label{awards-and-scholarships}
+
+\begin{description}
+\item[2014]
+Garfield Weston Foundation / BC Packers Ltd. Graduate Fellowship in
+Marine Sciences
+\item[2014]
+Graduate Fellowship (two semesters), Simon Fraser University
 \item[2012--13]
-\href{http://www.fulbright.ca/programs/canadian-students/2012-2013-canadian-fulbright-students.html}{Canadian
-Fulbright Scholar award} to the University of Washington in 2012--13
+Canadian Fulbright Scholar award to the University of Washington in
+2012--13
 \item[2011]
@@
 Environmental Programmes Honour Society Medal, Dalhousie University
-\item[2001--03]
-Academic All-Canadian, University of Waterloo
-\item[2001--02]
-Dean's Entrance Scholarship, University of Waterloo
 \end{description}
 
-\section{Invited talks and conference presentations}
+\section{Software}\label{software}
+
+\begin{description}
+\item[2014]
+\textbf{Anderson, S.C.}, J.W. Moore, M.M McClure, N.K. Dulvy, A.B.
+Cooper.\\ metafolio: Metapopulation simulations for conserving salmon
+through portfolio optimization.
+\url{http://cran.r-project.org/package=metafolio}
+\item[2013]
+\textbf{Anderson, S.C.}, C.C. Monnahan, K.F. Johnson, K. Ono, J.L.
+Valero, C.J Cunningham, F. Hurtado-Ferro, R. Licandeo, C.R. McGilliard,
+C.S. Szuwalski, K.A. Vert-pre, A.R. Whitten. ss3sim: Fisheries stock
+assessment simulation testing with Stock Synthesis.
+\url{http://cran.r-project.org/package=ss3sim}
+\item[2013]
+\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. ecofolio: Tools to
+quantify metapopulation portfolio effects.
+\url{https://github.com/seananderson/ecofolio}
+\end{description}
+
+\section{Invited talks and conference
+presentations}\label{invited-talks-and-conference-presentations}
 
@@
 challenges of quantifying ecological portfolios
-(\href{talks/2013/PE_SAFS_quantsem.pdf}{slides}), Quantitative Seminar,
-School of Aquatic and Fishery Sciences, University of Washington,
-Seattle, WA, United States.
+(\href{http://seananderson.ca/talks/2013/PE_SAFS_quantsem.pdf}{slides}),
+Quantitative Seminar, School of Aquatic and Fishery Sciences, University
+of Washington, Seattle, WA, United States.
 \item[2012]
@@
 \item[2009]
-\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
+\textbf{Anderson, S.C.}, J.E. Mills Flemming, R. Watson, H.K. Lotze.
 Ecosystem impacts of the global expansion of invertebrate fisheries.
@@
 \item[2009]
-\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
+\textbf{Anderson, S.C.}, J.E Mills Flemming, R. Watson, H.K. Lotze.
 Global invertebrate fisheries: trends and consequences. NCEAS (National
@@
 
-\section{Contract employment}
+\section{Teaching}\label{teaching}
 
 \begin{description}
-\item[2011--13]
-Database management, data analysis, and visualization for the project
-\emph{Global Shark Abundance Baselines} (Hopkins Marine Station,
-Stanford University, CA, United States) with F. Micheli, F. Ferretti,
-N.K. Dulvy, H.K. Lotze, and B. Worm.
+\item[2013--14]
+Organizer of Stats Beerz --- a statistical help group attended by
+graduate students and postdocs primarily in the Earth to Oceans research
+group, but also the wider SFU Biology and Geography Departments, and the
+School of Resource and Environmental Management (REM).
+\item[2014]
+Developed self-directed lecture and exercises on ggplot2 for FISH 554:
+Beautiful graphics in R, School of Aquatic and Fishery Sciences,
+University of Washington, Seattle, WA, United States.\\
+\url{https://github.com/seananderson/ggplot2-FISH554}
+\item[2013]
+Two-part workshop on data manipulation for Stats Beerz and Earth to
+Oceans groups at Simon Fraser University with approximately 25
+participants. An introduction to plyr, advanced concepts with plyr and
+function debugging, and an introduction to dplyr.\\
+\url{https://github.com/seananderson/plyr-statsbeerz}
+\item[2013]
+Instructor for BISC-888-1: Data Wrangling and Visualization in R, a
+graduate-level course at Simon Fraser University, Burnaby, BC, Canada
+with approximately 20 participants. Co-developed curriculum and
+developed/delivered lectures, exercises, notes, and assignments for
+three of six two-hour classes.\\
+\url{https://github.com/seananderson/datawranglR} (see classes 03, 04,
+05)
+\item[2012]
+Introduction to ggplot2.
+(\href{http://seananderson.ca/courses/12-ggplot2/ggplot2_notes.pdf}{notes},
+\href{http://seananderson.ca/courses/12-ggplot2/ggplot2_slides_with_examples.pdf}{slides})
+Lecture for FISH 507H: Beautiful Graphics in R, School of Aquatic and
+Fishery Sciences, University of Washington, Seattle, WA, United States
+\item[2012]
+Workshop on the R package plyr.
+(\href{http://seananderson.ca/courses/12-plyr/plyr_2012.pdf}{notes},
+\href{http://seananderson.ca/courses/12-plyr/plyr_2012_slides.pdf}{slides},
+\href{http://seananderson.ca/courses/12-plyr/plyr_2012_examples.html}{examples})
+Branch Lab, School of Aquatic and Fishery Sciences, University of
+Washington, Seattle, WA, United States
+\item[2011]
+Multipanel plotting in R with base graphics.
+(\href{http://seananderson.ca/courses/11-multipanel/multipanel.pdf}{notes},
+\href{http://seananderson.ca/courses/11-multipanel/multipanel-slides.pdf}{slides}).
+Earth2Ocean Research Group, Simon Fraser University, Burnaby, BC,
+Canada.
+\item[2011]
+A brief introduction to R.
+(\href{http://seananderson.ca/courses/11-rintro/RIntro.pdf}{notes},
+\href{http://seananderson.ca/courses/11-rintro/RIntro.R}{workshop
+code}). Earth2Ocean Research Group, Simon Fraser University, Burnaby,
+BC, Canada.
+\item[2011]
+Workshop on the R package plyr
+(\href{http://seananderson.ca/courses/11-plyr/plyr.pdf}{notes},
+\href{http://seananderson.ca/courses/11-plyr/plyr-slides.pdf}{slides}).
+Earth2Ocean Research Group, Simon Fraser University, Burnaby, BC,
+Canada.
+\item[2007--08]
+Teaching Assistant, Organismal Biology and Ecology, Dalhousie
+University, two semesters (BIOL 1021).
+\item[2007--08]
+Teaching Assistant, Marine Mammology, Dalhousie University (BIOL 4060).
+\end{description}
+
+\section{Contract employment}\label{contract-employment}
+
+\begin{description}
+\item[2013]
+Co-developed a set of maps showing trends in global fisheries status for
+the front page of \href{http://ramlegacy.org/}{ramlegacy.org}, the RAM
+Legacy Stock Assessment Database (for Dr.\ Ray Hilborn).
+\item[2011--12]
+Database management for the project \emph{Global Shark Abundance
+Baselines} (Hopkins Marine Station, Stanford University, CA, United
+States) with F. Micheli, F. Ferretti, N.K. Dulvy, H.K. Lotze, and B.
+Worm.
 \item[2010]
-Assessment of Arctic surfclam on Banquereau Bank using industry
-collected data for Clearwater Seafoods, Bedford, NS, Canada.
+Assessment of Arctic surfclam on Banquereau Bank using
+industry-collected data for Clearwater Seafoods, Bedford, NS, Canada.
 \item[2010]
@@
 
-\section{Working groups and workshops}
+\section{Working groups and
+workshops}\label{working-groups-and-workshops}
 
@@
 
-\section{Selected courses and training}
+\section{Reviews}\label{reviews}
 
-\begin{description}
-\item[Present]
-Proficient in statistical analyses and data management using R, WinBUGS,
-JAGS, Sweave, knitr, LaTeX, Perl, PostgreSQL, Git.
-\item[2013]
-Quantitative Modelling in Ecosystem Based Management, R. Hilborn, School
-of Aquatic and Fishery Sciences, University of Washington (auditing).
-\item[2013]
-Applied Time Series Analysis; E. Holmes, M. Scheuerell, and E. Ward;
-School of Aquatic and Fishery Sciences; University of Washington
-(audited).
-\item[2012]
-Numerical Computing for the Natural Resources, Dr.\ A.E. Punt, School of
-Aquatic and Fishery Sciences, University of Washington (audited).
-\item[2012]
-Risk Assessment and Decision Analysis for Management of Natural
-Resources, Dr.\ R.M. Peterman, School of Resource and Environmental
-Management, Simon Fraser University.
-\item[2011]
-Presenting Data and Information, Edward Tufte, Seattle, WA, United
-States.
-\item[2009]
-Technical Expertise in Stock Assessment: Model Fitting and
-Interpretation by Dr.\ C. Needle, chair of the International Council for
-the Exploration of the Sea working group on methods of fish stock
-assessment, Dalhousie University.
-\item[2008]
-Data Analysis, Dr.\ J. Mills Flemming, Department of Mathematics \&
-Statistics, Dalhousie University (audited).
-\end{description}
-
-\section{Teaching}
-
-\begin{description}
-\item[2012]
-Introduction to ggplot2.
-(\href{courses/12-ggplot2/ggplot2_notes.pdf}{notes},
-\href{courses/12-ggplot2/ggplot2_slides_with_examples.pdf}{slides})
-Lecture for FISH 507H: Beautiful Graphics in R, School of Aquatic and
-Fishery Sciences, University of Washington, Seattle, WA, United States
-\item[2012]
-Workshop on the R package plyr.
-(\href{courses/12-plyr/plyr_2012.pdf}{notes},
-\href{courses/12-plyr/plyr_2012_slides.pdf}{slides},
-\href{courses/12-plyr/plyr_2012_examples.html}{examples}) Branch Lab,
-School of Aquatic and Fishery Sciences, University of Washington,
-Seattle, WA, United States
-\item[2011]
-Multipanel plotting in R with base graphics.
-(\href{courses/11-multipanel/multipanel.pdf}{notes},
-\href{courses/11-multipanel/multipanel-slides.pdf}{slides}). Earth2Ocean
-Research Group, Simon Fraser University, Burnaby, BC, Canada.
-\item[2011]
-A brief introduction to R. (\href{courses/11-rintro/RIntro.pdf}{notes},
-\href{courses/11-rintro/RIntro.R}{workshop code}). Earth2Ocean Research
-Group, Simon Fraser University, Burnaby, BC, Canada.
-\item[2011]
-Workshop on the R package plyr (\href{courses/11-plyr/plyr.pdf}{notes},
-\href{courses/11-plyr/plyr-slides.pdf}{slides}). Earth2Ocean Research
-Group, Simon Fraser University, Burnaby, BC, Canada.
-\item[2007--08]
-Teaching Assistant, Organismal Biology and Ecology, Dalhousie
-University, two semesters (BIOL 1021).
-\item[2007--08]
-Teaching Assistant, Marine Mammology, Dalhousie University (BIOL 4060).
-\end{description}
-
-\section{Reviews}
-
-Reviewer for Ecology, Conservation Biology, Fish and Fisheries,
+Reviewer for Science, Ecology, Conservation Biology, Fish and Fisheries,
 International Journal of Tropical Biology and Conservation, Journal of
-Environmental Management, Aquatic Conservation
+Environmental Management, Endangered Species Research, Aquatic
+Conservation

### 2014-10-11 refs.R
Fix some silly code in refs.R

@@ -1,17 +1,17 @@
 # TODO J.E.M. Flemming should be J.E. Mills Flemming
 # TODO J.E. Mills Flemming needs to be fixed throughout refs.bib too
-# TODO Earth2Oceans Derby needs to be mannually fixed
+# TODO Earth2Oceans Derby needs to be manually fixed
 
 library("bibtex")
 d <- read.bib("refs.bib")
 
 ref <- list()
 
-for(i in seq_len(length(d))) {
+for(i in seq_along(d)) {
   name <- list()
   x <- d[[i]]
   # authors:
-  for(j in seq_len(length(x$author))) {
+  for(j in seq_along(x$author)) {
     init <- paste0(paste(substr(x$author$given[[j]], 1, 1), collapse = "."), ".")
     family <- x$author$family[[j]]
     if(j == 1) {

### 2014-10-13 _knitr-posts/2014-10-11-tmb.Rmd
Add draft TMB intro post

---
layout: post
title: The TMB R package
subtitle: Life beyond ADMB
published: false
---

I'd heard whispers of this new R package [TMB](https://github.com/kaskr/adcomp) and finally spent some time playing with it in the last few days. It's incredible — fast like ADMB (faster?), has built-in parallel functionality, has other advanced [statistical niceties](https://github.com/kaskr/adcomp/wiki), and it's built around R. What more could you ask for? (Well better documenation and the option to switch on Bayesian MCMC like you can in ADMB would be nice.)

The basic idea is to write a C++ template file and then compile and run it with some R functions. If you've used ADMB then the format will look similar. If you haven't, it's not too hard to pick up. It helps to [know a bit about C++](http://adv-r.had.co.nz/Rcpp.html), but for the most part it shouldn't be that hard for anyone familiar with R to pick it up.

The steps to running a TMB model are:

1. Write a C++ model template (a `.cpp` file).

2. Compile the model with `TMB::compile()`.

3. Load the model object with `dyn.load(TMB::dynlib())`

4. Make an object with data and parameters to pass to the TMB model with 
   `TMB::MakeADFun()`. If there are random effects / latent variables / unobserved
   states / coefficients that are allowed to vary (choose your preferred 
   terminology) then set the `random` argument in `TMB::MakeADFun()` to specify
   the name(s) of the parameters to integrate out of the likelihood.

5. Minimize the objective function with `optim()`.

6. Extract the estimates and standard errors from the model fit with 
   `TMB::sdreport()` and `TMB::summary.sdreport()`.

You can install the TMB package from GitHub:

```{r, eval=FALSE}
# install.packages("devtools") # if needed
devtools::install_github("kaskr/adcomp", subdir = "TMB")
```

The documentation is still very sparse, but you can find some instructions on [the R package GitHub wiki](https://github.com/kaskr/adcomp/wiki) and documentation of the underlying C++ functions and classes [here](http://folk.uib.no/hsk021/tmbdoc/).

## A linear regression example

Let's start with a really simple linear regression model.

We'll start by writing the TMB template file. Here, I've written it as inline R code that I then `write()` to a `.cpp` file. I have only done this so the code shows up here.

```{r tmb-regression}
tmb_model <- "
// linear regression
#include <TMB.hpp>

template<class Type>
Type objective_function<Type>::operator() () {
// data:
DATA_VECTOR(x);
DATA_VECTOR(y);

// parameters:
PARAMETER(a); // intercept
PARAMETER(b); // slope
PARAMETER(log_sigma); // log(residual SD)
// we fit sigma on a log scale to keep it > 0

// procedures: (transformed parameters)
Type sigma = exp(log_sigma);

int n = y.size(); // get number of data points to loop over

Type nll = 0.0; // initialize negative log likelihood

for(int i = 0; i < n; i++){ // C++ starts loops at 0!
  // get negative log likelihood (last argument is log = TRUE)
  nll -= dnorm(y[i], a + b * x[i], sigma, true);
}

return nll;
}"
write(tmb_model, file = "regression.cpp")
```

We'll load the model template:

```{r load, message=FALSE, results='hide'}
library(TMB)
compile("regression.cpp")
dyn.load(dynlib("regression"))
```

And simulate some data:

```{r tmb1, fig.height=5}
set.seed(1)
x <- runif(50, 1, 40)
y <- rnorm(50, mean = 2.5 + 0.3 * x, sd = exp(0.6))
plot(x, y)
```

Then we'll fit the model:

```{r, message=FALSE, results='hide', warning=FALSE}
obj <- MakeADFun(
  data = list(x = x, y = y), 
  parameters = list(a = 0, b = 0, log_sigma = 0),
  DLL = "regression")
opt <- do.call("optim", obj)
rep <- sdreport(obj)
```

We can check the estimates against the true values above (`a = 2.5, b = 0.3, log_sigma = 0.6`):

```{r}
rep
```

Which is exactly the same as we'd get from `stats::lm()`:

```{r lm}
arm::display(lm(y ~ x))
```

## Gompertz state space model

Now let's try something a bit more exciting — a state space Gompertz population dynamics model. We'll have TMB integrate out the unobserved "true" state that is masked by observation error. You can represent the Gompertz model as:

\\[\begin{aligned}
y_t &= \ln N\_t\\\      
U\_t &amp;= a + b U\_{t-1} + \epsilon\_t\\\
\epsilon_t &amp;\sim \mathrm{Normal}(0, \sigma^2\_\mathrm{proc})\\\
y\_t &amp;\sim \mathrm{Normal}(U\_t, \sigma^2\_\mathrm{obs}),
\end{aligned}\\]

where \\(y\_t\\) is the observed \\(\ln\\) abundance (\\(N\\)) at time \\(t\\). The model is density independent if \\(b = 1\\), maximally density dependent if \\(b = 0\\), and inversely density dependence if \\(b < 0\\). The parameter \\(a\\) represents the expected ln abundance at the next time step when \\(y\_t = 0\\). The process noise is modelled as a normal distribution with mean of \\(0\\), and a standard deviation of \\(\sigma\_\mathrm{proc}\\). \\(U\\) represents the unobserved state vector, and \\(\sigma\_\mathrm{obs}\\) represents the standard deviation of observation error (on a log scale).

The TMB model for this state-space Gompertz model is:
```{r tmb-model}
tmb_model <- "
// State-space Gompertz model
#include <TMB.hpp>

template<class Type>
Type objective_function<Type>::operator() () {
// data:
DATA_VECTOR(y);

// parameters:
PARAMETER(a); // population growth rate parameter
PARAMETER(b); // density dependence parameter
PARAMETER(log_sigma_proc); // log(process SD)
PARAMETER(log_sigma_obs); // log(observation SD)
PARAMETER_VECTOR(u); // unobserved state vector

// procedures: (transformed parameters)
Type sigma_proc = exp(log_sigma_proc);
Type sigma_obs = exp(log_sigma_obs);

// reports on transformed parameters:
ADREPORT(sigma_proc)
ADREPORT(sigma_obs)

int n = y.size(); // get time series length

Type nll = 0.0; // initialize negative log likelihood

// process model:
for(int i = 1; i < n; i++){
  Type m = a + b * u[i - 1]; // Gompertz
  nll -= dnorm(u[i], m, sigma_proc, true);
}

// observation model:
for(int i = 0; i < n; i++){
  nll -= dnorm(y[i], u[i], sigma_obs, true);
}

return nll;
}"
write(tmb_model, file = "gompertztmb.cpp")
```

We'll compile the model:

```{r compile, message=FALSE, results='hide'}
compile("gompertztmb.cpp")
dyn.load(dynlib("gompertztmb"))
```

And write a function that simulates some data, fits the model, and plots the estimates:

```{r fitfunction}
sim_gomp <- function(N = 100, seed = 123, sigma_obs = 0.2,
  sigma_proc = 0.2, a = 1.4, b = 0.7, y1 = 4) {

  set.seed(seed)
  N <- N
  ytrue <- numeric(length = N)
  ytrue[1] <- y1
  log_sigma_proc <- log(sigma_proc)
  proc_error <- rnorm(N, mean = 0, sd = sigma_proc)
  log_sigma_obs <- log(sigma_obs)
  for(i in 2:N) {
    ytrue[i] <- a + b * ytrue[i-1] + proc_error[i-1]
    }
  x <- seq_len(N)
  y <- rnorm(N, mean = ytrue, sd = sigma_obs)

  data <- list(y = y)
  parameters <- list(a = 1, b = 0.5, log_sigma_proc = -1,
    log_sigma_obs = -1, u = rep(mean(y), N))
  obj <- MakeADFun(data, parameters, random = "u", DLL = "gompertztmb")
  obj$hessian <- FALSE
  opt <- do.call("optim", obj)
  rep <- sdreport(obj)

  # extract estimated process:
  u <- summary(rep, "random")[, "Estimate"]
  u_se <- summary(rep, "random")[, "Std. Error"]
  # extract fixed effects:
  fixed <- summary(rep, "fixed")

  # We'll write a little helper function to plot fixed effect 
  # estimates:
  plot_fixed <- function(par, y, mat, trans = I) {
    points(trans(eval(parse(text = par))), y, pch = 21,
      bg = "#00000030", col = "#00000080")
    f <- mat[par, "Estimate"]
    se <- mat[par, "Std. Error"]
    points(trans(f), y, pch = 19, col = "red")
    segments(trans(f - 2 * se), y, trans(f + 2 * se), y, pch = 19,
      col = "red")
    }

  par(mfrow = c(2, 1), mar = c(4, 6, 0, 0), oma = c(.5, .5, .5, .5),
    mgp = c(2, 0.5, 0), cex = 1, tck = -0.02)
  plot(1, 1, xlim = c(0, 2), ylim = c(1, 4), type = "n",
    xlab = "Coefficient value", ylab = "", yaxt = "n")
  axis(2, at = 1:4, labels = c("a", "b", "sigma_obs", "sigma_proc"),
    las = 1)

  plot_fixed("a", 1, fixed)
  plot_fixed("b", 2, fixed)
  plot_fixed("log_sigma_obs", 3, fixed, trans = function(z) exp(z))
  plot_fixed("log_sigma_proc", 4, fixed, trans = function(z) exp(z))

  plot(x, y, col = "#00000080", pch = 21, bg = "#00000030", las = 1,
    ylab = "log(abundance)", xlab = "time")
  lines(x, ytrue, lty = 2, lwd = 2)
  lines(x, u, col = "red", lty = 1, lwd = 1.5)
  polygon(c(x, rev(x)), c(u - 2 * u_se, rev(u + 2 * u_se)), 
    border = NA, col = "#FF000050")
  legend("bottomright", legend = c("Observed", "True", "Estimated"),
    pch = c(21, NA, NA), lty = c(NA, 2, 1), lwd = c(NA, 2, 1.5),
    col = c("#00000080", "black", "red"), 
    pt.bg = c("#00000030", NA, NA))
  }
```

One good test that you've coded a model correctly is to throw a boat-load of data at it and make sure it comes back with unbiased parameter estimates. In the following plots, the first panel shows the coefficient estimates ("fixed effects") +/- two standard errors. The red dots are our estimates and the shaded grey dots are the true values. In the lower panels, the red lines are our estimated states, the black line (hidden behind the red in the first plot) is the true underlying state, and the grey dots are our observations.

```{r tmb-test1, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 5000, sigma_obs = 0.1)
```

(We just modelled a 5000 time step state space model in a couple of seconds!)

Now we can try a more realistic 50 data points:

```{r tmb-short, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 50)
```

And with a different seed value but more data:

```{r tmb-seed2, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 100, seed = 999)
```

### 2014-10-17 _knitr-posts/2014-10-17-tmb.Rmd
Tweak TMB post and publish it

--- 2014-10-13 _knitr-posts/2014-10-11-tmb.Rmd
@@
 title: The TMB R package
-subtitle: Life beyond ADMB
-published: false
+subtitle: Life beyond ADMB?
+published: true
 ---
@@
 What more could you ask for?
-(Well better documenation and the option to switch on Bayesian MCMC like you can in ADMB would be nice.)
+(Well better documenation and the option to switch on Bayesian MCMC like you can in ADMB would be nice.) This post just represents some of my initial steps as I learned and played with the package.
 
-The basic idea is to write a C++ template file and then compile and run it with some R functions.
+The basic idea with TMB is to write a C++ template file and then compile and run it with some R functions.
 If you've used ADMB then the format will look similar.
@@
 
-5. Minimize the objective function with `optim()`.
+5. Minimize the objective function with something like `stats::optim()` or `stats::nlminb()`.
 
@@
 
-You can install the TMB package from GitHub:
+You can install the TMB package from GitHub.
+You'll need a C++ compiler installed.
+If you're on Windows then you'll also need Rtools.
+See specific instructions on the [TMB GitHub page](https://github.com/kaskr/adcomp).
 
@@
 ```{r tmb1, fig.height=5}
-set.seed(1)
-x <- runif(50, 1, 40)
-y <- rnorm(50, mean = 2.5 + 0.3 * x, sd = exp(0.6))
+set.seed(123)
+x <- runif(20, 1, 10)
+y <- rnorm(20, mean = 1.8 + 2.4 * x, sd = exp(0.3))
 plot(x, y)
@@
 
-We can check the estimates against the true values above (`a = 2.5, b = 0.3, log_sigma = 0.6`):
+We can check the estimates against the true values above (`a = 1.8, b = 2.4, log_sigma = 0.3`):
 
@@
 
-Which is exactly the same as we'd get from `stats::lm()`:
+Which is about the same as we'd get from `stats::lm()`:
 
@@
 ```{r tmb-test1, results='hide', message=FALSE, warning=FALSE, fig.height=8}
-sim_gomp(N = 5000, sigma_obs = 0.1)
+sim_gomp(N = 5000, sigma_obs = 0.1, sigma_proc = 0.2, b = 0.7, a = 1.4)
 ```

### 2014-10-19 cv-body.tex
Update CV; pandoc has also been upgraded
Add author preprint of salmon paper

@@
 \begin{description}
+\itemsep1pt\parskip0pt\parsep0pt
 \item[2011--14]
@@
 \begin{description}
+\itemsep1pt\parskip0pt\parsep0pt
+\item[2014]
+Hurtado-Ferro, F., C.S. Szuwalski, J.L. Valero, \textbf{S.C. Anderson},
+C.J. Cunningham, K.F. Johnson, R.R. Licandeo, C.R. McGilliard, C.C.
+Monahan, M.L. Muradian, K. Ono, K.A. Vert-Pre, A.R. Whitten, A.E. Punt.
+Looking in the rear-view mirror: bias and retrospective patterns in
+integrated, age-structured stock assessment models. In press at ICES
+Journal of Marine Science.
 \item[2014]
@@
 Ecological Applications. In press.
-\url{http://doi.org/10.1890/14-0266.1}.
+\url{http://doi.org/10.1890/14-0266.1}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2014_salmon_portfolios.pdf}{PDF}).
 \item[2014]
@@
 \begin{description}
+\itemsep1pt\parskip0pt\parsep0pt
 \item[2014]
@@
 review at Proceedings of the Royal Society B.
-\item[2014]
-Hurtado-Ferro, F., C.S. Szuwalski, J.L. Valero, \textbf{S.C. Anderson},
-C.J. Cunningham, K.F. Johnson, R.R. Licandeo, C.R. McGilliard, C.C.
-Monahan, M.L. Muradian, K. Ono, K.A. Vert-Pre, A.R. Whitten, A.E. Punt.
-Looking in the rear-view mirror: bias and retrospective patterns in
-integrated, age-structured stock assessment models. In review at ICES
-Journal of Marine Science.
 \end{description}
@@
 \begin{description}
+\itemsep1pt\parskip0pt\parsep0pt
 \item[2013]
@@
 \begin{description}
+\itemsep1pt\parskip0pt\parsep0pt
 \item[2013--14]
@@
 \begin{description}
+\itemsep1pt\parskip0pt\parsep0pt
 \item[2011--13]

### 2015-01-13 _knitr-posts/2014-11-11-prediction-intervals.Rmd
Add a draft post on prediction intervals

---
layout: post
title: Confidence intervals vs. prediction intervals in R
subtitle: What do they mean?
published: true
---

In R speak, there are confidence intervals and prediction intervals. Using loose terminology, confidence intervals refer to uncertainty in the coefficients and prediction intervals refer to the expected variation in additional observations---prediction intervals include both uncertainty in the coefficients and the residual sampling noise.

If you recorded the heights of a group of students in a class and calculated the mean, you might want to know how certain you are that your sample mean (from your class) matches what you'd get if you went out and recorded the height of all students at the same age across many schools (the population). The "confidence" interval helps you with this. As you collect more observations, your certainty in the mean increases and the confidence interval decreases.

If, instead, you wanted to know the expected range of student heights you'd observe if you surveyed a second class room, you'd want to use a prediction interval. Even if you had sampled 10,000 students, the expected variability in observations you'd get in another class room would be about the same. Prediction intervals can be estimated more precisely as you gain more data, but they shouldn't systematically get smaller.

Let's use a small example in R to illustrate this. First, we'll simulate some data and fit a linear model with only an intercept. This is equivalent to estimating the population mean.

```{r}
x <- rnorm(50, mean = 1.5, sd = 0.6)
m <- lm(x ~ 1)
```

We can get the two kinds of intervals in R like this:

```{r, warning = FALSE}
conf <- predict(m, se.fit = TRUE, interval = "confidence")
pred <- predict(m, se.fit = TRUE, interval = "prediction")
```

If you want to see more details on these options, see `?predict.lm`.

Now we'll plot these intervals:

```{r}
plot(x)
abline(h = 1.5, col = "red")
abline(h = c(conf$fit[1, "lwr"], conf$fit[1, "upr"]))
abline(h = c(pred$fit[1, "lwr"], pred$fit[1, "upr"]), lty = 2)
```

Notice how much wider the prediction intervals (dashed lines) are. They encompass most of the observed data points.

Now, let's do that repeatedly to see more clearly what these intervals refer to. First, we'll write a little function to do what we did above. We'll simulate some data, fit a model, and get confidence intervals from that model. Later, we'll plot these.

```{r, warning = FALSE}
sim_intervals <- function(n = 20, mean = 2, sd = 1) {
  x <- rnorm(n = n, mean = mean, sd = sd)
  m <- lm(x ~ 1)
  conf <- as.data.frame(t(predict(m, se.fit = TRUE, interval = "confidence")$fit[1, ]))
  pred <- as.data.frame(t(predict(m, se.fit = TRUE, interval = "prediction")$fit[1, ]))
  names(conf) <- paste0("conf_", names(conf))
  names(pred) <- paste0("pred_", names(pred))
  data.frame(sample = seq_along(x), x, mean, sd, conf, pred)
}
```

And we'll write a ggplot2 function to plot the data:

```{r}
plot_intervals <- function(dat) {
  library("ggplot2")
  ggplot(dat, aes(sample, x)) + geom_point(cex = 1.1) + facet_wrap(~id) +
  geom_hline(aes(yintercept = mean), colour = "red", lwd = 1) +
  geom_hline(aes(yintercept = c(conf_lwr, conf_upr))) +
  geom_hline(aes(yintercept = c(pred_lwr, pred_upr)), colour = "grey80") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    strip.background = element_blank())
}
```

We'll use the plyr package to run our function 20 times:

```{r, warning = FALSE}
set.seed(1)
out_n20 <- plyr::ldply(setNames(seq_len(20), seq_len(20)), function(i) sim_intervals(n = 20, mean = 2, sd = 0.5), .id = "id")
```

And again with more data:

```{r, warning = FALSE}
set.seed(1)
out_n100 <- plyr::ldply(setNames(seq_len(20), seq_len(20)), function(i) sim_intervals(n = 100, mean = 2, sd = 0.5), .id = "id")
```

And plot the first example:

```{r}
plot_intervals(out_n20)
```

So, roughly 1 time out of 20 the confidence interval (black lines) should miss containing the true mean value (red line). We see that in panel 9.

And, approximately 95% of the sample dots should be inside the prediction interval (grey lines). In other words, on average 1 dot should be outside the grey lines in each panel. In general that seems about right too.

Using a prediction interval on existing data doesn't make a lot of sense. You already know the observed data after all. Typically, you'd use the prediction interval for 'future' or 'additional' observations, especially in a scenario where you're doing more than just estimating a mean. Maybe you want to predict the variability in your observations at a different level of a predictor variable. Here, I'm showing them on the existing data just to illustrate a point. Further, this can be a useful way of checking your calculations to make sure they match the original data.

Let's compare the second example we ran above. In the second example, we had 100 observations instead of 20:

```{r}
plot_intervals(out_n100)
```

You'll notice that the confidence intervals have gotten much tighter (although on average 1 out of 20 will still miss the mark). The prediction intervals are just as wide as before and on average encompass 95% of the observations.

### 2015-01-16 cv-body.tex
Update cv
Update doi on lobster paper
Update CV

@@
 Looking in the rear-view mirror: bias and retrospective patterns in
-integrated, age-structured stock assessment models. In press at ICES
-Journal of Marine Science.
+integrated, age-structured stock assessment models. ICES Journal of
+Marine Science. In press. \url{http://doi.org/10.1093/icesjms/fsu198}.
 \item[2014]
 Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down and bottom-up
-forces interact at thermal range extremes on American lobster. Accepted
-with minor revisions at Journal of Animal Ecology.
+forces interact at thermal range extremes on American lobster. Journal
+of Animal Ecology. In press.
+\url{http://doi.org/10.1111/1365-2656.12322}.
 \item[2014]
@@
 (\href{https://dl.dropboxusercontent.com/u/254940/papers/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf}{PDF}).
-(*\emph{Authors contributed equally, listed in reverse alphabetical
-order}).
+(*Authors contributed equally, listed in reverse alphabetical order).
 \item[2012]
@@
 \itemsep1pt\parskip0pt\parsep0pt
-\item[2014]
-Finnegan, S., \textbf{S.C. Anderson}, P.G. Harnik, C. Simpson, D.P.
+\item[2015]
+Orzechowski, E.A., R. Lockwood, J.E. Byrnes, \textbf{S.C. Anderson}, S.
+Finnegan, Z.V. Finkel, P.G. Harnik, D.R. Lindberg, L.H. Liow, H.K.
+Lotze, C.R. McClain, J.L. McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson,
+D.P. Tittensor. Determinants of extinction selectivity over the last 500
+million years: a meta-analysis of marine bivalves and gastropods.
+Submitted.
+\item[2015]
+Finnegan*, S., \textbf{S.C. Anderson}*, P.G. Harnik*, C. Simpson, D.P.
 Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
@@
 Pandolfi. Paleontological baselines for evaluating extinction risk in
-the modern oceans. In revision.
-\item[2014]
+the modern oceans. In review. (*Authors contributed equally).
+\item[2015]
 Trebilco, R., N.K. Dulvy, \textbf{S.C. Anderson}, A.K. Salomon. The
 paradox of inverted biomass pyramids in kelp forest fish communities. In
-review at Proceedings of the Royal Society B.
+revision.
 \end{description}

### 2015-03-12 _knitr-posts/2015-02-02-stan-matt-trick.Rmd
Add 2 draft posts

---
layout: post
title: The Matt Trick in Stan
subtitle: What do they mean?
published: false
---

If you read about Stan for a while you'll invariably hear about the 'Matt Trick'. In fact, most of the example models coded in the GitHub wiki are coded with the Matt Trick. The Matt Trick is a way of writing a multilevel/hierarchical/mixed-effects model in Stan to create less-correlated, better-mixing sample chains. Thing is, I can't find one good explanation anywhere what the trick is, and every time I do it I have to cobble together bits and pieces of information. Then I quickly forget what I was doing. This post is mostly to document the 'trick' for myself as a reference.

We'll use a 'simple' multilevel model with one predictor and varying intercepts. We'll start by simulating some data, then we'll fit the model normally and fit the model with the 'trick'. Then we'll look at run times and mixing properties of the chains.

```{r}
set.seed(123)
N <- 50 # number of observations per group
Nj <- 16 # number of groups
sigma_j <- 2.0 # standard deviation of intercepts across groups
d <- data.frame(j = 1:Nj, jname = letters[1:Nj], b0_j = rnorm(Nj, 0, sigma_j))
sigma <- 1.1 # residual standard deviation
b0 <- 0.3 # intercept
b1 <- 0.5 # slope
x <- lapply(seq_len(Nj), function(x) rnorm(n = N)) # predictor
dat <- list()
for(j in d$j) {
  dat[[j]] <- data.frame(j = j, jname = d$jname[j], x = x[[j]],
    y = rnorm(N, mean = b0 + x[[j]] * b1 + d$b0_j[j], sd = sigma))
}
dat <- as.data.frame(do.call("rbind", dat))
library("ggplot2")
ggplot(dat, aes(x, y)) + geom_point() + facet_wrap(~jname)
```

The model coded normally:

```{r}
model <- "
data {
  int<lower=1> N;        // rows of data
  int<lower=1> Nj;      // number of groups
  int<lower=1,upper=Nj> j_id[N];  // group ids
  vector[N] x;       // predictor
  real y[N]; // response
}
parameters {
  real<lower=0> sigma;  // residual standard deviation
  real b0;              // intercept
  real b1;              // slope
  real<lower=0> sigma_j; // group-level standard deviation
  vector[Nj] b0_j;
}
transformed parameters {
  vector[N] yhat;
  for (i in 1:N) {
    yhat[i] <- b0_j[j_id[i]] + b1 * x[i];
  }
}
model {
  b0_j ~ normal(b0, sigma_j); 
  // priors:
  sigma ~ cauchy(0, 2.5);
  sigma_j ~ cauchy(0, 2.5);
  b0 ~ normal(0, 2.5);
  b1 ~ normal(0, 2.5);
  y ~ normal(yhat, sigma);
}
"
sm <- rstan::stan_model(model_code = model)
```

The model with the Matt Trick:

```{r}
model <- "
data {
  int<lower=1> N;                 // rows of data
  int<lower=1> Nj;                // number of groups
  int<lower=1,upper=Nj> j_id[N];  // group ids
  vector[N] x;                    // predictor
  real y[N];                      // response
}
parameters {
  real<lower=0> sigma;            // residual standard deviation
  real b0;                        // intercept
  real b1;                        // slope
  real<lower=0> sigma_j;          // group-level standard deviation
  vector[Nj] e_b0_j;              // Matt Trick #1
}
transformed parameters {
  vector[N] yhat;
  vector[Nj] b0_j;                // Matt Trick #2
  b0_j <- b0 + sigma_j * e_b0_j;  // Matt Trick #3

  for (i in 1:N) {
    yhat[i] <- b0_j[j_id[i]] + b1 * x[i];
  }
}
model {
  e_b0_j ~ normal(0, 1);          // Matt Trick #4
  sigma ~ cauchy(0, 2.5);
  sigma_j ~ cauchy(0, 2.5);
  b0 ~ normal(0, 2.5);
  b1 ~ normal(0, 2.5);
  y ~ normal(yhat, sigma);
}
"
sm_matt <- rstan::stan_model(model_code = model)
```

Now let's fit the models:

```{r}
model_data <- list(N = nrow(dat), y = dat$y, x = dat$x, Nj = max(dat$j), 
  j_id = dat$j)
pars <- c("b0", "b1", "sigma", "sigma_j")

st <- system.time({
  m <- sampling(sm, data = model_data, iter = 400, chains = 3, 
    pars = pars, seed = 999)
})
st_matt <- system.time({
  m_matt <- sampling(sm_matt, data = model_data, iter = 400, chains = 3, 
    pars = pars, seed = 999)
})
```

### 2015-03-12 _knitr-posts/poisson.Rmd
Add 2 draft posts

---
layout: post
title: "Overdispersion and Poisson GLMs"
subtitle: "Is mean equals variance enough?"
published: false
---

> One assumption of Poisson regression is that mean(Y)≈var(Y).
> 
> I have two count datasets with means = 1.16 x & 1.37 x variance,
> respectively. I am attempting to use a Poisson regression (lme4 package). I
> have many zeroes, but no values >5 (histograms are below).
> 
> An AICc comparison between the global model with Poisson distribution and the
> global model with a Negative Binomial distribution yields a lower AICc value
> for Poisson.

> Does mean≈variance pass muster? Or do I need to do an official test to
> justify that there is:
>
> A) no overdispersion
> B) no zero-inflation

Checking that the sample mean approximately equals the sample variance is suggestive that a Poisson GLM may suffice but not sufficient unless you're fitting an intercept-only model.

The *mean* and *variance* here refer to the expected values (the mean model fit) and the variance of the residuals around those expected values. In anything beyond an intercept-only model the expectation is therefore that the mean *scales* with the variance.

One way to check if you have overdispersion is to look at the `summary()` model output and compare the residual deviance to the residual degrees of freedom. The ratio of these two should be ballpark 1. If the residual deviance is much larger than the degrees of freedom then you have evidence of overdispersion. There are statistical tests for this if you want to go that route.

Another quick way to check and potentially fix the issue is to try a quasipoisson family. This will calculate that overdispersion value and scale the standard errors appropriately. But be aware that you aren't in strict likelihood anymore. Among other things, this means you can't just use AIC any more. But there's an appropriate alphabet soup for that too (QAIC). To use a quasipoisson model as a check, compare the standard errors or confidence intervals on your coefficients. If they are much larger in the quasipoisson model then you have overdispersion.

The quasipoisson is a simpler model to fit than the negative binomial because it has one less parameter to estimate. Essentially, the quasipoisson fits a Poisson distribution, calculates the dispersion, and then makes a correction. The negative binomial distribution can scale to represent data that are less or more dispersed than a Poisson distribution, but requires an extra parameter to do so. The Poisson assumes that if you give it a mean at any point in the model, the data will have exactly that same variance around the mean. That's a big assumption that is unlikely to be precisely met, but often it's close enough to be useful.

The lower AIC for the Poisson model over the negative binomial is suggestive that their isn't overdispersion (the data don't support estimating an additional scale parameter), but not necessarily sufficient either. It's still possible that that Poisson (or negative binomial) model doesn't capture some important aspect of your data. The way to check this is to plot, plot, plot. Plot both model fits through the data. Plot residuals vs. fitted and residuals vs. predictors. If you really want to be thorough then simulate data from the models and make sure they create realistic data (see Gelman and Hill 2007 or the big Gelman et al. Bayesian Data Analysis book or Ben Bolker's book).

Let's try a quick example using a model from the help for `glm()`:

```{r}
counts <- c(18,17,15,20,10,20,25,13,12)
outcome <- gl(3,1,9)
treatment <- gl(3,3)
print(d.AD <- data.frame(treatment, outcome, counts))
glm.D93 <- glm(counts ~ outcome + treatment, family = poisson())
summary(glm.D93)
```

Note that the residual deviance and residual degrees of freedom are pretty close (5.1291 and 4). Let's try adding a single larger count value to create some overdispersion:

```{r}
counts[7] <- 40
d.AD <- data.frame(treatment, outcome, counts)
glm.D93 <- glm(counts ~ outcome + treatment, family = poisson())
summary(glm.D93)
```

So, now we have some overdispersion (11.987/4 = 3). Let's try a quasipoisson model:

```{r}
glm.D93.quasi <- glm(counts ~ outcome + treatment, family = quasipoisson)
summary(glm.D93.quasi)
```

Notice how much larger standard errors are in the quasipoisson model. In the first model we were assuming the data had less variability than they did and so underestimating the standard errors. In the plot below, the left panel is the Poisson distribution and the right panel is quasipoisson.

```{r}
par(mfrow = c(1, 2))
arm::coefplot(glm.D93)
arm::coefplot(glm.D93.quasi)
```

> Follow-up:
>
> I can't test quasi-poisson in lme4 because it gave unreliable
> estimates (so they took it out of the package). No other package seems
> to allow me to test quasi-poisson with a random effect (in my case
> Site), although I'm happy to be corrected there.

I think MASS::glmmPQL will.

> When I ran a quasi-poisson on the full model without a random effect,
> I got similar standard errors to my Poisson model (also tested w/out
> random effect). My residual deviance was 151.59  on 113  degrees of
> freedom (seems ok). This may mean nothing given I removed my random
> effect though.

This is probably conservative (overdispersion in the GLMM is probably
less) given that the random effect will likely absorb some of the
added variability, but yes, a bit hard to tell.

> The summary of my full lme4 model (with random effect) gives me
> deviance (but not residual deviance) and no report of degrees of
> freedom. Is there a way to extract this from a glmm?

One of the best resources on this issue and just about anything to do
with GLMMs in R is: <http://glmm.wikidot.com/faq#toc19>

I think that's largely maintained by Ben Bolker and says it better
than I ever could, so best to just read that. The wiki includes a
function to give an approximation of whether you have overdispersion
in a GLMM and advice on fixes. One approach that is mentioned there is
to try adding an individual (i.e. data-row-level) random effect and
see how that affects the model. If you think about it a bit this works
because you're allowing for and estimating some added stochastic
process that creates extra variability (overdispersion) at the
individual data point level. If the variance of this random effect is
zero then you're back to the original Poisson model.

> If I am satisfied that these data are neither overdispersed or require
> zero-inflation, is it statistically kosher to test all subsets of the
> global model and compare via AICc to determine the best model.

This is a whole other can of statistical worms. It is something that
people often do in ecology but it would make Burnam and Anderson,
among others, pretty squeamish. It also partly depends on how much the
goal of your modelling is prediction (out-of-sample performance is
then quite important and AIC or even better cross-validation is very
useful), how much is exploratory (maybe you can get away with this),
how much is confirmatory (run away from this), and how much is for
understanding (fit as complex a model as you can interpret and that a
power analysis indicates is reasonable).

> I will also then standardize the data to produce an RVI plot.

I don't think there's a need to standardize the data just for a
relative variable important plot (although there are other good
reasons to standardize predictors such as for comparing the magnitude
of effect sizes, for interpreting interactions, for comparing models
with and without interactions, and for computational reasons).

> How do people generally write about these things in their methods?
> i.e. We tested for overdispersion and the residual deviance was
> approximately the residual degrees of freedom.

I don't think people tend to say much more, and often don't say
anything at all, not that that's a good thing. If possible I would
probably quantify whatever statement you make so people don't have to
take your word for it and get an idea of what you mean by
'approximately'. Alternatively, fit a model with a negative binomial
distribution (or any of the other strategies mentioned or linked to
above) and mention if your conclusions were unchanged and the results
were qualitatively similar.

> There is generally a high chance that "we just don't know what these
> other processes are and are happy to wash it away in the error term".

Yep, that's true for much of ecology, which is fine as long as we
represent the uncertainty appropriately. And you're putting lots of
thought into that aspect, which is good!

### 2015-03-12 cv-body.tex
Update CV - yay Smith!
Fix typo in CV
Update CV
Candidate -> PhD

@@
 \itemsep1pt\parskip0pt\parsep0pt
-\item[2011--14]
-Ph.D.\ Candidate, Simon Fraser University, Burnaby, Canada\\ Supervisors:
+\item[2011--15]
+Ph.D.\ Biology, Simon Fraser University, Burnaby, Canada\\ Supervisors:
 Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper\\ Canadian Fulbright
@@
 \itemsep1pt\parskip0pt\parsep0pt
-\item[2014]
+\item[2015]
 Hurtado-Ferro, F., C.S. Szuwalski, J.L. Valero, \textbf{S.C. Anderson},
 C.J. Cunningham, K.F. Johnson, R.R. Licandeo, C.R. McGilliard, C.C.
-Monahan, M.L. Muradian, K. Ono, K.A. Vert-Pre, A.R. Whitten, A.E. Punt.
+Monnahan, M.L. Muradian, K. Ono, K.A. Vert-Pre, A.R. Whitten, A.E. Punt.
 Looking in the rear-view mirror: bias and retrospective patterns in
@@
 Marine Science. In press. \url{http://doi.org/10.1093/icesjms/fsu198}.
-\item[2014]
+\item[2015]
 Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down and bottom-up
@@
 \url{http://doi.org/10.1111/1365-2656.12322}.
-\item[2014]
+\item[2015]
 \textbf{Anderson, S.C.}, J.W. Moore, M.M. McClure, N.K. Dulvy, A.B.
 Cooper. Portfolio conservation of metapopulations under climate change.
-Ecological Applications. In press.
+Ecological Applications. 2015. 25(2): 559--572.
 \url{http://doi.org/10.1890/14-0266.1}
@@
 D.P. Tittensor. Determinants of extinction selectivity over the last 500
-million years: a meta-analysis of marine bivalves and gastropods.
-Submitted.
+million years: a meta-analysis of marine bivalves and gastropods. In
+revision.
 \item[2015]
@@
 \itemsep1pt\parskip0pt\parsep0pt
+\item[2015]
+David H. Smith Conservation Postdoctoal Research Fellowship
+\item[2015]
+Natural Sciences and Engineering Research Council of Canada Postdoctoral
+Fellowship
 \item[2014]

### 2015-03-24 cv-body.tex
CV tweaks

@@
 \item[2011--15]
-Ph.D.\ Biology, Simon Fraser University, Burnaby, Canada\\ Supervisors:
-Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper\\ Canadian Fulbright
-Scholar to the University of Washington in 2012--13
+Ph.D.\ Biology, Simon Fraser University, Burnaby, Canada\\ Variance and
+extreme events in population ecology\\ Supervisors: Dr.\ Nicholas K.
+Dulvy and Dr.\ Andrew B. Cooper\\ Canadian Fulbright Scholar to the
+University of Washington in 2012--13
 \item[2008--10]
-M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\ Thesis title:
-Trends, drivers, and ecosystem effects of expanding global invertebrate
-fisheries. Supervisor: Dr.\ Heike K. Lotze.
+M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\ Trends, drivers,
+and ecosystem effects of expanding global invertebrate fisheries\\
+Supervisor: Dr.\ Heike K. Lotze.
 (\href{http://seananderson.ca/papers/Anderson_MSc_2010.pdf}{PDF})\\
@@
 \item[2015]
-Hurtado-Ferro, F., C.S. Szuwalski, J.L. Valero, \textbf{S.C. Anderson},
-C.J. Cunningham, K.F. Johnson, R.R. Licandeo, C.R. McGilliard, C.C.
-Monnahan, M.L. Muradian, K. Ono, K.A. Vert-Pre, A.R. Whitten, A.E. Punt.
-Looking in the rear-view mirror: bias and retrospective patterns in
-integrated, age-structured stock assessment models. ICES Journal of
-Marine Science. In press. \url{http://doi.org/10.1093/icesjms/fsu198}.
+Finnegan*, S., \textbf{S.C. Anderson}*, P.G. Harnik*, C. Simpson, D.P.
+Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
+Lockwood, H.K. Lotze, C.M. McClain, J.L. McGuire, A. O'Dea, J.M.
+Pandolfi. Paleontological baselines for evaluating extinction risk in
+the modern oceans. In press. (\emph{*Authors contributed equally}).
 \item[2015]
@@
 \item[2015]
+Hurtado-Ferro, F., C.S. Szuwalski, J.L. Valero, \textbf{S.C. Anderson},
+C.J. Cunningham, K.F. Johnson, R.R. Licandeo, C.R. McGilliard, C.C.
+Monnahan, M.L. Muradian, K. Ono, K.A. Vert-Pre, A.R. Whitten, A.E. Punt.
+Looking in the rear-view mirror: bias and retrospective patterns in
+integrated, age-structured stock assessment models. ICES Journal of
+Marine Science. 72(1): 99--110.
+\url{http://doi.org/10.1093/icesjms/fsu198}.
+\item[2015]
 \textbf{Anderson, S.C.}, J.W. Moore, M.M. McClure, N.K. Dulvy, A.B.
@@
 (\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2014_salmon_portfolios.pdf}{PDF}).
+\item[2015]
+Johnson, K.F., C.C. Monnahan, C.R. McGilliard, K.A. Vert-pre,
+\textbf{S.C. Anderson}, C.J. Cunningham, F. Hurtado-Ferro, R.R.
+Licandeo, M.L. Muradian, K. Ono, C.S. Szuwalski, J.L. Valero, A.R.
+Whitten, A.E. Punt. Time-varying natural mortality in fisheries stock
+assessment models: identifying a default approach. ICES Journal of
+Marine Science. 72(1): 137--150.
+\url{http://doi.org/10.1093/icesjms/fsu055}
+(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
+\item[2015]
+Ono, K., R. Licandeo, M.L. Muradian, C.J. Cunningham, \textbf{S.C.
+Anderson}, F. Hurtado-Ferro, K.F. Johnson, C.R. McGilliard, C.C.
+Monnahan, C.S. Szuwalski, J.L. Valero, K.A. Vert-pre, A.R. Whitten, A.E.
+Punt. The importance of length and age composition data in statistical
+catch-at-age models for marine species. ICES Journal of Marine Science.
+72(1): 31--43. \url{http://doi.org/10.1093/icesjms/fsu007}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Ono_etal_2014_importance_of_length_and_age_composition_data.pdf}{PDF}).
 \item[2014]
@@
 \href{https://dl.dropboxusercontent.com/u/254940/ss3sim-vignette.pdf}{vignette}).
-\item[2014]
-Johnson, K.F., C.C. Monnahan, C.R. McGilliard, K.A. Vert-pre,
-\textbf{S.C. Anderson}, C.J. Cunningham, F. Hurtado-Ferro, R.R.
-Licandeo, M.L. Muradian, K. Ono, C.S. Szuwalski, J.L. Valero, A.R.
-Whitten, A.E. Punt. Time-varying natural mortality in fisheries stock
-assessment models: identifying a default approach. ICES Journal of
-Marine Science. In press. \url{http://doi.org/10.1093/icesjms/fsu055}
-(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
-\item[2014]
-Ono, K., R. Licandeo, M.L. Muradian, C.J. Cunningham, \textbf{S.C.
-Anderson}, F. Hurtado-Ferro, K.F. Johnson, C.R. McGilliard, C.C.
-Monnahan, C.S. Szuwalski, J.L. Valero, K.A. Vert-pre, A.R. Whitten, A.E.
-Punt. The importance of length and age composition data in statistical
-catch-at-age models for marine species. ICES Journal of Marine Science.
-In press. \url{http://doi.org/10.1093/icesjms/fsu007}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Ono_etal_2014_importance_of_length_and_age_composition_data.pdf}{PDF}).
 \item[2014]
@@
 \item[2015]
-Finnegan*, S., \textbf{S.C. Anderson}*, P.G. Harnik*, C. Simpson, D.P.
-Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
-Lockwood, H.K. Lotze, C.M. McClain, J.L. McGuire, A. O'Dea, J.M.
-Pandolfi. Paleontological baselines for evaluating extinction risk in
-the modern oceans. In review. (*Authors contributed equally).
-\item[2015]
 Trebilco, R., N.K. Dulvy, \textbf{S.C. Anderson}, A.K. Salomon. The
@@
 \item[2015]
-David H. Smith Conservation Postdoctoal Research Fellowship
+David H. Smith Conservation Postdoctoral Research Fellowship
 \item[2015]
@@
 \itemsep1pt\parskip0pt\parsep0pt
+\item[2014--15]
+Developed self-directed lecture and exercises on ggplot2 for FISH 554:
+Beautiful graphics in R, School of Aquatic and Fishery Sciences,
+University of Washington, Seattle, WA, United States.\\
+\url{http://seananderson.ca/ggplot2-FISH554/}
 \item[2013--14]
@@
 School of Resource and Environmental Management (REM).
-\item[2014]
-Developed self-directed lecture and exercises on ggplot2 for FISH 554:
-Beautiful graphics in R, School of Aquatic and Fishery Sciences,
-University of Washington, Seattle, WA, United States.\\
-\url{https://github.com/seananderson/ggplot2-FISH554}
 \item[2013]
@@
 \item[2013]
-Co-developed a set of maps showing trends in global fisheries status for
-the front page of \href{http://ramlegacy.org/}{ramlegacy.org}, the RAM
-Legacy Stock Assessment Database (for Dr.\ Ray Hilborn).
-\item[2011--12]
-Database management for the project \emph{Global Shark Abundance
-Baselines} (Hopkins Marine Station, Stanford University, CA, United
-States) with F. Micheli, F. Ferretti, N.K. Dulvy, H.K. Lotze, and B.
-Worm.
+Co-developed global fisheries status maps for the
+\href{http://ramlegacy.org/}{RAM Legacy Stock Assessment Database} for
+R. Hilborn
 \item[2010]

### 2015-04-11 cv-body.tex
Update cv

@@
 \itemsep1pt\parskip0pt\parsep0pt
+\item[2015]
+Orzechowski, E.A., R. Lockwood, J.E. Byrnes, \textbf{S.C. Anderson}, S.
+Finnegan, Z.V. Finkel, P.G. Harnik, D.R. Lindberg, L.H. Liow, H.K.
+Lotze, C.R. McClain, J.L. McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson,
+D.P. Tittensor. Marine extinction risk shaped by trait-environment
+interactions over 500 million years. In press at Global Change Biology.
 \item[2015]
@@
 \item[2015]
-Orzechowski, E.A., R. Lockwood, J.E. Byrnes, \textbf{S.C. Anderson}, S.
-Finnegan, Z.V. Finkel, P.G. Harnik, D.R. Lindberg, L.H. Liow, H.K.
-Lotze, C.R. McClain, J.L. McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson,
-D.P. Tittensor. Determinants of extinction selectivity over the last 500
-million years: a meta-analysis of marine bivalves and gastropods. In
-revision.
-\item[2015]
 Trebilco, R., N.K. Dulvy, \textbf{S.C. Anderson}, A.K. Salomon. The
 paradox of inverted biomass pyramids in kelp forest fish communities. In
-revision.
+review.
 \end{description}
@@
 David H. Smith Conservation Postdoctoral Research Fellowship
-\item[2015]
-Natural Sciences and Engineering Research Council of Canada Postdoctoral
-Fellowship
 \item[2014]

### 2015-05-08 cv-body.tex
Update CV

@@
 Ph.D.\ Biology, Simon Fraser University, Burnaby, Canada\\ Variance and
-extreme events in population ecology\\ Supervisors: Dr.\ Nicholas K.
-Dulvy and Dr.\ Andrew B. Cooper\\ Canadian Fulbright Scholar to the
-University of Washington in 2012--13
+extreme events in population ecology\\ Fulbright Scholar to the
+University of Washington in 2012--13\\ Dean of Graduate Studies
+Convocation Medal
 \item[2008--10]
@@
 and ecosystem effects of expanding global invertebrate fisheries\\
-Supervisor: Dr.\ Heike K. Lotze.
-(\href{http://seananderson.ca/papers/Anderson_MSc_2010.pdf}{PDF})\\
 Canadian \href{http://goo.gl/nA1zE}{Governor General's Academic Gold
-Medal} for the top-ranked Master's Natural Sciences and Engineering
-thesis at Dalhousie University in 2010
+Medal}
 \item[2004--07]
 B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
-Canada\\ Thesis title: How sustainable are emerging low-trophic level
-fisheries on the Scotian Shelf? Supervisors: Dr.\ Heike K. Lotze and
-Dr.\ Ransom A. Myers.
-(\href{http://seananderson.ca/papers/Anderson_HSc_2006.pdf}{PDF})\\
-First Class Honours \& Environmental Programmes Honour Society Medal
+Canada\\ First Class Honours \& Environmental Programmes Honour Society
+Medal
 \item[2001--03]
 B.Sc.\ Candidate (Hons.) Kinesiology, University of Waterloo, Waterloo,
-Canada Transferred to Dalhousie University in 2004.
+Canada; transferred to Dalhousie University in 2004.
 \end{description}
@@
 \item[2015]
+Finnegan*, S., \textbf{S.C. Anderson}*, P.G. Harnik*, C. Simpson, D.P.
+Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
+Lockwood, H.K. Lotze, C.M. McClain, J.L. McGuire, A. O'Dea, J.M.
+Pandolfi. Paleontological baselines for evaluating extinction risk in
+the modern oceans. Science. 348(6234): 567--570.
+\url{http://doi.org/10.1126/science.aaa6635} (\emph{*Authors contributed
+equally}).
+\item[2015]
 Orzechowski, E.A., R. Lockwood, J.E. Byrnes, \textbf{S.C. Anderson}, S.
@@
 \item[2015]
-Finnegan*, S., \textbf{S.C. Anderson}*, P.G. Harnik*, C. Simpson, D.P.
-Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
-Lockwood, H.K. Lotze, C.M. McClain, J.L. McGuire, A. O'Dea, J.M.
-Pandolfi. Paleontological baselines for evaluating extinction risk in
-the modern oceans. In press. (\emph{*Authors contributed equally}).
-\item[2015]
 Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down and bottom-up
 forces interact at thermal range extremes on American lobster. Journal
-of Animal Ecology. In press.
+of Animal Ecology. 83(3):840--850.
 \url{http://doi.org/10.1111/1365-2656.12322}.
@@
 \item[2015]
+Monnahan, C.C., K. Ono, S.C Anderson, M.B. Rudd, A.C. Hicks, F.
+Hurtado-Ferro, K.F. Johnson, P.T. Kuriyama, R.R. Licandeo, C.C. Stawitz,
+I.G. Taylor, J.L. Valero. The effect of length bin structures on growth
+estimation in integrated age-structured stock assessments. Submitted.
+\item[2015]
+Kuriyamaa, P.T., F. Hurtado-Ferro, K. Ono, A.C. Hicks, I.G. Taylor, R.R.
+Licandeo, K.F. Johnson, S.C. Anderson, C.C. Monnahan, M.B. Rudd, C.C.
+Stawitz, J.L. Valero. An empirical weight-at-age approach reduces
+estimation bias compared to modeling parametric growth in integrated,
+statistical stock assessment models when growth is time varying.
+Submitted.
+\item[2015]
 Trebilco, R., N.K. Dulvy, \textbf{S.C. Anderson}, A.K. Salomon. The
 paradox of inverted biomass pyramids in kelp forest fish communities. In
-review.
+revision.
 \end{description}

### 2015-05-25 _knitr-posts/2015-05-25-stan-ricker-missing-values.Rmd
Add draft post on state-space Stan models w missing values

---
layout: post
title: Stan state-space models with missing values
subtitle: Ricker-logistic models with imputation
published: true
---

```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(fig.align='center', fig.pos="htpb", cache=TRUE, echo=TRUE,
message=FALSE, autodep=TRUE, fig.width=7, fig.height=5)
opts_chunk$set(warning=FALSE, tidy=FALSE, refresh=TRUE, fig.retina=2)
```

The following is some code I played with to understand imputing missing values
with ecological state-space models in Stan. I simulate some data from a
Ricker-logistic model, show how to fit a basic model in R, translate that model
to Stan, develop a Stan model to impute missing values while fitting the model,
and finally turn the model into a spate-space model by also estimating
observation error.

Stan can handle missing value imputation quite well, although the syntax to
write the models is a bit awkward because it can't handle NA values
implicitely. You have to feed it fake data and then estimate the missing values
as parameters. I couldn't find a good example of this anywhere, so I thought
I'd play with some simulated data to figure it out. 
[This thread](https://groups.google.com/forum/#!searchin/stan-users/missing/stan-users/y9BalO9ILF0/ErbERTPsFbMJ)
on the Stan mailing list got me pointed in the right direction.

Please take everything I've written with some skepticism because I've written this code very quickly and could very well have made mistakes.

```{r stan-impute1}
set.seed(1)
N <- 50         # number of time steps
x <- 1:N        # time step values
y <- rep(NA, N) # vector to hold (scaled) abundance values
y[1] <- 1       # initial scaled abundance
sigma_p <- 0.6  # process noise standard deviation
r <- 0.8        # Ricker productivity parameter
K <- 50         # Carrying capacity
# Mean-bias corrected process noise:
proc_noise <- rnorm(N, -(sigma_p^2)/2, sigma_p)

# Ricker-logistic dynamics:
ricker <- function(spawners, r, K, v_t) {
  spawners * exp(r * (1 - spawners / K) + v_t)
}

for (i in seq(1, N - 1)) {
  y[i + 1] <- ricker(y[i], r = 0.8, K = 50, v_t = proc_noise[i])
}

par(yaxs = "i", las = 1)
plot(x, y, type = "o", pch = 20, ylim = c(0, max(y) * 1.1))
```

Let's start by looking at how we can quickly fit a Ricker-logistic model in R. We'll fit `log(R/S)` as the response and `S` as the predictor, where `R` is the recruits and `S` is the spawners. In this case, `R` and `S` are a time-shifted version of each other.

```{r stan-impute2}
plot(log(y[2:N] / y[1:(N - 1)]) ~ y[1:(N - 1)])
m1 <- lm(log(y[2:N] / y[1:(N - 1)]) ~ y[1:(N - 1)])
abline(m1)
coef(m1)[[1]]                # r
-coef(m1)[[1]]/coef(m1)[[2]] # K
```

Now, let's fit that same model in Stan:

```{r, results="hide"}
stan_model <-
'data {
  int<lower=0> N; // rows of data
  vector[N] r_obs; // observed growth rates for time t
  vector[N] Nt; // numbers at time t
}
parameters {
  real b;
  real a;
  real<lower=0> sigma_proc;
}
model {
  sigma_proc ~ cauchy(0, 5);
  a ~ cauchy(0, 5);
  b ~ cauchy(0, 5);
  r_obs ~ normal(a + b * Nt, sigma_proc);
}
generated quantities {
  real K;
  for (i in 1:N) {
    K <- -a/b;
  }
}
'
library("rstan")
stan_ricker <- stan_model(model_code = stan_model)
```

```{r, results="hide"}
r_obs <- c(NA, diff(log(y)))
m1_stan <- sampling(
  stan_ricker,
  data = list(N = length(y) - 1, r_obs = r_obs[-1], Nt = y[-length(y)]))
```

```{r}
print(m1_stan)
```

Let's write the same model in a format that makes the time-series nature of the data clearer. This will make it easier to include missing data in the model later. We're also going to switch to working on a log scale. We'll use the variable `y` to refer to `log(abundance)` inside the Stan model.

```{r, results="hide"}
stan_model2 <-
'data {
  int<lower=0> N; // rows of data
  vector[N] y; // log(numbers) at time t
  real<lower=0> K_upper; // upper boundary on K
}
parameters {
  real<lower=0, upper=K_upper> K;
  real r;
  real<lower=0> sigma_proc;
}
model {
  sigma_proc ~ cauchy(0, 5);
  r ~ cauchy(0, 5);
  for (i in 2:N) {
   y[i] ~ normal(y[i-1] + r * (1 - exp(y[i-1]) / K), sigma_proc);
  }
}
'
stan_ricker2 <- stan_model(model_code = stan_model2)
```

```{r, results="hide"}
m2_stan <- sampling(stan_ricker2, 
  data = list(N = N, y = log(y), K_upper = max(y) * 2))
```

It's a very slightly different model because of the priors and the starting point, but we get essentially the same posterior:

```{r}
print(m2_stan)
```

Now we'll introduce missing values. First, we'll write a Stan model. The
imputation part is relatively straightforward conceptually, but a bit nasty to
write out in Stan. We add `Nmissing`, which defines how many missing values we
have. We add `missing_id`, which is a vector of IDs indicating which values are
missing. For example, if the 10th and 20th values were missing then
`missing_id` would equal `c(10, 20)`. We add a vector to hold the estimated
missing values `y_missing`. We then feed all the observed data into a vector
called `y_impute` and fill in the values we need to impute with estimated
parameter values from `y_missing`. We then proceed with the same model as before.

```{r, results="hide"}
stan_model_missing <-
'data {
  int<lower=0> N; // rows of data
  vector[N] y; // log(numbers) at time t
  real<lower=0> K_upper; // upper boundary on K
  int<lower=0> Nmissing; // number of missing values
  int missing_id[Nmissing]; // IDs of missing values
}
parameters {
  real<lower=0, upper=K_upper> K;
  real r;
  real<lower=0> sigma_proc;
  real y_missing[Nmissing]; // imputed missing values
}
transformed parameters {
  vector[N] y_impute; // y with some imputed
  for (i in 1:N) {
    y_impute[i] <- y[i];
  }
  for (i in 1:Nmissing) {
    y_impute[missing_id[i]] <- y_missing[i];
  }
}
model {
  sigma_proc ~ cauchy(0, 5);
  r ~ cauchy(0, 5);
  for (i in 2:N) {
   y_impute[i] ~
     normal(y_impute[i-1] + r * (1 - exp(y_impute[i-1]) / K),
            sigma_proc);
  }
}
'
stan_ricker_missing <- stan_model(model_code = stan_model_missing)
```

```{r, results="hide"}
missing_id <- sample(seq_len(N), 4)
y_missing <- y
y_missing[missing_id] <- 9999 # Stan can't handle NAs; give it fake data
m_missing <- sampling(
  stan_ricker_missing,
  data = list(N = N, y = log(y_missing), 
    K_upper = max(y_missing[-missing_id]) * 2, 
    Nmissing = length(missing_id), missing_id = missing_id),
  iter = 4000,
  pars = c("K", "r", "sigma_proc", "y_impute"))
```

Let's plot the model output:

```{r stan-impute3, fig.width=8, fig.height=7}
p <- extract(m_missing)
layout(rbind(c(1, 2, 3), c(4, 4, 4)))
plot(density(p$r));abline(v = r, col = "blue", lwd = 2)
plot(density(p$K));abline(v = K, col = "blue", lwd = 2)
plot(density(p$sigma_proc));abline(v = sigma_p, col = "blue", lwd = 2)

pch <- rep(20, N)
col <- rep("black", N)
pch[missing_id] <- 21
col[missing_id] <- "blue"
plot(x, y, type = "o", pch = pch, ylim = c(0, max(y_missing[-missing_id]) * 1.1), col = col)
imputed <- exp(apply(p$y_impute, 2, median))
imputed_l <- exp(apply(p$y_impute, 2, quantile, prob = 0.1))
imputed_u <- exp(apply(p$y_impute, 2, quantile, prob = 0.9))
imputed_ll <- exp(apply(p$y_impute, 2, quantile, prob = 0.25))
imputed_uu <- exp(apply(p$y_impute, 2, quantile, prob = 0.75))
points(x[missing_id], imputed[missing_id], col = "red", pch = 21, cex = 1.5)
segments(x[missing_id], imputed_l[missing_id], 
  x[missing_id], imputed_u[missing_id], col = "red", lwd = 1)
segments(x[missing_id], imputed_ll[missing_id], 
  x[missing_id], imputed_uu[missing_id], col = "red", lwd = 2.5)
```

Now with observation error and missing values. We can add the estimation of observation error and make this a state-space model by adding only a few lines of Stan code. We'll add `sigma_obs`, which estimates the standard deviation of observation error. We'll then add a vector `U`, which represents the unobserved 'true' states. The observation error gets modelled as normally distributed (on a log scale) around `U`.

```{r, results="hide"}
stan_model_missing_obs <-
'data {
  int<lower=0> N; // rows of data
  vector[N] y; // numbers at time t
  real<lower=0> K_upper; // upper boundary on K
  int<lower=0> Nmissing; // number of missing values
  int missing_id[Nmissing];
}
parameters {
  real<lower=0, upper=K_upper> K;
  real r;
  real<lower=0> sigma_proc;
  real<lower=0> sigma_obs;
  real y_missing[Nmissing]; // imputed missing values
  vector[N] U; // states
}
transformed parameters {
  vector[N] y_impute; // Nt with some imputed
  for (i in 1:N) {
    y_impute[i] <- y[i];
  }
  for (i in 1:Nmissing) {
    y_impute[missing_id[i]] <- y_missing[i];
  }
}
model {
  sigma_proc ~ cauchy(0, 5);
  r ~ cauchy(0, 5);
  for (i in 2:N) {
   U[i] ~
     normal(U[i-1] + r * (1 - exp(U[i-1]) / K), sigma_proc);
  }
  y_impute ~ normal(U, sigma_obs);
}
'
stan_ricker_missing_obs <- stan_model(model_code = stan_model_missing_obs)
```

We'll need to run the model for some extra iterations, because parsing out process and observation error is challenging:

```{r, results="hide"}
# add some observation error to the data:
sigma_obs <- 0.2
y_missing_obs <- y_missing * rlnorm(N, 0, sigma_obs)
m_missing_obs <- sampling(
  stan_ricker_missing_obs,
  data = list(N = N, y = log(y_missing_obs), K_upper = max(y_missing_obs[-missing_id]) * 2, 
    Nmissing = length(missing_id), missing_id = missing_id),
  iter = 6000,
  pars = c("K", "r", "sigma_proc", "sigma_obs", "y_impute", "U"))
```

And plot the output:

```{r stan-impute5 stan-impute4, fig.width=8, fig.height=7}
p <- extract(m_missing_obs)
layout(rbind(c(1, 2, 3, 4), c(5, 5, 5, 5)))
plot(density(p$r));abline(v = r, col = "blue", lwd = 2)
plot(density(p$K));abline(v = K, col = "blue", lwd = 2)
plot(density(p$sigma_proc));abline(v = sigma_p, col = "blue", lwd = 2)
plot(density(p$sigma_obs));abline(v = sigma_obs, col = "blue", lwd = 2)

y_missing_obs[missing_id] <- NA
pch <- rep(20, N)
col <- rep("black", N)
pch[missing_id] <- 21
col[missing_id] <- "blue"
plot(x, y, type = "o", pch = pch, cex = 1.5, 
  ylim = c(0, max(y_missing_obs, na.rm = TRUE) * 1.1), col = "grey40")
points(x, y_missing_obs, col = "grey60", pch = 4)
U <- exp(apply(p$U, 2, median))
U_l <- exp(apply(p$U, 2, quantile, prob = 0.1))
U_u <- exp(apply(p$U, 2, quantile, prob = 0.9))
U_ll <- exp(apply(p$U, 2, quantile, prob = 0.25))
U_uu <- exp(apply(p$U, 2, quantile, prob = 0.75))

polygon(c(x, rev(x)), c(U_l, rev(U_u)), border = NA, col = "#FF000020")
polygon(c(x, rev(x)), c(U_ll, rev(U_uu)), border = NA, col = "#FF000030")
lines(x, U, col = "#FF000060", lwd = 3)
```

In the above, the black dots and lines are the true underlying data. The open circles were missing when the model was run. The grey crosses indicate the observed data. The red line and shaded regions indicate the posterior median, and 50% and 90% credible intervals of the estimated true abundance.

### 2015-05-26 cv-body.tex
Add GitHub repo to CV
Add GitHub repos to CV papers

@@
 \begin{description}
-\itemsep1pt\parskip0pt\parsep0pt
+\itemsep3pt\parskip0pt\parsep0pt
 \item[2011--15]
@@
 \begin{description}
-\itemsep1pt\parskip0pt\parsep0pt
+\itemsep3pt\parskip0pt\parsep0pt
 \item[2015]
@@
 the modern oceans. Science. 348(6234): 567--570.
-\url{http://doi.org/10.1126/science.aaa6635} (\emph{*Authors contributed
-equally}).
+\url{http://doi.org/10.1126/science.aaa6635}
+(\href{https://github.com/seananderson/paleobaselines}{GitHub
+repository}) (\emph{*Authors contributed equally}).
 \item[2015]
@@
 \url{http://doi.org/10.1111/1365-2656.12322}.
+(\href{https://github.com/seananderson/lobsters-predators}{GitHub
+repository})
 \item[2015]
@@
 \url{http://doi.org/10.1890/14-0266.1}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2014_salmon_portfolios.pdf}{PDF}).
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2014_salmon_portfolios.pdf}{PDF},
+\href{https://github.com/seananderson/metafolio}{GitHub repository},
+\href{http://cran.r-project.org/web/packages/metafolio/index.html}{CRAN
+R package}).
 \item[2015]
@@
 \url{http://doi.org/10.1093/icesjms/fsu055}
-(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
+(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF},
+\href{https://github.com/ss3sim/natural-mortality}{GitHub repository}).
 \item[2015]
@@
 Evolution. 4(10): 971--981. \url{http://doi.org/10.1111/2041-210X.12093}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2013_ecological_prophets_with_SOM.pdf}{PDF}).
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2013_ecological_prophets_with_SOM.pdf}{PDF},
+\href{https://github.com/seananderson/ecofolio}{GitHub R package}).
 \item[2013]
@@
 \url{http://doi.org/10.1111/j.1755-263X.2012.00294.x}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf}{PDF}).
-(*Authors contributed equally, listed in reverse alphabetical order).
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf}{PDF},
+\href{https://github.com/seananderson/conservation_pathways}{GitHub
+repository}). (*Authors contributed equally, listed in reverse
+alphabetical order).
 \item[2012]
@@
 \begin{description}
-\itemsep1pt\parskip0pt\parsep0pt
+\itemsep3pt\parskip0pt\parsep0pt
 \item[2014]
@@
 \begin{description}
-\itemsep1pt\parskip0pt\parsep0pt
+\itemsep3pt\parskip0pt\parsep0pt
 \item[2013]
@@
 \begin{description}
-\itemsep1pt\parskip0pt\parsep0pt
+\itemsep3pt\parskip0pt\parsep0pt
 \item[2014--15]
@@
 \begin{description}
-\itemsep1pt\parskip0pt\parsep0pt
+\itemsep3pt\parskip0pt\parsep0pt
 \item[2011--13]

### 2015-11-25 _knitr-posts/2015-05-25-stan-ricker-missing-values.Rmd
Quick update to CV

@@
 y[1] <- 1       # initial scaled abundance
-sigma_p <- 0.6  # process noise standard deviation
+sigma_p <- 0.3  # process noise standard deviation
 r <- 0.8        # Ricker productivity parameter
@@
 stan_ricker_missing <- stan_model(model_code = stan_model_missing)
+
+stan_model_missing1 <-
+'data {
+  int<lower=0> N; // rows of data
+  vector[N] y; // log(numbers) at time t
+  real<lower=0> K_upper; // upper boundary on K
+  int<lower=1> missing_id; // ID of missing values
+}
+parameters {
+  real<lower=0, upper=K_upper> K;
+  real r;
+  real<lower=0> sigma_proc;
+  real y_missing; // imputed missing values
+}
+transformed parameters {
+  vector[N] y_impute; // y with some imputed
+  for (i in 1:N) {
+    y_impute[i] <- y[i];
+  }
+  y_impute[missing_id] <- y_missing;
+}
+model {
+  sigma_proc ~ cauchy(0, 5);
+  r ~ cauchy(0, 5);
+  for (i in 2:N) {
+   y_impute[i] ~
+     normal(y_impute[i-1] + r * (1 - exp(y_impute[i-1]) / K),
+            sigma_proc);
+  }
+}
+'
+stan_ricker_missing1 <- stan_model(model_code = stan_model_missing1)
+
 ```
@@
 ```{r, results="hide"}
-missing_id <- sample(seq_len(N), 4)
+missing_id <- sample(seq_len(N), 1)
 y_missing <- y
@@
 m_missing <- sampling(
-  stan_ricker_missing,
+  stan_ricker_missing1,
   data = list(N = N, y = log(y_missing), 
     K_upper = max(y_missing[-missing_id]) * 2, 
-    Nmissing = length(missing_id), missing_id = missing_id),
-  iter = 4000,
+    # Nmissing = length(missing_id), 
+    missing_id = missing_id),
+  iter = 2000,
   pars = c("K", "r", "sigma_proc", "y_impute"))
@@
 col[missing_id] <- "blue"
-plot(x, y, type = "o", pch = pch, ylim = c(0, max(y_missing[-missing_id]) * 1.1), col = col)
+plot(x, y, type = "o", pch = pch, ylim = c(min(y_missing), max(y_missing[-missing_id]) * 1.1), col = col, log = "y")
 imputed <- exp(apply(p$y_impute, 2, median))

### 2015-11-25 cv-body.tex
Quick update to CV

@@
 \itemsep3pt\parskip0pt\parsep0pt
+\item[2015]
+Monnahan, C.C., K. Ono, S.C Anderson, M.B. Rudd, A.C. Hicks, F.
+Hurtado-Ferro, K.F. Johnson, P.T. Kuriyama, R.R. Licandeo, C.C. Stawitz,
+I.G. Taylor, J.L. Valero. The effect of length bin width on growth
+estimation in integrated age-structured stock assessments. In press.
+\url{http://doi.org/10.1016/j.fishres.2015.09.007}
+\item[2015]
+Kuriyamaa, P.T., F. Hurtado-Ferro, K. Ono, A.C. Hicks, I.G. Taylor, R.R.
+Licandeo, K.F. Johnson, S.C. Anderson, C.C. Monnahan, M.B. Rudd, C.C.
+Stawitz, J.L. Valero. An empirical weight-at-age approach reduces
+estimation bias compared to modeling parametric growth in integrated,
+statistical stock assessment models when growth is time varying. In
+press. \url{http://doi.org/10.1016/j.fishres.2015.11.002}
 \item[2015]
@@
 Eighth most-read CJFAS article in 2008.
-\end{description}
-
-\section{Manuscripts in review}\label{manuscripts-in-review}
-
-\begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
-\item[2015]
-Monnahan, C.C., K. Ono, S.C Anderson, M.B. Rudd, A.C. Hicks, F.
-Hurtado-Ferro, K.F. Johnson, P.T. Kuriyama, R.R. Licandeo, C.C. Stawitz,
-I.G. Taylor, J.L. Valero. The effect of length bin structures on growth
-estimation in integrated age-structured stock assessments. Submitted.
-\item[2015]
-Kuriyamaa, P.T., F. Hurtado-Ferro, K. Ono, A.C. Hicks, I.G. Taylor, R.R.
-Licandeo, K.F. Johnson, S.C. Anderson, C.C. Monnahan, M.B. Rudd, C.C.
-Stawitz, J.L. Valero. An empirical weight-at-age approach reduces
-estimation bias compared to modeling parametric growth in integrated,
-statistical stock assessment models when growth is time varying.
-Submitted.
-\item[2015]
-Trebilco, R., N.K. Dulvy, \textbf{S.C. Anderson}, A.K. Salomon. The
-paradox of inverted biomass pyramids in kelp forest fish communities. In
-revision.
 \end{description}

### 2015-11-25 _knitr-posts/impute.R

@@ -1,0 +1,58 @@
+# Let's try imputing missing values
+
+
+# now we're going to impute with Stan:
+
+stan_model <-
+'data {
+  int<lower=0> N; // rows of data
+  vector[N] x;
+  int<lower=0> Nmissing; // number of missing values
+  int missing_id[Nmissing]; // IDs of missing values
+}
+parameters {
+  real<lower=-0.99, upper=0.99> phi;
+  real<lower=0> sigma;
+  real x_missing[Nmissing]; // imputed missing values
+}
+transformed parameters {
+  vector[N] x_impute; // y with some imputed
+  for (i in 1:N) {
+    x_impute[i] <- x[i];
+  }
+  for (i in 1:Nmissing) {
+    x_impute[missing_id[i]] <- x_missing[i];
+  }
+
+}
+model {
+  sigma ~ cauchy(0, 2.5);
+  phi ~ normal(0, 2);
+  for (i in 2:N) {
+    x_impute[i] ~ normal(x_impute[i-1] * phi, sigma);
+  }
+}
+'
+library(rstan)
+stan_missing_ar1 <- stan_model(model_code = stan_model)
+
+x <- arima.sim(n = 100L, list(ar = 0.8, ma = 0), sd = 0.05)
+missing_id <- c(50, 80)
+erased <- x[missing_id]
+x[missing_id] <- NA
+plot(x)
+x_missing <- x
+x_missing[missing_id] <- 9999 # Stan can't handle NAs; give it fake data
+m_missing <- sampling(
+  stan_missing_ar1,
+  data = list(N = length(x), x = x_missing,
+    Nmissing = length(missing_id), missing_id = missing_id),
+  iter = 500,
+  pars = c("sigma", "x_impute", "sigma"))
+
+imp <- apply(extract(m_missing)$x_impute[,missing_id], 2, median)
+l <- apply(extract(m_missing)$x_impute[,missing_id], 2, quantile, probs = 0.1)
+u <- apply(extract(m_missing)$x_impute[,missing_id], 2, quantile, probs = 0.9)
+points(missing_id, imp)
+segments(missing_id, l, missing_id, u)
+points(missing_id, erased, col = "red")

### 2016-05-20 cv-body.tex
Update CV

@@
 \itemsep3pt\parskip0pt\parsep0pt
+\item[2016]
+Trebilco, R., N.K. Dulvy, \textbf{S.C. Anderson}, A.K. Salomon. The
+paradox of inverted biomass pyramids in kelp forest fish communities.
+Proceedings of the Royal Society B: Biological Sciences. In press.
+\item[2016]
+Artelle, K.A., \textbf{S.C. Anderson}, J.D. Reynolds, A.B. Cooper, P.C.
+Paquet, C.T. Darimont. Ecology of conflict: marine food supply affects
+human-wildlife interactions on land. Scientific Reports. 6: 25936.
+\url{http://doi.org/10.1038/srep25936}
+(\href{http://www.nature.com/articles/srep25936.pdf}{PDF}).
+\item[2016]
+Kuriyama, P.T., K. Ono, F. Hurtado-Ferro, A.C. Hicks, I.G. Taylor, R.R.
+Licandeo, K.F. Johnson, \textbf{S.C. Anderson}, C.C. Monnahan, M.B.
+Rudd, C.C. Stawitz, J.L. Valero. An empirical weight-at-age approach
+reduces estimation bias compared to modeling parametric growth in
+integrated, statistical stock assessment models when growth is time
+varying. Fisheries Research. 180: 119--127.
+\url{http://doi.org/10.1016/j.fishres.2015.09.007}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Kuriyama_etal_2016_empirical.pdf}{PDF}).
+\item[2016]
+Monnahan, C.C., K. Ono, \textbf{S.C. Anderson}, M.B. Rudd, A.C. Hicks,
+F. Hurtado-Ferro, K.F. Johnson, P.T. Kuriyama, R.R. Licandeo, C.C.
+Stawitz, I.G. Taylor, J.L. Valero. The effect of length bin width on
+growth estimation in integrated age-structured stock assessments.
+Fisheries Research. 180: 103--112.
+\url{http://doi.org/10.1016/j.fishres.2015.11.002}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Monnahan_etal_2016_binwidth.pdf}{PDF}).
 \item[2015]
-Monnahan, C.C., K. Ono, S.C Anderson, M.B. Rudd, A.C. Hicks, F.
-Hurtado-Ferro, K.F. Johnson, P.T. Kuriyama, R.R. Licandeo, C.C. Stawitz,
-I.G. Taylor, J.L. Valero. The effect of length bin width on growth
-estimation in integrated age-structured stock assessments. In press.
-\url{http://doi.org/10.1016/j.fishres.2015.09.007}
+Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down and bottom-up
+forces interact at thermal range extremes on American lobster. Journal
+of Animal Ecology. 84(3): 840--850.
+\url{http://doi.org/10.1111/1365-2656.12322}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Boudreau_etal_2015_lobster.pdf}{PDF}).
 \item[2015]
-Kuriyamaa, P.T., F. Hurtado-Ferro, K. Ono, A.C. Hicks, I.G. Taylor, R.R.
-Licandeo, K.F. Johnson, S.C. Anderson, C.C. Monnahan, M.B. Rudd, C.C.
-Stawitz, J.L. Valero. An empirical weight-at-age approach reduces
-estimation bias compared to modeling parametric growth in integrated,
-statistical stock assessment models when growth is time varying. In
-press. \url{http://doi.org/10.1016/j.fishres.2015.11.002}
+\textbf{Anderson, S.C.}, J.W. Moore, M.M. McClure, N.K. Dulvy, A.B.
+Cooper. Portfolio conservation of metapopulations under climate change.
+Ecological Applications. 25(2): 559--572.
+\url{http://doi.org/10.1890/14-0266.1}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2015_salmonportfolios.pdf}{PDF}).
 \item[2015]
-Finnegan*, S., \textbf{S.C. Anderson}*, P.G. Harnik*, C. Simpson, D.P.
+Orzechowski, E.A., R. Lockwood, J.E.K. Byrnes, \textbf{S.C. Anderson},
+S. Finnegan, Z.V. Finkel, P.G. Harnik, D.R. Lindberg, L.H. Liow, H.K.
+Lotze, C.R. McClain, J.L. McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson,
+D.P. Tittensor. Marine extinction risk shaped by trait-environment
+interactions over 500 million years. Global Change Biology. 21(10):
+3595--3607. \url{http://doi.org/10.1111/gcb.12963}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Orzechowski_etal_2015_paleometa.pdf}{PDF}).
+\item[2015]
+Finnegan, S., \textbf{S.C. Anderson}, P.G. Harnik, C. Simpson, D.P.
 Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
-Lockwood, H.K. Lotze, C.M. McClain, J.L. McGuire, A. O'Dea, J.M.
+Lockwood, H.K. Lotze, C.R. McClain, J.L. McGuire, A. O'Dea, J.M.
 Pandolfi. Paleontological baselines for evaluating extinction risk in
@@
 \url{http://doi.org/10.1126/science.aaa6635}
-(\href{https://github.com/seananderson/paleobaselines}{GitHub
-repository}) (\emph{*Authors contributed equally}).
-\item[2015]
-Orzechowski, E.A., R. Lockwood, J.E. Byrnes, \textbf{S.C. Anderson}, S.
-Finnegan, Z.V. Finkel, P.G. Harnik, D.R. Lindberg, L.H. Liow, H.K.
-Lotze, C.R. McClain, J.L. McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson,
-D.P. Tittensor. Marine extinction risk shaped by trait-environment
-interactions over 500 million years. In press at Global Change Biology.
-\item[2015]
-Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down and bottom-up
-forces interact at thermal range extremes on American lobster. Journal
-of Animal Ecology. 83(3):840--850.
-\url{http://doi.org/10.1111/1365-2656.12322}.
-(\href{https://github.com/seananderson/lobsters-predators}{GitHub
-repository})
 \item[2015]
@@
 integrated, age-structured stock assessment models. ICES Journal of
-Marine Science. 72(1): 99--110.
-\url{http://doi.org/10.1093/icesjms/fsu198}.
-\item[2015]
-\textbf{Anderson, S.C.}, J.W. Moore, M.M. McClure, N.K. Dulvy, A.B.
-Cooper. Portfolio conservation of metapopulations under climate change.
-Ecological Applications. 2015. 25(2): 559--572.
-\url{http://doi.org/10.1890/14-0266.1}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2014_salmon_portfolios.pdf}{PDF},
-\href{https://github.com/seananderson/metafolio}{GitHub repository},
-\href{http://cran.r-project.org/web/packages/metafolio/index.html}{CRAN
-R package}).
-\item[2015]
-Johnson, K.F., C.C. Monnahan, C.R. McGilliard, K.A. Vert-pre,
-\textbf{S.C. Anderson}, C.J. Cunningham, F. Hurtado-Ferro, R.R.
-Licandeo, M.L. Muradian, K. Ono, C.S. Szuwalski, J.L. Valero, A.R.
-Whitten, A.E. Punt. Time-varying natural mortality in fisheries stock
-assessment models: identifying a default approach. ICES Journal of
-Marine Science. 72(1): 137--150.
-\url{http://doi.org/10.1093/icesjms/fsu055}
-(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF},
-\href{https://github.com/ss3sim/natural-mortality}{GitHub repository}).
-\item[2015]
-Ono, K., R. Licandeo, M.L. Muradian, C.J. Cunningham, \textbf{S.C.
-Anderson}, F. Hurtado-Ferro, K.F. Johnson, C.R. McGilliard, C.C.
-Monnahan, C.S. Szuwalski, J.L. Valero, K.A. Vert-pre, A.R. Whitten, A.E.
-Punt. The importance of length and age composition data in statistical
-catch-at-age models for marine species. ICES Journal of Marine Science.
-72(1): 31--43. \url{http://doi.org/10.1093/icesjms/fsu007}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Ono_etal_2014_importance_of_length_and_age_composition_data.pdf}{PDF}).
+Marine Sciences. 72(1): 99--110.
+\url{http://doi.org/10.1093/icesjms/fsu198}
+(\href{https://dl.dropboxusercontent.com/u/254940/papers/Hurtado-Ferro_etal_2014_retrospective.pdf}{PDF}).
 \item[2014]
@@
 \href{https://dl.dropboxusercontent.com/u/254940/ss3sim-vignette.pdf}{vignette}).
+\item[2014]
+Johnson, K.F., C.C. Monnahan, C.R. McGilliard, K.A. Vert-pre,
+\textbf{S.C. Anderson}, C.J. Cunningham, F. Hurtado-Ferro, R.R.
+Licandeo, M.L. Muradian, K. Ono, C.S. Szuwalski, J.L. Valero, A.R.
+Whitten, A.E. Punt. Time-varying natural mortality in fisheries stock
+assessment models: identifying a default approach. ICES Journal of
+Marine Science. 72(1): 137--150.
+\url{http://doi.org/10.1093/icesjms/fsu055}
+(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
+\item[2015]
+Ono, K., R. Licandeo, M.L. Muradian, C.J. Cunningham, \textbf{S.C.
+Anderson}, F. Hurtado-Ferro, K.F. Johnson, C.R. McGilliard, C.C.
+Monnahan, C.S. Szuwalski, J.L. Valero, K.A. Vert-Pre, A.R. Whitten, A.E.
+Punt. The importance of length and age composition data in statistical
+age-structured models for marine species. ICES Journal of Marine
+Science. 72(1): 31--43. \url{http://doi.org/10.1093/icesjms/fsu007}
+(\href{http://icesjms.oxfordjournals.org/content/early/2014/02/20/icesjms.fsu007.full.pdf}{PDF}).
 \item[2014]
@@
 \itemsep3pt\parskip0pt\parsep0pt
+\item[2015]
+\textbf{Anderson, S.C.}, A.B. Cooper, O.P. Jensen, C. Minto, J.T.
+Thorson, J.C. Walsh, M. Dickey-Collas, K.M. Kleisner, C. Longo, G.C.
+Osio, D. Ovando, I. Mosqueira, A.A. Rosenberg, E.R. Selig. Improving
+estimates of population status and trend with superensemble models.
+American Fisheries Society Annual Meeting, Portland, OR, United States.
 \item[2013]
@@
 
-Reviewer for Science, Ecology, Conservation Biology, Fish and Fisheries,
-International Journal of Tropical Biology and Conservation, Journal of
-Environmental Management, Endangered Species Research, Aquatic
-Conservation
+Reviewer for Science, Ecology, Ecological Applications, Conservation
+Biology, Fish and Fisheries, International Journal of Tropical Biology
+and Conservation, Journal of Environmental Management, Endangered
+Species Research, Aquatic Conservation

### 2016-06-13 cv-body.tex
Fix order of references
Updates to CV

@@
+\section{Professional experience}\label{professional-experience}
+
+\begin{description}
+\itemsep3pt\parskip0pt\parsep0pt
+\item[2015--16]
+David H. Smith Conservation Research Fellow, School of Aquatic and
+Fishery Sciences, University of Washington and Northwest Fisheries
+Science Center, Seattle, USA
+\item[2015]
+Post-doctoral Fellow, School of Resource and Environmental Management,
+Simon Fraser University, Burnaby, Canada
+\item[2011--15]
+Graduate Research Assistant, Department of Biological Sciences, Simon
+Fraser University, Burnaby, Canada
+\item[2008--11]
+Graduate Research Assistant, Department of Biology, Dalhousie
+University, Halifax, Canada
+\end{description}
+
 \section{Education}\label{education}
@@
 extreme events in population ecology\\ Fulbright Scholar to the
-University of Washington in 2012--13\\ Dean of Graduate Studies
-Convocation Medal
+University of Washington in 2012--13\\
+\href{https://www.sfu.ca/dean-gradstudies/blog/year/2015/06/SeanAnderson.html}{Dean
+of Graduate Studies Convocation Medal}
 \item[2008--10]
@@
 and ecosystem effects of expanding global invertebrate fisheries\\
-Canadian \href{http://goo.gl/nA1zE}{Governor General's Academic Gold
+\href{http://goo.gl/nA1zE}{Canadian Governor General's Academic Gold
 Medal}
@@
 \item[2015]
-Finnegan, S., \textbf{S.C. Anderson}, P.G. Harnik, C. Simpson, D.P.
+Finnegan, S.*, \textbf{S.C. Anderson}*, P.G. Harnik*, C. Simpson, D.P.
 Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
@@
 the modern oceans. Science. 348(6234): 567--570.
-\url{http://doi.org/10.1126/science.aaa6635}
+\url{http://doi.org/10.1126/science.aaa6635} (\emph{*Authors contributed
+equally}).
 \item[2015]
@@
 (\href{https://dl.dropboxusercontent.com/u/254940/papers/Hurtado-Ferro_etal_2014_retrospective.pdf}{PDF}).
+\item[2015]
+Johnson, K.F., C.C. Monnahan, C.R. McGilliard, K.A. Vert-pre,
+\textbf{S.C. Anderson}, C.J. Cunningham, F. Hurtado-Ferro, R.R.
+Licandeo, M.L. Muradian, K. Ono, C.S. Szuwalski, J.L. Valero, A.R.
+Whitten, A.E. Punt. Time-varying natural mortality in fisheries stock
+assessment models: identifying a default approach. ICES Journal of
+Marine Science. 72(1): 137--150.
+\url{http://doi.org/10.1093/icesjms/fsu055}
+(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
+\item[2015]
+Ono, K., R. Licandeo, M.L. Muradian, C.J. Cunningham, \textbf{S.C.
+Anderson}, F. Hurtado-Ferro, K.F. Johnson, C.R. McGilliard, C.C.
+Monnahan, C.S. Szuwalski, J.L. Valero, K.A. Vert-Pre, A.R. Whitten, A.E.
+Punt. The importance of length and age composition data in statistical
+age-structured models for marine species. ICES Journal of Marine
+Science. 72(1): 31--43. \url{http://doi.org/10.1093/icesjms/fsu007}
+(\href{http://icesjms.oxfordjournals.org/content/early/2014/02/20/icesjms.fsu007.full.pdf}{PDF}).
 \item[2014]
@@
 \item[2014]
-Johnson, K.F., C.C. Monnahan, C.R. McGilliard, K.A. Vert-pre,
-\textbf{S.C. Anderson}, C.J. Cunningham, F. Hurtado-Ferro, R.R.
-Licandeo, M.L. Muradian, K. Ono, C.S. Szuwalski, J.L. Valero, A.R.
-Whitten, A.E. Punt. Time-varying natural mortality in fisheries stock
-assessment models: identifying a default approach. ICES Journal of
-Marine Science. 72(1): 137--150.
-\url{http://doi.org/10.1093/icesjms/fsu055}
-(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
-\item[2015]
-Ono, K., R. Licandeo, M.L. Muradian, C.J. Cunningham, \textbf{S.C.
-Anderson}, F. Hurtado-Ferro, K.F. Johnson, C.R. McGilliard, C.C.
-Monnahan, C.S. Szuwalski, J.L. Valero, K.A. Vert-Pre, A.R. Whitten, A.E.
-Punt. The importance of length and age composition data in statistical
-age-structured models for marine species. ICES Journal of Marine
-Science. 72(1): 31--43. \url{http://doi.org/10.1093/icesjms/fsu007}
-(\href{http://icesjms.oxfordjournals.org/content/early/2014/02/20/icesjms.fsu007.full.pdf}{PDF}).
-\item[2014]
 O'Regan, S.M., W.J. Palen, \textbf{S.C. Anderson}. Climate warming
@@
 (\href{http://www.esajournals.org/doi/pdf/10.1890/13-0916.1}{PDF}).
-\href{http://www.esajournals.org/userimages/ContentEditor/1398099670397/i0012-9658-95-4-cov.pdf}{Cover
-Article}.
 \item[2013]
@@
 Phillis*, C.C., S.M. O'Regan*, S.J. Green*, J.E. Bruce*, \textbf{S.C.
-Anderson}, J.N. Linton, E.R. Derby, B. Favaro. Multiple pathways to
-conservation success. Conservation Letters. 6(2): 98--106.
+Anderson}, J.N. Linton, Earth2Ocean Research Derby, B. Favaro. Multiple
+pathways to conservation success. Conservation Letters. 6(2): 98--106.
 \url{http://doi.org/10.1111/j.1755-263X.2012.00294.x}
@@
 David H. Smith Conservation Postdoctoral Research Fellowship
+(\href{http://conbio.org/mini-sites/smith-fellows/meet-the-fellows/2015-fellows/sean-anderson}{URL})
+\item[2015]
+Simon Fraser University Dean of Graduate Studies Convocation Medal
+(\href{https://www.sfu.ca/dean-gradstudies/blog/year/2015/06/SeanAnderson.html}{URL})
 \item[2014]
@@
 \item[2012--13]
-Canadian Fulbright Scholar award to the University of Washington in
-2012--13
+Canadian Fulbright Scholar award to the University of Washington
 \item[2011]
@@
 \itemsep3pt\parskip0pt\parsep0pt
+\item[2016]
+\textbf{Anderson, S.C.}, A.B. Cooper, O.P. Jensen, C. Minto, J.T.
+Thorson, J.C. Walsh, M. Dickey-Collas, K.M. Kleisner, C. Longo, G.C.
+Osio, D. Ovando, I. Mosqueira, A.A. Rosenberg, E.R. Selig. Improving
+estimates of population status and trend with superensemble models.
+World Fisheries Conference, Busan, South Korea. (Presented \emph{in
+absentia} by J.T. Thorson.)
 \item[2015]
@@
 \itemsep3pt\parskip0pt\parsep0pt
-\item[2014--15]
+\item[2014--16]
 Developed self-directed lecture and exercises on ggplot2 for FISH 554:
@@
 \itemsep3pt\parskip0pt\parsep0pt
+\item[2014--15]
+Data analysis in R and C++ for Ecoscape Environmental Consultants Ltd.
 \item[2013]
@@
 \itemsep3pt\parskip0pt\parsep0pt
+\item[2015-16]
+NCEAS (National Center for Ecological Analysis and Synthesis, Santa
+Barbara, CA) ``Applying portfolio effects to the Gulf of Alaska
+ecosystem: Did multi-scale diversity buffer against the Exxon Valdez oil
+spill?'' (\href{https://www.nceas.ucsb.edu/featured/marshall}{URL})
+\item[2015-16]
+Gordon and Betty Moore Foundation funded working group ``Applying
+data-limited stock status models and developing management guidance for
+unassessed fish stocks''.
 \item[2011--13]

### 2016-06-13 refs.R

@@ -2,6 +2,8 @@
 # TODO J.E. Mills Flemming needs to be fixed throughout refs.bib too
 # TODO Earth2Oceans Derby needs to be manually fixed
 
+# Need to fix all authors contributed equally and asterisks 
+
 library("bibtex")
 d <- read.bib("refs.bib")
 

### 2016-06-13 cv-body.tex
Copyediting CV

@@
 NCEAS (National Center for Ecological Analysis and Synthesis, Santa
-Barbara, CA) ``Applying portfolio effects to the Gulf of Alaska
+Barbara, CA) \emph{Applying portfolio effects to the Gulf of Alaska
 ecosystem: Did multi-scale diversity buffer against the Exxon Valdez oil
-spill?'' (\href{https://www.nceas.ucsb.edu/featured/marshall}{URL})
+spill?} (\href{https://www.nceas.ucsb.edu/featured/marshall}{URL})
 \item[2015-16]
-Gordon and Betty Moore Foundation funded working group ``Applying
+Gordon and Betty Moore Foundation funded working group \emph{Applying
 data-limited stock status models and developing management guidance for
-unassessed fish stocks''.
+unassessed fish stocks}.
 \item[2011--13]

### 2016-06-13 cv-body.tex
Add two journals

@@
 Reviewer for Science, Ecology, Ecological Applications, Conservation
-Biology, Fish and Fisheries, International Journal of Tropical Biology
-and Conservation, Journal of Environmental Management, Endangered
-Species Research, Aquatic Conservation
+Biology, Fish and Fisheries, Marine Policy, Population Ecology,
+International Journal of Tropical Biology and Conservation, Journal of
+Environmental Management, Endangered Species Research, Aquatic
+Conservation

### 2016-09-06 _knitr-posts/2015-02-02-stan-matt-trick.Rmd
Play with Stan Matt trick draft

@@
 
-```{r}
-set.seed(123)
-N <- 50 # number of observations per group
-Nj <- 16 # number of groups
-sigma_j <- 2.0 # standard deviation of intercepts across groups
-d <- data.frame(j = 1:Nj, jname = letters[1:Nj], b0_j = rnorm(Nj, 0, sigma_j))
-sigma <- 1.1 # residual standard deviation
-b0 <- 0.3 # intercept
-b1 <- 0.5 # slope
-x <- lapply(seq_len(Nj), function(x) rnorm(n = N)) # predictor
+```{r dat}
+library("dplyr")
+library("ggplot2")
+library("rstan")
+library("purrr")
+# rstan_options(auto_write = TRUE)
+# options(mc.cores = parallel::detectCores())
+# options(mc.cores = 1)
+
+sim_data <- function(N = 10, Nj = 10, sigma_j = 0.6, sigma = 1.1, 
+  seed = 1, b0 = -0.3, b1 = 0.9, scale = FALSE) {
+  set.seed(seed)
+  # N <- 4 # number of observations per group
+  # Nj <- 10 # number of groups
+  # sigma_j <- 0.6 # standard deviation of intercepts across groups
+  d <- tibble(j = 1:Nj, jname = seq_len(Nj), b0_j = rnorm(Nj, 0, sigma_j))
+  # sigma <- 1.1 # residual standard deviation
+  # b0 <- -0.3 # intercept
+  # b1 <- 0.9 # slope
+  x <- map(seq_len(Nj), function(x) runif(N, 0, 2)) # predictor
   dat <- list()
   for(j in d$j) {
-  dat[[j]] <- data.frame(j = j, jname = d$jname[j], x = x[[j]],
-    y = rnorm(N, mean = b0 + x[[j]] * b1 + d$b0_j[j], sd = sigma))
+    dat[[j]] <- data.frame(j = j, jname = d$jname[[j]], x = x[[j]],
+      y = rnorm(N, mean = b0 + x[[j]] * b1 + d$b0_j[j], sd = sigma), 
+      stringsAsFactors = FALSE)
   }
-dat <- as.data.frame(do.call("rbind", dat))
-library("ggplot2")
-ggplot(dat, aes(x, y)) + geom_point() + facet_wrap(~jname)
+  dat <- bind_rows(dat)
+  if (scale) dat <- mutate(dat, x = arm::rescale(x))
+  dat
+}
+# ggplot(dat, aes(x, y)) + geom_point() + facet_wrap(~jname)
 ```
@@
 
-```{r}
+```{r model}
 model <- "
@@
   for (i in 1:N) {
-    yhat[i] <- b0_j[j_id[i]] + b1 * x[i];
+    yhat[i] = b0_j[j_id[i]] + b1 * x[i];
   }
@@
 "
-sm <- rstan::stan_model(model_code = model)
 ```
@@
 
-```{r}
-model <- "
+```{r model_matt}
+model_matt <- "
 data {
@@
   vector[Nj] b0_j;                // Matt Trick #2
-  b0_j <- b0 + sigma_j * e_b0_j;  // Matt Trick #3
+  b0_j = b0 + sigma_j * e_b0_j;  // Matt Trick #3
 
   for (i in 1:N) {
-    yhat[i] <- b0_j[j_id[i]] + b1 * x[i];
+    yhat[i] = b0_j[j_id[i]] + b1 * x[i];
   }
@@
 "
-sm_matt <- rstan::stan_model(model_code = model)
 ```
@@
 
-```{r}
+```{r fits, dependson=c("model", "model_matt", "dat"), cache=TRUE}
+dat <- sim_data(N = 20, Nj = 8, seed = 999)
 model_data <- list(N = nrow(dat), y = dat$y, x = dat$x, Nj = max(dat$j), 
   j_id = dat$j)
-pars <- c("b0", "b1", "sigma", "sigma_j")
+pars <- c("b0", "b1", "sigma", "sigma_j", "b0_j")
 
 st <- system.time({
-  m <- sampling(sm, data = model_data, iter = 400, chains = 3, 
-    pars = pars, seed = 999)
+  m <- stan(model_code = model, data = model_data, iter = 600, 
+    chains = 4, pars = pars, seed = 999)
 })
 st_matt <- system.time({
-  m_matt <- sampling(sm_matt, data = model_data, iter = 400, chains = 3, 
-    pars = pars, seed = 999)
+  m_matt <- stan(model_code = model_matt, data = model_data, iter = 600, 
+    chains = 4, pars = pars, seed = 999)
 })
 ```
+
+```{r, dependson=fits}
+# print(st)
+# print(st_matt)
+broom::tidyMCMC(m, rhat = TRUE, ess = TRUE) %>%
+  mutate(ess_per_sec = round(ess / st[[1]], 0), time = st[[1]])
+broom::tidyMCMC(m_matt, rhat = TRUE, ess = TRUE) %>%
+  mutate(ess_per_sec = round(ess / st_matt[[1]], 0), time = st_matt[[1]])
+```
+
+```{r}
+pars <- c("b0", "b1", "sigma", "sigma_j", "b0_j")
+
+out <- plyr::ldply(c(2, 3, 5, 7, 10, 15, 20, 50), function(i) {
+  dat <- sim_data(N = i, Nj = 10, seed = 2)
+  model_data <- list(N = nrow(dat), y = dat$y, x = dat$x, Nj = max(dat$j), 
+    j_id = dat$j)
+  st <- system.time({
+    m <- stan(model_code = model, data = model_data, iter = 300, 
+      chains = 2, pars = pars, seed = 999)
+  })
+  st_matt <- system.time({
+    m_matt <- stan(model_code = model_matt, data = model_data, iter = 300, 
+      chains = 2, pars = pars, seed = 999)
+  })
+  b1 <- broom::tidyMCMC(m, rhat = TRUE, ess = TRUE) %>%
+  mutate(ess_per_sec = round(ess / st[[1]], 0), time = st[[1]])
+  b2 <- broom::tidyMCMC(m_matt, rhat = TRUE, ess = TRUE) %>%
+  mutate(ess_per_sec = round(ess / st_matt[[1]], 0), time = st_matt[[1]])
+  data.frame(m = min(b1$ess_per_sec), m_matt = min(b2$ess_per_sec), N = i,
+    time = b1$time[1], time_matt = b2$time[1])
+})
+select(out, N, m, m_matt) %>% 
+  reshape2::melt(id.vars = "N") %>%
+  ggplot(aes(N, value, colour = variable)) + geom_line()
+```

### 2016-10-25 _knitr-posts/2014-07-30-centering-interactions.Rmd
Change subheading on centering post
Fix fig

@@
 title: Centering factor interactions
-subtitle: Why and how you must center interactions for model averaging
+subtitle: Why and how to center interactions for model averaging
 published: true
@@
 
-<img src="/images/interactions.png" alt="Illustration of centering interactions" width=580 height=436>
+![Illustration of centering interactions](/images/interactions.png)
 

### 2016-12-01 cv-body.tex
Update CV + newer Pandoc

@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
 \item[2015--16]
@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
 \item[2011--15]
-Ph.D.\ Biology, Simon Fraser University, Burnaby, Canada\\ Variance and
-extreme events in population ecology\\ Fulbright Scholar to the
-University of Washington in 2012--13\\
+Ph.D.\ Biology, Simon Fraser University, Burnaby, Canada\\
+Variance and extreme events in population ecology\\
+Fulbright Scholar to the University of Washington in 2012--13\\
 \href{https://www.sfu.ca/dean-gradstudies/blog/year/2015/06/SeanAnderson.html}{Dean
@@
 \item[2008--10]
-M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\ Trends, drivers,
-and ecosystem effects of expanding global invertebrate fisheries\\
+M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\
+Trends, drivers, and ecosystem effects of expanding global invertebrate
+fisheries\\
 \href{http://goo.gl/nA1zE}{Canadian Governor General's Academic Gold
@@
 B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
-Canada\\ First Class Honours \& Environmental Programmes Honour Society
-Medal
+Canada\\
+First Class Honours \& Environmental Programmes Honour Society Medal
 \item[2001--03]
@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
+\item[2016]
+\textbf{Anderson, S.C.}, A.B. Cooper, O.P. Jensen, C. Minto, J.T.
+Thorson, J.C. Walsh, M. Dickey-Collas, K.M. Kleisner, C. Longo, G.C.
+Osio, D. Ovando, I. Mosqueira, A.A. Rosenberg, E.R. Selig. Improving
+estimates of population status and trend with superensemble models. Fish
+and Fisheries. In press.
 \item[2016]
@@
 paradox of inverted biomass pyramids in kelp forest fish communities.
-Proceedings of the Royal Society B: Biological Sciences. In press.
+Proceedings of the Royal Society B: Biological Sciences. 283: 20160816.
+\url{http://doi.org/10.1098/rspb.2016.0816}
 \item[2016]
@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
 \item[2015]
@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
 \item[2014]
 \textbf{Anderson, S.C.}, J.W. Moore, M.M McClure, N.K. Dulvy, A.B.
-Cooper.\\ metafolio: Metapopulation simulations for conserving salmon
-through portfolio optimization.
+Cooper.\\
+metafolio: Metapopulation simulations for conserving salmon through
+portfolio optimization.
 \url{http://cran.r-project.org/package=metafolio}
@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
+\item[2016]
+\textbf{Anderson, S.C.}, T.A. Branch, A.B. Cooper, N.K. Dulvy.
+Black-swan events in animal populations. North America Congress for
+Conservation Biology, Madison, WI, United States.
 \item[2016]
@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
+\item[2016]
+Two two-day (14 hour) workshops on advanced generalized linear
+mixed-effects modeling in R for the Simon Fraser University Library
+Research Commons. \url{https://github.com/seananderson/glmm-course}
 \item[2014--16]
@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
 \item[2014--15]
@@
 \begin{description}
-\itemsep3pt\parskip0pt\parsep0pt
+\tightlist
 \item[2015-16]
@@
 
+\section{Selected courses and
+training}\label{selected-courses-and-training}
+
+\begin{description}
+\tightlist
+\item[Present]
+Proficient in statistical analyses and data management using R, Stan,
+JAGS, Sweave, knitr, Shiny, LaTeX, Git, TMB. Familiar with ADMB,
+PostgreSQL.
+\item[2016]
+Diversity in Conservation Biology, facilitated by Yolanda L. Comedy,
+Director of the American Association for the Advancement of Science
+(AAAS) Center for Advancing Science \& Engineering Capacity
+\item[2015--16]
+Powerful Presentation Skills and Media Training with Teresa Erickson and
+Tim Ward of Intermedia Communications, authors of
+\href{https://www.amazon.com/Master-Communicators-Handbook-Teresa-Erickson/dp/1785351532}{\emph{The
+Master Communicator's Handbook}}
+\item[2013]
+Quantitative Modelling in Ecosystem Based Management. R. Hilborn, School
+of Aquatic and Fishery Sciences, University of Washington (audited).
+\item[2013]
+Applied Time Series Analysis. E. Holmes, M. Scheuerell, and E. Ward;
+School of Aquatic and Fishery Sciences, University of Washington
+(audited).
+\item[2012]
+Numerical Computing for the Natural Resources. A.E. Punt, School of
+Aquatic and Fishery Sciences, University of Washington (audited).
+\item[2012]
+Risk Assessment and Decision Analysis for Management of Natural
+Resources. R.M. Peterman, School of Resource and Environmental
+Management, Simon Fraser University.
+\item[2011]
+Presenting Data and Information. Edward Tufte, Seattle, WA, United
+States.
+\item[2009]
+Technical Expertise in Stock Assessment: Model Fitting and
+Interpretation. C. Needle, chair of the International Council for the
+Exploration of the Sea working group on methods of fish stock
+assessment, Dalhousie University.
+\end{description}
+
 \section{Reviews}\label{reviews}

### 2016-12-01 refs.R

@@ -2,7 +2,7 @@
 # TODO J.E. Mills Flemming needs to be fixed throughout refs.bib too
 # TODO Earth2Oceans Derby needs to be manually fixed
 
-# Need to fix all authors contributed equally and asterisks 
+# Need to fix all authors contributed equally and asterisks
 
 library("bibtex")
 d <- read.bib("refs.bib")
@@ -91,3 +91,8 @@ for(i in seq_along(d)) {
 
 ref <- paste(ref, collapse = "\n\n")
 writeLines(ref, con = "papers.md")
+
+# authors:
+library(magrittr)
+lapply(d, function(x) x$author$family) %>% unlist() %>%
+  gsub("\\\\\\*", "", .) %>% unique() %>% sort()

### 2017-02-17 _knitr-posts/2017-02-17-negative-binomial.Rmd
Add draft negbin post

---
layout: post
title: 
subtitle: 
published: false
---

> A group of us [...] are in a "Bayesian club" and trying to use our collective
> intelligence to learn more about Bayesian modelling. [...] Today we decided to
> overview probability distributions [...] identifying the different main
> distribution types and how they are parameterized. It all went pretty good and
> we were all pretty clear on things... until we got to the negative binomial
> distribution. The bit of research/reading and googling that we had done
> suggested this is a notoriously nebulous distribution in that there are
> multiple ways it can be parameterized? [...]

There certainly are a lot of ways to parameterize the negative binomial. I've
seen claims of up to 7 parameterizations. Not only are there different ways to
parameterize it but there are at least 2 different versions of the negative
binomial distribution itself. One where the variance scales quadratically with
the mean and another where it scales linearly. By far the most common in
ecology is the version where the variance scales quadratically with the mean.
This is often referred to as the NB2 in negative binomial. I think that name
was coined in this book:
https://books.google.com/books/about/Negative_Binomial_Regression.html?id=DDxEGQuqkJoC

The other version where the variance and mean scale linearly is sometimes
called NB1. In general, I wouldn't worry too much about this besides being able
to describe the version of the distribution that you have used. I'd focus on
the NB2.

As with any distribution, you need to read the manual or help files if you are
using a probability density function written by someone else. A common
parameterization in ecology is one that is parameterized in terms of a mean and
a dispersion parameter, sometimes called phi, theta, or size.

In that case, you have NB2(mu, phi), and the mean and the variance at a given
mean value are:
mean = mu
variance = mu + (mu^2)/phi

So as phi becomes large the variance approaches the mean and the distribution
approaches the Poisson distribution.

This is how MASS::rnegbin is parameterized. Although they call the dispersion
parameter theta.

stats::dnbinom() can also be parameterized this way. Although they call the
dispersion parameter size.

This is also how the negative binomial is parameterized in glmmTMB and
glmmADMB.

And this is one of the ways it is parameterized in Stan:
neg_binomial_2(mu, phi)

Note that sometimes the distribution is parameterized as:
variance = mu + (mu^2) * phi

In which case, phi has simply been inverted so that as phi become small, the
variance approaches the mean and the distribution approaches the Poisson
distribution.

It's been a while since I worked in JAGS/BUGS, but after a quick look it seems
like JAGS is parameterized with 2 parameters we will call p and r. Then:
dnegbin(p, r)
mean = r(1-p)/p
variance = r(1-p)/p^2

Simple, right? :)  This matches the stats::dnbinom distribution if you use the
size and prob arguments... except that the order of the 2 parameters is
reversed. They wouldn't want to make it too easy after all. Details here:
https://journal.r-project.org/archive/2013-1/lebauer-dietze-bolker.pdf

Now if you want to write a GLM or GLMM yourself, you'll be working with a log
link and then exponentiating before passing the values to the density function.

I think, based on a bit of Googling, you could fit a model in JAGS with
something like this:

mu[i] <- b0 + b1 * x1[i] ...
lambda[i] <- exp(mu[i])
p[i] <- r/(r+lambda[i])
y[i] ~ dnegbin(p[i],r)

Then,
mean = lambda
variance = lambda + (lambda^2)/r

Stan has a built in parameterization on the log scale so that you can avoid the
exponentiating. This is a bit faster, cleaner, and less likely to run into
numerical issues with big or small numbers:

mu = b0 + b1 * x1...
y ~ neg_binomial_2_log(mu, phi)

If you're interested in fitting negative binomial GLMs and GLMMs in Stan,
I wrote these examples a couple years ago:
https://github.com/seananderson/negbin-stan/blob/master/negbin-glm.md

(You can also now fit these easily with the rstanarm package.)

As with any model you write yourself, it's important to check your model by
either comparing it with some other version that you trust from a different
package or software or by simulating some data and recovering the true values
with your model (as I was doing in the above link). This is especially so if
you're unsure about the the parameterization of a distribution.

### 2017-03-09 cv-body.tex
Tweak CV
Update CV
CV

@@
 \tightlist
-\item[2015--16]
+\item[2015--17]
 David H. Smith Conservation Research Fellow, School of Aquatic and
@@
 \item[2015]
-Post-doctoral Fellow, School of Resource and Environmental Management,
+Postdoctoral Fellow, School of Resource and Environmental Management,
 Simon Fraser University, Burnaby, Canada
@@
 \tightlist
-\item[2016]
+\item[2017]
+\textbf{Anderson, S.C.}, T.A. Branch, A.B. Cooper, N.K. Dulvy.
+Black-swan events in animal populations. Proceedings of the National
+Academy of Sciences. In press.
+\url{http://doi.org/10.1073/pnas.1611525114} (Open access).
+\item[2017]
 \textbf{Anderson, S.C.}, A.B. Cooper, O.P. Jensen, C. Minto, J.T.
@@
 estimates of population status and trend with superensemble models. Fish
-and Fisheries. In press.
+and Fisheries. In press. \url{http://doi.org/10.1111/faf.12200}.
 \item[2016]

### 2017-05-26 cv-body.tex
Update CV
Fix cv
Update cv
Start updating CV

@@
 \tightlist
+\item[2017--present]
+Aquatic Science Biologist, Offshore Section, Aquatic Resource Research
+and Assessment Division, Pacific Biological Station, Fisheries and
+Oceans Canada, Nanaimo, BC, Canada
 \item[2015--17]
@@
 Fishery Sciences, University of Washington and Northwest Fisheries
-Science Center, Seattle, USA
+Science Center, Seattle, WA, USA
 \item[2015]
 Postdoctoral Fellow, School of Resource and Environmental Management,
-Simon Fraser University, Burnaby, Canada
+Simon Fraser University, Burnaby, BC, Canada
 \item[2011--15]
 Graduate Research Assistant, Department of Biological Sciences, Simon
-Fraser University, Burnaby, Canada
+Fraser University, Burnaby, BC, Canada
 \item[2008--11]
 Graduate Research Assistant, Department of Biology, Dalhousie
-University, Halifax, Canada
+University, Halifax, NS, Canada
 \end{description}
@@
 \item[2017]
+Rosenberg, A.A., K.M. Kleisner, J. Afflerbach, \textbf{S.C. Anderson},
+M. Dickey-Collas, A.B. Cooper, M.J. Fogarty, E.A. Fulton, N.L.
+Gutiérrez, K.J.W. Hyde, E. Jardim, O.P. Jensen, T. Kristiansen, C.
+Longo, C.V. Minte-Vera, C. Minto, I. Mosqueira, G.C. Osio, D. Ovando,
+E.R. Selig, J.T. Thorson, J.C. Walsh, Y. Ye. Applying a new ensemble
+approach to estimating stock status of marine fisheries around the
+world. Conservation Letters. In press.
+\item[2017]
 \textbf{Anderson, S.C.}, T.A. Branch, A.B. Cooper, N.K. Dulvy.
 Black-swan events in animal populations. Proceedings of the National
-Academy of Sciences. In press.
+Academy of Sciences. 114(12): 3252---3257.
 \url{http://doi.org/10.1073/pnas.1611525114} (Open access).
@@
 Proceedings of the Royal Society B: Biological Sciences. 283: 20160816.
-\url{http://doi.org/10.1098/rspb.2016.0816}
+\url{http://doi.org/10.1098/rspb.2016.0816}.
 \item[2016]
@@
 \tightlist
+\item[2017]
+Canadian Society for Ecology and Evolution (CSEE) Early Career Award
+(\href{http://csee-scee.ca/?p=5183}{URL})
 \item[2015]
@@
 \tightlist
+\item[2017]
+rrfields: Robust random fields for spatiotemporal modelling.
+\url{https://github.com/seananderson/rrfields}
+\item[2017]
+datalimited: Stock assessment methods for data-limited fisheries.
+\url{https://github.com/datalimited/datalimited}
 \item[2014]
@@
 \tightlist
+\item[2017]
+\textbf{Anderson, S.C.} Data-driven approaches to quantifying population
+status and extremes. Early Career Award keynote lecture at the Canadian
+Society for Ecology and Evolution (CSEE), Victoria, BC, Canada.
 \item[2016]
@@
 \tightlist
+\item[2017]
+Half-day workshop on generalized linear modeling in R for the David H.
+Smith Conservation Postdoctoral Research Fellows.
+\url{https://github.com/seananderson/glmm-course}
+\item[2017]
+One-day workshop on linear mixed-effects modeling in R for the David H.
+Smith Conservation Postdoctoral Research Fellows.
+\url{https://github.com/seananderson/glmm-course}
 \item[2016]
@@
 Proficient in statistical analyses and data management using R, Stan,
-JAGS, Sweave, knitr, Shiny, LaTeX, Git, TMB. Familiar with ADMB,
-PostgreSQL.
+JAGS, Sweave, knitr, Shiny, LaTeX, Git, TMB, SQL. Familiar with ADMB.
 \item[2016]
@@
 States.
-\item[2009]
-Technical Expertise in Stock Assessment: Model Fitting and
-Interpretation. C. Needle, chair of the International Council for the
-Exploration of the Sea working group on methods of fish stock
-assessment, Dalhousie University.
 \end{description}
@@
 Reviewer for Science, Ecology, Ecological Applications, Conservation
-Biology, Fish and Fisheries, Marine Policy, Population Ecology,
-International Journal of Tropical Biology and Conservation, Journal of
-Environmental Management, Endangered Species Research, Aquatic
-Conservation
+Biology, Fish and Fisheries, Marine Policy, Population Ecology, Journal
+of Applied Ecology, International Journal of Tropical Biology and
+Conservation, Journal of Environmental Management, Endangered Species
+Research, Aquatic Conservation

### 2017-05-31 cv-body.tex
Update CV and pages

@@
 \tightlist
-\item[2017--present]
+\item[2017]
 Aquatic Science Biologist, Offshore Section, Aquatic Resource Research

### 2017-07-18 cv-body.tex
Fix en dash
Update CV

@@
 \item[2017]
-Aquatic Science Biologist, Offshore Section, Aquatic Resource Research
-and Assessment Division, Pacific Biological Station, Fisheries and
-Oceans Canada, Nanaimo, BC, Canada
+Aquatic Science Biologist and Program Head in Spatial Stock Assessment,
+Offshore Section, Aquatic Resource Research and Assessment Division,
+Pacific Biological Station, Fisheries and Oceans Canada, Nanaimo, BC,
+Canada
 \item[2015--17]
@@
 \item[2017]
+Hughes, B., S. Lummis, \textbf{S.C. Anderson}, K. Kroeker. Unexpected
+resilience of a seagrass system exposed to multiple stressors. Global
+Change Biology. In press.
+\item[2017]
+Megias, D.A., \textbf{S.C. Anderson}, R.J. Smith, D. Veríssimo.
+Investigating the impact of media on demand for wildlife: a case study
+of Harry Potter and the UK trade in owls. PLOS ONE. In press.
+\item[2017]
+Chezik, K.A., \textbf{S.C. Anderson}, J.W. Moore. River networks dampen
+long-term hydrological signals of climate change. Geophysical Research
+Letters. In press. \url{http://doi.org/10.1002/2017GL074376}
+\item[2017]
 Rosenberg, A.A., K.M. Kleisner, J. Afflerbach, \textbf{S.C. Anderson},
@@
 world. Conservation Letters. In press.
+\url{http://doi.org/10.1111/conl.12363}
 \item[2017]
@@
 Black-swan events in animal populations. Proceedings of the National
-Academy of Sciences. 114(12): 3252---3257.
+Academy of Sciences. 114(12): 3252--3257.
 \url{http://doi.org/10.1073/pnas.1611525114} (Open access).
@@
 \textbf{Anderson, S.C.}, A.B. Cooper, O.P. Jensen, C. Minto, J.T.
-Thorson, J.C. Walsh, M. Dickey-Collas, K.M. Kleisner, C. Longo, G.C.
-Osio, D. Ovando, I. Mosqueira, A.A. Rosenberg, E.R. Selig. Improving
-estimates of population status and trend with superensemble models. Fish
-and Fisheries. In press. \url{http://doi.org/10.1111/faf.12200}.
+Thorson, J.C. Walsh, J. Afflerbach, M. Dickey-Collas, K.M. Kleisner, C.
+Longo, G.C. Osio, D. Ovando, I. Mosqueira, A.A. Rosenberg, E.R. Selig.
+Improving estimates of population status and trend with superensemble
+models. Fish and Fisheries. 18(4): 732--741.
+\url{http://doi.org/10.1111/faf.12200}
 \item[2016]

### 2017-10-06 cv-body.tex
Add DOI and in press commercial portfolio paper
Update CV and add workshops
Update CV

@@
 \begin{description}
-\tightlist
+\item[2017]
+\textbf{Anderson, S.C.}, E.J. Ward, A.O. Shelton, M.D.\ Adkison, A.H.
+Beaudreau, R.E. Brenner, A.C. Haynie, J.C. Shriver, J.T. Watson, B.C.
+Williams. Benefits and risks of diversification for individual fishers.
+In press at Proceedings of the National Academy of Sciences.
+\url{https://doi.org/10.1073/pnas.1702506114} (Open access).
 \item[2017]
@@
 resilience of a seagrass system exposed to multiple stressors. Global
-Change Biology. In press.
+Change Biology. In press. \url{https://doi.org/10.1111/gcb.13854}
 \item[2017]
@@
 Investigating the impact of media on demand for wildlife: a case study
-of Harry Potter and the UK trade in owls. PLOS ONE. In press.
+of Harry Potter and the UK trade in owls. PLoS ONE. 12(10): e0182368.
+\url{https://doi.org/10.1371/journal.pone.0182368} (Open access).
 \item[2017]
@@
 long-term hydrological signals of climate change. Geophysical Research
-Letters. In press. \url{http://doi.org/10.1002/2017GL074376}
+Letters. 44(14):7256--7264. \url{https://doi.org/10.1002/2017GL074376}
 \item[2017]
@@
 world. Conservation Letters. In press.
-\url{http://doi.org/10.1111/conl.12363}
+\url{https://doi.org/10.1111/conl.12363}
 \item[2017]
@@
 Academy of Sciences. 114(12): 3252--3257.
-\url{http://doi.org/10.1073/pnas.1611525114} (Open access).
+\url{https://doi.org/10.1073/pnas.1611525114} (Open access).
 \item[2017]
@@
 models. Fish and Fisheries. 18(4): 732--741.
-\url{http://doi.org/10.1111/faf.12200}
+\url{https://doi.org/10.1111/faf.12200}
 \item[2016]
@@
 Proceedings of the Royal Society B: Biological Sciences. 283: 20160816.
-\url{http://doi.org/10.1098/rspb.2016.0816}.
+\url{https://doi.org/10.1098/rspb.2016.0816}.
 \item[2016]
@@
 human-wildlife interactions on land. Scientific Reports. 6: 25936.
-\url{http://doi.org/10.1038/srep25936}
+\url{https://doi.org/10.1038/srep25936}
 (\href{http://www.nature.com/articles/srep25936.pdf}{PDF}).
@@
 varying. Fisheries Research. 180: 119--127.
-\url{http://doi.org/10.1016/j.fishres.2015.09.007}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Kuriyama_etal_2016_empirical.pdf}{PDF}).
+\url{https://doi.org/10.1016/j.fishres.2015.09.007}
+(\href{https://sean.updog.co/papers/Kuriyama_etal_2016_empirical.pdf}{PDF}).
+
+;drop
 \item[2016]
@@
 Fisheries Research. 180: 103--112.
-\url{http://doi.org/10.1016/j.fishres.2015.11.002}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Monnahan_etal_2016_binwidth.pdf}{PDF}).
+\url{https://doi.org/10.1016/j.fishres.2015.11.002}
+(\href{https://sean.updog.co/papers/Monnahan_etal_2016_binwidth.pdf}{PDF}).
 \item[2015]
@@
 of Animal Ecology. 84(3): 840--850.
-\url{http://doi.org/10.1111/1365-2656.12322}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Boudreau_etal_2015_lobster.pdf}{PDF}).
+\url{https://doi.org/10.1111/1365-2656.12322}
+(\href{https://sean.updog.co/papers/Boudreau_etal_2015_lobster.pdf}{PDF}).
 \item[2015]
@@
 Ecological Applications. 25(2): 559--572.
-\url{http://doi.org/10.1890/14-0266.1}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2015_salmonportfolios.pdf}{PDF}).
+\url{https://doi.org/10.1890/14-0266.1}
+(\href{https://sean.updog.co/papers/Anderson_etal_2015_salmonportfolios.pdf}{PDF}).
 \item[2015]
@@
 interactions over 500 million years. Global Change Biology. 21(10):
-3595--3607. \url{http://doi.org/10.1111/gcb.12963}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Orzechowski_etal_2015_paleometa.pdf}{PDF}).
+3595--3607. \url{https://doi.org/10.1111/gcb.12963}
+(\href{https://sean.updog.co/papers/Orzechowski_etal_2015_paleometa.pdf}{PDF}).
 \item[2015]
@@
 the modern oceans. Science. 348(6234): 567--570.
-\url{http://doi.org/10.1126/science.aaa6635} (\emph{*Authors contributed
-equally}).
+\url{https://doi.org/10.1126/science.aaa6635} (\emph{*Authors
+contributed equally}).
 \item[2015]
@@
 Marine Sciences. 72(1): 99--110.
-\url{http://doi.org/10.1093/icesjms/fsu198}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Hurtado-Ferro_etal_2014_retrospective.pdf}{PDF}).
+\url{https://doi.org/10.1093/icesjms/fsu198}
+(\href{https://sean.updog.co/papers/Hurtado-Ferro_etal_2014_retrospective.pdf}{PDF}).
 \item[2015]
@@
 Marine Science. 72(1): 137--150.
-\url{http://doi.org/10.1093/icesjms/fsu055}
+\url{https://doi.org/10.1093/icesjms/fsu055}
 (\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
@@
 age-structured models for marine species. ICES Journal of Marine
-Science. 72(1): 31--43. \url{http://doi.org/10.1093/icesjms/fsu007}
+Science. 72(1): 31--43. \url{https://doi.org/10.1093/icesjms/fsu007}
 (\href{http://icesjms.oxfordjournals.org/content/early/2014/02/20/icesjms.fsu007.full.pdf}{PDF}).
@@
 Observer aging and long-term avian survey data quality. Ecology and
-Evolution. 4(12): 2563--2576. \url{http://doi.org/10.1002/ece3.1101}
+Evolution. 4(12): 2563--2576. \url{https://doi.org/10.1002/ece3.1101}
 (\href{http://onlinelibrary.wiley.com/doi/10.1002/ece3.1101/pdf}{PDF}).
@@
 Valero. ss3sim: An R package for fisheries stock assessment simulation
-with Stock Synthesis. PLOS ONE. 9(4): e92725.
-\url{http://doi.org/10.1371/journal.pone.0092725}
+with Stock Synthesis. PLoS ONE. 9(4): e92725.
+\url{https://doi.org/10.1371/journal.pone.0092725}
 (\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0092725\&representation=PDF}{PDF}).
@@
 package}, \href{https://github.com/ss3sim/ss3sim}{code},
-\href{https://dl.dropboxusercontent.com/u/254940/ss3sim-vignette.pdf}{vignette}).
+\href{https://sean.updog.co/ss3sim-vignette.pdf}{vignette}).
 \item[2014]
@@
 species. Ecology. 95(4): 845--855.
-\url{http://doi.org/10.1890/13-0916.1}
+\url{https://doi.org/10.1890/13-0916.1}
 (\href{http://www.esajournals.org/doi/pdf/10.1890/13-0916.1}{PDF}).
@@
 Ecology and Evolution. Ideas in Ecology and Evolution. 6: 40--46.
-\url{http://doi.org/10.4033/iee.2013.6.9.n}
+\url{https://doi.org/10.4033/iee.2013.6.9.n}
 (\href{http://library.queensu.ca/ojs/index.php/IEE/article/download/4931/4899}{PDF}).
@@
 Quantifying metapopulation portfolio effects. Methods in Ecology and
-Evolution. 4(10): 971--981. \url{http://doi.org/10.1111/2041-210X.12093}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2013_ecological_prophets_with_SOM.pdf}{PDF},
+Evolution. 4(10): 971--981.
+\url{https://doi.org/10.1111/2041-210X.12093}
+(\href{https://sean.updog.co/papers/Anderson_etal_2013_ecological_prophets_with_SOM.pdf}{PDF},
 \href{https://github.com/seananderson/ecofolio}{GitHub R package}).
@@
 Reynolds, C.T. Darimont. Confronting uncertainty in wildlife management:
-performance of grizzly bear management. PLOS ONE. 8(11): e78041.
-\url{http://doi.org/10.1371/journal.pone.0078041}
+performance of grizzly bear management. PLoS ONE. 8(11): e78041.
+\url{https://doi.org/10.1371/journal.pone.0078041}
 (\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0078041\&representation=PDF}{PDF}).
@@
 pathways to conservation success. Conservation Letters. 6(2): 98--106.
-\url{http://doi.org/10.1111/j.1755-263X.2012.00294.x}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf}{PDF},
+\url{https://doi.org/10.1111/j.1755-263X.2012.00294.x}
+(\href{https://sean.updog.co/papers/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf}{PDF},
 \href{https://github.com/seananderson/conservation_pathways}{GitHub
@@
 Extinctions in ancient and modern seas. Trends in Ecology and Evolution.
-27(11): 608--617. \url{http://doi.org/10.1016/j.tree.2012.07.010}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Harnik_etal_2012_Extinctions_in_ancient_and_modern_seas.pdf}{PDF}).
+27(11): 608--617. \url{https://doi.org/10.1016/j.tree.2012.07.010}
+(\href{https://sean.updog.co/papers/Harnik_etal_2012_Extinctions_in_ancient_and_modern_seas.pdf}{PDF}).
 \item[2012]
@@
 and stock-assessment reference points. ICES Journal of Marine Science.
-69(8): 1491--1500. \url{http://doi.org/10.1093/icesjms/fss105}
+69(8): 1491--1500. \url{https://doi.org/10.1093/icesjms/fss105}
 (\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF}).
@@
 BioScience. 61(7): 538--549.
-\url{http://doi.org/10.1525/bio.2011.61.7.8}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF}).
+\url{https://doi.org/10.1525/bio.2011.61.7.8}
+(\href{https://sean.updog.co/papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF}).
 \item[2011]
@@
 Fisheries. 12(3): 317--339.
-\url{http://doi.org/10.1111/j.1467-2979.2010.00397.x}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF}).
+\url{https://doi.org/10.1111/j.1467-2979.2010.00397.x}
+(\href{https://sean.updog.co/papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF}).
 \href{http://www.sciencemag.org/content/331/6014/129.1.full}{Featured in
@@
 Ocean. Marine Ecology Progress Series. 429: 169--183.
-\url{http://doi.org/10.3354/meps09081}
+\url{https://doi.org/10.3354/meps09081}
 (\href{http://www.int-res.com/articles/meps_oa/m429p169.pdf}{PDF}).
@@
 Rapid global expansion of invertebrate fisheries: trends, drivers, and
-ecosystem effects. PLOS ONE. 6(3): e14735.
-\url{http://doi.org/10.1371/journal.pone.0014735}
+ecosystem effects. PLoS ONE. 6(3): e14735.
+\url{https://doi.org/10.1371/journal.pone.0014735}
 (\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0014735\&representation=PDF}{PDF}).
@@
 Canada. Canadian Journal of Fisheries and Aquatic Sciences. 65(12):
-2553--2571. \url{http://doi.org/10.1139/F08-156}
-(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2008_knowledge.pdf}{PDF}).
+2553--2571. \url{https://doi.org/10.1139/F08-156}
+(\href{https://sean.updog.co/papers/Anderson_etal_2008_knowledge.pdf}{PDF}).
 Eighth most-read CJFAS article in 2008.

### 2017-10-06 cv-body.tex
Fix broken links

@@
 (\href{http://cran.r-project.org/web/packages/ss3sim/index.html}{R
-package}, \href{https://github.com/ss3sim/ss3sim}{code},
-\href{https://sean.updog.co/ss3sim-vignette.pdf}{vignette}).
+package}, \href{https://github.com/ss3sim/ss3sim}{code}.
 \item[2014]
@@
 \url{https://doi.org/10.1890/13-0916.1}
-(\href{http://www.esajournals.org/doi/pdf/10.1890/13-0916.1}{PDF}).
+(\href{http://onlinelibrary.wiley.com/doi/10.1890/13-0916.1/epdf}{PDF}).
 \item[2013]
@@
 Ecology and Evolution. Ideas in Ecology and Evolution. 6: 40--46.
-\url{https://doi.org/10.4033/iee.2013.6.9.n}
-(\href{http://library.queensu.ca/ojs/index.php/IEE/article/download/4931/4899}{PDF}).
+\url{https://doi.org/10.4033/iee.v6i1.4931} (Open access).
 \item[2013]
@@
 69(8): 1491--1500. \url{https://doi.org/10.1093/icesjms/fss105}
-(\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF}).
+(\href{https://sean.updog.co/Anderson_etal_2012_Assessing_global_marine_fishery.pdf}{PDF}).
 \item[2011]

### 2017-10-06 cv-body.tex
Fix typo

@@
 \begin{description}
+\tightlist
 \item[2017]
@@
 (\href{https://sean.updog.co/papers/Kuriyama_etal_2016_empirical.pdf}{PDF}).
-
-;drop
 \item[2016]

### 2017-11-16 cv-body.tex
Add to cv
Not open access
Fix conflicts
Update cv

@@
 \item[2017]
-Aquatic Science Biologist and Program Head in Spatial Stock Assessment,
-Offshore Section, Aquatic Resource Research and Assessment Division,
-Pacific Biological Station, Fisheries and Oceans Canada, Nanaimo, BC,
-Canada
+Senior Aquatic Science Biologist and Program Head in Spatial Stock
+Assessment, Offshore Section, Aquatic Resource Research and Assessment
+Division, Pacific Biological Station, Fisheries and Oceans Canada,
+Nanaimo, BC, Canada
 \item[2015--17]
@@
 \item[2017]
+Ward, E.J., \textbf{S.C. Anderson}, A.O. Shelton, R.E. Brenner,
+M.D.\ Adkison, A.H. Beaudreau,.T. Watson, J.C. Shriver, A.C. Haynie, B.C.
+Williams. Effects of increased specialization and hatchery production on
+revenue variability of Alaskan salmon fishers. In press at Journal of
+Applied Ecology.
+\item[2017]
+\textbf{Anderson, S.C.}, E.J. Ward, A.O. Shelton, M.D.\ Adkison, A.H.
+Beaudreau, R.E. Brenner, A.C. Haynie, J.C. Shriver, J.T. Watson, B.C.
+Williams. Benefits and risks of diversification for individual fishers.
+In press at Proceedings of the National Academy of Sciences.
+\url{https://doi.org/10.1073/pnas.1702506114} (Open access).
+\item[2017]
+\textbf{Anderson, S.C.}, T.A. Branch, A.B. Cooper, N.K. Dulvy. Reply to
+Youngflesh and Lynch: Migration and population growth rate in animal
+black-swan events. Proceedings of the National Academy of Sciences. In
+press at Proceedings of the National Academy of Sciences.
+\url{https://doi.org/10.1073/pnas.1714157114}
+\item[2017]
 \textbf{Anderson, S.C.}, E.J. Ward, A.O. Shelton, M.D.\ Adkison, A.H.
@@
 Hughes, B., S. Lummis, \textbf{S.C. Anderson}, K. Kroeker. Unexpected
-resilience of a seagrass system exposed to multiple stressors. Global
-Change Biology. In press. \url{https://doi.org/10.1111/gcb.13854}
+resilience of a seagrass system exposed to multiple stressors. In press
+at Global Change Biology. \url{https://doi.org/10.1111/gcb.13854}
 \item[2017]
@@
 (\href{https://sean.updog.co/papers/Anderson_etal_2008_knowledge.pdf}{PDF}).
-Eighth most-read CJFAS article in 2008.
 \end{description}
@@
 \item[2017]
-rrfields: Robust random fields for spatiotemporal modelling.
-\url{https://github.com/seananderson/rrfields}
+E.J. Ward, \textbf{Anderson, S.C.} bayesdfa: Bayesian Dynamic Factor
+Analysis (DFA) with Stan. \url{https://github.com/fate-ewi/bayesdfa}
 \item[2017]
-datalimited: Stock assessment methods for data-limited fisheries.
+\textbf{Anderson, S.C.}, E.J. Ward. glmmfields: GLMMs with robust random
+fields for spatial and spatiotemporal modelling.
+\url{https://github.com/seananderson/glmmfields}
+\item[2016]
+\textbf{Anderson, S.C.}, A.B. Cooper, O.P. Jensen, C. Minto, J.T.
+Thorson, J.C. Walsh, J. Afflerbach, M. Dickey-Collas, K.M. Kleisner, C.
+Longo, G.C. Osio, D. Ovando, I. Mosqueira, A.A. Rosenberg, E.R. Selig.
+datalimited: Stock Assessment Methods for Data-limited Fisheries.
 \url{https://github.com/datalimited/datalimited}
@@
 assessment simulation testing with Stock Synthesis.
-\url{http://cran.r-project.org/package=ss3sim}
+\url{http://cran.r-project.org/package=ss3sim},
+\url{https://github.com/ss3sim/ss3sim}
 \item[2013]
@@
 \tightlist
+\item[2017]
+\textbf{Anderson, S.C.}, E.J. Ward, A.O. Shelton. Benefits and risks of
+diversification for individual fishers in Alaska, 1975--2015. NOAA
+National Ocean Service Science Seminar. Presented remotely to
+Washington, DC and other NOAA participants.
 \item[2017]
@@
 \tightlist
+\item[2016--17]
+NOAA Fisheries and the Environment (FATE) working group \emph{An
+early-warning index for abrupt change in northeast Pacific ecosystems}.
 \item[2015-16]
@@
 \tightlist
-\item[Present]
-Proficient in statistical analyses and data management using R, Stan,
-JAGS, Sweave, knitr, Shiny, LaTeX, Git, TMB, SQL. Familiar with ADMB.
 \item[2016]
@@
 Master Communicator's Handbook}}
-\item[2013]
-Quantitative Modelling in Ecosystem Based Management. R. Hilborn, School
-of Aquatic and Fishery Sciences, University of Washington (audited).
 \item[2013]
@@
 Biology, Fish and Fisheries, Marine Policy, Population Ecology, Journal
-of Applied Ecology, International Journal of Tropical Biology and
-Conservation, Journal of Environmental Management, Endangered Species
-Research, Aquatic Conservation
+of Applied Ecology, ICES Journal of Marine Science, International
+Journal of Tropical Biology and Conservation, Journal of Environmental
+Management, Endangered Species Research, Aquatic Conservation

### 2018-07-04 cv-body.tex
Fix cv
Update

--- 2017-10-06 cv-body.tex
@@
-\section{Professional experience}\label{professional-experience}
+\hypertarget{professional-experience}{%
+\section{Professional experience}\label{professional-experience}}
 
@@
 
-\section{Education}\label{education}
+\hypertarget{education}{%
+\section{Education}\label{education}}
 
@@
 
-\section{Publications}\label{publications}
+\hypertarget{publications}{%
+\section{Publications}\label{publications}}
 
@@
 \tightlist
+\item[2017]
+\textbf{Anderson, S.C.}, T.A. Branch, A.B. Cooper, N.K. Dulvy. Reply to
+Youngflesh and Lynch: Migration and population growth rate in animal
+black-swan events. Proceedings of the National Academy of Sciences. In
+press at Proceedings of the National Academy of Sciences.
+\url{https://doi.org/10.1073/pnas.1714157114}
+\item[2017]
+Ward, E.J., \textbf{S.C. Anderson}, A.O. Shelton, R.E. Brenner,
+M.D.\ Adkison, A.H. Beaudreau, J.T. Watson, J.C. Shriver, A.C. Haynie,
+B.C. Williams. Effects of increased specialization on revenue of Alaskan
+salmon fishers over four decades. In press at Journal of Applied
+Ecology.
 \item[2017]
@@
 
-\section{Awards and scholarships}\label{awards-and-scholarships}
+\hypertarget{awards-and-scholarships}{%
+\section{Awards and scholarships}\label{awards-and-scholarships}}
 
@@
 \item[2014]
-Garfield Weston Foundation / BC Packers Ltd. Graduate Fellowship in
+Garfield Weston Foundation / BC Packers Ltd.\ Graduate Fellowship in
 Marine Sciences
@@
 
-\section{Software}\label{software}
+\hypertarget{software}{%
+\section{Software}\label{software}}
 
@@
 
+\hypertarget{invited-talks-and-conference-presentations}{%
 \section{Invited talks and conference
-presentations}\label{invited-talks-and-conference-presentations}
+presentations}\label{invited-talks-and-conference-presentations}}
 
@@
 
-\section{Teaching}\label{teaching}
+\hypertarget{teaching}{%
+\section{Teaching}\label{teaching}}
 
@@
 
-\section{Contract employment}\label{contract-employment}
+\hypertarget{contract-employment}{%
+\section{Contract employment}\label{contract-employment}}
 
@@
 
+\hypertarget{working-groups-and-workshops}{%
 \section{Working groups and
-workshops}\label{working-groups-and-workshops}
+workshops}\label{working-groups-and-workshops}}
 
@@
 
+\hypertarget{selected-courses-and-training}{%
 \section{Selected courses and
-training}\label{selected-courses-and-training}
+training}\label{selected-courses-and-training}}
 
@@
 
-\section{Reviews}\label{reviews}
+\hypertarget{reviews}{%
+\section{Reviews}\label{reviews}}
 

### 2018-07-04 content/blog/2013-04-28-cross-referencing-reviewer-replies-in-latex.md
Merge hugo site

--- 2013-10-20 _posts/2013-04-28-cross-referencing-reviewer-replies-in-latex.md
@@
 ---
-layout: post
 title: Cross-Referencing Line Numbers in LaTeX
 subtitle: How to make writing reviewer response documents easier.
-published: true
+date: '2013-04-28'
+slug: cross-referencing-reviewer-replies-in-latex
 ---
@@
 
+```tex
 \usepackage[pagewise]{lineno}
+```
 
@@
 
+```tex
 \linenumbers
+```
 
@@
 
+```tex
 \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
+```
 
@@
 
+```tex
 \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
+```
 
@@
 
+```tex
 \R{A1} A sentenced we have revised...
+```
 
@@
 
+```tex
 We have revised the text. Please see \lr{A1}.
+```
 
@@
 
+```tex
 We have revised the text. Please see page 3, line 23.
+```
 

### 2018-07-04 content/blog/2013-12-01-plyr.Rmd
Merge hugo site

--- 2013-12-10 _knitr-posts/2013-12-01-plyr.Rmd
@@
 ---
-layout: post
-title: "plyr: Split-Apply-Combine for Mortals"
+title: 'plyr: Split-Apply-Combine for Mortals'
 subtitle: An introduction to the plyr R package.
-published: true
+date: '2013-12-01'
+slug: plyr
 ---
@@
 
-```{r d_ply_plot, fig.width=6, fig.height=2.5, fig.cap=""}
+```{r d_ply_plot, fig.width=6, fig.height=2.5, fig.cap="", tidy=FALSE}
 par(mfrow = c(1, 3), mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0))

### 2018-07-04 content/blog/2014-05-18-gamma-hurdle.Rmd
Merge hugo site

--- 2014-05-20 _knitr-posts/2014-05-18-gamma-hurdle.Rmd
@@
 ---
-layout: post
 title: Gamma Hurdle Models
 subtitle: Fitting and interpreting Gamma hurdle models
-published: true
+date: '2014-05-18'
+slug: gamma-hurdle
 ---
@@
 
-```{r hurdle1}
+```{r hurdle1, cache=FALSE}
 set.seed(1)

### 2018-07-04 content/blog/2014-10-17-tmb.Rmd
Merge hugo site

--- 2014-10-17 _knitr-posts/2014-10-17-tmb.Rmd
@@
 ---
-layout: post
 title: The TMB R package
 subtitle: Life beyond ADMB?
-published: true
+date: '2014-10-17'
+slug: tmb
 ---
+
+```{r}
+knitr::opts_chunk$set(cache = FALSE, autodep = TRUE)
+```
 
@@
   DLL = "regression")
-opt <- do.call("optim", obj)
+opt <- nlminb(start = obj$par, obj = obj$fn, gr = obj$gr)
 rep <- sdreport(obj)
@@
 
-Which is about the same as we'd get from `stats::lm()`:
+Which is about the same as we'd get from `lm()`:
 
@@
 
-Now let's try something a bit more exciting — a state space Gompertz population dynamics model.
+Now let's try something a bit more exciting --- a state space Gompertz population dynamics model.
 We'll have TMB integrate out the unobserved "true" state that is masked by observation error.
@@
 
-\\[\begin{aligned}
-y_t &= \ln N\_t\\\      
-U\_t &amp;= a + b U\_{t-1} + \epsilon\_t\\\
-\epsilon_t &amp;\sim \mathrm{Normal}(0, \sigma^2\_\mathrm{proc})\\\
-y\_t &amp;\sim \mathrm{Normal}(U\_t, \sigma^2\_\mathrm{obs}),
-\end{aligned}\\]
+$$
+\begin{aligned}
+y_t &= \ln N_t\\
+U_t &= a + b U_{t-1} + \epsilon_t\\
+\epsilon_t &\sim \mathrm{Normal}(0, \sigma^2_\mathrm{proc})\\\
+y_t &\sim \mathrm{Normal}(U_t, \sigma^2_\mathrm{obs}),
+\end{aligned}
+$$
 
-where \\(y\_t\\) is the observed \\(\ln\\) abundance (\\(N\\)) at time \\(t\\).
-The model is density independent if \\(b = 1\\), maximally density dependent if \\(b = 0\\), and inversely density dependence if \\(b < 0\\).
-The parameter \\(a\\) represents the expected ln abundance at the next time step when \\(y\_t = 0\\).
-The process noise is modelled as a normal distribution with mean of \\(0\\), and a standard deviation of \\(\sigma\_\mathrm{proc}\\).
-\\(U\\) represents the unobserved state vector, and \\(\sigma\_\mathrm{obs}\\) represents the standard deviation of observation error (on a log scale).
+where $y_t$ is the observed $\ln$ abundance ($N$) at time $t$.
+The model is density independent if $b = 1$, maximally density dependent if $b = 0$, and inversely density dependence if $b < 0$.
+The parameter $a$ represents the expected ln abundance at the next time step when $y_t = 0$.
+The process noise is modelled as a normal distribution with mean of $0$, and a standard deviation of $\sigma_\mathrm{proc}$.
+$U$ represents the unobserved state vector, and $\sigma_\mathrm{obs}$ represents the standard deviation of observation error (on a log scale).
 
 The TMB model for this state-space Gompertz model is:
+
 ```{r tmb-model}
@@
   obj$hessian <- FALSE
-  opt <- do.call("optim", obj)
+  opt <- nlminb(start = obj$par, obj = obj$fn, gr = obj$gr)
   rep <- sdreport(obj)
@@
 
-(We just modelled a 5000 time step state space model in a couple of seconds!)
+We just modelled a 5000 time step state space model in a couple seconds!
 
@@
 
-```{r tmb-short, results='hide', message=FALSE, warning=FALSE, fig.height=8}
+```{r tmb-short, results='hide', message=FALSE, warning=FALSE, fig.height=8, tidy=FALSE}
 sim_gomp(N = 50)

### 2018-07-04 cv/cv-body.tex
Merge hugo site

--- 2018-07-04 cv-body.tex
@@
 \tightlist
-\item[2017]
-Aquatic Science Biologist and Program Head in Spatial Stock Assessment,
-Offshore Section, Aquatic Resource Research and Assessment Division,
-Pacific Biological Station, Fisheries and Oceans Canada, Nanaimo, BC,
-Canada
+\item[2017-18]
+Program Head in Analytical Approaches and Spatial Stock Assessment,
+Groundfish Section, Stock Assessment and Research Division, Pacific
+Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
 \item[2015--17]
@@
 \tightlist
-\item[2017]
-\textbf{Anderson, S.C.}, T.A. Branch, A.B. Cooper, N.K. Dulvy. Reply to
-Youngflesh and Lynch: Migration and population growth rate in animal
-black-swan events. Proceedings of the National Academy of Sciences. In
-press at Proceedings of the National Academy of Sciences.
-\url{https://doi.org/10.1073/pnas.1714157114}
-\item[2017]
+\item[2018]
+\textbf{Anderson, S.C.}, Ward, E. J. Black swans in space: modelling
+spatiotemporal processes with extremes. In press at Ecology.
+\url{https://doi.org/10.1002/ecy.2403}. glmmfields R package:
+\href{https://cran.r-project.org/package=glmmfields}{CRAN},
+\href{https://github.com/seananderson/glmmfields}{GitHub}.
+\item[2018]
 Ward, E.J., \textbf{S.C. Anderson}, A.O. Shelton, R.E. Brenner,
@@
 B.C. Williams. Effects of increased specialization on revenue of Alaskan
-salmon fishers over four decades. In press at Journal of Applied
-Ecology.
-\item[2017]
-\textbf{Anderson, S.C.}, E.J. Ward, A.O. Shelton, M.D.\ Adkison, A.H.
-Beaudreau, R.E. Brenner, A.C. Haynie, J.C. Shriver, J.T. Watson, B.C.
-Williams. Benefits and risks of diversification for individual fishers.
-In press at Proceedings of the National Academy of Sciences.
-\url{https://doi.org/10.1073/pnas.1702506114} (Open access).
-\item[2017]
+salmon fishers over four decades. Journal of Applied Ecology.
+55(3):1082--1091. \url{https://doi.org/10.1111/1365-2664.13058}.
+\item[2018]
 Hughes, B., S. Lummis, \textbf{S.C. Anderson}, K. Kroeker. Unexpected
 resilience of a seagrass system exposed to multiple stressors. Global
-Change Biology. In press. \url{https://doi.org/10.1111/gcb.13854}
-\item[2017]
-Megias, D.A., \textbf{S.C. Anderson}, R.J. Smith, D. Veríssimo.
-Investigating the impact of media on demand for wildlife: a case study
-of Harry Potter and the UK trade in owls. PLoS ONE. 12(10): e0182368.
-\url{https://doi.org/10.1371/journal.pone.0182368} (Open access).
-\item[2017]
-Chezik, K.A., \textbf{S.C. Anderson}, J.W. Moore. River networks dampen
-long-term hydrological signals of climate change. Geophysical Research
-Letters. 44(14):7256--7264. \url{https://doi.org/10.1002/2017GL074376}
-\item[2017]
+Change Biology. 24(1) 224--234. \url{https://doi.org/10.1111/gcb.13854}.
+\item[2018]
 Rosenberg, A.A., K.M. Kleisner, J. Afflerbach, \textbf{S.C. Anderson},
@@
 approach to estimating stock status of marine fisheries around the
-world. Conservation Letters. In press.
-\url{https://doi.org/10.1111/conl.12363}
+world. Conservation Letters. 11(1): e12363.
+\url{https://doi.org/10.1111/conl.12363}.
+\item[2017]
+\textbf{Anderson, S.C.}, T.A. Branch, A.B. Cooper, N.K. Dulvy. Reply to
+Youngflesh and Lynch: Migration and population growth rate in animal
+black-swan events. Proceedings of the National Academy of Sciences.
+114(43): E8955--E8956. \url{https://doi.org/10.1073/pnas.1714157114}.
+\item[2017]
+\textbf{Anderson, S.C.}, E.J. Ward, A.O. Shelton, M.D.\ Adkison, A.H.
+Beaudreau, R.E. Brenner, A.C. Haynie, J.C. Shriver, J.T. Watson, B.C.
+Williams. Benefits and risks of diversification for individual fishers.
+Proceedings of the National Academy of Sciences. 114(40):10797--10802.
+\url{https://doi.org/10.1073/pnas.1702506114}.
+\item[2017]
+Megias, D.A., \textbf{S.C. Anderson}, R.J. Smith, D. Veríssimo.
+Investigating the impact of media on demand for wildlife: a case study
+of Harry Potter and the UK trade in owls. PLoS ONE. 12(10): e0182368.
+\url{https://doi.org/10.1371/journal.pone.0182368}.
+\item[2017]
+Chezik, K.A., \textbf{S.C. Anderson}, J.W. Moore. River networks dampen
+long-term hydrological signals of climate change. Geophysical Research
+Letters. 44(14):7256--7264. \url{https://doi.org/10.1002/2017GL074376}.
 \item[2017]
@@
 Academy of Sciences. 114(12): 3252--3257.
-\url{https://doi.org/10.1073/pnas.1611525114} (Open access).
+\url{https://doi.org/10.1073/pnas.1611525114}.
 \item[2017]
@@
 models. Fish and Fisheries. 18(4): 732--741.
-\url{https://doi.org/10.1111/faf.12200}
+\url{https://doi.org/10.1111/faf.12200}.
 \item[2016]
@@
 \item[2017]
-rrfields: Robust random fields for spatiotemporal modelling.
-\url{https://github.com/seananderson/rrfields}
+glmmfields: Generalized linear mixed models with robust random fields
+for spatiotemporal modeling.
+\url{https://github.com/seananderson/glmmfields}
 \item[2017]
@@
 \tightlist
+\item[2018]
+\textbf{Anderson, S.C.}, Ward, E. J. Black swans in space: modelling
+spatiotemporal processes with extremes. International Statistical
+Ecology Conference, St.\ Andrews, Scotland.
+\href{https://www.dropbox.com/s/z61ssz9xgymtzsg/anderson-ward-isec-2018.pdf?dl=1}{slides}
 \item[2017]
@@
 \tightlist
-\item[2017]
-Half-day workshop on generalized linear modeling in R for the David H.
-Smith Conservation Postdoctoral Research Fellows.
+\item[2018]
+One-day workshop on Bayesian data analysis for ecologists with Stan.
+\url{https://github.com/seananderson/bayes-course}
+\item[2018]
+Two two-day workshops on advanced generalized linear mixed-effects
+modeling in R for the Simon Fraser University Library Research Commons.
 \url{https://github.com/seananderson/glmm-course}
 \item[2017]
-One-day workshop on linear mixed-effects modeling in R for the David H.
-Smith Conservation Postdoctoral Research Fellows.
+One-and-a-half-day workshop on linear mixed-effects modeling in R and
+generalized linear modeling in R for the David H. Smith Conservation
+Postdoctoral Research Fellows.
 \url{https://github.com/seananderson/glmm-course}
 \item[2016]
-Two two-day (14 hour) workshops on advanced generalized linear
-mixed-effects modeling in R for the Simon Fraser University Library
-Research Commons. \url{https://github.com/seananderson/glmm-course}
+Two two-day workshops on advanced generalized linear mixed-effects
+modeling in R for the Simon Fraser University Library Research Commons.
+\url{https://github.com/seananderson/glmm-course}
 \item[2014--16]
@@
 \tightlist
+\item[2016-17]
+NOAA (National Oceanic and Atmospheric Administration) FATE (Fisheries
+and the Environment) working group \emph{An early-warning index for
+abrupt change in northeast Pacific ecosystems}
 \item[2015-16]
@@
 Proficient in statistical analyses and data management using R, Stan,
-JAGS, Sweave, knitr, Shiny, LaTeX, Git, TMB, SQL. Familiar with ADMB.
+JAGS, knitr, Shiny, Git, LaTeX, TMB, SQL.
 \item[2016]
@@
 Reviewer for Science, Ecology, Ecological Applications, Conservation
-Biology, Fish and Fisheries, Marine Policy, Population Ecology, Journal
-of Applied Ecology, International Journal of Tropical Biology and
-Conservation, Journal of Environmental Management, Endangered Species
-Research, Aquatic Conservation
+Biology, Fish and Fisheries, Marine Policy, Canadian Journal of
+Fisheries and Aquatic Sciences, Population Ecology, Journal of Applied
+Ecology, International Journal of Tropical Biology and Conservation,
+Journal of Environmental Management, Endangered Species Research,
+Aquatic Conservation

### 2018-07-04 static/courses/11-multipanel/multipanel.Rnw
Merge hugo site

% Sean Anderson, 2011, sean@seananderson.ca
\documentclass[12pt]{article}
\usepackage{geometry} 
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{Sweave} 
\setkeys{Gin}{width=0.65\textwidth}

\title{Multipanel plotting in R\\(with base graphics)}
\author{Sean Anderson}

\textheight 21cm
\setlength\parskip{0.1in}
\setlength\parindent{0in}

\begin{document}
\maketitle

Edward Tufte, \emph{Envisioning Information}:
\begin{quote}
``At the heart of quantitative reasoning is a single question: \emph{Compared to what?} Small multiple designs, multivariate and data bountiful, answer directly by visually enforcing comparison of changes, of difference among objects, of the scope of alternatives. For a wide range of problems in data presentation, small multiples are the best design solution.''
\end{quote}

\section{Multipanel approaches in R}
To my knowledge, there are five main approaches to multipanel layouts in R.
\begin{description}
  \item[Do them by hand] Manually combine your plots in graphics software outside of R. Advantages: you get complete control over your layout. Disadvantages: just about everything else. Your figure is no longer reproducible. This can become increasingly annoying as analyses inevitably get re-run. It can also be time-consuming to perfectly line up your panels. I try and avoid this at all costs, but occasionally it's your only or best choice.
  \item[grid graphics, \texttt{lattice}, \texttt{ggplot2}]
Packages like \texttt{ggplot2} and \texttt{lattice} are great. Where I think they excel is in exploratory data analysis. You might be able to generate ten \texttt{ggplot} figures in the time it would take you to do the same in base graphics. Data analysis involves a lot of exploratory data plotting, so don't underestimate the value of this. Base graphics shine when it comes to plot customization. Data presentation for publication often consists of making highly-customized plots tailored to your specific situation. I use both, but almost always base graphics for publication. Learning a grid graphics package can be very helpful, but you still need to learn base graphics. This workshop will focus on base graphics. We'll cover grid graphics another time.
\item[\texttt{par(mfrow)}] The simplest method in base graphics. Works well for simple grid layouts where each panel is the same size.
\item[\texttt{layout()}] In addition to what you can do with \texttt{par(mfrow)}, \texttt{layout()} lets you combine panels.
\item[\texttt{split.screen()}] Lets you specify the co-ordinates of your panels. Panels no longer have to be simple ratios of each other.
\end{description}

\section{Where I make a silly analogy to explain the increasing levels of complexity}
\texttt{par(mfrow)}, \texttt{layout()}, and \texttt{split.screen()} are all capable of basic equal-sized-panel grid layouts. If you think of creating a small multiple layout in R to be like putting screws into a wall: \texttt{par(mfrow)} would be the equivalent of grabbing your Leatherman to hang one picture frame --- it's all you need and it's fast. \texttt{layout()} would be the equivalent of hunting around for a proper screwdriver to hang a bunch of picture frames. \texttt{split.screen} would be the equivalent of finding and plugging in your power drill --- more of a hassle to set up, but much more powerful in the end.
 Don't grab a tool that's more complex than it needs to be, but don't try and build a house with a Leatherman.

\section{Questions to ask yourself when making a multipanel plot}
\begin{enumerate}
  \item What comparison do I want to emphasize?
  \item How can I use order to enhance the comparison?
  \item Is this a series of plots or does the grid layout matter? (\texttt{facet\_wrap} vs.\ \texttt{facet\_grid} in \texttt{ggplot2} terminology)
  \item What's a reasonable number of panels to show? Everything? A sample?
  \item Which axes can I fix and which need to vary? Would a log transformation be appropriate and allow the axes to be combined?
  \item What chart junk can I remove?
  \item What's important in my plots and what necessary but less-important elements do I want to de-emphasize?
  \item Can I make it all smaller and increase the information density without detracting from readability? (Almost always, yes.)
  \item If the layout is complicated, have I drawn it out on paper first?
\end{enumerate}

\section{Margin space}
Extra margins are usually wasted space and a break in the comparisons between panels. You will almost always want to shrink your margins.  Set your margins for each panel with \texttt{mar} and your outer margins with \texttt{oma}. If all the axes can be shared then set \texttt{mar = c(0,0,0,0)}. These numbers refer to the space on the bottom, left, top, and right. Then you can use \texttt{par(oma)} to set your outer margins to create the necessary space for axes. If your content won't show up in the outer margins, you'll need to set \texttt{par(xpd = NA)}.

\section{Ways to iterate through your data}
Common approaches are to use a \texttt{for} loop with subsetting or an \texttt{apply} function. You could also manually make all your plots, but unless you were only making a few plots you wouldn't do that would you?

A favourite approach of mine is to use \texttt{d\_ply()} from the \texttt{plyr} package. This takes a data frame, splits it up, does something with it (plots it), but doesn't return a value.

\section{Basic multipanel layouts with \texttt{par(mfrow)}}
For most basic grid layouts, \texttt{par(mfcol)} or \texttt{par(mfrow)} are your simplest option. \texttt{mfrow} plots row by row and \texttt{mfcol} plots column by column. \texttt{mfrow} is therefore likely the most commonly used option. You're going to give \texttt{mfrow} vector of length two corresponding to the number of rows followed by the number of columns.

Let's try a basic example with 2 rows and 3 columns:

<<echo = true, fig = true, height = 3, width = 3>>=
par(mfrow = c(2, 3))
par(cex = 0.6)
par(mar = c(3, 3, 0, 0), oma = c(1,1,1,1))
for(i in 1:6) {
  plot(1, 1, type = "n")
  mtext(letters[i], side = 3, line = -1, adj = 0.1, cex = 0.6)
}
@

We can eliminate the redundant axes, remove margin space, and reduce the emphasis on the structural (non-data) elements of the figure. These are some of the frequent ``tricks'' you can use to create a basic multipanel layout that will focus the reader's attention on trends in the data. If you aren't familiar with an option for \texttt{par()}, look up the help: \texttt{?par}.

<<echo = true, fig = true, height = 3, width = 3>>=
par(mfrow = c(2, 3))
par(cex = 0.6)
par(mar = c(0,0,0,0), oma = c(4, 4, .5, .5))
par(tcl = -0.25)
par(mgp = c(2, 0.6, 0))
for(i in 1:6) {
  plot(1, axes = FALSE, type = "n")
  mtext(letters[i], side = 3, line = -1, adj = 0.1, cex = 0.6, col = "grey40")
  if(i %in% c(4, 5, 6)) axis(1, col = "grey40", col.axis = "grey20", at = seq(0.6, 1.2, 0.2))
  if(i %in% c(1, 4)) axis(2, col = "grey40", col.axis = "grey20", at = seq(0.6, 1.2, 0.2))
  box(col = "grey60")
}
  mtext("x axis", side = 1, outer = TRUE, cex = 0.7, line = 2.2, col = "grey20")
  mtext("y axis", side = 2, outer = TRUE, cex = 0.7, line = 2.2, col = "grey20")
@

\section{Fancy multipanel layouts with \texttt{layout()}}

Say you wanted to make a figure with one wide panel on top and two smaller panels underneath.
We can't do that with \texttt{par(mfrow)}, now can we?  This is where \mbox{\texttt{layout()}} becomes useful. It takes a matrix and turns it into a layout. The shape of the matrix corresponds to individual cells. The numbers in the matrix correspond to the order of the plots.  Cells with the same number represent a single panel. I often find it easiest to create a matrix with \texttt{rbind()} and a series of vectors representing the rows.

Let's generate a layout like the previous example. First I will use the \texttt{layout.show()} command to see how the figure will look. The numbers correspond to the order that the panels will be plotted in.

\setkeys{Gin}{width=0.30\textwidth}
<<echo = true, fig = true, results = verbatim, height = 2, width = 2>>=
m <- rbind(c(1, 1), c(2, 3))
print(m)
layout(m)
layout.show(3)
@
\setkeys{Gin}{width=0.65\textwidth}

<<echo = true, fig = true, results = hide>>=
layout(m)
par(mar = c(3, 3, 0, 0))
for(i in 1:3) plot(1, 1, type = "n")
@

You can create some fairly complex layouts with \texttt{layout()} once you realize that numbers that are higher than the number of plots you make will be left empty. They can be used to generate margin space. For example, take this matrix:
<<echo = false, results = verbatim>>=
load("cod.crab.layout.rda")
row.names(m) <- NULL
print(m)
@

This matrix can be turned into the layout shown below. Note that there are only 30 panels that are plotted. The rest of the panels make up margin space.

\setkeys{Gin}{width=0.50\textwidth}
<<echo = false, fig = true, height = 2, width = 2>>=
layout(m)
par(cex = 0.3)
layout.show(50)
@
\setkeys{Gin}{width=0.65\textwidth}

\vspace{5mm}
Here's the finished product:\\
\includegraphics[width=5.2in]{cod-crab-temp.pdf}

This is about as complicated as it's worth getting with \texttt{layout()}. In fact, this would probably be simpler to accomplish with \texttt{split.screen()}, which I will now discuss.

\section{Crazy multipanel layouts with \texttt{split.screen()}}
What if you want different panels to be different sizes and the sizes don't correspond to some simple ratio that you could divide up with \texttt{layout()}? One common use for this is to keep the unit scales in different panels the same, without using extra margin space. \texttt{split.screen()} is only limited by your imagination and ability to figure out the co-ordinates of your plots.

Split screen takes a set of vectors or a matrix to specify the layout. I typically use the matrix notation. In this case each row describes a \texttt{screen} with values for left, right, bottom, and top.\footnote{Note that this is different than the order used in \texttt{par} statements!} 0 corresponds to the left or bottom of the plot. 1 corresponds to the top or right of the plot. Sometimes it's easiest to figure out this matrix in a spreadsheet before and then read it in as a .csv file.

First let's generate a ``simple'' two-by-two layout with \texttt{split.screen()}.

\begin{verbatim}
> m <- rbind(c(0.1, 0.55, 0.55, 1), c(0.55, 1, 0.55, 1), 
+   c(0.1, 0.55, 0.1, 0.55), c(0.55, 1, 0.1, 0.55))
> split.screen(m)
> for(i in 1:4) {
+   screen(i)
+   par(mar = c(0, 0, 0, 0))
+   plot(1, axes = FALSE, type = "n")
+   box()
+   if(i %in% c(3, 4)) axis(1)
+   if(i %in% c(1, 3)) axis(2)
+ }
> close.screen(all.screens = TRUE)
\end{verbatim}

\includegraphics[width=3.5in]{split-basic.pdf}

Here's an example of setting up the panels to arbitrary sizes so that the units are equal on both x-axes. I figured out these values in a spreadsheet first.
\begin{verbatim}
> split.screen(rbind(c(0.1,0.292,0.1, 0.98), c(0.312, 0.95, 0.1, 0.98)))
> screen(1)
> par(mar = c(0, 0, 0, 0))
> plot(1:30, rnorm(30), xaxs = "i", ylim = c(-3, 3), xaxt = "n")
> axis(1, at = seq(0, 30, 20))
> 
> screen(2)
> par(mar = c(0, 0, 0, 0))
> plot(1:100, rnorm(100), xaxs = "i", ylim = c(-3, 3), yaxt = "n", col = "red")
> close.screen(all.screens = TRUE)
\end{verbatim}

\includegraphics[width=2.4in]{split-screen-fancy.pdf}

With \texttt{split.screen()} you can apply the same concept to vertical axes. With a bit of planning you can make fairly elaborate layouts.\footnote{See Branch et al.\ 2010, Nature, The trophic fingerprint of marine fisheries, for a great example of this layout.}\\
\includegraphics[width=3.7in]{staggered.pdf}

\end{document}

### 2018-07-04 static/courses/11-plyr/plyr.Rnw
Merge hugo site

% Sean Anderson, 2011, sean@seananderson.ca
\documentclass[12pt]{article}
\usepackage{geometry} 
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{Sweave} 
\usepackage{url}
\usepackage[usenames,dvipsnames]{color}
\usepackage[pdftex,colorlinks=true,urlcolor=MidnightBlue]{hyperref}
\usepackage{booktabs}
\usepackage{listings}

\title{A quick introduction to plyr}
\author{Sean Anderson}
%\date{}

\setlength\parskip{0.1in}
\setlength\parindent{0in}

\begin{document}
\maketitle

<<echo = false, results = hide>>=
library(plyr)
@

What is \texttt{plyr}? It's a bundle of awesomeness (i.e.\ an R package) that makes it simple to split apart data, do stuff to it, and mash it back together. This is a common data manipulation step.

Or, from the documentation:

``\texttt{plyr} is a set of tools that solves a common set of problems: you need to break a big problem down into manageable pieces, operate on each pieces and then put all the pieces back together. It's already possible to do this with split and the apply functions, but \texttt{plyr} just makes it all a bit easier\ldots''

This is a very quick introduction to \texttt{plyr}. For more details have a look at the \texttt{plyr} site: \url{http://had.co.nz/plyr/} and particularly Hadley Wickham's introductory guide \emph{The split-apply-combine strategy for data analysis}.

\url{http://had.co.nz/plyr/plyr-intro-090510.pdf}

\section{Why use \texttt{apply} functions instead of \texttt{for} loops?}
\begin{enumerate}
  \item the code is cleaner -- easier to code and read, and less error prone because:
    \begin{enumerate}
  \item you don't have to deal with subsetting
  \item you don't have to deal with saving your results
    \end{enumerate}
  \item apply functions are often faster, sometimes dramatically 
\end{enumerate}

\section{Why use \texttt{plyr} over base \texttt{apply} functions?}
\begin{enumerate}
  \item \texttt{plyr} has a common syntax --- easier to remember
  \item \texttt{plyr} requires less code since it takes care of the input and output format
  \item \texttt{plyr} can be run in parallel --- faster
\end{enumerate}

\section{The basic idea behind apply functions}
\texttt{apply} functions work by applying a function to a set of values and returning the output in some format. Here's about as simple an example as possible:

<<<echo = true>>=
y <- c(1, 2, 3)
f <- function(x) x^2
sapply(y, f)
@
Here, I have applied the function \texttt{f} to the values of \texttt{y}. Note that the \texttt{sapply} function was unnecessary here. This would have been better done in a vectorized format. If \texttt{y} was large, and the function more complex, the vectorized format could be noticeably faster. 
<<<echo = true>>=
f(y)
@
But, it isn't always possible (or easy) to vectorize a function, particularly when you're dealing with groupings of data as in the following examples.

\section{\texttt{plyr} basics}
\texttt{plyr} builds on the built in \texttt{apply} functions by giving you control over the input and output formats and keeping the syntax consistent across all variations. It also adds some niceties like error processing, parallel processing, and progress bars.

The basic format is 2 letters followed by  \texttt{ply()}. The first letter refers to the format in and the second to the format out.

The 3 main letters are: 
\begin{enumerate}
  \item \texttt{d} = data frame
  \item \texttt{a} = array (includes matrices)
  \item \texttt{l} = list
\end{enumerate}

So, \texttt{ddply} means: take a data frame, split it up, do something to it, and return a data frame. I find I use this the majority of the time since I often work with data frames.

\texttt{ldply} means: take a list, split it up, do something to it, and return a data frame. This extends to all combinations. The columns are the input formats and the rows are the output format:

\begin{tabular}{llll}
  \toprule
 & data frame & list & array\\
 \midrule
 data frame & \texttt{ddply} & \texttt{ldply} & \texttt{adply}\\
 list & \texttt{dlply} & \texttt{llply} & \texttt{alply}\\
 array & \texttt{daply} & \texttt{laply} & \texttt{aaply}\\
\bottomrule
\end{tabular}

I've ignored a couple other format options. One that you might find useful is the underscore (\texttt{\_}) which will throw away the output (e.g.,\ \texttt{d\_ply()}). This can be useful when plotting.

\section{A general example with \texttt{plyr}}
Let's take a simple example. Take a data frame, split it up (by \texttt{year}), calculate the coefficient of variation of the \texttt{count}, and return a data frame. This could easily be done on one line, but I'm expanding it here to show the format a more complex function could take.
<<echo = true>>=
set.seed(1)
d <- data.frame(year = rep(2000:2002, each = 3), count = round(runif(9, 0, 20)))
print(d)
library(plyr)
ddply(d, "year", function(x) {
  mean.count <- mean(x$count)
  sd.count <- sd(x$count)
  cv <- sd.count/mean.count
  data.frame(cv.count = cv)
})
@

\section{\texttt{transform} and \texttt{summarise}} 
It is often convenient to use these functions within \texttt{plyr}. \texttt{transform} acts as it would normally as the base R function and modifies an existing data frame. \texttt{summarise} creates a new (usually) condensed data frame.
<<echo = true>>=
ddply(d, "year", summarise, mean.count = mean(count))
ddply(d, "year", transform, total.count = sum(count))
@

%\section{Common usage examples}
%You can use \texttt{plyr} to plot data by throwing away the ouput with an underscore (\texttt{\_}). This is a bit cleaner than a for loop since you don't have to subset the data manually. Often you'd be better off using a grid graphics package for this such as \texttt{ggplot2} or \texttt{lattice}, but here's an example:
%
%<<echo = true, fig = true, height = 2, width = 5>>=
%par(mfrow = c(1, 3), mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0))
%d_ply(d, "year", summarise, hist(count, main = unique(year)))
%mtext("count", side = 1, outer = TRUE, line = 1)
%mtext("frequency", side = 2, outer = TRUE, line = 1)
%@

\section{Other useful options}
\subsection{Dealing with errors}
You can use the \texttt{failwith} function to control how errors are dealt with.
<<echo = true>>=
f <- function(x) if (x == 1) stop("Error!") else 1
safe.f <- failwith(NA, f, quiet = TRUE)
#llply(1:2, f)
llply(1:2, safe.f)
@

\subsection{Parallel processing}
In conjunction with \texttt{doMC} (or \texttt{doSMP} on Windows) you can run your function separately on each core of your computer. On a dual core machine this could double your speed in some situations. Set \texttt{.parallel = TRUE}.
<<<echo = true>>=
  x <- c(1:10)
  wait <- function(i) Sys.sleep(0.1)
  system.time(llply(x, wait))
  system.time(sapply(x, wait))

  library(doMC)
  registerDoMC(2) 
  system.time(llply(x, wait, .parallel = TRUE))
@

\end{document}

### 2018-07-04 static/courses/12-ggplot2/ggplot2.Rnw
Merge hugo site

% Sean Anderon, 2012, sean@seananderson.ca
\documentclass{tufte-handout}
\usepackage{geometry} 
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{Sweave} 
\usepackage{url} 
\urlstyle{rm} 
\setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}
\usepackage{fancyvrb}
\fvset{fontsize=\normalsize}
\bibliographystyle{apalike}   
\usepackage{booktabs}

\title{Lecture 9: An introduction to ggplot2}
\author{Sean C. Anderson}

\begin{document}
\maketitle

<<echo = false, results = hide>>=
library(ggplot2)
@

% Components
% 
% Data
% Geometric object (geom) Statistical transformation (stat) Scales
% Coordinate system
% (plus Position adjustment, facetting)

% Previous description is a bit of a simplification
% Actually have: defaults + layers + scales + coordinate system
% Layer = data + mapping + geom + stat + position

\section{The ggplot philosophy: rapid data exploration}
\texttt{ggplot2} is an R package that implements Wilkinson's Grammar of
Graphics.\sidenote{Wilkinson, L. (2005). \textit{The Grammar of Graphics}. Springer, 2\textsuperscript{2nd} edition.} Hadley Wickham wrote the package as a chapter of his PhD thesis. Many people now participate in developing the package.

The emphasis of \texttt{ggplot} is on rapid exploration of data, and especially high-dimensional data. Think of base graphics functions as drawing with data. You have complete control over every pixel in a plot (once you learn the arcane world of \texttt{par}) but it can take a lot of time and code to produce a complex plot.

Although \texttt{ggplot} can be fully customized, I find it reaches a point of diminishing returns. I tend to use \texttt{ggplot} and base graphics for what they excel at: \texttt{ggplot} for rapid data exploration and base graphics for polished and fully-customized plots for publication.

The idea is simple: good graphical displays data require rapid iteration and lots of exploration. If it takes you hours to code a plot in base graphics, you're unlikely to throw it out and try something else. If it takes you hours to code a plot in base graphics, you're unlikely to explore other ways of visualizing the data or all the dimensions of the data.

% \section{Why ggplot?}
% 
% OK, so \texttt{ggplot} lets you visualize data quickly.
% 
% \begin{enumerate}
%   \item Automatic assignment of colours, sizes
%   \item Automatic creation of legends
%   \item Complex figures with one line of code
% \end{enumerate}

%More thought, less coding

\section{qplot vs.\ ggplot}     
There are two main plotting functions in the \texttt{ggplot2} package: \texttt{qplot} and \texttt{ggplot}. \texttt{qplot} is short for ``quick plot'' and is made to mimic the format of \texttt{plot} from base R. \texttt{qplot} requires less syntax for many common tasks, but has limitations --- it's essentially a wrapper for \texttt{ggplot}. The \texttt{ggplot} function itself isn't complicated and will work in all cases. I prefer to work with just the \texttt{ggplot} syntax and will focus on it here.

\section{Basics of the grammar}
Let's look at some illustrative \texttt{ggplot} code:

<<echo = true, fig = true, height = 4, width = 8>>=
 d <- data.frame(x = c(1:10, 1:10), y = runif(20), 
                 group1 = rep(gl(2, 5, labels = c("a", "b")), 2), 
                 group2 = gl(2, 10))
 ggplot(d) + geom_point(aes(x, y, colour = group1)) + facet_grid(~group2)
@

The basic format in this example is:

\begin{enumerate}
  \item \texttt{ggplot()}: start a \texttt{ggplot} object and specify the data
  \item \texttt{geom\_point()}: we want a scatter plot; this is called a \texttt{geom}
  \item \texttt{aes()}: specifies the ``aesthetic'' elements; a legend is automatically created
  \item \texttt{facet\_grid()}: specifies the panel layout
\end{enumerate}

There are also statistics, scales, and annotation options, among others.
% plot = data + aesthetics + geoms + facets + scales + statistics + annotation
At a minimum, you must specify the data, some aesthetics, and a geom. I will elaborate on these below. Yes, \texttt{ggplot} combines elements with \texttt{+} symbols!\sidenote{This may seem non-standard, although it has the advantage of allowing \texttt{ggplot} plots to be proper R objects, which can modified, inspected, and re-used.}

\section{Geoms}
\texttt{geom} refers to a geometric object. It determines the ``shape'' of the plot elements. Some common geoms:

\begin{tabular}{llll}
  \toprule
  \texttt{geom} & description \\
 \midrule
 \texttt{geom\_point} & Points, e.g.\ a scatterplot\\
 \texttt{geom\_line} & Lines\\
 \texttt{geom\_ribbon} & Ribbons, y range with continuous x values\\
 \texttt{geom\_polygon} & Polygon, a filled path\\
 \texttt{geom\_pointrange} & Vertical line with a point in the middle\\ % todo - flip axes
 \texttt{geom\_linerange} & An interval represented by a vertical line\\
 \texttt{geom\_path} & Connect observations in original order\\
 \texttt{geom\_histogram} & Histograms \\
 \texttt{geom\_text} & Textual annotations\\
 \texttt{geom\_violin} & Violin plot\\
 \texttt{geom\_map} & Polygons from a map\\
\bottomrule
\end{tabular}

\section{Aesthetics}
Aesthetics refer to the attributes of the data you want to display. They map the data to an attribute (such as the size or shape of a symbol) and generate an appropriate legend. Aesthetics are specified with the \texttt{aes} function.

As an example, the aesthetics available for \texttt{geom\_point} are: \texttt{x}, \texttt{y}, \texttt{alpha}, \texttt{colour}, \texttt{fill}, \texttt{shape}, and \texttt{size}.\sidenote{Note that \texttt{ggplot} tries to accommodate the user who's never ``suffered'' through base graphics before by using intuitive terms like \texttt{colour}, \texttt{size}, and \texttt{linetype}, but \texttt{ggplot} will also accept terms such as \texttt{col}, \texttt{cex}, and \texttt{lty}.} Read the help files to see the aesthetic options for the geom you're using. They're generally self explanatory.

Aesthetics can be specified within the data function or within a geom. If they're specified within the data function then they apply to all geoms you specify.

Note the important difference between specifying characteristics like colour and shape inside or outside the \texttt{aes} function --- those inside the \texttt{aes} function are assigned the colour or shape automatically based on the data. If characteristics like colour or shape are defined outside the \texttt{aes} function, then the characteristic is not mapped to data. Example:

<<echo = true, fig = true, height = 4, width = 6>>=
library(ggplot2)
ggplot(mpg, aes(cty, hwy)) + geom_point(aes(colour = class))
@

<<echo = true, fig = true, height = 4, width = 6>>=
ggplot(mpg, aes(cty, hwy)) + geom_point(colour = "red")
@

\section{Small multiples}
In \texttt{ggplot} parlance, small multiples are referred to as facets. There are two kinds: \texttt{facet\_wrap} and \texttt{facet\_grid}.
This is where \texttt{ggplot} really shines.

<<echo = true, fig = true, height = 5, width = 6.5>>=
ggplot(mpg, aes(cty, hwy)) + geom_point() + facet_wrap(~class)
@

\clearpage
<<echo = true, fig = true, height = 3, width = 8>>=
 ggplot(mpg, aes(cty, hwy)) + geom_point() + facet_grid(year~class)
@

\texttt{face\_wrap} plots the panels in the order of the factor levels. When it gets to the end of a column it wraps to the next column. You can specify the number of columns and rows with \texttt{nrow} and \texttt{ncol}. \texttt{facet\_grid} lays out the panels in a grid with an explicit x and y position.

By default all x and y axes will be shared among panels. You could, for example, specify ``free'' y axes with \texttt{face\_wrap(scales = "free\_y")}.

% advantages
% automatic assignment and legend for "aesthetics"
% 
% discrete vs. continuous variables
% 
% experiment with aesthetics

% todo \sidenote{Use with \texttt{axis\_flip} if you want horizontal lines with a point in the middle.}

% position
% jitter
% dodge

% \section{stats}
% stat\_smooth
% 
% \section{coordinate systems}
% coord\_flip
% Flipped cartesian coordinates.
% 
% \section{position adjustments}
% dodge jitter

% binwidth for histograms

% \texttt{coord\_cartesian} vs zooming with xlim, ylim

% combining geoms with qplot

\section{Themes}
A useful theme built into \texttt{ggplot} is \texttt{theme\_bw}:
<<echo = true, fig = true, height = 2.5, width = 4.5>>=
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_bw()
@

A powerful aspect of \texttt{ggplot} is that you can write your own themes. This feature of \texttt{ggplot} was recently expanded substantially, and I imagine we'll see more themes developed and shared in the future. See the \texttt{ggthemes} package for some examples.\sidenote{Install the R package from:\\ \url{https://github.com/jrnold/ggthemes}}

\clearpage
A Tufte-like theme:
<<echo = true, fig = true, height = 2.5, width = 4.5>>=
library("ggthemes")
ggplot(mtcars, aes(wt, mpg)) + geom_point() + geom_rangeframe() +
 theme_tufte()
@

Just what you wanted:
<<echo = true, fig = true, height = 2.5, width = 5>>=
qplot(carat, price, data = dsamp, colour = cut) + theme_excel2003() +
 scale_colour_excel2003()
@

You can customize just about every aspect of a \texttt{ggplot} plot. We won't get into that today.

\section{ggplot's dirty little secret}
\begin{enumerate}
  \item \texttt{ggplot} is easy to learn, however \ldots
  \item you need to be a data-manipulation ninja to effectively use it
\end{enumerate}

With base plot, you can work with almost any data structure you'd like providing you can write the code to work with it. \texttt{ggplot} requires you to think carefully about the data structure and then write one line of code.

\texttt{ggplot} works with ``long'' format data with each aesthetic or facet variable in its own column.  So, for example, if we wanted a panel for each level of a factor called ``fishstock'' then we'd need a column named ``fishstock'' with all the different values of ``fishstock'' in that column.

With the \texttt{reshape} and \texttt{plyr} packages you can get almost any dataset into shape for \texttt{ggplot} in a few lines. Sometimes this will take some serious thought.

\section{Help}
The R help files were notoriously bad but are rapidly improving.

The best (and nearly essential) help source is the website.\sidenote{\url{http://docs.ggplot2.org/}} There's also an active \texttt{ggplot} discussion group.\sidenote{\url{http://groups.google.com/group/ggplot2}} 
\texttt{ggplot} is heavily featured on stackoverflow.\sidenote{\url{http://stackoverflow.com/questions/tagged/ggplot2}}

Hadley wrote a book on \texttt{ggplot}. It's quite thorough but doesn't feature some of the newer additions to \texttt{ggplot}.

\section{Random tips}

\begin{description}
  \item[Jittering and statistics] \hfill \\
    \texttt{ggplot2} has lots of built in niceties like automatic position jittering and the addition of basic statistical models to your plots. Have a look through the website.
  \item[Axis labels] \hfill \\
    \texttt{xlab("Your x-axis label")}
  \item[Suppressing the legend] \hfill \\ 
    \texttt{theme(legend.position = "none")}
  \item[Exploiting the object-oriented nature of \texttt{ggplot2}] \hfill \\
    Save the basic plot to an object and then experiment with different aesthetics, geoms, and theme adjustments.
<<<echo = true, figure = false>>=
 p <- ggplot(d) + geom_point(aes(x, y, colour = group1)) 
 p <- p + facet_grid(~group2)
@
  \item[Horizontal error bars]\hfill \\  
    Say you want to make a coefficient plot with the coefficients down the y-axis. You can either build the plot piece by piece with points and segments, or you can use \texttt{point\_range()} and then rotate the axes with \texttt{+ coord\_flip()}.
  \item[Axis limits and zooming] \hfill \\
    \texttt{ggplot2} has two ways of adjusting the axis limits: \texttt{+ xlim(2, 5)} will ``cut'' the data at \texttt{2} and \texttt{5} and plot the data. \texttt{coord\_cartesian(xlim = c(2, 5))} will zoom in on the plot while retaining the original data. This will, for example, affect colour scales.
  \item[Displaying and saving ggplot2 plots] \hfill \\
    If \texttt{ggplot2} plots are generated in a function, they will be need to be wrapped in \texttt{print()} to display. There is a convenience function \texttt{ggsave("filename.pdf")}, which will save the last plot to a pdf and guess at reasonable dimensions. You can, of course specify which plot to save and the dimensions of the pdf. You can also use all the same methods of saving \texttt{ggplot2} plots that you can use for base graphics.
\end{description}

% \bibliography{/Users/seananderson/Dropbox/tex/jshort.bib,/Users/seananderson/Dropbox/tex/ref.bib}
\end{document}

% Change legend label: + labs(colour = "My Custom Legend")
% Hide all legends: + opts(legend.position = "none")
% Angle axis labels: opts(axis.text.x = theme text(angle = 45, hjust = 1, colour
% = "grey50")
% Reorder factors: Attributes$Dexterity <- factor(Attributes$Dexterity, levels =
% c("Low", "Mid", "Hi"))

### 2018-07-04 static/courses/12-plyr/plyr_2012.Rnw
Merge hugo site

--- 2018-07-04 static/courses/11-plyr/plyr.Rnw
@@
-% Sean Anderson, 2011, sean@seananderson.ca
+% Sean Anderson, 2011-12, sean@seananderson.ca
 \documentclass[12pt]{article}
@@
 \author{Sean Anderson}
-%\date{}
 
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 \maketitle
@@
 
-What is \texttt{plyr}? It's a bundle of awesomeness (i.e.\ an R package) that makes it simple to split apart data, do stuff to it, and mash it back together. This is a common data manipulation step.
+\texttt{plyr} is an R package that makes it simple to split data apart, do stuff to it, and mash it back together.
+This is a common data-manipulation step.
+Importantly, \texttt{plyr} makes it easy to control the input and output data format from a syntactically consistent set of functions.
 
@@
 This is a very quick introduction to \texttt{plyr}.
-For more details have a look at the \texttt{plyr} site: \url{http://had.co.nz/plyr/} and particularly Hadley Wickham's introductory guide \emph{The split-apply-combine strategy for data analysis}.
-
-\url{http://had.co.nz/plyr/plyr-intro-090510.pdf}
+For more details see Hadley Wickham's introductory guide \emph{\href{http://www.jstatsoft.org/v40/i01}{The split-apply-combine strategy for data analysis}} (2011, Journal of Statistical Software, Vol 40).
+There's quite a bit of discussion online in general, and especially on \href{http://stackoverflow.com/questions/tagged/plyr}{stackoverflow.com}.
 
 \section{Why use \texttt{apply} functions instead of \texttt{for}  loops?}
+
 \begin{enumerate}
-  \item the code is cleaner -- easier to code and read, and less error prone because:
+  \item The code is cleaner (once you're familiar with the concept). The code can be easier to code and read, and less error prone because:
     \begin{enumerate}
@@
     \end{enumerate}
-  \item apply functions are often faster, sometimes dramatically 
+  \item Apply functions can be faster than for loops, sometimes dramatically.
 \end{enumerate}
@@
 \section{Why use \texttt{plyr} over base \texttt{apply} functions?}
+
 \begin{enumerate}
@@
   \item \texttt{plyr} requires less code since it takes care of the input and output format
-  \item \texttt{plyr} can be run in parallel --- faster
+  \item \texttt{plyr} can easily be run in parallel --- faster
 \end{enumerate}
 
-\section{The basic idea behind apply functions}
-\texttt{apply} functions work by applying a function to a set of values and returning the output in some format. Here's about as simple an example as possible:
+\section{\texttt{plyr} basics}
 
-<<<echo = true>>=
-y <- c(1, 2, 3)
-f <- function(x) x^2
-sapply(y, f)
-@
-Here, I have applied the function \texttt{f} to the values of \texttt{y}.
-Note that the \texttt{sapply} function was unnecessary here.
-This would have been better done in a vectorized format.
-If \texttt{y} was large, and the function more complex, the vectorized format could be noticeably faster.
-<<<echo = true>>=
-f(y)
-@
-But, it isn't always possible (or easy) to vectorize a function, particularly when you're dealing with groupings of data as in the following examples.
-
-\section{\texttt{plyr} basics}
-\texttt{plyr} builds on the built in \texttt{apply} functions by giving you control over the input and output formats and keeping the syntax consistent across all variations.
+\texttt{plyr} builds on the built-in \texttt{apply} functions by giving you control over the input and output formats and keeping the syntax consistent across all variations.
 It also adds some niceties like error processing, parallel processing, and progress bars.
@@
 
-I've ignored a couple other format options.
-One that you might find useful is the underscore (\texttt{\_}) which will throw away the output (e.g.,\ \texttt{d\_ply()}).
-This can be useful when plotting.
+I've ignored some less common format options:
+
+\begin{enumerate}
+  \item \texttt{m} = multi-argument function input
+  \item \texttt{r} = replicate a function \texttt{n} times.
+  \item \texttt{\_} = throw away the output
+\end{enumerate}
+
+For plotting, you might find the underscore (\texttt{\_}) option useful. It will do something with the data (say add line segments to a plot) and then throw away the output (e.g.,\ \texttt{d\_ply()}).
+
+\section{Base R \texttt{apply} functions and \texttt{plyr}}
+
+\texttt{plyr} provides a consistent and easy-to-work-with format for \texttt{apply} functions with control over the input and output formats.
+Some of the functionality can be duplicated with base R functions (but with less consistent syntax).
+Also, few R \texttt{apply} functions work directly with data frames as input and output and data frames are a common object class to work with.
+
+Base R \texttt{apply} functions (from a presentation given by Hadley):
+
+\begin{tabular}{lllll}
+\toprule
+ & array & data frame & list & nothing\\
+\midrule
+array & \texttt{apply} & . & . & .\\
+data frame & . & \texttt{aggregate} & \texttt{by} & .\\
+list & \texttt{sapply} & . & \texttt{lapply} & .\\
+n replicates & \texttt{replicate} & . & \texttt{replicate} & .\\
+function arguments & \texttt{mapply} & . & \texttt{mapply} & .\\
+\bottomrule
+\end{tabular}
 
 \section{A general example with \texttt{plyr}}
+
 Let's take a simple example.
@@
 This could easily be done on one line, but I'm expanding it here to show the format a more complex function could take.
+
 <<echo = true>>=
 set.seed(1)
-d <- data.frame(year = rep(2000:2002, each = 3), count = round(runif(9, 0, 20)))
+d <- data.frame(year = rep(2000:2002, each = 3), 
+count = round(runif(9, 0, 20)))
 print(d)
@@
 \section{\texttt{transform} and \texttt{summarise}}
+
 It is often convenient to use these functions within \texttt{plyr}.
@@
 \texttt{summarise} creates a new (usually) condensed data frame.
+
 <<echo = true>>=
@@
 
-%\section{Common usage examples}
-%You can use \texttt{plyr} to plot data by throwing away the ouput with an underscore (\texttt{\_}).
-This is a bit cleaner than a for loop since you don't have to subset the data manually.
-Often you'd be better off using a grid graphics package for this such as \texttt{ggplot2} or \texttt{lattice}, but here's an example:
-%
-%<<echo = true, fig = true, height = 2, width = 5>>=
-%par(mfrow = c(1, 3), mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0))
-%d_ply(d, "year", summarise, hist(count, main = unique(year)))
-%mtext("count", side = 1, outer = TRUE, line = 1)
-%mtext("frequency", side = 2, outer = TRUE, line = 1)
-%@
+Bonus function: \texttt{mutate}. \texttt{mutate} works like \texttt{transform} but lets you build on columns you build.
+
+<<echo = true>>=
+ddply(d, "year", mutate, mu = mean(count), sigma = sd(count), 
+  cv = sigma/mu)
+@
+
+\section{Plotting with \texttt{plyr}}
+
+You can use \texttt{plyr} to plot data by throwing away the output with an underscore (\texttt{\_}). This is a bit cleaner than a for loop since you don't have to subset the data manually.
+
+<<echo = true, fig = true, height = 2, width = 5>>=
+par(mfrow = c(1, 3), mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0))
+d_ply(d, "year", transform, plot(count, main = unique(year), type = "o"))
+mtext("count", side = 1, outer = TRUE, line = 1)
+mtext("frequency", side = 2, outer = TRUE, line = 1)
+@
+
+\section{Nested chunking of the data}
+The basic syntax can be easily extended to break apart the data based on multiple columns:
+
+<<echo = true>>=
+baseball.dat <- subset(baseball, year > 2000) # data from the plyr package
+x <- ddply(baseball.dat, c("year", "team"), summarize,
+  homeruns = sum(hr))
+head(x)
+@
 
@@
 \subsection{Dealing with errors}
+
 You can use the \texttt{failwith} function to control how errors are dealt with.
+
 <<echo = true>>=
@@
 \subsection{Parallel processing}
+
 In conjunction with \texttt{doMC} (or \texttt{doSMP} on Windows) you can run your function separately on each core of your computer.
@@
 Set \texttt{.parallel = TRUE}.
+
 <<<echo = true>>=
@@
 
+\section{So, why would I \textit{not} want to use \texttt{plyr}?}
+
+\texttt{plyr} can be slow --- particularly if you are working with very large datasets that involve a lot of subsetting.
+Hadley is working on this and recent development versions of \texttt{plyr} run much faster.
+
+It's important to remember that typically the speed that you can write code and understand it later is the rate-limiting step.
+
+Three faster options:
+
+(1) Use a base R apply function:
+
+<<<echo = true>>=
+system.time(ddply(baseball, "id", summarize, length(year)))
+system.time(tapply(baseball$year, baseball$id, function(x) length(x)))
+@
+
+(2) Use an immutable data frame.
+An immutable data frame (\texttt{idata.frame}) returns pointers to the original object when subset instead of creating a copy of itself each time.
+This is often the rate-limiting step in an apply function.
+
+<<<echo = true>>=
+system.time(ddply(idata.frame(baseball), "id", summarize, length(year)))
+@
+
+(3) Use the \texttt{data.table} package:
+
+<<<echo = true>>=
+library(data.table)
+dt <- data.table(baseball, key="id")
+system.time(dt[, length(year), by=list(id)])
+@
+
+The next version of \texttt{plyr} lets you work with objects of class \texttt{data.table}:
+
+\url{https://gist.github.com/4007552}
+
 \end{document}

### 2018-07-04 static/courses/12-plyr/plyr_2012_examples.Rmd
Merge hugo site

# plyr examples

Let's load plyr and the data:

```{r}
library(plyr)
load("imdb.rda")
```

Let's start with a basic example. How does the mean movie length change through time?

```{r}
x0 <- ddply(imdb, "year", summarize, mean.length = mean(length, na.rm = TRUE))
plot(x0)
```

We can also look at the mean length by genre and year. We can work group by multiple columns using the c("column1", "column2") format.

```{r}
 x1 <- ddply(imdb, c("genre", "year"), summarize, mean.length = mean(length, na.rm = TRUE))

par(mfrow = c(5,5), mar = c(0,1.2,0,1), oma = c(3,3,1,1), cex = 0.6)
d_ply(x1, "genre", transform, {
  plot(year, mean.length, type = "l", xaxt = "n", xlim = c(1920, 2011))
  mtext(unique(genre), line = -1.5)
})
```

What if we wanted to detrend each time series with a linear model?

```{r}
x2 <- ddply(subset(x1, !is.na(mean.length)), "genre", transform, residual = residuals(lm(mean.length~year)))
head(x2)
```

How about running a regression of movie length and rating for each year and storing the models?

```{r}
x3 <- dlply(subset(imdb, year > 1950 & genre != "Short"), "year", function(x) 
  lm(length~imdb_rating, data = x))
x3[[1]]
```

Now what if we wanted to extract the slopes and their confidence intervals from those models and return them in a data frame? You might want to step into the function with browser() to help write the code.

```{r}
x4 <- ldply(x3, function(x) {
      est <- coef(x)[2]
      ci <- confint(x)
      ci.l <- ci[2, 1]
      ci.u <- ci[2, 2]
      data.frame(est, ci.l, ci.u)
})
library(ggplot2)
ggplot(subset(imdb, year >= 1990 & genre != "Short"), aes(imdb_rating, length)) + geom_point(alpha = 0.05) + facet_wrap(~year) + stat_smooth(method = "lm")
par(mfrow = c(1, 1))
with(x4, plot(year, est, ylim = c(min(ci.l), max(ci.u))))
d_ply(x4, "year", transform, segments(year, ci.l, year, ci.u))
```

### 2018-07-04 _knitr-posts/impute.R

@@ -1,58 +1,0 @@
-# Let's try imputing missing values
-
-
-# now we're going to impute with Stan:
-
-stan_model <-
-'data {
-  int<lower=0> N; // rows of data
-  vector[N] x;
-  int<lower=0> Nmissing; // number of missing values
-  int missing_id[Nmissing]; // IDs of missing values
-}
-parameters {
-  real<lower=-0.99, upper=0.99> phi;
-  real<lower=0> sigma;
-  real x_missing[Nmissing]; // imputed missing values
-}
-transformed parameters {
-  vector[N] x_impute; // y with some imputed
-  for (i in 1:N) {
-    x_impute[i] <- x[i];
-  }
-  for (i in 1:Nmissing) {
-    x_impute[missing_id[i]] <- x_missing[i];
-  }
-
-}
-model {
-  sigma ~ cauchy(0, 2.5);
-  phi ~ normal(0, 2);
-  for (i in 2:N) {
-    x_impute[i] ~ normal(x_impute[i-1] * phi, sigma);
-  }
-}
-'
-library(rstan)
-stan_missing_ar1 <- stan_model(model_code = stan_model)
-
-x <- arima.sim(n = 100L, list(ar = 0.8, ma = 0), sd = 0.05)
-missing_id <- c(50, 80)
-erased <- x[missing_id]
-x[missing_id] <- NA
-plot(x)
-x_missing <- x
-x_missing[missing_id] <- 9999 # Stan can't handle NAs; give it fake data
-m_missing <- sampling(
-  stan_missing_ar1,
-  data = list(N = length(x), x = x_missing,
-    Nmissing = length(missing_id), missing_id = missing_id),
-  iter = 500,
-  pars = c("sigma", "x_impute", "sigma"))
-
-imp <- apply(extract(m_missing)$x_impute[,missing_id], 2, median)
-l <- apply(extract(m_missing)$x_impute[,missing_id], 2, quantile, probs = 0.1)
-u <- apply(extract(m_missing)$x_impute[,missing_id], 2, quantile, probs = 0.9)
-points(missing_id, imp)
-segments(missing_id, l, missing_id, u)
-points(missing_id, erased, col = "red")

### 2018-07-04 clean.R

@@ -1,0 +1,44 @@
+files = list.files('content/',
+  '[.](md|markdown|Rmd)$',
+  full.names = TRUE,
+  recursive = TRUE)
+
+for (f in files) {
+  blogdown:::process_file(f, function(x) {
+    x <- gsub("published: false", "draft: true", x)
+    x <- gsub("published: true", "", x)
+    x
+  })
+}
+
+for (f in files) {
+  blogdown:::modify_yaml(
+    f,
+    slug = function(old, yaml) {
+      gsub("content\\/\\/post\\/([0-9]+-[0-9]+-[0-9]+)-([a-z0-9A-Z-]+)\\.(md|Rmd)$",
+        "\\2", f)
+    },
+    date = function(old, yaml) {
+      gsub("content\\/\\/post\\/([0-9]+-[0-9]+-[0-9]+)-([a-z0-9A-Z-]+)\\.(md|Rmd)$",
+        "\\1", f)
+    },
+    .keep_fields = c('title', 'subtitle', 'author', 'date', 'categories', 'tags', 'slug', 'draft'),
+    .keep_empty = FALSE
+  )
+}
+
+for (f in files) {
+  blogdown:::modify_yaml(
+    f,
+    aliases = function(old, yaml) {
+      .date <- gsub("content\\/\\/post\\/([0-9]+-[0-9]+-[0-9]+)-([a-z0-9A-Z-]+)\\.(md|Rmd)$",
+        "\\1", f)
+      .slug <- gsub("content\\/\\/post\\/([0-9]+-[0-9]+-[0-9]+)-([a-z0-9A-Z-]+)\\.(md|Rmd)$",
+        "\\2", f)
+      paste0(gsub("-", "/", .date), "/", .slug, ".html")
+    },
+    .keep_fields = c('title', 'subtitle', 'author', 'date', 'categories',
+      'tags', 'slug', 'aliases', 'draft'),
+    .keep_empty = FALSE
+  )
+}

### 2018-07-04 refs.R

@@ -1,98 +1,0 @@
-# TODO J.E.M. Flemming should be J.E. Mills Flemming
-# TODO J.E. Mills Flemming needs to be fixed throughout refs.bib too
-# TODO Earth2Oceans Derby needs to be manually fixed
-
-# Need to fix all authors contributed equally and asterisks
-
-library("bibtex")
-d <- read.bib("refs.bib")
-
-ref <- list()
-
-for(i in seq_along(d)) {
-  name <- list()
-  x <- d[[i]]
-  # authors:
-  for(j in seq_along(x$author)) {
-    init <- paste0(paste(substr(x$author$given[[j]], 1, 1), collapse = "."), ".")
-    family <- x$author$family[[j]]
-    if(j == 1) {
-      name[[j]] <- paste0(family, ", ", init)
-      if(name[[j]] == "Anderson, S.C.") name[[j]] <- "**Anderson, S.C.**"
-    } else {
-      name[[j]] <- paste(init, family)
-      if(name[[j]] == "S.C. Anderson") name[[j]] <- "**S.C. Anderson**"
-    }
-  }
-  name0 <- paste0(paste0(name, collase = ""), collapse = ", ")
-
-  vol_num_p <- ""
-  if(!is.null(x$number) & !is.null(x$volume) & !is.null(x$pages)) {
-    vol_num_p <- paste0(x$volume, "(", x$number, "): ", x$pages)
-  }
-  # no number:
-  if(is.null(x$number) & !is.null(x$volume) & !is.null(x$pages)) {
-    vol_num_p <- paste0(x$volume, ": ", x$pages)
-  }
-  # no pages:
-  if(is.null(x$number) & !is.null(x$volume) & is.null(x$pages)) {
-    vol_num_p <- paste0(x$volume, "(", x$number, ")")
-  }
-  # no number or pages:
-  if(is.null(x$number) & !is.null(x$volume) & is.null(x$pages)) {
-    vol_num_p <- x$volume
-  }
-  # no vol, number, or pages:
-  if(is.null(x$number) & is.null(x$volume) & is.null(x$pages)) {
-    vol_num_p <- ""
-  }
-
-  if(vol_num_p != "" & !vol_num_p %in% c("In review", "In revision", "Submitted")) {
-    main <- paste0(paste0(name0, ". ", x$title, ". ", x$journal, ". ", vol_num_p), ".")
-  } else {
[…]

### 2018-07-04 static/ggplot2-FISH554/ggplot2-notes.Rnw
Add back ggplot files

% Sean Anderson, 2012-14, sean "at" seananderson.ca
\documentclass[11pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{graphicx}
\textheight 21.0cm

\usepackage{url}
\usepackage{fancyvrb}
\fvset{fontsize=\normalsize}
\usepackage{booktabs}

\usepackage{titling}
\newcommand{\subtitle}[1]{%
  \posttitle{%
    \par\end{center}
    \begin{center}\large#1\end{center}}%
}

\let\proglang=\textsf
\let\pkg=\textbf

\newcommand{\ggtwo}{\pkg{ggplot2}}
\newcommand{\ggplot}{\texttt{ggplot()}}
\newcommand{\qplot}{\texttt{qplot()}}

\title{An introduction to \ggtwo}
\subtitle{FISH 544: Beautiful Graphics in \proglang{R}}
\author{Sean C. Anderson\\ \texttt{sean@seananderson.ca}}

\setlength\parskip{0.1in}
\setlength\parindent{0in}

\begin{document}

<<set-knitr-options, echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.align='center', fig.pos="htpb", cache=TRUE, echo=TRUE,
message=FALSE, autodep=TRUE, fig.path='figure/', fig.width=5, par=TRUE)
opts_chunk$set(warning=FALSE, message=FALSE, tidy=FALSE, refresh=TRUE)
opts_chunk$set(dev = 'pdf')
thm <- knit_theme$get("solarized-light.css")
knit_theme$set(thm)
@

\maketitle

\section{\ggtwo\ for rapid data exploration}

\ggtwo\ is an \proglang{R} package by Hadley Wickham that implements Wilkinson's
Grammar of Graphics.\footnote{Wilkinson, L. (2005). \textit{The Grammar of
Graphics}. Springer, 2\textsuperscript{2nd} edition.} The emphasis of \ggtwo\ is
on rapid exploration of data, and especially high-dimensional data. Think of
base graphic functions as drawing with data\footnote{Examples of base graphic
functions are \texttt{plot()}, \texttt{points()}, and \texttt{lines()}.}. With
base graphics, you have complete control over every pixel in a plot but it can
take a lot of time and code to produce a plot.

Although \ggtwo\ can be fully customized, it reaches a point of diminishing
returns. I tend to use \ggtwo\ and base graphics for what they excel at: \ggtwo\
for rapid data exploration and base graphics for polished and fully-customized
plots for publication (Figure~\ref{fig:gg-time}).

\begin{figure}[htb]
  \centering
    \includegraphics[height=2.15in]{figure/gg-dimensions.pdf}
    \includegraphics[height=2.15in]{figure/gg-vs-base-pub.pdf}
  \caption{Creation time vs.\ data dimensions and customization level for base
  graphics (blue) and \ggtwo\ (red). (Left panel) It's remarkably easy to plot
  high-dimensional data in \ggtwo\ with, for example, colours, sizes, shapes,
  and panels. (Right panel) \ggtwo\ excels at rapid visual exploration of
  data, but has some limitations in how it can be customized. Base graphics
  are fully customizable but can take longer to set up. I try and exploit the
  grey shaded areas: I use \ggtwo\ for data exploration and once I've decided
  on a small number of key plots, I'll use base graphics to make
  fully-customized plots if needed.}
  \label{fig:gg-time}
\end{figure}

Good graphical displays of data require rapid iteration and lots of exploration.
If it takes you hours to code a plot in base graphics, you're unlikely to throw
it out and explore other ways of visualizing the data, and you're unlikely to
explore all the dimensions of the data.

\section{\qplot\ vs.\ \ggplot}

There are two main plotting functions in \ggtwo: \qplot\ and \ggplot. \qplot\ is
short for ``quick plot'' and is made to mimic the format of \texttt{plot()} from
base \proglang{R}. \qplot\ requires less syntax for many common tasks, but has
limitations --- it's essentially a wrapper for \ggplot. The \ggplot\ function
itself isn't complicated and will work in all cases. I prefer to work with just
the \ggplot\ syntax and will focus on it here. I find it easier to master one
function that can do everything.

\section{Basics of the grammar}
Let's look at some illustrative \ggtwo\ code:

<<echo=TRUE, fig.height=3.5, fig.width=7, out.width="4.5in">>=
library(ggplot2)
theme_set(theme_bw()) # use the black and white theme throughout
# fake data:
d <- data.frame(x = c(1:8, 1:8), y = runif(16),
  group1 = rep(gl(2, 4, labels = c("a", "b")), 2),
  group2 = gl(2, 8))
head(d)
ggplot(data = d) + geom_point(aes(x, y, colour = group1)) +
  facet_grid(~group2)
@

The basic format in this example is:

\begin{enumerate}
  \item \texttt{ggplot()}: start a \ggplot\ object and specify the data
  \item \texttt{geom\_point()}: we want a scatter plot; this is called a
    ``geom''
  \item \texttt{aes()}: specifies the ``aesthetic'' elements; a legend is
    automatically created
  \item \texttt{facet\_grid()}: specifies the ``faceting'' or panel layout
\end{enumerate}

There are also statistics, scales, and annotation options, among others. At a
minimum, you must specify the data, some aesthetics, and a geom. I will
elaborate on these below. Yes, \ggtwo\ combines elements with \texttt{+}
symbols! This may seem non-standard, although it has the advantage of allowing
\ggtwo\ plots to be proper \proglang{R} objects, which can modified, inspected,
and re-used (I provide some examples at the end).

\clearpage
\section{Geoms}

\texttt{geom} refers to a geometric object. It determines the ``shape'' of the
plot elements. Some common geoms:

\begin{tabular}{llll}
  \toprule
  \texttt{geom} & description \\
 \midrule
 \texttt{geom\_point()} & Points\\
 \texttt{geom\_line()} & Lines\\
 \texttt{geom\_ribbon()} & Ribbons, y range with continuous x values\\
 \texttt{geom\_polygon()} & Polygon, a filled path\\
 \texttt{geom\_pointrange()} & Vertical line with a point in the middle\\
 \texttt{geom\_linerange()} & An interval represented by a vertical line\\
 \texttt{geom\_path()} & Connect observations in original order\\
 \texttt{geom\_histogram()} & Histograms \\
 \texttt{geom\_text()} & Text annotations\\
 \texttt{geom\_violin()} & Violin plot (another name for a beanplot)\\
 \texttt{geom\_map()} & Polygons from a map\\
\bottomrule
\end{tabular}

\section{Aesthetics}

Aesthetics refer to the attributes of the data you want to display. They map the
data to an attribute (such as the size or shape of a symbol) and generate an
appropriate legend. Aesthetics are specified with the \texttt{aes()} function.

As an example, the aesthetics available for \texttt{geom\_point()} are:
\texttt{x}, \texttt{y}, \texttt{alpha}, \texttt{colour}, \texttt{fill},
\texttt{shape}, and \texttt{size}.\footnote{Note that \ggtwo\ tries to
accommodate the user who's never ``suffered'' through base graphics before by
using intuitive arguments like \texttt{colour}, \texttt{size}, and
\texttt{linetype}, but \ggtwo\ will also accept arguments such as \texttt{col},
\texttt{cex}, and \texttt{lty}.} Read the help files to see the aesthetic
options for the geom you're using. They're generally self explanatory.
Aesthetics can be specified within the data function or within a geom. If
they're specified within the data function then they apply to all geoms you
specify.

Note the important difference between specifying characteristics like colour and
shape inside or outside the \texttt{aes()} function: those inside the
\texttt{aes()} function are assigned the colour or shape automatically based on
the data. If characteristics like colour or shape are defined outside the
\texttt{aes()} function, then the characteristic is not mapped to data. Here's
an example:

\clearpage

<<echo=TRUE, fig.height=3, fig.width=5, out.width="3.5in">>=
ggplot(mpg, aes(cty, hwy)) + geom_point(aes(colour = class))
@

<<echo=TRUE, fig.height=3, fig.width=4, out.width="2.75in">>=
ggplot(mpg, aes(cty, hwy)) + geom_point(colour = "red")
@

\section{Small multiples}

In \ggtwo\ parlance, small multiples are referred to as ``facets''. There are
two kinds: \texttt{facet\_wrap()} and \texttt{facet\_grid()}.
\texttt{facet\_wrap()} plots the panels in the order of the factor levels. When
it gets to the end of a column it wraps to the next column. You can specify the
number of columns and rows with \texttt{nrow} and \texttt{ncol}.
\texttt{facet\_grid()} lays out the panels in a grid with an explicit x and y
position. By default all x and y axes will be shared among panels. However, you
could, for example, allow the y axes to vary with \texttt{facet\_wrap(scales =
"free\_y")} or allow all axes to vary with \texttt{facet\_wrap(scales =
"free")}.

\clearpage

<<echo=TRUE, fig.height=4.5, fig.width=5, out.width="3.5in">>=
ggplot(mpg, aes(cty, hwy)) + geom_point() + facet_wrap(~class)
@

<<echo=TRUE, fig.height=3, fig.width=8, out.width="5in">>=
 ggplot(mpg, aes(cty, hwy)) + geom_point() + facet_grid(year~class)
@

\section{Themes}

A useful theme built into \ggtwo\ is \texttt{theme\_bw()}. You'll notice that I
set it as the default in this document back when I first loaded \pkg{ggplot2}.
You can specify it for a specific plot like this:

\clearpage

<<echo=TRUE, fig.height=3.5, fig.width=4, out.width="2.75in">>=
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_bw()
@

Alternatively, the following plot shows the default theme. The grey is designed
to match the typical darkness of a page filled with text to keep the plot from
drawing too much attention.

<<echo=TRUE, fig.height=3.5, fig.width=4, out.width="2.75in">>=
ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_grey()
@

A powerful aspect of \ggtwo\ is that you can write your own themes. See the
\pkg{ggthemes} package for some examples.

An Edward Tufte-like theme:
<<echo=TRUE, fig.height=2.5, fig.width=3.0, out.width="3.00in">>=
library("ggthemes")
ggplot(mtcars, aes(wt, mpg)) + geom_point() + geom_rangeframe() +
 theme_tufte()
@

Just what you wanted:
<<echo=TRUE, fig.height=2.5, fig.width=5, out.width="3.70in">>=
ggplot(dsamp, aes(carat, price, colour = cut)) + geom_point() +
  theme_excel() + scale_colour_excel()
@

Here's an example custom theme:
<<custom-theme, fig.height=3.5, fig.height=4, out.width="2.75in">>=
theme_example <- function (base_size = 12, base_family = "") {
  theme_gray(base_size = base_size, base_family = base_family) %+replace%
    theme(axis.text = element_text(colour = "grey50"),
          axis.title.x = element_text(colour = "grey50"),
          axis.title.y = element_text(colour = "grey50", angle = 90),
          panel.background = element_rect(fill = "white"),
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major = element_blank(),
          plot.background = element_blank()
  )
}
ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_example()
@

You can customize just about every aspect of a \ggtwo\ plot. We won't get into
that today, but see the additional help resources at the end of this document.

\section{\ggtwo's dirty little secret}
\begin{enumerate}
  \item \ggtwo\ is easy to learn, but\ldots
  \item You need to be a data-manipulation ninja to use it effectively.
\end{enumerate}

With base graphics, you can work with almost any data structure you'd like,
providing you can write the code to work with it. For example, you data could be
in a list or a data frame. Or maybe it's scattered across multiple objects.
\ggtwo, on the other hand, requires you to think carefully about the data
structure and then write one (possibly long) line of code. So, with base
plotting you'll tend to spend lots of time writing plotting code. With \ggtwo\
you'll tend to spend time thinking beforehand, and then quickly write the code.

\ggplot\ works with ``long'' format data or ``tidy data'' with each aesthetic or
facet variable in its own column\footnote{See Hadley Wickham's pre-print on tidy
data here: \url{http://vita.had.co.nz/papers/tidy-data.pdf}}.  So, for example,
if we wanted a panel or colour for each level of a variable called
\texttt{fishstock} then we'd need a column named \texttt{fishstock} with all the
different values of \texttt{fishstock} in that column.

With the \pkg{reshape2} and \pkg{plyr} packages, combined with \texttt{merge()}
(or \texttt{plyr::join()}), you can get almost any dataset into shape for
\ggtwo\ in a few lines. Sometimes this will take some serious thought along with
pen and paper.

\section{Random tips}

\begin{description}
  \item[Jittering and statistics] \hfill \\
    \ggtwo\ has lots of built-in niceties like automatic position jittering
    and the addition of basic statistical models to your plots. Have a look
    through the help resources listed at the end of this document.
  \item[Axis labels] \hfill \\
    \texttt{xlab("Your x-axis label")}
  \item[Suppressing the legend] \hfill \\
    \texttt{theme(legend.position = "none")}
  \item[Exploiting the object-oriented nature of \ggtwo] \hfill \\
    Save the basic plot to an object and then experiment with different
    aesthetics, geoms, and theme adjustments.

<<echo=TRUE, fig.keep='none'>>=
# Build the basic object:
p <- ggplot(d) + geom_point(aes(x, y, colour = group1))
p + facet_grid(~group1) # try one way
p + facet_grid(~group2) # try another way
@

  \item[Horizontal error bars]\hfill \\
    Say you want to make a coefficient plot with the coefficients down the
    y-axis. You can either build the plot piece by piece with points and
    segments, or you can use \texttt{geom\_pointrange()} and then rotate the axes
    with \texttt{+ coord\_flip()}.
  \item[Axis limits and zooming] \hfill \\
    \ggtwo\ has two ways of adjusting the axis limits: \texttt{+ xlim(2, 5)}
    will ``cut'' the data at \texttt{2} and \texttt{5} and plot the data.
    \texttt{coord\_cartesian(xlim = c(2, 5))} will zoom in on the plot while
    retaining the original data. This will, for example, affect colour scales.
  \item[Displaying and saving \ggtwo\ plots] \hfill \\
    If \ggtwo\ plots are generated in a function, they will be need to be
    wrapped in \texttt{print()} to display. There is a convenience function
    \texttt{ggsave("filename.pdf")}, which will save the last plot to a PDF
    and guess at reasonable dimensions. You can, of course specify which plot
    to save and the dimensions of the PDF. You can also use all the same
    methods of saving \ggtwo\ plots that you can use for base graphics.
  \item[Combining multiple \ggtwo\ panels] \hfill \\
    \ggtwo\ makes it easy to create multipanel figures with
    \texttt{facet\_grid()} and \texttt{facet\_wrap()}. But, sometimes you need
    to combine different kinds of plots in one figure or plots that require
    different data. One easy way to do this is with the \texttt{grid.arrange()}
    function from the \pkg{gridExtra} package. For example:

<<grid.arrange, fig.width=4.2, fig.height=5.2, out.width="2.75in">>=
df <- data.frame(x = 1:10, y = rnorm(100))
p1 <- ggplot(df, aes(x, y)) + geom_point()
p2 <- ggplot(df, aes(y)) + geom_histogram()
gridExtra::grid.arrange(p1, p2)
@

\item[\pkg{ggvis}]\hfill \\
  \pkg{ggvis} is a partial successor to \ggtwo\ and is under rapid development:
  \url{https://github.com/rstudio/ggvis}. It focuses on implementing a similar
  grammar of graphics for web-based and possibly interactive output. Take a look at the vignettes ---
  start with \texttt{vignette("ggvis-basics")} --- to see some examples.
  You'll need to install the \pkg{devtools} \proglang{R} package first.

\end{description}

\clearpage

\section{Help}

The best (and nearly essential) help source is the website, which is largely
based off the package help files. However, the website also shows the example
plots and is easier to navigate:\\ \url{http://docs.ggplot2.org/}\\ Don't be
afraid to keep it open whenever you're using \ggtwo.

There's also an active \ggtwo\ discussion group:\\
\url{http://groups.google.com/group/ggplot2}

\ggtwo\ is heavily featured on stackoverflow:\\
\url{http://stackoverflow.com/questions/tagged/ggplot2}

Hadley wrote a book on \ggtwo: \url{http://ggplot2.org/book/}. It's quite
thorough but doesn't feature some of the newer additions to the package.

\end{document}

### 2018-07-04 static/ggplot2-FISH554/index.Rmd
Add back ggplot files

---
author: Sean C. Anderson
date: February 2015
output:
  knitrBootstrap::bootstrap_document:
    title: "An introduction to ggplot2"
    theme: Default
    highlight: 'Tomorrow'
---

```{r knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, message=FALSE, tidy=FALSE, fig.height=4, fig.width=6, 
  fig.align = "center")
```

# An introduction to ggplot2

*February 2015, by Sean C. Anderson*

This page will help you teach yourself how to rapidly explore data with the ggplot2 R package. It was written as a self-study component for the FISH554 class at the University of Washington.

Work your way through this page. When you want to check your work, or if you get  stuck on an exercise, just click on the button `R source` and you'll see the R source code that created a figure.

# Learning objectives

By the end you should be able to:

- Understand the basic grammar of ggplot2 (data, geoms, aesthetics, facets).
- Make quick exploratory plots of your multidimensional data.
- Have an idea about how to start customizing ggplot2 figures.
- Know how to find help on ggplot2 when you run into problems.

# ggplot2 for rapid data exploration

ggplot2 is an R package by Hadley Wickham and Winston Chang that implements Wilkinson's Grammar of Graphics. The emphasis of ggplot2 is on rapid exploration of data, and especially high-dimensional data. Think of base graphic functions as drawing with data (examples of base graphic functions are `plot()`, `points()`, and `lines()`. With base graphics, you have complete control over every pixel in a plot but it can take a lot of time and code to produce a plot.

Although ggplot2 can be fully customized, it reaches a point of diminishing
returns. I tend to use ggplot2 and base graphics for what they excel at: ggplot2 
for rapid data exploration and base graphics for polished and fully-customized
plots for publication.

The following figure shows figure creation time vs. data dimensions and customization level for base graphics (blue) and ggplot2 (red). Left panel: It's remarkably easy to plot high-dimensional data in ggplot2 with, for example, colours, sizes, shapes, and panels. Right panel: ggplot2 excels at rapid visual exploration of data, but has some limitations in how it can be customized. Base graphics are fully customizable but can take longer to set up. I try and exploit the grey shaded areas: I use ggplot2 for data exploration and once I've decided on a small number of key plots, I'll use base graphics to make fully-customized plots if needed. Some people get really good at customizing ggplot and stick with it for all their plots, but since you've already learned the ways of base graphics in FISH554 I think you'll benefit from the strategy in this figure:

![](figure/gg-vs-base.png)

Good graphical displays of data require rapid iteration and lots of exploration. If it takes you hours to code a plot in base graphics, you're unlikely to throw it out and explore other ways of visualizing the data, and you're unlikely to explore all the dimensions of the data.

# Basics of the grammar

Let's look at some illustrative ggplot2 code:

```{r basics1}
library("ggplot2")
theme_set(theme_bw()) # use the black and white theme throughout
# fake data:
d <- data.frame(x = c(1:8, 1:8), y = runif(16),
  group1 = rep(gl(2, 4, labels = c("a", "b")), 2),
  group2 = gl(2, 8))
head(d)
ggplot(data = d) + geom_point(aes(x, y, colour = group1)) +
  facet_grid(~group2)
```

The basic format in this example is:

1.  `ggplot()`: start an object and specify the data

2.  `geom_point()`: we want a scatter plot; this is called a “geom”

3.  `aes()`: specifies the “aesthetic” elements; a legend is automatically created

4.  `facet_grid()`: specifies the “faceting” or panel layout

There are also statistics, scales, and annotation options, among others. At a minimum, you must specify the data, some aesthetics, and a geom. I will elaborate on these below. Yes, ggplot2 combines elements with `+` symbols! This may seem non-standard, although it has the advantage of allowing ggplot2 plots to be proper R objects, which can modified, inspected, and re-used (I provide some examples at the end).

There are two main plotting functions in ggplot2: `qplot` and `ggplot`. `qplot` is short for "quick plot" and is made to mimic the format of `plot` from base R. `qplot` requires less syntax for many common tasks, but has limitations — it's essentially a wrapper for `ggplot`. The `ggplot` function itself isn't complicated and will work in all cases. I prefer to work with just the `ggplot` syntax and will focus on it here. I find it easier to master one function that can do everything.

# Loading the data

We're going to work with morphological data from Galapagos finches, which is available from BIRDD: Beagle Investigation Return with Darwinian Data at <http://bioquest.org/birdd/morph.php>. It is originally from Sato et al. 2000 Mol. Biol. Evol. <http://mbe.oxfordjournals.org/content/18/3/299.full>.

I've saved a `.csv` version of this dataset for you. **Download the data `Morph_for_Sato.csv` [here](https://raw.githubusercontent.com/seananderson/ggplot2-FISH554/master/Morph_for_Sato.csv) and put it in your R working directory.**

Before we get started, we're going to clean the data up a bit. I've removed some columns and made the column names lower case. I've also removed all but one island. You can do that with this code:

```{r}
morph <- read.csv("Morph_for_Sato.csv")
names(morph) <- tolower(names(morph)) # make columns names lowercase
morph <- subset(morph, islandid == "Flor_Chrl") # take only one island
morph <- morph[,c("taxonorig", "sex", "wingl", "beakh", "ubeakl")] # only keep these columns
names(morph)[1] <- "taxon"
morph <- data.frame(na.omit(morph)) # remove all rows with any NAs to make this simple
morph$taxon <- factor(morph$taxon) # remove extra remaining factor levels
morph$sex <- factor(morph$sex) # remove extra remaining factor levels
row.names(morph) <- NULL # tidy up the row names
```

Take a look at the data. There are columns for taxon, sex, wing length, beak height, and upper beak length:

```{r look-data, eval=FALSE}
head(morph)
str(morph)
```

# Geoms

`geom` refers to a geometric object. It determines the “shape” of the plot elements. Some common geoms:

`geom`              Description
------              -----------
`geom_point()`      Points
`geom_line()`       Lines
`geom_ribbon()`     Ribbons, y range with continuous x values
`geom_polygon()`    Polygon, a filled path
`geom_pointrange()` Vertical line with a point in the middle
`geom_linerange()`  An interval represented by a vertical line
`geom_path()`       Connect observations in original order
`geom_histogram()`  Histograms
`geom_text()`       Text annotations
`geom_violin()`     Violin plot (another name for a beanplot)
`geom_map()`        Polygons on a map

Open the ggplot2 web documentation <http://docs.ggplot2.org/> and keep it open to refer back to it throughout these exercises.

First, let's experiment with some geoms using the `morph` dataset that you downloaded and cleaned up above. I'll start by setting up a basic scatterplot of beak height vs. wing length:

```{r geom1}
library("ggplot2")
ggplot(morph, aes(wingl, beakh)) + geom_point(alpha = 0.4)
```

Because there's lots of overplotting, I've set the alpha (opacity) value to 40%. Alternatively, we could have added some jittering. We'll do that below.

### Your turn

Experiment with ggplot2 geoms. Try applying at least 3 different geoms to the `morph` dataset. For example, try showing the distribution of wing length with `geom_histogram()` and `geom_density()`. You could also try showing the distribution of wing length for male and female birds by using `geom_violin()`. After you've experimented for a while, click on `R Source` to see the code behind the plots below.

```{r geom-examples, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl)) + geom_histogram(binwidth=1)
ggplot(morph, aes(wingl)) + geom_density()
ggplot(morph, aes(sex, wingl)) + geom_violin()
ggplot(morph, aes(taxon, wingl)) + geom_violin() + coord_flip() # coord_flip() rotates 90 degrees
ggplot(morph, aes(taxon, wingl)) + geom_boxplot() + coord_flip()
```

# Aesthetics

Aesthetics refer to the attributes of the data you want to display. They map the data to an attribute (such as the size or shape of a symbol) and generate an appropriate legend. Aesthetics are specified with the `aes()` function.

As an example, the aesthetics available for `geom_point()` are: `x`, `y`, `alpha`, `colour`, `fill`, `shape`, and `size`. (Note that ggplot tries to accommodate the user who’s never “suffered” through base graphics before by using intuitive arguments like `colour`, `size`, and `linetype`, but ggplot will also accept arguments such as `col`, `cex`, and `lty`.) Read the help files to see the aesthetic options for the geom you’re using. They’re generally self explanatory. Aesthetics can be specified within the data function or within a geom. If they’re specified within the data function then they apply to all geoms you specify.

Note the important difference between specifying characteristics like colour and shape inside or outside the `aes()` function: those inside the `aes()` function are assigned the colour or shape automatically based on the data. If characteristics like colour or shape are defined outside the `aes()` function, then the characteristic is not mapped to data. Here’s an example:

```{r echo=TRUE, bootstrap.show.code=TRUE}
ggplot(mpg, aes(cty, hwy)) + geom_point(aes(colour = class))
```

```{r echo=TRUE, bootstrap.show.code=TRUE}
ggplot(mpg, aes(cty, hwy)) + geom_point(colour = "red")
```

Let's play with mapping some of our data to aesthetics. I'll start with one example. I'm going to map the male/female value to a colour in our scatterplot of wing length and beak height. This time I'll use jittering instead of transparency to deal with overplotting:

```{r aes-example, bootstrap.show.code=TRUE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(aes(colour = sex),
    position = position_jitter(width = 0.3, height = 0))
```

### Your turn

Explore the `morph` dataset yourself by applying some aesthetics. You can see all the available aesthetics for a give geom by looking at the documentation. Either see [the website](http://docs.ggplot2.org/current/) or, for example, `?geom_point`

Some suggestions:

- try the same scatterplot but show upper beak length (`ubeakl`) with size
- try the same scatterplot but show the taxon with colour
- try the same scatterplot but show the upper beak length with colour (note how ggplot treats `ubeakl` differently than `taxon` when it picks a colour scale)
- try the same scatterplot but show the sex with a different shape
- combine all these: colour for taxon, shape for sex, and size for upper beak length

This last version is a bit silly, but it illustrates how quickly you can explore multiple dimensions with ggplot2.

```{r aes1, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(aes(size = ubeakl), alpha = 0.4)
```

```{r aes2, fig.width=9, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(aes(colour = taxon),
    position = position_jitter(width = 0.3, height = 0))
```

```{r aes3, fig.width=9, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(aes(colour = ubeakl),
    position = position_jitter(width = 0.3, height = 0))
```

```{r aes4, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(aes(shape = sex),
    position = position_jitter(width = 0.3, height = 0))

```{r aes5, fig.width=9, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(aes(shape = sex, size = ubeakl, colour = taxon),
    alpha = 0.4)
```

# Facets (small multiples)

In ggplot2 parlance, small multiples are referred to as "facets". There are two kinds: `facet_wrap()` and `facet_grid()`. `facet_wrap()` plots the panels in the order of the factor levels. When it gets to the end of a column it wraps to the next column. You can specify the number of columns and rows with `nrow` and `ncol`. `facet_grid()` lays out the panels in a grid with an explicit x and y position. By default all x and y axes will be shared among panels. However, you could, for example, allow the y axes to vary with `facet_wrap(scales = "free_y")` or allow all axes to vary with `facet_wrap(scales = free)`.

To specify the data frame columns that are mapped to the rows and columns of facets, separate them with a tilde. For example: `+ facet_grid(row_name~column_name)`. Usually you'll only give a row or column to `facet_wrap()`, but try and see what happens if you give it both. See the help `?facet_wrap`.

### Your turn

Try a scatterplot of beak height against wing length with a different panel for each taxon. Use `facet_wrap()`:

```{r facet1, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(alpha = 0.4) + facet_wrap(~taxon)
```

In some cases, it's useful to let the x or y axes have different scales for each panel. Try giving each panel a different axis here using `scales = "free"` in your call to `facet_wrap()`:

```{r facet2, fig.width=10, fig.height=8, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(alpha = 0.4) +
  facet_wrap(~taxon, scales = "free")
```

Now try using `facet_grid` to explore the same scatterplot for each combination of sex and taxa. (Remove the `scales = "free"` code for simplicity.)

```{r facet3, fig.width=10, fig.height=7, bootstrap.show.code=FALSE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(alpha = 0.4) + facet_grid(sex~taxon)
```

As another example, let's look at the distribution of wing length by sex with different panels for each taxa. Use a boxplot or violin plot to show the distributions.

```{r facet4, fig.width=9, fig.height=9, bootstrap.show.code=FALSE}
ggplot(morph, aes(sex, wingl)) + geom_violin() +
  facet_wrap(~taxon)
```

# Customizing ggplot2

So far we've used the default settings. Let's try applying a theme and adjust some of the annotation of our plots.

A useful theme built into ggplot2 is `theme_bw()`. You’ll notice that I set it as the default in this document back when I first loaded ggplot2. You can specify it for a specific plot like this:

```{r echo=TRUE, bootstrap.show.code=TRUE}
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_bw()
```

Alternatively, the following plot shows the default theme. The grey is designed to match the typical darkness of a page filled with text to keep the plot from drawing too much attention.

```{r echo=TRUE, bootstrap.show.code=TRUE}
ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_grey()
```

A powerful aspect of ggplot2 is that you can write your own themes. See the ggthemes package for some examples.

An Edward Tufte-like theme:

```{r echo=TRUE, bootstrap.show.code=TRUE}
library("ggthemes")
ggplot(mtcars, aes(wt, mpg)) + geom_point() + geom_rangeframe() +
 theme_tufte()
```

Just what you wanted:

```{r echo=TRUE, bootstrap.show.code=TRUE}
ggplot(dsamp, aes(carat, price, colour = cut)) + geom_point() +
  theme_excel() + scale_colour_excel()
```

Here's an example custom theme:

```{r custom-theme, bootstrap.show.code=TRUE}
theme_example <- function (base_size = 12, base_family = "") {
  theme_gray(base_size = base_size, base_family = base_family) %+replace%
    theme(axis.text = element_text(colour = "grey50"),
          axis.title.x = element_text(colour = "grey50"),
          axis.title.y = element_text(colour = "grey50", angle = 90),
          panel.background = element_rect(fill = "white"),
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major = element_blank(),
          plot.background = element_blank()
  )
}
ggplot(mtcars, aes(wt, mpg)) + geom_point() + theme_example()
```

You can customize just about every aspect of a \ggtwo\ plot. We won't get into that today, but see the additional help resources at the end of this document.

### Your turn

Let's start by applying the black and white theme, `theme_bw()`. Try adding that on to the end of a plot showing the distribution of wing length by sex using `geom_violin()`. Save your plot to an object named `p` and `print()` it to show it. We'll re-use this plot in the next step.

```{r customizing1, bootstrap.show.code=FALSE}
p <- ggplot(morph, aes(sex, wingl)) + geom_violin() + theme_bw()
print(p)
```

Let's go one step further and remove all grid lines. See the included notes and the help for `?theme`. Hint: setting an argument to `element_blank()` will remove it. Add these elements to the object `p` and print `p` again. Hint: see the note "Exploiting the object-oriented nature of ggplot2" in the "Random tips" section of the notes.

```{r customizing2, bootstrap.show.code=FALSE}
p <- p + theme(panel.grid.major = element_blank(),
  panel.grid.minor = element_blank())
print(p)
```

And now let's set the x and y axis labels ourselves. Name them something more appropriate.

```{r customizing3, bootstrap.show.code=FALSE}
p <- p + xlab("Sex") + ylab("Wing length")
print(p)
```

Use the function `ggsave()` to save your plot to a PDF file.

```{r customizing4, bootstrap.show.code=FALSE}
ggsave(p, file = "wingl-violin.pdf", width = 6, height = 6)
```

# More examples

## Dot and line plots

ggplot2 can be useful for quickly making dot and line plots. For example, this is useful for coefficient plots. To illustrate how to make this style of plot, let's make a plot that shows a dot for the median and line segment for the quantiles. First, we'll calculate these values using the [dplyr package](http://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html) and [data pipes](http://cran.r-project.org/web/packages/magrittr/vignettes/magrittr.html). Run the following code:

```{r manipulate, bootstrap.show.code=TRUE}
# install.packages("dplyr")
library("dplyr")
morph_quant <- 
  morph %>%
  group_by(taxon) %>%
  summarise(
    l = quantile(wingl, 0.25)[[1]],
    m = median(wingl),
    u = quantile(wingl, 0.75)[[1]]) %>%
  # re-order factor levels by median for plotting:
  mutate(taxon = reorder(taxon, m, function(x) x)) 
```

### Your turn

Now make the plot. The final plot should have the taxa listed down the y-axis and the wing length value on the x axis. Use the geom `geom_pointrange()`. Because this geom only works for vertical line segments, you'll need to rotate the whole plot by adding `+ coord_flip()`. So, `ymax` and `ymin` refer to the maximum and minimum line segment values and `x` to the taxa, even though they will appear rotated in the end.

```{r pointrange-example, bootstrap.show.code=FALSE}
ggplot(morph_quant, aes(x = taxon, y = m, ymin = l, ymax = u)) +
  geom_pointrange() + coord_flip() +
  ylab("Wing length") + xlab("")
```

## Adding model fits to the data

ggplot2 can add model fits to the data to help visualize patterns. For example, it can quickly add linear regression lines, GLMs, GAMs, and loess curves. Let's add loess curves to scatter plots of beak height and wing length with a panel (facets) for male and female. See `?stat_smooth`.

```{r loess-example, bootstrap.show.code=TRUE}
ggplot(morph, aes(wingl, beakh)) +
  geom_point(alpha = 0.4) + facet_wrap(~sex) +
  stat_smooth(method = "loess")
```

You can add any model fit by specifying the function and the formula (see the examples in `?stat_smooth`). Alternatively, you can fit your own model, predict from it, and add the data with `geom_ribbon()`.

# Some final notes

## ggplot2's dirty little secret

1.  ggplot2 is easy to learn, but...

2.  You need to be a data-manipulation ninja to use it effectively.

With base graphics, you can work with almost any data structure you'd like, providing you can write the code to work with it. For example, you data could be in a list or a data frame. Or maybe it's scattered across multiple objects. ggplot2, on the other hand, requires you to think carefully about the data structure and then write one (possibly long) line of code. So, with base plotting you'll tend to spend lots of time writing plotting code. With ggplot2 you'll tend to spend time thinking beforehand, and then quickly write the code.

ggplot2 works with “long” format data or “tidy data” with each aesthetic or facet variable in its own column (see Hadley Wickham’s [pre-print on tidy data](http://vita.had.co.nz/papers/tidy-data.pdf)). So, for example, if we wanted a panel or colour for each level of a variable called `fishstock` then we’d need a column named `fishstock` with all the different values of `fishstock` in that column.

With the tidyr (or reshape2) and dplyr packages, you can get almost any dataset into shape for ggplot2 in a few lines. Sometimes this will take some serious thought along with pen and paper.

## Random tips

Jittering and statistics

:   ggplot2 has lots of built-in niceties like automatic position jittering and the addition of basic statistical models to your plots. Have a look through the help resources listed at the end of this document.

Axis labels

:   `xlab("Your x-axis label")`

Suppressing the legend

:   `theme(legend.position = none)`

Exploiting the object-oriented nature of ggplot2

:   Save the basic plot to an object and then experiment with different aesthetics, geoms, and theme adjustments.

```{r}
p <- ggplot(d) + geom_point(aes(x, y, colour = group1))
p + facet_grid(~group1) # try one way
p + facet_grid(~group2) # try another way
```

Horizontal error bars

:   Say you want to make a coefficient plot with the coefficients down the y-axis. You can either build the plot piece by piece with points and segments, or you can use `geom_pointrange()` and then rotate the axes with `+ coord_flip()`.

Axis limits and zooming

:   ggplot has two ways of adjusting the axis limits: `+ xlim(2, 5)` will “cut” the data at `2` and `5` and plot the data. `coord_cartesian(xlim = c(2, 5))` will zoom in on the plot while retaining the original data. This will, for example, affect colour scales.

Displaying and saving ggplot2 plots

:   If ggplot plots are generated in a function, they will be need to be wrapped in `print()` to display. There is a convenience function `ggsave("filename.pdf")`, which will save the last plot to a PDF and guess at reasonable dimensions. You can, of course specify which plot to save and the dimensions of the PDF. You can also use all the same methods of saving ggplot plots that you can use for base graphics.

Combining multiple ggplot panels

:   ggplot makes it easy to create multipanel figures with `facet_grid()` and `facet_wrap()`. But, sometimes you need to combine different kinds of plots in one figure or plots that require different data. One easy way to do this is with the `grid.arrange()` function from the gridExtra package. For example:

```{r grid.arrange}
df <- data.frame(x = 1:10, y = rnorm(100))
p1 <- ggplot(df, aes(x, y)) + geom_point()
p2 <- ggplot(df, aes(y)) + geom_histogram()
gridExtra::grid.arrange(p1, p2)
```

ggvis

:   [ggvis](https://github.com/rstudio/ggvis) is a partial successor to ggplot2 and is under rapid development. It focuses on implementing a similar grammar of graphics for web-based and possibly interactive output. Install it with `devtools::install_github("rstudio/ggvis")`.

## Finding help

The best (and nearly essential) help source is the website, which is largely based off the package help files. However, the website also shows the example plots and is easier to navigate: <http://docs.ggplot2.org/>. Don’t be afraid to keep it open whenever you’re using ggplot2.

There's also an [active ggplot2 discussion group](http://groups.google.com/group/ggplot2).

ggplot2 is [heavily featured](http://stackoverflow.com/questions/tagged/ggplot2) on stackoverflow.

Hadley wrote a [book on ggplot2](http://ggplot2.org/book/) and it was recently updated.

# Feedback

If you have any questions on the notes or exercises, or want to provide feedback, feel free to email me at sean "at" seananderson.ca.

<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/deed.en_US"><img alt="Creative Commons License" style="border-width:0" src="http://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/deed.en_US">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

### 2018-07-19 content/blog/2014-09-13-dplyr-intro.Rmd
Merge hugo site
Fix typo

--- 2014-09-15 _knitr-posts/2014-09-13-dplyr-intro.Rmd
@@
 ---
-layout: post
-title: "dplyr and pipes: the basics"
-subtitle: "Revolutionize your data analysis with dplyr and pipes"
-published: true
+title: 'dplyr and pipes: the basics'
+subtitle: Revolutionize your data analysis with dplyr and pipes
+date: '2014-09-13'
+slug: dplyr-intro
 ---
@@
 plyr 2.0 if you will.
-It does less than dplyr, but what it does it does more elegantly and much more quickly.
+It does less than plyr, but what it does it does more elegantly and much more quickly.
 

### 2018-07-19 content/blog/2014-09-13-dplyr-intro.Rmd
Not a draft

@@
 slug: dplyr-intro
+draft: no
 ---

### 2018-08-07 cv/cv-body.tex
Switch ISEC slides
Update CV

@@
 \tightlist
+\item[2018]
+Walsh, J.C., C. Minto, E. Jardim, \textbf{S.C. Anderson}, O.P. Jensen,
+J. Afflerbach, K.M. Kleisner, M. Dickey-Collas, C. Longo1, G.C. Osio,
+E.R. Selig, J.T. Thorson, M.B. Rudd, K.J. Papacostas, J.N. Kittinger,
+A.A. Rosenberg, A.B. Cooper. Trade-offs for data-limited fisheries when
+using harvest strategies based on catch-only models. Fish and Fisheries.
+In press.
 \item[2018]
@@
 Ecology Conference, St.\ Andrews, Scotland.
-\href{https://www.dropbox.com/s/z61ssz9xgymtzsg/anderson-ward-isec-2018.pdf?dl=1}{slides}
+(\href{https://www.dropbox.com/s/ufce52ocpycftk1/anderson-ward-isec-2018.pdf?dl=1}{slides})
 \item[2017]
@@
 ecological portfolio effect overestimates the benefit of diversity.
-Branch Lab, School of Aquatic and Fishery Sciences, University of
-Washington, Seattle, WA, United States.
+School of Aquatic and Fishery Sciences, University of Washington,
+Seattle, WA, United States.
 \item[2012]

### 2018-08-07 cv/refs.R

@@ -1,0 +1,98 @@
+# TODO J.E.M. Flemming should be J.E. Mills Flemming
+# TODO J.E. Mills Flemming needs to be fixed throughout refs.bib too
+# TODO Earth2Oceans Derby needs to be manually fixed
+
+# Need to fix all authors contributed equally and asterisks
+
+library("bibtex")
+d <- read.bib("refs.bib")
+
+ref <- list()
+
+for(i in seq_along(d)) {
+  name <- list()
+  x <- d[[i]]
+  # authors:
+  for(j in seq_along(x$author)) {
+    init <- paste0(paste(substr(x$author$given[[j]], 1, 1), collapse = "."), ".")
+    family <- x$author$family[[j]]
+    if(j == 1) {
+      name[[j]] <- paste0(family, ", ", init)
+      if(name[[j]] == "Anderson, S.C.") name[[j]] <- "**Anderson, S.C.**"
+    } else {
+      name[[j]] <- paste(init, family)
+      if(name[[j]] == "S.C. Anderson") name[[j]] <- "**S.C. Anderson**"
+    }
+  }
+  name0 <- paste0(paste0(name, collase = ""), collapse = ", ")
+
+  vol_num_p <- ""
+  if(!is.null(x$number) & !is.null(x$volume) & !is.null(x$pages)) {
+    vol_num_p <- paste0(x$volume, "(", x$number, "): ", x$pages)
+  }
+  # no number:
+  if(is.null(x$number) & !is.null(x$volume) & !is.null(x$pages)) {
+    vol_num_p <- paste0(x$volume, ": ", x$pages)
+  }
+  # no pages:
+  if(is.null(x$number) & !is.null(x$volume) & is.null(x$pages)) {
+    vol_num_p <- paste0(x$volume, "(", x$number, ")")
+  }
+  # no number or pages:
+  if(is.null(x$number) & !is.null(x$volume) & is.null(x$pages)) {
+    vol_num_p <- x$volume
+  }
+  # no vol, number, or pages:
+  if(is.null(x$number) & is.null(x$volume) & is.null(x$pages)) {
+    vol_num_p <- ""
+  }
+
+  if(vol_num_p != "" & !vol_num_p %in% c("In review", "In revision", "Submitted")) {
+    main <- paste0(paste0(name0, ". ", x$title, ". ", x$journal, ". ", vol_num_p), ".")
+  } else {
[…]

### 2018-11-18 cv/cv-body.tex
Render CV

@@
 \tightlist
-\item[2017-18]
+\item[2017--18]
 Program Head in Analytical Approaches and Spatial Stock Assessment,
@@
 \item[2018]
+Afflerbach, J.C., M. Frazier, H.E. Froehlich, \textbf{S.C. Anderson},
+B.S. Halpern. Uncertainty in a global fisheries indicator. In press at
+Fish and Fisheries.
+\item[2018]
+Marshall, K.N., J.T. Duffy-Anderson, E.J. Ward, \textbf{S.C. Anderson},
+M.E. Hunsicker, B.C. Williams. Long-term trends in ichthyoplankton
+assemblage structure, biodiversity, and synchrony in the Gulf of Alaska
+and their relationships to climate. In press at Progress in
+Oceanography. \url{https://doi.org/10.1016/j.pocean.2018.11.002}.
+\item[2018]
 Walsh, J.C., C. Minto, E. Jardim, \textbf{S.C. Anderson}, O.P. Jensen,
@@
 A.A. Rosenberg, A.B. Cooper. Trade-offs for data-limited fisheries when
-using harvest strategies based on catch-only models. Fish and Fisheries.
-In press.
+using harvest strategies based on catch-only models. In press at Fish
+and Fisheries. \url{https://doi.org/10.1111/faf.12316}.
 \item[2018]
@@
 (\href{https://sean.updog.co/papers/Anderson_etal_2008_knowledge.pdf}{PDF}).
-Eighth most-read CJFAS article in 2008.
+\# Peer-reviewed reports
+\item[2018]
+Forrest, R.E., \textbf{S.C. Anderson}, C.J. Grandin, P.J. Starr.
+Assessment of Pacific Cod (\emph{Gadus macrocephalus}) for Hecate Strait
+and Queen Charlotte Sound (Area 5ABCD), and West Coast Vancouver Island
+(Area 3CD) in 2018. Canadian Science Advisory Secretariat (CSAS)
+Research Document. Accepted at regional peer-review meeting October
+2018.
 \end{description}
@@
 \tightlist
+\item[2018--20]
+Principal Investigator of a two-year grant from Fisheries and Oceans
+Canada's Aquatic Climate Change Adaptation Services Program (ACCASP).
+Project title: Characterizing range shifts in British Columbia
+groundfish species in response to local climate velocities.
 \item[2017]
-Canadian Society for Ecology and Evolution (CSEE) Early Career Award
-(\href{http://csee-scee.ca/?p=5183}{URL})
+Canadian Society for Ecology and Evolution (CSEE) Early Career Award.\\
+\url{http://csee-scee.ca/2017-early-career-awards/}
 \item[2015]
-David H. Smith Conservation Postdoctoral Research Fellowship
-(\href{http://conbio.org/mini-sites/smith-fellows/meet-the-fellows/2015-fellows/sean-anderson}{URL})
+David H. Smith Conservation Postdoctoral Research Fellowship\\
+\url{http://conbio.org/mini-sites/smith-fellows/meet-the-fellows/2015-fellows/sean-anderson}
 \item[2015]
-Simon Fraser University Dean of Graduate Studies Convocation Medal
-(\href{https://www.sfu.ca/dean-gradstudies/blog/year/2015/06/SeanAnderson.html}{URL})
+Simon Fraser University Dean of Graduate Studies Convocation Medal\\
+\url{https://www.sfu.ca/dean-gradstudies/blog/year/2015/06/SeanAnderson.html}
 \item[2014]
@@
 \tightlist
+\item[2018]
+\textbf{gfplot} \& \textbf{gfsynopsis}: R packages to generate an annual
+Research Document synopsis report showing all available fishery and
+biological data and basic model fits for 109 Pacific groundfish species
+for publication through the Canadian Science Advisory Secretariat
+(CSAS). To be reviewed via regional peer-review meeting November 22--23
+2018.\\
+\url{https://github.com/pbs-assess/gfplot} and
+\url{https://github.com/pbs-assess/gfsynopsis}
+\item[2018]
+\textbf{csasdown}: Reproducible Canadian Science Advisory Secretariat
+Research Documents via R Markdown.\\
+\url{https://github.com/pbs-assess/csasdown}
 \item[2017]
-glmmfields: Generalized linear mixed models with robust random fields
-for spatiotemporal modeling.
-\url{https://github.com/seananderson/glmmfields}
+\textbf{glmmfields}: Generalized linear mixed models with robust random
+fields for spatiotemporal modeling.\\
+\url{https://cran.r-project.org/package=glmmfields}
 \item[2017]
-datalimited: Stock assessment methods for data-limited fisheries.
-\url{https://github.com/datalimited/datalimited}
+\textbf{datalimited}: Stock assessment methods for data-limited
+fisheries. \url{https://github.com/datalimited/datalimited}
 \item[2014]
@@
 Cooper.\\
-metafolio: Metapopulation simulations for conserving salmon through
-portfolio optimization.
+\textbf{metafolio}: Metapopulation simulations for conserving salmon
+through portfolio optimization.
 \url{http://cran.r-project.org/package=metafolio}
@@
 Valero, C.J Cunningham, F. Hurtado-Ferro, R. Licandeo, C.R. McGilliard,
-C.S. Szuwalski, K.A. Vert-pre, A.R. Whitten. ss3sim: Fisheries stock
-assessment simulation testing with Stock Synthesis.
+C.S. Szuwalski, K.A. Vert-pre, A.R. Whitten. \textbf{ss3sim}: Fisheries
+stock assessment simulation testing with Stock Synthesis.
 \url{http://cran.r-project.org/package=ss3sim}
 \item[2013]
-\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. ecofolio: Tools to
-quantify metapopulation portfolio effects.
+\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. \textbf{ecofolio}:
+Tools to quantify metapopulation portfolio effects.
 \url{https://github.com/seananderson/ecofolio}
@@
 \tightlist
-\item[2016-17]
+\item[2016--17]
 NOAA (National Oceanic and Atmospheric Administration) FATE (Fisheries
@@
 abrupt change in northeast Pacific ecosystems}
-\item[2015-16]
+\item[2015--16]
 NCEAS (National Center for Ecological Analysis and Synthesis, Santa
@@
 spill?} (\href{https://www.nceas.ucsb.edu/featured/marshall}{URL})
-\item[2015-16]
+\item[2015--16]
 Gordon and Betty Moore Foundation funded working group \emph{Applying
@@
 Reviewer for Science, Ecology, Ecological Applications, Conservation
-Biology, Fish and Fisheries, Marine Policy, Canadian Journal of
-Fisheries and Aquatic Sciences, Population Ecology, Journal of Applied
-Ecology, International Journal of Tropical Biology and Conservation,
-Journal of Environmental Management, Endangered Species Research,
-Aquatic Conservation
+Biology, Canadian Journal of Fisheries and Aquatic Sciences, Proceedings
+B, Fish and Fisheries, Marine Policy, Population Ecology, Journal of
+Applied Ecology, International Journal of Tropical Biology and
+Conservation, Journal of Environmental Management, Endangered Species
+Research, Aquatic Conservation

### 2018-11-27 cv/cv-body.tex
Fix reports

@@
 (\href{https://sean.updog.co/papers/Anderson_etal_2008_knowledge.pdf}{PDF}).
-\# Peer-reviewed reports
+\end{description}
+
+\hypertarget{peer-reviewed-reports}{%
+\section{Peer-reviewed reports}\label{peer-reviewed-reports}}
+
+\begin{description}
+\tightlist
+\item[2018]
+\textbf{Anderson, S.C.}, E.A. Keppel, and A.M. Edwards. A reproducible
+data synopsis for over 100 species of British Columbia groundfish.
+Canadian Science Advisory Secretariat (CSAS) Research Document. Accepted
+at regional peer-review meeting November 2018.
 \item[2018]

### 2019-01-08 cv/cv-body.tex
Typo
Update CV

@@
 \tightlist
-\item[2017--18]
+\item[2017--19]
 Program Head in Analytical Approaches and Spatial Stock Assessment,
@@
 \tightlist
-\item[2018]
+\item[2019]
+Harvey, B.J., R.A. Andrus, \textbf{S.C. Anderson}. Incorporating
+biophysical gradients and uncertainty into burn severity maps in a
+temperate fire-prone forested region. In press at Ecosphere.
+\item[2019]
 Afflerbach, J.C., M. Frazier, H.E. Froehlich, \textbf{S.C. Anderson},
@@
 Fish and Fisheries.
-\item[2018]
+\item[2019]
 Marshall, K.N., J.T. Duffy-Anderson, E.J. Ward, \textbf{S.C. Anderson},
@@
 assemblage structure, biodiversity, and synchrony in the Gulf of Alaska
-and their relationships to climate. In press at Progress in
-Oceanography. \url{https://doi.org/10.1016/j.pocean.2018.11.002}.
-\item[2018]
-Walsh, J.C., C. Minto, E. Jardim, \textbf{S.C. Anderson}, O.P. Jensen,
-J. Afflerbach, K.M. Kleisner, M. Dickey-Collas, C. Longo1, G.C. Osio,
-E.R. Selig, J.T. Thorson, M.B. Rudd, K.J. Papacostas, J.N. Kittinger,
-A.A. Rosenberg, A.B. Cooper. Trade-offs for data-limited fisheries when
-using harvest strategies based on catch-only models. In press at Fish
-and Fisheries. \url{https://doi.org/10.1111/faf.12316}.
-\item[2018]
+and their relationships to climate. Progress in Oceanography.
+170:134--145. \url{https://doi.org/10.1016/j.pocean.2018.11.002}.
+\item[2019]
 \textbf{Anderson, S.C.}, Ward, E. J. Black swans in space: modelling
-spatiotemporal processes with extremes. In press at Ecology.
+spatiotemporal processes with extremes. Ecology. 100(1):e02403.
 \url{https://doi.org/10.1002/ecy.2403}. glmmfields R package:
@@
 \href{https://github.com/seananderson/glmmfields}{GitHub}.
+\item[2018]
+Walsh, J.C., C. Minto, E. Jardim, \textbf{S.C. Anderson}, O.P. Jensen,
+J. Afflerbach, K.M. Kleisner, M. Dickey-Collas, C. Longo, G.C. Osio,
+E.R. Selig, J.T. Thorson, M.B. Rudd, K.J. Papacostas, J.N. Kittinger,
+A.A. Rosenberg, A.B. Cooper. Trade-offs for data-limited fisheries when
+using harvest strategies based on catch-only models. Fish and Fisheries.
+19(6):1130--1146. \url{https://doi.org/10.1111/faf.12316}.
 \item[2018]
@@
 \tightlist
+\item[2019]
+Canadian Journal of Fisheries and Aquatic Sciences Stevenson
+Lectureship. \url{http://www.nrcresearchpress.com/page/cjfas/award}
 \item[2018--20]

### 2019-06-01 cv/cv-body.tex
Update CV
Fixed name formatting
Add editing/reviewing; remove courses from CV

@@
 \item[2019]
-Harvey, B.J., R.A. Andrus, \textbf{S.C. Anderson}. Incorporating
-biophysical gradients and uncertainty into burn severity maps in a
-temperate fire-prone forested region. In press at Ecosphere.
+Beaudreau, A.H., E.J. Ward, R.E. Brenner, A.O. Shelton, J.T. Watson,
+J.C. Womack, \textbf{S.C. Anderson}, A.C. Haynie, K.N. Marshall, B.C.
+Williams. Thirty years of change and the future of Alaskan fisheries:
+Shifts in fishing participation and diversification in response to
+environmental, regulatory, and economic pressures. In press at Fish and
+Fisheries.
 \item[2019]
@@
 Fish and Fisheries.
+\item[2019]
+Harvey, B.J., R.A. Andrus, \textbf{S.C. Anderson}. Incorporating
+biophysical gradients and uncertainty into burn severity maps in a
+temperate fire-prone forested region. Ecosphere. 10(2):02600.
+\url{https://doi.org/10.1002/ecs2.2600}.
 \item[2019]
@@
 
-\hypertarget{selected-courses-and-training}{%
-\section{Selected courses and
-training}\label{selected-courses-and-training}}
+\hypertarget{editing-and-reviewing}{%
+\section{Editing and Reviewing}\label{editing-and-reviewing}}
 
-\begin{description}
-\tightlist
-\item[Present]
-Proficient in statistical analyses and data management using R, Stan,
-JAGS, knitr, Shiny, Git, LaTeX, TMB, SQL.
-\item[2016]
-Diversity in Conservation Biology, facilitated by Yolanda L. Comedy,
-Director of the American Association for the Advancement of Science
-(AAAS) Center for Advancing Science \& Engineering Capacity
-\item[2015--16]
-Powerful Presentation Skills and Media Training with Teresa Erickson and
-Tim Ward of Intermedia Communications, authors of
-\href{https://www.amazon.com/Master-Communicators-Handbook-Teresa-Erickson/dp/1785351532}{\emph{The
-Master Communicator's Handbook}}
-\item[2013]
-Quantitative Modelling in Ecosystem Based Management. R. Hilborn, School
-of Aquatic and Fishery Sciences, University of Washington (audited).
-\item[2013]
-Applied Time Series Analysis. E. Holmes, M. Scheuerell, and E. Ward;
-School of Aquatic and Fishery Sciences, University of Washington
-(audited).
-\item[2012]
-Numerical Computing for the Natural Resources. A.E. Punt, School of
-Aquatic and Fishery Sciences, University of Washington (audited).
-\item[2012]
-Risk Assessment and Decision Analysis for Management of Natural
-Resources. R.M. Peterman, School of Resource and Environmental
-Management, Simon Fraser University.
-\item[2011]
-Presenting Data and Information. Edward Tufte, Seattle, WA, United
-States.
-\end{description}
+\href{https://www.nrcresearchpress.com/page/cjfas/editors}{Associate
+Editor} at Canadian Journal of Fisheries and Aquatic Sciences
+(2019--present)
 
-\hypertarget{reviews}{%
-\section{Reviews}\label{reviews}}
+\vspace{4mm}
 
-Reviewer for Science, Ecology, Ecological Applications, Conservation
-Biology, Canadian Journal of Fisheries and Aquatic Sciences, Proceedings
-B, Fish and Fisheries, Marine Policy, Population Ecology, Journal of
+Reviewer for Science, Ecology, Canadian Journal of Fisheries and Aquatic
+Sciences, Fish and Fisheries, Ecological Applications, Conservation
+Biology, Proceedings B, Marine Policy, Population Ecology, Journal of
 Applied Ecology, International Journal of Tropical Biology and
-Conservation, Journal of Environmental Management, Endangered Species
-Research, Aquatic Conservation
+Conservation, Biological Conservation, Oikos, Journal of Environmental
+Management, Endangered Species Research, Aquatic Conservation

### 2019-06-25 cv/cv-body.tex
Update CV

@@
 \item[2019]
+Freshwater, C., \textbf{S.C. Anderson}, K.R. Holt, A.-M. Huang, C.A.
+Holt. Weakened portfolio effects constrain management effectiveness for
+population aggregates. In press at Ecological Applications.
+\item[2019]
+Ward, E.J., \textbf{S.C. Anderson}, L.A. Damiano, M.E. Hunsicker,
+M.A.\ Litzow. Modeling regimes with extremes: the `bayesdfa' package for
+identifying and forecasting common trends and anomalies in multivariate
+time-series data. In press at The R Journal.
+\item[2019]
+Afflerbach, J.C., M. Frazier, H.E. Froehlich, \textbf{S.C. Anderson},
+B.S. Halpern. Uncertainty in a global fisheries indicator. In press at
+Fish and Fisheries.
+\item[2019]
 Beaudreau, A.H., E.J. Ward, R.E. Brenner, A.O. Shelton, J.T. Watson,
@@
 environmental, regulatory, and economic pressures. In press at Fish and
-Fisheries.
-\item[2019]
-Afflerbach, J.C., M. Frazier, H.E. Froehlich, \textbf{S.C. Anderson},
-B.S. Halpern. Uncertainty in a global fisheries indicator. In press at
-Fish and Fisheries.
+Fisheries. \url{https://doi.org/10.1111/faf.12364}.
 \item[2019]
@@
 data synopsis for over 100 species of British Columbia groundfish.
-Canadian Science Advisory Secretariat (CSAS) Research Document. Accepted
-at regional peer-review meeting November 2018.
+Canadian Science Advisory Secretariat (CSAS) Research Document 2019/041.
+In press.
 \item[2018]

### 2019-06-25 cv/cv-body.tex
More updating CV

@@
 Canadian Journal of Fisheries and Aquatic Sciences Stevenson
-Lectureship. \url{http://www.nrcresearchpress.com/page/cjfas/award}
+Lectureship. \url{http://www.nrcresearchpress.com/page/cjfas/award} A
+``lectureship conferred upon a young, energetic and creative researcher
+at the cutting edge of an aquatic discipline''.
 \item[2018--20]
@@
 \tightlist
+\item[2019]
+NOAA (National Oceanic and Atmospheric Administration) FATE (Fisheries
+and the Environment) working group \emph{Spatial ecosystem state and
+pressure indicators and the scale-dependence of temporal indicator
+performance in the California Current and beyond}
+\item[2018--19]
+European Commission Joint Research Council working group \emph{A new
+approach to strengthen evidence-based fisheries management}
+investigating approaches to providing fisheries management scientific
+advice based on model ensembles.
 \item[2016--17]

### 2019-08-15 cv/cv-body.tex
Update CV

@@
 \tightlist
-\item[2017--19]
+\item[2017--present]
 Program Head in Analytical Approaches and Spatial Stock Assessment,
 Groundfish Section, Stock Assessment and Research Division, Pacific
-Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
+Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada\\
+Associate Editor at the Canadian Journal of Fisheries and Aquatic
+Sciences
 \item[2015--17]
@@
 \item[2019]
+Veríssimo, D., \textbf{S.C. Anderson}, M. Tlusty. Did the movie Finding
+Dory increase demand for blue tang fish? In press at Ambio.
+\url{https://doi.org/10.1007/s13280-019-01233-7}.
+\item[2019]
 Freshwater, C., \textbf{S.C. Anderson}, K.R. Holt, A.-M. Huang, C.A.
@@
 population aggregates. In press at Ecological Applications.
+\url{https://doi.org/10.1002/eap.1966}.
 \item[2019]
@@
 time-series data. In press at The R Journal.
+\url{https://doi.org/10.32614/RJ-2019-007}.
 \item[2019]
@@
 \tightlist
-\item[2018]
+\item[2019]
+\textbf{rosettafish}: An R package for translating fish- and
+fisheries-related terms.\\
+\url{https://github.com/pbs-assess/rosettafish}
+\item[2019]
+\textbf{sdmTMB}: An R package for fitting spatiotemporal species
+distribution 9GLMMs (generalized linear mixed effects models) with TMB
+(Template Model Builder)\\
+\url{https://github.com/pbs-assess/sdmTMB}
+\item[2019]
 \textbf{gfplot} \& \textbf{gfsynopsis}: R packages to generate an annual
@@
 \url{https://github.com/pbs-assess/gfsynopsis}
-\item[2018]
+\item[2019]
 \textbf{csasdown}: Reproducible Canadian Science Advisory Secretariat
@@
 \url{https://github.com/pbs-assess/csasdown}
-\item[2017]
+\item[2018]
 \textbf{glmmfields}: Generalized linear mixed models with robust random
@@
 \tightlist
+\item[2019]
+\textbf{Anderson, S.C.}, E.A. Keppel, A.M. Edwards, P.A. English, E.J.
+Ward. An automated synopsis of the state of Pacific Canadian groundfish
+and climate impacts. North Pacific Marine Science Organization (PICES)
+conference. Victoria, BC, Canada. Invited talk to plenary session (to be
+given October 2019).
 \item[2018]
@@
 
-\hypertarget{editing-and-reviewing}{%
-\section{Editing and Reviewing}\label{editing-and-reviewing}}
-
-\href{https://www.nrcresearchpress.com/page/cjfas/editors}{Associate
-Editor} at Canadian Journal of Fisheries and Aquatic Sciences
-(2019--present)
-
-\vspace{4mm}
+\hypertarget{reviewing}{%
+\section{Reviewing}\label{reviewing}}
 

### 2020-03-05 cv/cv-body.tex
Fix link
Update CV
Update cv

@@
 \tightlist
-\item[2019]
+\item[2020]
+Huynh, Q.C., A.R. Hordyk, R.E. Forrest, C.E. Porch, \textbf{S.C.
+Anderson}, T.R. Carruthers. The interim management procedure approach
+for assessed stocks: Responsive management advice and lower assessment
+frequency. In press at Fish and Fisheries.
+\url{https://doi.org/10.1111/faf.12453}.
+\item[2020]
+Free, C.M., O.P. Jensen, \textbf{S.C. Anderson}, N.L. Gutierrez, K.M.
+Kleisner, C. Longo, C. Minto, G.C. Osio, J.C. Walsh. Blood from a stone:
+performance of catch-only methods in estimating stock biomass status.
+Fisheries Research. 223:105452.
+\url{https://doi.org/10.1016/j.fishres.2019.105452}.
+\item[2020]
 Veríssimo, D., \textbf{S.C. Anderson}, M. Tlusty. Did the movie Finding
-Dory increase demand for blue tang fish? In press at Ambio.
+Dory increase demand for blue tang fish? Ambio. 49:903--911.
 \url{https://doi.org/10.1007/s13280-019-01233-7}.
@@
 Holt. Weakened portfolio effects constrain management effectiveness for
-population aggregates. In press at Ecological Applications.
+population aggregates. Ecological Applications. 29(7):e01966.
 \url{https://doi.org/10.1002/eap.1966}.
@@
 time-series data. In press at The R Journal.
-\url{https://doi.org/10.32614/RJ-2019-007}.
+\href{https://journal.r-project.org/archive/2019/RJ-2019-007/RJ-2019-007.pdf}{https://doi.org/10.32614/RJ-2019-007}.
 \item[2019]
 Afflerbach, J.C., M. Frazier, H.E. Froehlich, \textbf{S.C. Anderson},
-B.S. Halpern. Uncertainty in a global fisheries indicator. In press at
-Fish and Fisheries.
+B.S. Halpern. Quantifying uncertainty in the wild-caught fisheries goal
+of the Ocean Health Index. Fish and Fisheries. 20(2):343--354.
+\url{https://doi.org/10.1111/faf.12346}.
 \item[2019]
@@
 Shifts in fishing participation and diversification in response to
-environmental, regulatory, and economic pressures. In press at Fish and
-Fisheries. \url{https://doi.org/10.1111/faf.12364}.
+environmental, regulatory, and economic pressures. Fish and Fisheries.
+20(4):601--619. \url{https://doi.org/10.1111/faf.12364}.
 \item[2019]

### 2020-04-15 cv/cv-body.tex
Quick update of CV and PDF links

@@
 \tightlist
+\item[2020]
+\textbf{Anderson, S.C.}, E.A. Keppel, A.M. Edwards. Reproducible
+visualization of raw fisheries data for 113 species improves
+transparency, assessment efficiency, and monitoringy. In press at
+Fisheries.
 \item[2020]
@@
 \url{https://doi.org/10.1016/j.fishres.2015.09.007}
-(\href{https://sean.updog.co/papers/Kuriyama_etal_2016_empirical.pdf}{PDF}).
+(\href{https://www.dropbox.com/s/425a18a4xm9kq8l/Kuriyama_etal_2016_empirical.pdf?dl=1}{PDF}).
 \item[2016]
@@
 \url{https://doi.org/10.1016/j.fishres.2015.11.002}
-(\href{https://sean.updog.co/papers/Monnahan_etal_2016_binwidth.pdf}{PDF}).
+(\href{https://www.dropbox.com/s/uksmdtptby9w0ku/Monnahan_etal_2016_binwidth.pdf?dl=1}{PDF}).
 \item[2015]
@@
 \url{https://doi.org/10.1111/1365-2656.12322}
-(\href{https://sean.updog.co/papers/Boudreau_etal_2015_lobster.pdf}{PDF}).
+(\href{https://www.dropbox.com/s/w027swfx2o8hgvl/Boudreau_etal_2015_lobster.pdf?dl=1}{PDF}).
 \item[2015]
@@
 \url{https://doi.org/10.1890/14-0266.1}
-(\href{https://sean.updog.co/papers/Anderson_etal_2015_salmonportfolios.pdf}{PDF}).
+(\href{https://www.dropbox.com/s/141rsnv5rc7mi5i/Anderson_etal_2015_salmonportfolios.pdf?dl=1}{PDF}).
 \item[2015]
@@
 \url{https://doi.org/10.1111/2041-210X.12093}
-(\href{https://sean.updog.co/papers/Anderson_etal_2013_ecological_prophets_with_SOM.pdf}{PDF},
+(\href{https://www.dropbox.com/s/7tx1h1pkmmp222j/Anderson_etal_2013_ecological_prophets_with_SOM.pdf?dl=1}{PDF},
 \href{https://github.com/seananderson/ecofolio}{GitHub R package}).
@@
 \url{https://doi.org/10.1111/j.1755-263X.2012.00294.x}
-(\href{https://sean.updog.co/papers/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf}{PDF},
+(\href{https://www.dropbox.com/s/cxt848ng5x4hc4t/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf?dl=1}{PDF},
 \href{https://github.com/seananderson/conservation_pathways}{GitHub
@@
 27(11): 608--617. \url{https://doi.org/10.1016/j.tree.2012.07.010}
-(\href{https://sean.updog.co/papers/Harnik_etal_2012_Extinctions_in_ancient_and_modern_seas.pdf}{PDF}).
+(\href{https://www.dropbox.com/s/rdn9685viqr37qy/Harnik_etal_2012_Extinctions_in_ancient_and_modern_seas.pdf?dl=1}{PDF}).
 \item[2012]
@@
 and stock-assessment reference points. ICES Journal of Marine Science.
-69(8): 1491--1500. \url{https://doi.org/10.1093/icesjms/fss105}
-(\href{https://sean.updog.co/Anderson_etal_2012_Assessing_global_marine_fishery.pdf}{PDF}).
+69(8): 1491--1500. \url{https://doi.org/10.1093/icesjms/fss105}.
 \item[2011]
@@
 \url{https://doi.org/10.1525/bio.2011.61.7.8}
-(\href{https://sean.updog.co/papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF}).
+(\href{https://www.dropbox.com/s/bt16dvi3idw3gdx/Anderson_etal_2011_BioScience_with_supplement.pdf?dl=1}{PDF}).
 \item[2011]
@@
 \url{https://doi.org/10.1111/j.1467-2979.2010.00397.x}
-(\href{https://sean.updog.co/papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF}).
+(\href{https://www.dropbox.com/s/d8id4zxe9xv6jau/Anderson_etal_2011_seacucumbers_with_supplement.pdf?dl=1}{PDF}).
 \href{http://www.sciencemag.org/content/331/6014/129.1.full}{Featured in
@@
 2553--2571. \url{https://doi.org/10.1139/F08-156}
-(\href{https://sean.updog.co/papers/Anderson_etal_2008_knowledge.pdf}{PDF}).
+(\href{https://www.dropbox.com/s/b0la81jbqitib6u/Anderson_etal_2008_knowledge.pdf?dl=1}{PDF}).
 \end{description}

### 2020-04-27 cv/cv-body.tex
Update a report
No photos; small tweaks
Add preprint

@@
 
+\hypertarget{preprints}{%
+\section{Preprints}\label{preprints}}
+
+\begin{description}
+\tightlist
+\item[2020]
+\textbf{Anderson, S.C.}, A.M. Edwards, M. Yerlanov, N. Mulberry, J.
+Stockdale, S.A. Iyaniwura, R.C. Falcao, M.C. Otterstatter, M.A.\ Irvine,
+N.Z. Janjua, D. Coombs, C. Colijn. Estimating the impact of COVID-19
+control measures using a Bayesian model of physical distancing. medRxiv
+2020.04.17.20070086. \url{https://doi.org/10.1101/2020.04.17.20070086}.
+\end{description}
+
 \hypertarget{publications}{%
@@
 transparency, assessment efficiency, and monitoringy. In press at
-Fisheries.
+Fisheries. \url{https://doi.org/10.1002/fsh.10441}.
 \item[2020]
@@
 \tightlist
-\item[2018]
-\textbf{Anderson, S.C.}, E.A. Keppel, and A.M. Edwards. A reproducible
-data synopsis for over 100 species of British Columbia groundfish.
-Canadian Science Advisory Secretariat (CSAS) Research Document 2019/041.
-In press.
+\item[2019]
+\textbf{Anderson, S.C.}, Keppel, E.A., Edwards, A.M. 2019. A
+reproducible data synopsis for over 100 species of British Columbia
+groundfish. DFO Can. Sci. Advis. Sec. Res. Doc. 2019/041. vii + 321 p.
+(\href{http://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2019/2019_041-eng.html}{English
+version},
+\href{http://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2019/2019_041-fra.html}{French
+version})
 \item[2018]
@@
 Research Document. Accepted at regional peer-review meeting October
-2018.
+2018. In press.
 \end{description}

### 2020-06-25 cv/cv-body.tex
Update CV
Quick partial cv update

@@
 Groundfish Section, Stock Assessment and Research Division, Pacific
-Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada\\
-Associate Editor at the Canadian Journal of Fisheries and Aquatic
-Sciences
+Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
+\end{description}
+
+2020--present Adjunct Professor, Department of Mathematics, Simon Fraser
+University
+
+2019--present Associate Editor, the Canadian Journal of Fisheries and
+Aquatic Sciences
+
+\begin{description}
+\tightlist
 \item[2015--17]
@@
 Fishery Sciences, University of Washington and Northwest Fisheries
-Science Center, Seattle, WA, USA
+Science Center, Seattle, WA
 \item[2015]
 Postdoctoral Fellow, School of Resource and Environmental Management,
-Simon Fraser University, Burnaby, BC, Canada
+Simon Fraser University
 \item[2011--15]
 Graduate Research Assistant, Department of Biological Sciences, Simon
-Fraser University, Burnaby, BC, Canada
+Fraser University
 \item[2008--11]
-Graduate Research Assistant, Department of Biology, Dalhousie
-University, Halifax, NS, Canada
+Graduate Research Assistant, Department of Biology, Dalhousie University
 \end{description}
@@
 \tightlist
+\item[2020]
+\textbf{Anderson, S.C.}, Mulberry, N., Edwards, A.M., Stockdale, J.E.,
+Iyaniwura, S.A., Falcao, R.C., Otterstatter, M.C., Janjua, N.Z., Coombs,
+D., and Colijn, C. 2020. How much leeway is there to relax COVID-19
+control measures? medRxiv 2020.06.12.20129833.
+\url{https://doi.org/10.1101/2020.06.12.20129833}.
 \item[2020]
@@
 \tightlist
+\item[2019]
+Two-day workshop on Advanced R topics (functional programming,
+debugging, profiling, benchmarking, parallel processing, integrating C++
+with R). University of British Columbia.
+\url{https://github.com/seananderson/adv-r-course}
 \item[2018]
-One-day workshop on Bayesian data analysis for ecologists with Stan.
+One-day workshop on Bayesian data analysis for ecologists with Stan for
+the David H. Smith Conservation Research Fellowship Program.
 \url{https://github.com/seananderson/bayes-course}

### 2020-06-25 cv/cv-body.tex
Formatting fixes to cv

@@
 \tightlist
-\item[2017--present]
+\item[2017--20]
 Program Head in Analytical Approaches and Spatial Stock Assessment,
@@
 Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
-\end{description}
-
-2020--present Adjunct Professor, Department of Mathematics, Simon Fraser
-University
-
-2019--present Associate Editor, the Canadian Journal of Fisheries and
-Aquatic Sciences
-
-\begin{description}
-\tightlist
+\item[2020--20]
+Adjunct Professor, Department of Mathematics, Simon Fraser University\\
+\item[2019--20]
+Associate Editor, the Canadian Journal of Fisheries and Aquatic Sciences
 \item[2015--17]

### 2020-09-07 cv/cv-body.tex
Fix f1000
Fix typo
Add Litzow et al. 2020
Add pubs

@@
 \url{https://doi.org/10.1101/2020.06.12.20129833}.
-\item[2020]
-\textbf{Anderson, S.C.}, A.M. Edwards, M. Yerlanov, N. Mulberry, J.
-Stockdale, S.A. Iyaniwura, R.C. Falcao, M.C. Otterstatter, M.A.\ Irvine,
-N.Z. Janjua, D. Coombs, C. Colijn. Estimating the impact of COVID-19
-control measures using a Bayesian model of physical distancing. medRxiv
-2020.04.17.20070086. \url{https://doi.org/10.1101/2020.04.17.20070086}.
 \end{description}
@@
 \item[2020]
+\textbf{Anderson, S.C.}, A.M. Edwards, M. Yerlanov, N. Mulberry, J.
+Stockdale, S.A. Iyaniwura, R.C. Falcao, M.C. Otterstatter, M.A.\ Irvine,
+N.Z. Janjua, D. Coombs, C. Colijn. Estimating the impact of COVID-19
+control measures using a Bayesian model of physical distancing. In press
+at PLOS Computational Biology. Preprint: medRxiv 2020.04.17.20070086.
+\url{https://doi.org/10.1101/2020.04.17.20070086}.
+\item[2020]
+Cahill, C.L., \textbf{S.C. Anderson}, A.J. Paul, L. MacPherson, M.G.
+Sullivan, B.T. van Poorten, C.J. Walters, and J.R. Post. A
+spatial-temporal approach to modeling somatic growth across inland
+recreational fisheries landscapes. In press at Canadian Journal of
+Fisheries and Aquatic Sciences.
+\url{https://doi.org/10.1139/cjfas-2019-0434}.
+\item[2020]
+Litzow, M.A., M.E. Hunsicker, E.J. Ward, \textbf{S.C. Anderson}, J. Gao,
+S. Zador, S. Batten, S. Dressel, J. Duffy-Anderson, E. Fergusson, R.
+Hopcroft, B.J. Laurel, and R. O'Malley. Evaluating ecosystem change as
+Gulf of Alaska temperature exceeds the limits of preindustrial
+variability. Progress in Oceanography. 186:102393.
+\url{https://doi.org/10.1016/j.pocean.2020.102393}.
+\item[2020]
 \textbf{Anderson, S.C.}, E.A. Keppel, A.M. Edwards. Reproducible
 visualization of raw fisheries data for 113 species improves
-transparency, assessment efficiency, and monitoringy. In press at
+transparency, assessment efficiency, and monitoring. In press at
 Fisheries. \url{https://doi.org/10.1002/fsh.10441}.
@@
 (\href{http://onlinelibrary.wiley.com/doi/10.1890/13-0916.1/epdf}{PDF}).
+\href{https://facultyopinions.com/prime/718498042}{F1000 Prime
+recommended}.
 \item[2013]
@@
 (\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0014735\&representation=PDF}{PDF}).
+\href{https://f1000.com/prime/9542957}{F1000 Prime recommended}.
 \item[2008]

### 2020-11-11 cv/cv-body.tex
Update CV

@@
 Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
-\item[2020--20]
-Adjunct Professor, Department of Mathematics, Simon Fraser University\\
 \item[2019--20]
 Associate Editor, the Canadian Journal of Fisheries and Aquatic Sciences
+\item[2020--20]
+Adjunct Professor, Department of Mathematics, Simon Fraser University
 \item[2015--17]
@@
 identifying and forecasting common trends and anomalies in multivariate
-time-series data. In press at The R Journal.
-\href{https://journal.r-project.org/archive/2019/RJ-2019-007/RJ-2019-007.pdf}{https://doi.org/10.32614/RJ-2019-007}.
+time-series data. The R Journal. 11(2):46--55.
+\url{https://doi.org/10.32614/RJ-2019-007}.
 \item[2019]
@@
 \tightlist
+\item[2020]
+Haggarty D.R., Q.C. Huynh, R.E. Forrest, \textbf{S.C. Anderson}, M.J.
+Bresch, E.A. Keppel. Evaluation of potential rebuilding strategies for
+Inside Yelloweye Rockﬁsh (\emph{Sebastes ruberrimus}) in British
+Columbia. Accepted as a DFO Can. Sci. Advis. Sec. Res. Doc. on June 11
+2020.
+(\href{https://www.dropbox.com/s/pll8axzsxqka8j8/resdoc-iye.pdf?dl=1}{English
+version}).
+\item[2020]
+\textbf{Anderson, S.C.}, R.E. Forrest, Q.C. Huynh, E.A. Keppel. A
+management procedure framework for groundﬁsh in British Columbia.
+Accepted as a DFO Can. Sci. Advis. Sec. Res. Doc. on June 9 2020.
+(\href{https://www.dropbox.com/s/rgxk2wco5v5zp0e/gfmp-2020-03-13.pdf?dl=1}{English
+version}).
+\item[2020]
+Forrest, R.E., \textbf{S.C. Anderson}, C.J. Grandin, P.J. Starr.
+Assessment of Pacific Cod (\emph{Gadus macrocephalus}) for Hecate Strait
+and Queen Charlotte Sound (Area 5ABCD), and West Coast Vancouver Island
+(Area 3CD) in 2018. DFO Can. Sci. Advis. Sec. Res. Doc. 2020/070 iv +
+204 p.
+(\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2020/2020_070-eng.html}{English
+version},
+\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2020/2020_070-fra.html}{French
+version}).
 \item[2019]
-\textbf{Anderson, S.C.}, Keppel, E.A., Edwards, A.M. 2019. A
-reproducible data synopsis for over 100 species of British Columbia
-groundfish. DFO Can. Sci. Advis. Sec. Res. Doc. 2019/041. vii + 321 p.
+\textbf{Anderson, S.C.}, E.A. Keppel, A.M. Edwards, 2019. A reproducible
+data synopsis for over 100 species of British Columbia groundfish. DFO
+Can. Sci. Advis. Sec. Res. Doc. 2019/041. vii + 321 p.
 (\href{http://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2019/2019_041-eng.html}{English
@@
 \href{http://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2019/2019_041-fra.html}{French
+version}).
+\end{description}
+
+\hypertarget{dfo-technical-reports}{%
+\section{DFO Technical Reports}\label{dfo-technical-reports}}
+
+\begin{description}
+\tightlist
+\item[2019]
+\textbf{Anderson, S.C.}, and Workman, G.D. A review of groundfish
+surveys in 2018 and an introduction to the groundfish data synopsis
+report. In State of the physical, biological and selected fishery
+resources of Pacific Canadian marine ecosystems in 2018. Edited by J.L.
+Boldt, J. Leonard, and P.C. Chandler. pp. 90--94. Can. Tech. Rep.\ Fish.
+Aquat. Sci. 3314, Sidney, BC.
+(\href{https://www.dfo-mpo.gc.ca/oceans/publications/soto-rceo/2018/index-eng.html}{English
+version})
+\item[2020]
+English, P.A., \textbf{Anderson, S.C.}, and Workman, G.D. A review of
+groundfish surveys in 2019. In State of the physical, biological and
+selected fishery resources of Pacific Canadian Marine Ecosystems in
+2019. Edited by J.L. Boldt, A. Javorski, and P.C. Chandler.
+pp.\ 102--106. Can. Tech. Rep.\ Fish. Aquat. Sci. 3377, Sidney, BC.
+(\href{https://www.dfo-mpo.gc.ca/oceans/publications/soto-rceo/2019/index-eng.html}{English
 version})
 \item[2018]
-Forrest, R.E., \textbf{S.C. Anderson}, C.J. Grandin, P.J. Starr.
-Assessment of Pacific Cod (\emph{Gadus macrocephalus}) for Hecate Strait
-and Queen Charlotte Sound (Area 5ABCD), and West Coast Vancouver Island
-(Area 3CD) in 2018. Canadian Science Advisory Secretariat (CSAS)
-Research Document. Accepted at regional peer-review meeting October
-2018. In press.
+Edwards, A.M., Duplisea, D.E., Grinnell, M.H., \textbf{Anderson, S.C.},
+Grandin, C.J., Ricard, D., Keppel, E.A., Anderson, E.D., Baker, K.D.,
+Benoit, H.P., Cleary, J.S., Connors, B.M., Desgagnes, M., English, P.A.,
+Fishman, D.F., Freshwater, C., Hedges, K.J., Holt, C.A., Holt, K.R.,
+Kronlund, R., Mariscak, A., Obradovich, S.G., Patten, B.A., Rogers, B.,
+Rooper, C.N., Simpson, M.R., Surette, T.J., Tallman, R.F., Wheeland,
+L.J., Wor, C., and Zhu, X. 2018. Proceedings of the Technical Expertise
+in Stock Assessment (TESA) national workshop on ``Tools for transparent,
+traceable, and transferable assessments'' 27--30 November 2018 in
+Nanaimo, British Columbia. Can. Tech. Rep.\ Fish. Aquat. Sci. 3290,
+Nanaimo, BC.
+(\href{https://waves-vagues.dfo-mpo.gc.ca/Library/40750152.pdf}{English
+version})
 \end{description}
@@
 \tightlist
+\item[2020--21]
+Co-principal Investigator on NOAA Office of Science and Technology
+NMFS-DFO grant ``Characterizing range shifts in groundfish species in
+response to local climate velocities''. An extension to apply methods
+from a similarly titled ACCASP grant below to NOAA waters. (\$30K USD)
+\item[2020--21]
+Collaborator on NOAA Office of Science and Technology NMFS-DFO grant
+``Groundfish distribution shifts between British Columbia, Canada, and
+Southeast Alaska, USA: climate drivers, ecosystem impacts and management
+implications''. (\$30K USD)
 \item[2019]
@@
 
-\begin{description}
+\setlist[itemize]{topsep=2pt}
+
+\textbf{covidseir}: \url{https://github.com/seananderson/covidseir}
+
+\begin{itemize}
 \tightlist
-\item[2019]
-\textbf{rosettafish}: An R package for translating fish- and
-fisheries-related terms.\\
-\url{https://github.com/pbs-assess/rosettafish}
-\item[2019]
-\textbf{sdmTMB}: An R package for fitting spatiotemporal species
-distribution 9GLMMs (generalized linear mixed effects models) with TMB
-(Template Model Builder)\\
-\url{https://github.com/pbs-assess/sdmTMB}
-\item[2019]
-\textbf{gfplot} \& \textbf{gfsynopsis}: R packages to generate an annual
-Research Document synopsis report showing all available fishery and
-biological data and basic model fits for 109 Pacific groundfish species
-for publication through the Canadian Science Advisory Secretariat
-(CSAS). To be reviewed via regional peer-review meeting November 22--23
-2018.\\
-\url{https://github.com/pbs-assess/gfplot} and
-\url{https://github.com/pbs-assess/gfsynopsis}
-\item[2019]
-\textbf{csasdown}: Reproducible Canadian Science Advisory Secretariat
-Research Documents via R Markdown.\\
-\url{https://github.com/pbs-assess/csasdown}
-\item[2018]
-\textbf{glmmfields}: Generalized linear mixed models with robust random
-fields for spatiotemporal modeling.\\
-\url{https://cran.r-project.org/package=glmmfields}
-\item[2017]
-\textbf{datalimited}: Stock assessment methods for data-limited
-fisheries. \url{https://github.com/datalimited/datalimited}
-\item[2014]
-\textbf{Anderson, S.C.}, J.W. Moore, M.M McClure, N.K. Dulvy, A.B.
-Cooper.\\
-\textbf{metafolio}: Metapopulation simulations for conserving salmon
-through portfolio optimization.
-\url{http://cran.r-project.org/package=metafolio}
-\item[2013]
-\textbf{Anderson, S.C.}, C.C. Monnahan, K.F. Johnson, K. Ono, J.L.
-Valero, C.J Cunningham, F. Hurtado-Ferro, R. Licandeo, C.R. McGilliard,
-C.S. Szuwalski, K.A. Vert-pre, A.R. Whitten. \textbf{ss3sim}: Fisheries
-stock assessment simulation testing with Stock Synthesis.
-\url{http://cran.r-project.org/package=ss3sim}
-\item[2013]
-\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. \textbf{ecofolio}:
-Tools to quantify metapopulation portfolio effects.
-\url{https://github.com/seananderson/ecofolio}
-\end{description}
+\item
+  Primary statistical model for COVID-19 forecasts for Canada (Public
+  Health Agency of Canada) since August 2020 and British Columbia since
+  April 2020
+\item
+  Fits a Bayesian SEIR (Susceptible, Exposed, Infectious, Recovered)
+  model to daily COVID-19 case data
+\item
+  Can accommodate multiple types of case data at once (e.g., reported
+  cases, hospitalizations, ICU admissions) and accounts for delays
+  between symptom onset and case appearance.
+\item
+  Can evaluate epidemic forecasts given changes to population contact
+  rates (e.g., levels of social distancing)
+\item
+  Published: Anderson et al.\ (2020; PLOS Computational Biology)
+\end{itemize}
 
-\hypertarget{invited-talks-and-conference-presentations}{%
-\section{Invited talks and conference
-presentations}\label{invited-talks-and-conference-presentations}}
+\textbf{sdmTMB}: \url{https://github.com/pbs-assess/sdmTMB}
+
+\begin{itemize}
+\tightlist
+\item
+  An R package for fitting spatiotemporal species distribution GLMMs
+  (generalized linear mixed effects models) with TMB (Template Model
+  Builder)
+\end{itemize}
+
+\textbf{glmmfields}: \url{https://cran.r-project.org/package=glmmfields}
+
+\begin{itemize}
+\tightlist
+\item
+  Bayesian generalized linear mixed models (GLMMs) with robust random
+  fields for spatiotemporal modeling
+\item
+  Allows identification of extreme changes in spatial distribution
+  (e.g., rapid range contractions)
+\item
+  Published: Anderson and Ward (2019; Ecology)
+\end{itemize}
+
+\textbf{bayesdfa}: \url{https://cran.r-project.org/package=bayesdfa}
+
+\begin{itemize}
+\tightlist
+\item
+  R package to fit Bayesian Dynamic Factor Analysis (DFA) time series
+  models (co-developed with E.J. Ward)
+\item
+  Can identify common trends, extreme events, or regime shifts in
+  multiple ecological timeseries
+\item
+  Published: Ward et al.\ (2019; The R Journal)
+\end{itemize}
+
+\textbf{csasdown}: \url{https://github.com/pbs-assess/csasdown}
+
+\begin{itemize}
+\tightlist
+\item
+  R package for reproducible Canadian Science Advisory Secretariat
+  Research Documents, Science Responses, and Technical Reports via R
+  Markdown
+\item
+  Enables fully reproducible CSAS reports going directly from databases
+  to analysis to written reports to bring CSAS reports in line with
+  modern data analysis best practices
+\end{itemize}
+
+\textbf{rosettafish}: \url{https://github.com/pbs-assess/rosettafish}
+
+\begin{itemize}
+\tightlist
+\item
+  Enables the seamless translation of fish- and fisheries-related terms
+  in R code to enable reproducible and efficient production of bilingual
+  CSAS Research Documents and Science Responses
+\end{itemize}
+
+\textbf{gfplot, gfdata, gfsynopsis, and gfdlm}:
+\url{https://github.com/pbs-assess/}
+
+\begin{itemize}
+\tightlist
+\item
+  R packages to facilitate rapid and reproducible data extraction, model
+  fitting, and plotting of BC groundfish data
+\item
+  gfsynopsis is used to generate an annual Research Document synopsis
+  report showing all available fishery and biological data and model
+  fits for 113 species
+\item
+  gfdlm facilitates applying the management strategy evaluation toolkit
+  `MSEtool' to BC groundfish
+\end{itemize}
+
+\textbf{ss3sim}: \url{https://github.com/ss3sim/ss3sim}
+
+\begin{itemize}
+\tightlist
+\item
+  R package for flexible, rapid, and reproducible fisheries stock
+  assessment simulation testing with the widely used Stock Synthesis
+  statistical age-structured stock assessment framework
+\item
+  Published: Anderson et al.\ (2014; PLOS ONE)
+\end{itemize}
+
+\hypertarget{selected-first-author-invited-conference-presentations}{%
+\section{Selected first-author invited conference
+presentations}\label{selected-first-author-invited-conference-presentations}}
 
@@
 American Fisheries Society Annual Meeting, Portland, OR, United States.
-\item[2013]
-\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. False prophets: The
-challenges of quantifying ecological portfolios
-(\href{http://seananderson.ca/talks/2013/PE_SAFS_quantsem.pdf}{slides}),
-Quantitative Seminar, School of Aquatic and Fishery Sciences, University
-of Washington, Seattle, WA, United States.
-\item[2012]
-\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. False prophets: The
-ecological portfolio effect overestimates the benefit of diversity.
-School of Aquatic and Fishery Sciences, University of Washington,
-Seattle, WA, United States.
-\item[2012]
-\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Metapopulation
-dynamics and the generality of the ecological portfolio effect. Les
-Ecologistes Departmental Seminar, Simon Fraser University, Burnaby, BC,
-Canada.
-\item[2010]
-\textbf{Anderson, S.C.} The rise of invertebrates, the fall of sea
-cucumbers, and the risk of maturing late. Earth2Ocean Research Group,
-Simon Fraser University, Burnaby, BC, Canada.
-\item[2009]
-\textbf{Anderson, S.C.}, J.E. Mills Flemming, R. Watson, H.K. Lotze.
-Ecosystem impacts of the global expansion of invertebrate fisheries.
-International Oceans Past II Conference \emph{Multidisciplinary
-Perspectives on the History and Future of Marine Animal Populations}.
-Vancouver, BC, Canada.
-\item[2009]
-\textbf{Anderson, S.C.}, J.E Mills Flemming, R. Watson, H.K. Lotze.
-Global invertebrate fisheries: trends and consequences. NCEAS (National
-Center for Ecological Analysis and Synthesis) Working Group
-\emph{Finding Common Ground in Marine Conservation and Management}.
-Santa Barbara, CA, USA.
-\item[2009]
-\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
-knowledge base for expanding low-trophic level fisheries in Atlantic
-Canada. Harvest Fisheries Seminar Series. Bedford Institute of
-Oceanography, Fisheries and Oceans, Dartmouth, NS, Canada.
-\item[2008]
-\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
-knowledge base for expanding low-trophic level fisheries in Atlantic
-Canada. Departmental Seminar, Biology Department, Dalhousie University,
-Canada.
 \end{description}
@@
 \tightlist
+\item[2020]
+Two webinars for the national TESA (Technical Expertise in Stock
+Assessment) 2019--2020 webinar series: (1) data visualization with
+ggplot2 and (2) efficiently iterating with code using the purrr package
 \item[2019]
@@
 \url{https://github.com/seananderson/adv-r-course}
+\item[2019]
+Organizer of \href{https://adv-r.hadley.nz/}{Advanced R} book club at
+Pacific Biological Station. Organized seven twice-a-month meetings of
+DFO Biologists and Research Scientists to work through the book and
+clarify exercise material.
 \item[2018]
@@
 
-\hypertarget{contract-employment}{%
-\section{Contract employment}\label{contract-employment}}
-
-\begin{description}
-\tightlist
-\item[2014--15]
-Data analysis in R and C++ for Ecoscape Environmental Consultants Ltd.
-\item[2013]
-Co-developed global fisheries status maps for the
-\href{http://ramlegacy.org/}{RAM Legacy Stock Assessment Database} for
-R. Hilborn
-\item[2010]
-Assessment of Arctic surfclam on Banquereau Bank using
-industry-collected data for Clearwater Seafoods, Bedford, NS, Canada.
-\item[2010]
-Modelling length-frequency distribution changes of Atlantic halibut for
-NAFO Divisions 4VWX5Z and 3NOP for Dr.\ K. Trzcinski, Population Ecology
-Division, Bedford Institute of Oceanography, Fisheries and Oceans,
-Dartmouth, NS, Canada.
-\end{description}
-
 \hypertarget{working-groups-and-workshops}{%
@@
 \tightlist
-\item[2019]
+\item[2020--23]
+Joint ICES/PICES Working Group on impacts of warming on growth rates and
+fisheries yields (WGGRAFY).
+\item[2020--22]
+TNC (The Nature Conservancy, based in the California office) working
+group on climate change impacts on North American west coast fisheries
+\item[2020]
+Ocean Frontier Institute/Canadian Statistical Sciences Institute
+(CANSSI) working group on \emph{Spatial Modelling of Fishery Survey Data
+to Improve Stock Assessment}; invited keynote speaker
+\item[2019--20]
 NOAA (National Oceanic and Atmospheric Administration) FATE (Fisheries
@@
 advice based on model ensembles.
+\item[2018]
+Galway-Mayo Institute of Technology (Ireland) working group on ecosystem
+tipping points
 \item[2016--17]
@@
 (\href{http://www.nceas.ucsb.edu/projects/12307}{URL})
-\item[2008]
-Workshop on Canadian Science and Management Strategies for Sea Cucumber
-(\emph{Cucumaria frondosa}), Fisheries and Oceans, Dartmouth, NS, Canada
-(work presented \emph{in absentia}).
-\item[2007]
-Workshop on Canadian Science and Management Strategies for Atlantic
-Hagfish, Fisheries and Oceans, Dartmouth, NS, Canada.
-(\href{http://www.dfo-mpo.gc.ca/CSAS/Csas/Publications/Pro-CR/2009/2009_009_e.htm}{URL})
 \end{description}
@@
 
-Reviewer for Science, Ecology, Canadian Journal of Fisheries and Aquatic
-Sciences, Fish and Fisheries, Ecological Applications, Conservation
-Biology, Proceedings B, Marine Policy, Population Ecology, Journal of
-Applied Ecology, International Journal of Tropical Biology and
-Conservation, Biological Conservation, Oikos, Journal of Environmental
-Management, Endangered Species Research, Aquatic Conservation
+Reviewer for Science, Canadian Journal of Fisheries and Aquatic
+Sciences, Fish and Fisheries, Fisheries Research, ICES Journal of Marine
+Science, Ecological Applications, Conservation Biology, Proceedings B,
+Ecology, Marine Policy, Population Ecology, Journal of Applied Ecology,
+International Journal of Tropical Biology and Conservation, Biological
+Conservation, Diversity and Distributions, Oikos, Journal of
+Environmental Management, Endangered Species Research, Aquatic
+Conservation, NSERC Discovery Grants

### 2020-11-11 cv/cv-body.tex
Update CV

@@
 \item[2020]
+\textbf{Anderson, S.C.}, P.R. Elsen, B.B. Hughes, R.K. Tonietto, M.C.
+Bletz, D.A. Gill, M.A.\ Holgerson, S.E. Kuebbing, C. McDonough MacKenzie,
+M.H. Meek, D. Veríssimo. Trends in ecology and conservation over eight
+decades. In press at Frontiers in Ecology and the Environment.
+\item[2020]
 \textbf{Anderson, S.C.}, A.M. Edwards, M. Yerlanov, N. Mulberry, J.
 Stockdale, S.A. Iyaniwura, R.C. Falcao, M.C. Otterstatter, M.A.\ Irvine,
-N.Z. Janjua, D. Coombs, C. Colijn. Estimating the impact of COVID-19
+N.Z. Janjua, D. Coombs, C. Colijn. Quantifying the impact of COVID-19
 control measures using a Bayesian model of physical distancing. In press
-at PLOS Computational Biology. Preprint: medRxiv 2020.04.17.20070086.
+at PLOS Computational Biology. medRxiv preprint:
 \url{https://doi.org/10.1101/2020.04.17.20070086}.
 \item[2020]
-Cahill, C.L., \textbf{S.C. Anderson}, A.J. Paul, L. MacPherson, M.G.
-Sullivan, B.T. van Poorten, C.J. Walters, and J.R. Post. A
-spatial-temporal approach to modeling somatic growth across inland
-recreational fisheries landscapes. In press at Canadian Journal of
-Fisheries and Aquatic Sciences.
-\url{https://doi.org/10.1139/cjfas-2019-0434}.
+Maureaud, A., R. Frelat, L. Pecuchet, N.L. Shackell, B. Merigot, M.L.
+Pinsky, K. Amador, \textbf{S.C. Anderson}, \ldots{} (67 co-authors),
+J.T. Thorson. Are we ready to track climate-driven shifts in marine
+species across international boundaries? A global survey of scientific
+bottom trawl data. In press at Global Change Biology.
+\url{https://doi.org/10.1111/gcb.15404}.
 \item[2020]
-Litzow, M.A., M.E. Hunsicker, E.J. Ward, \textbf{S.C. Anderson}, J. Gao,
-S. Zador, S. Batten, S. Dressel, J. Duffy-Anderson, E. Fergusson, R.
-Hopcroft, B.J. Laurel, and R. O'Malley. Evaluating ecosystem change as
-Gulf of Alaska temperature exceeds the limits of preindustrial
-variability. Progress in Oceanography. 186:102393.
-\url{https://doi.org/10.1016/j.pocean.2020.102393}.
+Cahill, C.L., \textbf{S.C. Anderson}, A.J. Paul. L. Macphearson, M.G.
+Sullivan, B. van Poorten, C. Walters, J.R. Post. A spatial-temporal
+approach to modeling somatic growth across recreational fisheries
+landscapes. Canadian Journal of Fisheries and Aquatic Science.
+77(11):1822--1835. \url{https://doi.org/10.1139/cjfas-2019-0434}.
+\item[2020]
+Litzow, M.A., M.E. Hunsicker, E.J. Ward, \textbf{S.C. Anderson}, J. Gao
+, S. Zador, S. Batten , S. Dressel, J. Duffy-Anderson, E. Fergusson, R.
+Hopcroft, B.J. Laurel, R. O'Malley Evaluating ecosystem change as Gulf
+of Alaska temperature exceeds the limits of preindustrial variability.
+Progress in Oceanography. 186:102393.
+\url{https://doi.org/10.1016/j.pocean.2020.102393}
 \item[2020]
@@
 visualization of raw fisheries data for 113 species improves
-transparency, assessment efficiency, and monitoring. In press at
-Fisheries. \url{https://doi.org/10.1002/fsh.10441}.
+transparency, assessment efficiency, and monitoringy. Fisheries.
+45:535--543. \url{https://doi.org/10.1002/fsh.10441}.
+(\href{https://www.dropbox.com/s/y8b0rlxvnodngf8/Anderson_etal_2020_reproducible_visualization.pdf?dl=1}{PDF})
 \item[2020]

### 2021-01-20 cv/cv-body.tex
Add preprint link
Add paper
Add doi
Add MP SAR

@@
 \item[2020]
+Barnett, L.A.K., E.J. Ward, \textbf{S.C. Anderson}. Improving estimates
+of species distribution change by incorporating local trends. In press
+at Ecography. \url{https://doi.org/10.1111/ecog.05176}.
+\item[2020]
 \textbf{Anderson, S.C.}, P.R. Elsen, B.B. Hughes, R.K. Tonietto, M.C.
@@
 control measures using a Bayesian model of physical distancing. In press
-at PLOS Computational Biology. medRxiv preprint:
+at PLOS Computational Biology:
+\url{https://doi.org/10.1371/journal.pcbi.1008274}. medRxiv preprint:
 \url{https://doi.org/10.1101/2020.04.17.20070086}.
@@
 45:535--543. \url{https://doi.org/10.1002/fsh.10441}.
-(\href{https://www.dropbox.com/s/y8b0rlxvnodngf8/Anderson_etal_2020_reproducible_visualization.pdf?dl=1}{PDF})
+(\href{https://www.dropbox.com/s/4mmnomvmpg0dbky/Anderson_etal_2020_reproducible_visualization_preprint.pdf?dl=1}{Preprint
+PDF})
 \item[2020]
@@
 version}).
+\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2021/2021_002-eng.html}{Associated
+Science Advisory Report}.
 \item[2020]

### 2021-02-12 cv/cv-body.tex
Add preprint
CV and typo

@@
 \tightlist
+\item[2021]
+Research Scientist, Stock Assessment and Research Division, Pacific
+Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
 \item[2017--20]
@@
 Fulbright Scholar to the University of Washington in 2012--13\\
-\href{https://www.sfu.ca/dean-gradstudies/blog/year/2015/06/SeanAnderson.html}{Dean
-of Graduate Studies Convocation Medal}
+Dean of Graduate Studies Convocation Medal
 \item[2008--10]
@@
 fisheries\\
-\href{http://goo.gl/nA1zE}{Canadian Governor General's Academic Gold
-Medal}
+Canadian Governor General's Academic Gold Medal
 \item[2004--07]
@@
 \tightlist
+\item[2021]
+Stockdale, J.E., \textbf{Anderson, S.C.}, Edwards, A.M., Iyaniwura,
+S.A., Mulberry, N., Otterstatter, M.C., Janjua, N.Z., Coombs, D.,
+Colijn, C., and Irvine, M.A.\ 2021. Quantifying transmissibility of
+COVID-19 and impact of intervention within long-term health care
+facilities. medRxiv: 2021.02.01.21249903.
+\url{https://doi.org/10.1101/2021.02.01.21249903}.
 \item[2020]
@@
 visualization of raw fisheries data for 113 species improves
-transparency, assessment efficiency, and monitoringy. Fisheries.
+transparency, assessment efficiency, and monitoring. Fisheries.
 45:535--543. \url{https://doi.org/10.1002/fsh.10441}.

### 2021-02-12 cv/cv-body.tex
CV years

@@
 Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
-\item[2019--20]
+\item[2019--21]
 Associate Editor, the Canadian Journal of Fisheries and Aquatic Sciences
-\item[2020--20]
+\item[2020--21]
 Adjunct Professor, Department of Mathematics, Simon Fraser University

### 2021-05-04 cv/cv-body.tex
Update CV
Add DOI

@@
 \item[2021]
+English, P., E.J. Ward, C.N. Rooper, R.E. Forrest, L.A. Rogers, K.L.
+Hunter, A.M. Edwards, B.M. Connors, S.C. Anderson. Contrasting climate
+velocity impacts in warm and cool locations: A meta-analysis across 38
+demersal fish species in the northeast Pacific. EcoEvoRxiv.
+\url{https://doi.org/10.32942/osf.io/b87ng}.
+\item[2021]
 Stockdale, J.E., \textbf{Anderson, S.C.}, Edwards, A.M., Iyaniwura,
 S.A., Mulberry, N., Otterstatter, M.C., Janjua, N.Z., Coombs, D.,
-Colijn, C., and Irvine, M.A.\ 2021. Quantifying transmissibility of
-COVID-19 and impact of intervention within long-term health care
-facilities. medRxiv: 2021.02.01.21249903.
+Colijn, C., and Irvine, M.A.\ Quantifying transmissibility of COVID-19
+and impact of intervention within long-term health care facilities.
+medRxiv: 2021.02.01.21249903.
 \url{https://doi.org/10.1101/2021.02.01.21249903}.
-\item[2020]
-\textbf{Anderson, S.C.}, Mulberry, N., Edwards, A.M., Stockdale, J.E.,
-Iyaniwura, S.A., Falcao, R.C., Otterstatter, M.C., Janjua, N.Z., Coombs,
-D., and Colijn, C. 2020. How much leeway is there to relax COVID-19
-control measures? medRxiv 2020.06.12.20129833.
-\url{https://doi.org/10.1101/2020.06.12.20129833}.
 \end{description}
@@
 \tightlist
+\item[2021]
+Freshwater C., \textbf{S.C. Anderson}, T.D. Beacham, W. Luedke, C. Wor,
+J. King. An integrated model of seasonal changes in stock composition
+and abundance with an application to Chinook salmon. PeerJ 9:e11163.
+\url{https://doi.org/10.7717/peerj.11163}.
+\item[2021]
+\textbf{Anderson, S.C.}, P.R. Elsen, B.B. Hughes, R.K. Tonietto, M.C.
+Bletz, D.A. Gill, M.A.\ Holgerson, S.E. Kuebbing, C. McDonough MacKenzie,
+M.H. Meek, D. Veríssimo. Trends in ecology and conservation over eight
+decades. In press at Frontiers in Ecology and the Environment.
+\url{http://doi.org/10.1002/fee.2320}.
+\item[2021]
+\textbf{Anderson, S.C.}, Mulberry, N., Edwards, A.M., Stockdale, J.E.,
+Iyaniwura, S.A., Falcao, R.C., Otterstatter, M.C., Janjua, N.Z., Coombs,
+D., and Colijn, C. 2020. How much leeway is there to relax COVID-19
+control measures? In press at Epidemics.
+\url{https://doi.org/10.1016/j.epidem.2021.100453}.
+\end{description}
+
+\begin{description}
+\tightlist
 \item[2020]
@@
 at Ecography. \url{https://doi.org/10.1111/ecog.05176}.
-\item[2020]
-\textbf{Anderson, S.C.}, P.R. Elsen, B.B. Hughes, R.K. Tonietto, M.C.
-Bletz, D.A. Gill, M.A.\ Holgerson, S.E. Kuebbing, C. McDonough MacKenzie,
-M.H. Meek, D. Veríssimo. Trends in ecology and conservation over eight
-decades. In press at Frontiers in Ecology and the Environment.
 \item[2020]

### 2021-05-25 cv/cv-body.tex
Add to CV

@@
 English, P., E.J. Ward, C.N. Rooper, R.E. Forrest, L.A. Rogers, K.L.
-Hunter, A.M. Edwards, B.M. Connors, S.C. Anderson. Contrasting climate
-velocity impacts in warm and cool locations: A meta-analysis across 38
-demersal fish species in the northeast Pacific. EcoEvoRxiv.
+Hunter, A.M. Edwards, B.M. Connors, \textbf{S.C. Anderson}. Contrasting
+climate velocity impacts in warm and cool locations: A meta-analysis
+across 38 demersal fish species in the northeast Pacific. EcoEvoRxiv.
 \url{https://doi.org/10.32942/osf.io/b87ng}.
@@
 \tightlist
+\item[2021]
+Evans R., P.A. English , \textbf{S.C. Anderson}, S. Gauthier, C.L.K.
+Robinson. Factors affecting the seasonal distribution and biomass of
+\emph{E. pacifica} and \emph{T. spinifera} along the Pacific coast of
+Canada: A spatiotemporal modelling approach. PLOS ONE. 2021 May
+14;16(5):e0249818. \url{https://doi.org/10.1371/journal.pone.0249818}.
 \item[2021]
@@
 \url{https://doi.org/10.1016/j.epidem.2021.100453}.
-\end{description}
-
-\begin{description}
-\tightlist
 \item[2020]

### 2021-10-14 cv/cv-body.tex
Move preprint to published

@@
 
-\hypertarget{preprints}{%
-\section{Preprints}\label{preprints}}
+\hypertarget{publications}{%
+\section{Publications}\label{publications}}
 
@@
 Hunter, A.M. Edwards, B.M. Connors, \textbf{S.C. Anderson}. Contrasting
-climate velocity impacts in warm and cool locations: A meta-analysis
-across 38 demersal fish species in the northeast Pacific. EcoEvoRxiv.
-\url{https://doi.org/10.32942/osf.io/b87ng}.
-\item[2021]
-Stockdale, J.E., \textbf{Anderson, S.C.}, Edwards, A.M., Iyaniwura,
-S.A., Mulberry, N., Otterstatter, M.C., Janjua, N.Z., Coombs, D.,
-Colijn, C., and Irvine, M.A.\ Quantifying transmissibility of COVID-19
-and impact of intervention within long-term health care facilities.
-medRxiv: 2021.02.01.21249903.
-\url{https://doi.org/10.1101/2021.02.01.21249903}.
-\end{description}
-
-\hypertarget{publications}{%
-\section{Publications}\label{publications}}
-
-\begin{description}
-\tightlist
+climate velocity impacts in warm and cool locations show that effects of
+marine warming are worse in already warmer temperate waters. In press at
+Fish and Fisheries. \url{https://doi.org/10.1111/faf.12613}. Open access
+preprint: \url{https://doi.org/10.32942/osf.io/b87ng}.
 \item[2021]

### 2022-01-13 cv/cv-body.tex
Update some pubs

@@
 
+\hypertarget{preprints}{%
+\section{Preprints}\label{preprints}}
+
+\begin{description}
+\tightlist
+\item[2021]
+\textbf{Anderson, S.C}, B.M. Connors, P.A. English, R.E. Forrest, R.
+Haigh, K.R. Holt. Trends in Pacific Canadian groundfish stock status.
+bioRxiv 2021.12.13.472502.
+\url{https://doi.org/10.1101/2021.12.13.472502}.
+\end{description}
+
 \hypertarget{publications}{%
@@
 \tightlist
+\item[2022]
+Stockdale, J.E., \textbf{S.C. Anderson}, A.M Edwards, S.A. Iyaniwura, N.
+Mulberry, M.C. Otterstatter, N.Z. Janjua, D. Coombs, C. Colijn, and
+M.A.\ Irvine. Quantifying transmissibility of SARS-CoV-2 and impact of
+intervention within long-term healthcare facilities. Royal Society Open
+Science 9(1): 211710. \url{https://doi.org/10.1098/rsos.211710}.
+\item[2022]
+Commander C.J.C, L.A.K. Barnett, E.J. Ward, \textbf{S.C. Anderson}, T.E.
+Essington. The shadow model: how and why small choices in spatially
+explicit species distribution models affect prediction. In press at
+PeerJ.
+\item[2022]
+Ward, E.J., \textbf{S.C. Anderson}, M.D.\ Hunsicker, and M.A.\ Litzow.
+Smoothed dynamic factor analysis for identifying trends in multivariate
+time series. Methods in Ecology and Evolution. In press.
+\url{https://doi.org/10.1111/2041-210X.13788}.
 \item[2021]
@@
 climate velocity impacts in warm and cool locations show that effects of
-marine warming are worse in already warmer temperate waters. In press at
-Fish and Fisheries. \url{https://doi.org/10.1111/faf.12613}. Open access
-preprint: \url{https://doi.org/10.32942/osf.io/b87ng}.
+marine warming are worse in already warmer temperate waters. Fish and
+Fisheries. 23(1) 239--255. \url{https://doi.org/10.1111/faf.12613}. Open
+access preprint: \url{https://doi.org/10.32942/osf.io/b87ng}.
 \item[2021]
@@
 M.H. Meek, D. Veríssimo. Trends in ecology and conservation over eight
-decades. In press at Frontiers in Ecology and the Environment.
+decades. Frontiers in Ecology and the Environment. 19(5): 274--282.
 \url{http://doi.org/10.1002/fee.2320}.
@@
 \url{https://doi.org/10.1016/j.epidem.2021.100453}.
-\item[2020]
+\item[2021]
 Barnett, L.A.K., E.J. Ward, \textbf{S.C. Anderson}. Improving estimates
-of species distribution change by incorporating local trends. In press
-at Ecography. \url{https://doi.org/10.1111/ecog.05176}.
+of species distribution change by incorporating local trends. Ecography.
+44: 427--439. \url{https://doi.org/10.1111/ecog.05176}.
 \item[2020]
@@
 N.Z. Janjua, D. Coombs, C. Colijn. Quantifying the impact of COVID-19
-control measures using a Bayesian model of physical distancing. In press
-at PLOS Computational Biology:
+control measures using a Bayesian model of physical distancing. PLOS
+Computational Biology. 16(12): e1008274.
 \url{https://doi.org/10.1371/journal.pcbi.1008274}. medRxiv preprint:
@@
 species across international boundaries? A global survey of scientific
-bottom trawl data. In press at Global Change Biology.
+bottom trawl data. Global Change Biology. 27: 220--236.
 \url{https://doi.org/10.1111/gcb.15404}.
@@
 \tightlist
+\item[2021]
+\textbf{Anderson, S.C.}, R.E. Forrest, Q.C. Huynh, E.A. Keppel. A
+management procedure framework for groundﬁsh in British Columbia. DFO
+Can. Sci. Advis. Sec. Res. Doc. 2021/007.
+(\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2021/2021_007-eng.html}{Research
+Document}).
+\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2021/2021_002-eng.html}{Associated
+Science Advisory Report}.
 \item[2020]
@@
 version}).
-\item[2020]
-\textbf{Anderson, S.C.}, R.E. Forrest, Q.C. Huynh, E.A. Keppel. A
-management procedure framework for groundﬁsh in British Columbia.
-Accepted as a DFO Can. Sci. Advis. Sec. Res. Doc. on June 9 2020.
-(\href{https://www.dropbox.com/s/rgxk2wco5v5zp0e/gfmp-2020-03-13.pdf?dl=1}{English
-version}).
-\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2021/2021_002-eng.html}{Associated
-Science Advisory Report}.
 \item[2020]
@@
 \tightlist
+\item[2020]
+English, P.A., \textbf{Anderson, S.C.}, and Workman, G.D. A review of
+groundfish surveys in 2019. In State of the physical, biological and
+selected fishery resources of Pacific Canadian Marine Ecosystems in
+2019. Edited by J.L. Boldt, A. Javorski, and P.C. Chandler.
+pp.\ 102--106. Can. Tech. Rep.\ Fish. Aquat. Sci. 3377, Sidney, BC.
+(\href{https://www.dfo-mpo.gc.ca/oceans/publications/soto-rceo/2019/index-eng.html}{English
+version})
 \item[2019]
@@
 (\href{https://www.dfo-mpo.gc.ca/oceans/publications/soto-rceo/2018/index-eng.html}{English
-version})
-\item[2020]
-English, P.A., \textbf{Anderson, S.C.}, and Workman, G.D. A review of
-groundfish surveys in 2019. In State of the physical, biological and
-selected fishery resources of Pacific Canadian Marine Ecosystems in
-2019. Edited by J.L. Boldt, A. Javorski, and P.C. Chandler.
-pp.\ 102--106. Can. Tech. Rep.\ Fish. Aquat. Sci. 3377, Sidney, BC.
-(\href{https://www.dfo-mpo.gc.ca/oceans/publications/soto-rceo/2019/index-eng.html}{English
 version})

### 2022-01-13 cv/cv-body.tex
Update dates in CV

@@
 \tightlist
-\item[2021]
+\item[2021--22]
 Research Scientist, Stock Assessment and Research Division, Pacific
 Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
+\item[2019--22]
+Associate Editor, the Canadian Journal of Fisheries and Aquatic Sciences
+\item[2020--22]
+Adjunct Professor, Department of Mathematics, Simon Fraser University
 \item[2017--20]
@@
 Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
-\item[2019--21]
-Associate Editor, the Canadian Journal of Fisheries and Aquatic Sciences
-\item[2020--21]
-Adjunct Professor, Department of Mathematics, Simon Fraser University
 \item[2015--17]

### 2022-04-20 cv/cv-body.tex
Update CV

@@
 \tightlist
+\item[2022]
+\textbf{Anderson, S.C.}, E.J. Ward, P.A. English, L.A.K. Barnett.
+sdmTMB: an R package for fast, flexible, and user-friendly generalized
+linear mixed effects models with spatial and spatiotemporal random
+fields. bioRxiv 2022.03.24.485545.
+\url{https://doi.org/10.1101/2022.03.24.485545}.
 \item[2021]
@@
 \item[2022]
+Essington, T. E. \textbf{S.C. Anderson}, L.A.K. Barnett, H.M. Berger,
+S.A. Siedlecki, E.J. Ward. Advancing statistical models to reveal the
+effect of dissolved oxygen on the spatial distribution of marine taxa
+using thresholds and a physiologically based index. In press at
+Ecography. \url{https://doi.org/10.1111/ecog.06249}.
+\item[2022]
+Commander, C.J.C, L.A.K. Barnett, E.J. Ward, \textbf{S.C. Anderson},
+T.E. Essington. The shadow model: how and why small choices in spatially
+explicit species distribution models affect predictions. PeerJ 10:
+e12783. \url{https://doi.org/10.7717/peerj.12783}.
+\item[2022]
 Stockdale, J.E., \textbf{S.C. Anderson}, A.M Edwards, S.A. Iyaniwura, N.
-Mulberry, M.C. Otterstatter, N.Z. Janjua, D. Coombs, C. Colijn, and
-M.A.\ Irvine. Quantifying transmissibility of SARS-CoV-2 and impact of
+Mulberry, M.C. Otterstatter, N.Z. Janjua, D. Coombs, C. Colijn, and M.A.
+Irvine. Quantifying transmissibility of SARS-CoV-2 and impact of
 intervention within long-term healthcare facilities. Royal Society Open
 Science 9(1): 211710. \url{https://doi.org/10.1098/rsos.211710}.
-\item[2022]
-Commander C.J.C, L.A.K. Barnett, E.J. Ward, \textbf{S.C. Anderson}, T.E.
-Essington. The shadow model: how and why small choices in spatially
-explicit species distribution models affect prediction. In press at
-PeerJ.
 \item[2022]
@@
 \item[2021]
+DFO. Status Update of Pacifc Cod (\emph{Gadus macrocephalus}) for West
+Coast Vancouver Island (Area 3CD), and Hecate Strait and Queen Charlotte
+Sound (Area 5ABCD) in 2020. DFO Can. Sci. Advis. Sec. Sci. Resp.
+2021/002.
+(\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2021/2021_007-eng.html}{English
+version},
+\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2021/2021_007-fra.html}{French
+version}).
+\item[2021]
 \textbf{Anderson, S.C.}, R.E. Forrest, Q.C. Huynh, E.A. Keppel. A
@@
 Bresch, E.A. Keppel. Evaluation of potential rebuilding strategies for
-Inside Yelloweye Rockﬁsh (\emph{Sebastes ruberrimus}) in British
-Columbia. Accepted as a DFO Can. Sci. Advis. Sec. Res. Doc. on June 11
-2020.
-(\href{https://www.dropbox.com/s/pll8axzsxqka8j8/resdoc-iye.pdf?dl=1}{English
-version}).
+Inside Yelloweye Rockfish (\emph{Sebastes ruberrimus}) in British
+Columbia. In Press. Accepected as a DFO Can. Sci. Advis. Sec. Res. Doc.
+on June 11 2020.
 \item[2020]
@@
 \tightlist
+\item[2021]
+English, P.A., \textbf{Anderson, S.C.}, and Workman, G.D. A review of
+groundfish surveys in 2019. In State of the physical, biological and
+selected fishery resources of Pacific Canadian Marine Ecosystems in
+2019. Edited by J.L. Boldt, A. Javorski, and P.C. Chandler.
+pp.\ 127--133. Can. Tech. Rep.\ Fish. Aquat. Sci. 3434, Sidney, BC.
+(\href{https://publications.gc.ca/site/eng/9.900827/publication.html}{English
+version}).
 \item[2020]
@@
 (\href{https://www.dfo-mpo.gc.ca/oceans/publications/soto-rceo/2019/index-eng.html}{English
-version})
+version}).
 \item[2019]
@@
 (\href{https://www.dfo-mpo.gc.ca/oceans/publications/soto-rceo/2018/index-eng.html}{English
-version})
+version}).
 \item[2018]
@@
 (\href{https://waves-vagues.dfo-mpo.gc.ca/Library/40750152.pdf}{English
-version})
+version}).
 \end{description}
@@
 \tightlist
+\item[2022]
+Two workshops on geostatistical modelling with sdmTMB for the NOAA
+(National Oceanic and Atmospheric Administration) PSAW (Protected
+Species Assessment Workshop) series.
+\href{https://pbs-assess.github.io/sdmTMB-teaching/noaa-psaw-2022/}{Slides
+and exercises}, Lecture recordings:
+\href{https://youtu.be/DIXa7ngVVL0}{Week 1},
+\href{https://youtu.be/VxnqgiAAjfk}{Week 2}.
 \item[2020]

### 2023-01-03 cv/cv-body.tex
Add preprint
Add CCE paper
Add paper
Update CV

@@
 \tightlist
-\item[2021--22]
+\item[2021--23]
 Research Scientist, Stock Assessment and Research Division, Pacific
 Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
-\item[2019--22]
+\item[2019--23]
 Associate Editor, the Canadian Journal of Fisheries and Aquatic Sciences
-\item[2020--22]
+\item[2020--23]
 Adjunct Professor, Department of Mathematics, Simon Fraser University
@@
 \item[2022]
+Lindmark, M., \textbf{S.C. Anderson}, M. Gogina, M. Casini. Evaluating
+drivers of spatiotemporal individual condition of a bottom-associated
+marine fish. bioRxiv 2022.04.19.488709.
+\url{https://doi.org/10.1101/2022.04.19.488709}.
+\item[2022]
 \textbf{Anderson, S.C.}, E.J. Ward, P.A. English, L.A.K. Barnett.
@@
 \item[2022]
-Essington, T. E. \textbf{S.C. Anderson}, L.A.K. Barnett, H.M. Berger,
+Ward, E.J., L.A.K. Barnett, \textbf{S.C. Anderson}, C.J.C. Commander,
+T.E. Essington. Incorporating non-stationary spatial variability into
+dynamic species distribution models. ICES Journal of Marine Science.
+79(9): 2422-2429. \url{https://doi.org/10.1093/icesjms/fsac179}.
+\item[2022]
+Essington, T.E. \textbf{S.C. Anderson}, L.A.K. Barnett, H.M. Berger,
 S.A. Siedlecki, E.J. Ward. Advancing statistical models to reveal the
 effect of dissolved oxygen on the spatial distribution of marine taxa
-using thresholds and a physiologically based index. In press at
-Ecography. \url{https://doi.org/10.1111/ecog.06249}.
+using thresholds and a physiologically based index. Ecography 2022(8):
+e06249. \url{https://doi.org/10.1111/ecog.06249}.
+\item[2022]
+Thompson, P.L., \textbf{S.C. Anderson}, J. Nephin, D.R. Haggarty, M.
+Angelica Peña, P.A. English, K.S.P. Gale, E. Rubidge. Disentangling the
+impacts of environmental change and commercial fishing on demersal fish
+biodiversity in a northeast Pacific ecosystem. Marine Ecology Progress
+Series. 689:137-154. \url{https://doi.org/10.3354/meps14034}.
+\item[2022]
+Hunsicker, M.E., E.J. Ward, M.A.\ Litzow, \textbf{S.C. Anderson}, C.J.
+Harvey, J.C. Field, J. Gao, M.G. Jacox, S. Melin, A.R. Thompson, P.
+Warzybok. Tracking and forecasting community responses to climate
+perturbations in the California Current Ecosystem. PLOS Climate. 1(3):
+e0000014. \url{https://doi.org/10.1371/journal.pclm.0000014}.
 \item[2022]
@@
 T.E. Essington. The shadow model: how and why small choices in spatially
-explicit species distribution models affect predictions. PeerJ 10:
+explicit species distribution models affect predictions. PeerJ. 10:
 e12783. \url{https://doi.org/10.7717/peerj.12783}.
@@
 intervention within long-term healthcare facilities. Royal Society Open
-Science 9(1): 211710. \url{https://doi.org/10.1098/rsos.211710}.
+Science. 9(1): 211710. \url{https://doi.org/10.1098/rsos.211710}.
 \item[2022]
@@
 Smoothed dynamic factor analysis for identifying trends in multivariate
-time series. Methods in Ecology and Evolution. In press.
+time series. Methods in Ecology and Evolution. 13(4):908-918.
 \url{https://doi.org/10.1111/2041-210X.13788}.
@@
 marine warming are worse in already warmer temperate waters. Fish and
-Fisheries. 23(1) 239--255. \url{https://doi.org/10.1111/faf.12613}. Open
-access preprint: \url{https://doi.org/10.32942/osf.io/b87ng}.
+Fisheries. 23(1) 239--255. \url{https://doi.org/10.1111/faf.12613}.
+{[}\href{https://www.dropbox.com/s/cmkzlxgiez6rzc0/English_etal_2022_contrasting_climate_velocity.pdf?dl=1}{PDF}{]}
 \item[2021]
@@
 D., and Colijn, C. 2020. How much leeway is there to relax COVID-19
-control measures? In press at Epidemics.
+control measures? Epidemics. 35: 100453.
 \url{https://doi.org/10.1016/j.epidem.2021.100453}.
@@
 for assessed stocks: Responsive management advice and lower assessment
-frequency. In press at Fish and Fisheries.
+frequency. Fish and Fisheries. 21(3):663-679.
 \url{https://doi.org/10.1111/faf.12453}.
@@
 (\href{http://cran.r-project.org/web/packages/ss3sim/index.html}{R
-package}, \href{https://github.com/ss3sim/ss3sim}{code}.
+package}, \href{https://github.com/ss3sim/ss3sim}{code}).
 \item[2014]
@@
 
+2022 Grandin, C.J., \textbf{S.C. Anderson}, P.A. English. Arrowtooth
+Flounder (\emph{Atheresthes stomias}) Stock Assessment for the West
+Coast of British Columbia in 2021. Accepted DFO Can. Sci. Advis. Sec.
+Res. Doc.
+
+2022 DFO. A data synopsis for British Columbia groundfish: 2021 data
+update. DFO Can. Sci. Advis. Sec. Sci. Resp. 2022/020.
+(\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ScR-RS/2022/2022_020-eng.html}{English
+version},
+\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ScR-RS/2022/2022_020-fra.html}{French
+version}).
+
 \begin{description}
@@
 Science Advisory Report}.
-\item[2020]
+\item[2021]
 Haggarty D.R., Q.C. Huynh, R.E. Forrest, \textbf{S.C. Anderson}, M.J.
@@
 Inside Yelloweye Rockfish (\emph{Sebastes ruberrimus}) in British
-Columbia. In Press. Accepected as a DFO Can. Sci. Advis. Sec. Res. Doc.
-on June 11 2020.
+Columbia. DFO Can. Sci. Advis. Sec. Res. Doc. 2021/008.
+(\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2021/2021_008-eng.html}{Research
+Document}).
 \item[2020]
@@
 response to local climate velocities''. An extension to apply methods
-from a similarly titled ACCASP grant below to NOAA waters. (\$30K USD)
+from a similarly titled ACCASP grant below to NOAA waters.
 \item[2020--21]
@@
 Southeast Alaska, USA: climate drivers, ecosystem impacts and management
-implications''. (\$30K USD)
+implications''.
 \item[2019]
@@
 
+\textbf{sdmTMB}: \url{https://github.com/pbs-assess/sdmTMB}
+
+\begin{itemize}
+\tightlist
+\item
+  An R package for fitting spatiotemporal species distribution GLMMs
+  (generalized linear mixed effects models) with TMB (Template Model
+  Builder)
+\end{itemize}
+
 \textbf{covidseir}: \url{https://github.com/seananderson/covidseir}
@@
   Published: Anderson et al.\ (2020; PLOS Computational Biology)
-\end{itemize}
-
-\textbf{sdmTMB}: \url{https://github.com/pbs-assess/sdmTMB}
-
-\begin{itemize}
-\tightlist
-\item
-  An R package for fitting spatiotemporal species distribution GLMMs
-  (generalized linear mixed effects models) with TMB (Template Model
-  Builder)
 \end{itemize}

### 2023-01-04 cv/cv-body.tex
Fix CV formatting

@@
 
-\hypertarget{peer-reviewed-reports}{%
-\section{Peer-reviewed reports}\label{peer-reviewed-reports}}
+\hypertarget{peer-reviewed-government-reports}{%
+\section{Peer-reviewed government
+reports}\label{peer-reviewed-government-reports}}
 
-2022 Grandin, C.J., \textbf{S.C. Anderson}, P.A. English. Arrowtooth
-Flounder (\emph{Atheresthes stomias}) Stock Assessment for the West
-Coast of British Columbia in 2021. Accepted DFO Can. Sci. Advis. Sec.
-Res. Doc.
-
-2022 DFO. A data synopsis for British Columbia groundfish: 2021 data
-update. DFO Can. Sci. Advis. Sec. Sci. Resp. 2022/020.
+\begin{description}
+\tightlist
+\item[2022]
+Grandin, C.J., \textbf{S.C. Anderson}, P.A. English. Arrowtooth Flounder
+(\emph{Atheresthes stomias}) Stock Assessment for the West Coast of
+British Columbia in 2021. Accepted DFO Can. Sci. Advis. Sec. Res. Doc.
+\item[2022]
+DFO. A data synopsis for British Columbia groundfish: 2021 data update.
+DFO Can. Sci. Advis. Sec. Sci. Resp. 2022/020.
 (\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ScR-RS/2022/2022_020-eng.html}{English
@@
 version}).
-
-\begin{description}
-\tightlist
 \item[2021]
@@
 (\href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/ResDocs-DocRech/2021/2021_007-eng.html}{Research
-Document}).
+Document},
 \href{https://www.dfo-mpo.gc.ca/csas-sccs/Publications/SAR-AS/2021/2021_002-eng.html}{Associated
-Science Advisory Report}.
+Science Advisory Report}).
 \item[2021]
@@
 
-\hypertarget{dfo-technical-reports}{%
-\section{DFO Technical Reports}\label{dfo-technical-reports}}
+\hypertarget{government-technical-reports}{%
+\section{Government Technical
+Reports}\label{government-technical-reports}}
 

### 2023-04-28 cv/cv-body.tex
Add missing paper
Update cv

@@
 \tightlist
-\item[2022]
-Lindmark, M., \textbf{S.C. Anderson}, M. Gogina, M. Casini. Evaluating
-drivers of spatiotemporal individual condition of a bottom-associated
-marine fish. bioRxiv 2022.04.19.488709.
-\url{https://doi.org/10.1101/2022.04.19.488709}.
+\item[2023]
+Liu, O.R. E.J. Ward, \textbf{S.C. Anderson}, K.S. Andrews, L.A.K.
+Barnett, S. Brodie, G. Carroll, J. Fiechter, M.A.\ Haltuch, C.J. Harvey,
+E.L. Hazen, P.Y. Hernvann, M. Jacox, I.C. Kaplan, S. Matson, K. Norman,
+M.P. Buil, R.L. Selden, A.O. Shelton, J.F. Samhouri. Species
+redistribution creates unequal outcomes for multispecies fisheries under
+projected climate change.
+\url{https://doi.org/10.21203/rs.3.rs-2399110/v1}.
 \item[2022]
@@
 \url{https://doi.org/10.1101/2022.03.24.485545}.
+\item[2022]
+Lindmark, M., \textbf{S.C. Anderson}, M. Gogina, M. Casini. Evaluating
+drivers of spatiotemporal individual condition of a bottom-associated
+marine fish. bioRxiv 2022.04.19.488709.
+\url{https://doi.org/10.1101/2022.04.19.488709}.
 \item[2021]
@@
 \tightlist
+\item[2023]
+Free, C.M., \textbf{S.C. Anderson}, E.A. Hellmers, B.A. Muhling, M.O.
+Navarro, K. Richerson, L.A. Rogers, W.H. Satterthwaite, A.R. Thompson,
+J.M. Burt, S.D. Gaines, K.N. Marshall, J.W. White, and L.F. Bellquist.
+Impact of the 2014--2016 marine heatwave on US and Canada West Coast
+fisheries: Surprises and lessons from key case studies. Fish and
+Fisheries. In press. \url{https://doi.org/10.1111/faf.12753}.
+\item[2023]
+Thompson, P.L., \textbf{S.C. Anderson}, J. Nephin, C.K. Robb, B.
+Proudfoot, A.E. Park, D.R. Haggarty, E.M. Rubidge. Integrating trawl and
+longline surveys across British Columbia improves groundfish
+distribution predictions. Can. J. Fish. Aquat. Sci. 80(1): 195--210.
+\url{https://doi.org/10.1139/cjfas-2022-0108}.
 \item[2022]
@@
 Progress in Oceanography. 186:102393.
-\url{https://doi.org/10.1016/j.pocean.2020.102393}
+\url{https://doi.org/10.1016/j.pocean.2020.102393}.
 \item[2020]
@@
 \tightlist
+\item[2023]
+Barrett, T.J., J.R. Marentette, R.E. Forrest, S.C. Anderson, C.A. Holt,
+D.W. Ings, M.E. Thiess. Technical Considerations for Stock Status and
+Limit Reference Points under the Fish Stocks Provisions. Accepted DFO
+Can. Sci. Advis. Sec. Res. Doc.
 \item[2022]
@@
 
-\hypertarget{government-technical-reports}{%
-\section{Government Technical
-Reports}\label{government-technical-reports}}
+\hypertarget{technical-reports}{%
+\section{Technical Reports}\label{technical-reports}}
 
@@
 \tightlist
+\item[2023]
+ICES. Workshop on Unavoidable Survey Effort Reduction 2 (WKUSER2). ICES
+Scientific Report, 5(13).
+\url{https://doi.org/10.17895/ices.pub.22086845.v1}. Co-lead author of
+chapter ``TOR III. Modeling and simulations: Further develop model
+performance evaluation through simulations, use of auxiliary information
+to improve survey data products, including appropriate propagation of
+uncertainty''
 \item[2021]
@@
 Canadian Journal of Fisheries and Aquatic Sciences Stevenson
-Lectureship. \url{http://www.nrcresearchpress.com/page/cjfas/award} A
+Lectureship. \url{https://cdnsciencepub.com/journal/cjfas/about} A
 ``lectureship conferred upon a young, energetic and creative researcher

### 2023-08-18 cv/cv-body.tex
Max paper
Add doi
Move pub to published

@@
 \tightlist
+\item[2022]
+\textbf{Anderson, S.C.}, E.J. Ward, P.A. English, L.A.K. Barnett.
+sdmTMB: an R package for fast, flexible, and user-friendly generalized
+linear mixed effects models with spatial and spatiotemporal random
+fields. bioRxiv 2022.03.24.485545.
+\url{https://doi.org/10.1101/2022.03.24.485545}.
+\item[2021]
+\textbf{Anderson, S.C}, B.M. Connors, P.A. English, R.E. Forrest, R.
+Haigh, K.R. Holt. Trends in Pacific Canadian groundfish stock status.
+bioRxiv 2021.12.13.472502.
+\url{https://doi.org/10.1101/2021.12.13.472502}.
+\end{description}
+
+\hypertarget{publications}{%
+\section{Publications}\label{publications}}
+
+\begin{description}
+\tightlist
 \item[2023]
@@
 redistribution creates unequal outcomes for multispecies fisheries under
-projected climate change.
-\url{https://doi.org/10.21203/rs.3.rs-2399110/v1}.
-\item[2022]
-\textbf{Anderson, S.C.}, E.J. Ward, P.A. English, L.A.K. Barnett.
-sdmTMB: an R package for fast, flexible, and user-friendly generalized
-linear mixed effects models with spatial and spatiotemporal random
-fields. bioRxiv 2022.03.24.485545.
-\url{https://doi.org/10.1101/2022.03.24.485545}.
-\item[2022]
+projected climate change. Science Advances 9(33): eadg5468.
+\url{https://doi.org/10.1126/sciadv.adg5468}.
+\item[2023]
 Lindmark, M., \textbf{S.C. Anderson}, M. Gogina, M. Casini. Evaluating
-drivers of spatiotemporal individual condition of a bottom-associated
-marine fish. bioRxiv 2022.04.19.488709.
-\url{https://doi.org/10.1101/2022.04.19.488709}.
-\item[2021]
-\textbf{Anderson, S.C}, B.M. Connors, P.A. English, R.E. Forrest, R.
-Haigh, K.R. Holt. Trends in Pacific Canadian groundfish stock status.
-bioRxiv 2021.12.13.472502.
-\url{https://doi.org/10.1101/2021.12.13.472502}.
-\end{description}
-
-\hypertarget{publications}{%
-\section{Publications}\label{publications}}
-
-\begin{description}
-\tightlist
+drivers of spatiotemporal variability in individual condition of a
+bottom-associated marine fish, Atlantic cod (\emph{Gadus morhua}). ICES
+Journal of Marine Science. In press.
+\url{https://doi.org/10.1093/icesjms/fsad084}.
 \item[2023]

### 2023-09-14 cv/cv-body.tex
Update CV
Update bib
Add pubs

@@
 \tightlist
+\item[2023]
+Yalcin, S., \textbf{S.C. Anderson}, P.M. Regular, P.A. English.
+Exploring the limits of spatiotemporal and design-based index
+standardization under reduced survey coverage. In press at ICES Journal
+of Marine Science.
+\item[2023]
+K.L. Blake, D. Veríssimo, \textbf{S.C. Anderson}, A. Gleave. Do species
+receive more attention when named after celebrities? In press at
+Conservation Biology. Preprint:
+\url{https://doi.org/10.31235/osf.io/wgbpv}.
+\item[2023]
+J. Nephin, P.L. Thompson, \textbf{S.C. Anderson}, A.E. Park, C.N.
+Rooper, B. Aulthouse, J. Watson. Integrating disparate survey data in
+species distribution models demonstrates the need for robust model
+evaluation. In press at Canadian Journal of Fisheries and Aquatic
+Science.
+\item[2023]
+Davies S.C., P.L. Thompson, C. Gomez, J. Nephin, A. Knudby, A.E. Park,
+S.K. Friesen, L.J. Pollock, E.M. Rubidge, \textbf{S.C. Anderson}, J.C.
+Iacarella, D.A. Lyons, A.A. MacDonald, A. McMillan, E.J. Ward, A.M.
+Holdsworth, N. Swart, J. Price, K.L. Hunter. Addressing uncertainty when
+projecting marine species' distributions under climate change.
+Ecography: e06731. \url{https://doi.org/10.1111/ecog.06731}.
 \item[2023]

### 2023-10-17 cv/cv-body.tex
DOI

@@
-\hypertarget{professional-experience}{%
-\section{Professional experience}\label{professional-experience}}
+\section{Professional experience}\label{professional-experience}
 
@@
 
-\hypertarget{education}{%
-\section{Education}\label{education}}
+\section{Education}\label{education}
 
@@
 
-\hypertarget{preprints}{%
-\section{Preprints}\label{preprints}}
+\section{Preprints}\label{preprints}
 
@@
 
-\hypertarget{publications}{%
-\section{Publications}\label{publications}}
+\section{Publications}\label{publications}
 
@@
 standardization under reduced survey coverage. In press at ICES Journal
-of Marine Science.
+of Marine Science. \url{https://doi.org/10.1093/icesjms/fsad155}.
 \item[2023]
@@
 \url{https://doi.org/10.1093/icesjms/fsu055}
-(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF}).
+(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref&ijkey=NEXmZIkz3289u3z}{PDF}).
 \item[2015]
@@
 \url{https://doi.org/10.1371/journal.pone.0092725}
-(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0092725\&representation=PDF}{PDF}).
+(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0092725&representation=PDF}{PDF}).
 (\href{http://cran.r-project.org/web/packages/ss3sim/index.html}{R
@@
 \url{https://doi.org/10.1371/journal.pone.0078041}
-(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0078041\&representation=PDF}{PDF}).
+(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0078041&representation=PDF}{PDF}).
 \item[2013]
@@
 \url{https://doi.org/10.1371/journal.pone.0014735}
-(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0014735\&representation=PDF}{PDF}).
+(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0014735&representation=PDF}{PDF}).
 \href{https://f1000.com/prime/9542957}{F1000 Prime recommended}.
@@
 
-\hypertarget{peer-reviewed-government-reports}{%
 \section{Peer-reviewed government
-reports}\label{peer-reviewed-government-reports}}
+reports}\label{peer-reviewed-government-reports}
 
@@
 
-\hypertarget{technical-reports}{%
-\section{Technical Reports}\label{technical-reports}}
+\section{Technical Reports}\label{technical-reports}
 
@@
 
-\hypertarget{awards-and-scholarships}{%
-\section{Awards and scholarships}\label{awards-and-scholarships}}
+\section{Awards and scholarships}\label{awards-and-scholarships}
 
@@
 
-\hypertarget{software}{%
-\section{Software}\label{software}}
+\section{Software}\label{software}
 
@@
 
-\hypertarget{selected-first-author-invited-conference-presentations}{%
 \section{Selected first-author invited conference
-presentations}\label{selected-first-author-invited-conference-presentations}}
+presentations}\label{selected-first-author-invited-conference-presentations}
 
@@
 
-\hypertarget{teaching}{%
-\section{Teaching}\label{teaching}}
+\section{Teaching}\label{teaching}
 
@@
 
-\hypertarget{working-groups-and-workshops}{%
 \section{Working groups and
-workshops}\label{working-groups-and-workshops}}
+workshops}\label{working-groups-and-workshops}
 
@@
 
-\hypertarget{reviewing}{%
-\section{Reviewing}\label{reviewing}}
+\section{Reviewing}\label{reviewing}
 

### 2023-11-07 cv/cv-body.tex
Add paper + doi
Add cjfas paper

@@
 \item[2023]
+Nephin, J., P.L. Thompson, \textbf{S.C. Anderson}, A.E. Park, C.N.
+Rooper, B. Aulthouse, and J. Watson. Integrating disparate survey data
+in species distribution models demonstrate the need for robust model
+evaluation. In press at Can. J. Fish. Aquat. Sci.
+\url{https://doi.org/10.1139/cjfas-2022-0279}.
+\item[2023]
+Baker, K.D., \textbf{S.C. Anderson}, D.R.J. Mullowney, W. Walkusz, and
+K.R. Skanes. Moving away from a scale mismatch: Spatiotemporal modelling
+of striped shrimp (\emph{Pandalus montagui}) density in Canada's
+subarctic. Fisheries Research. 270: 106898.
+\url{https://doi.org/10.1016/j.fishres.2023.106898}.
+\item[2023]
 Yalcin, S., \textbf{S.C. Anderson}, P.M. Regular, P.A. English.
@@
 receive more attention when named after celebrities? In press at
-Conservation Biology. Preprint:
-\url{https://doi.org/10.31235/osf.io/wgbpv}.
-\item[2023]
-J. Nephin, P.L. Thompson, \textbf{S.C. Anderson}, A.E. Park, C.N.
-Rooper, B. Aulthouse, J. Watson. Integrating disparate survey data in
-species distribution models demonstrates the need for robust model
-evaluation. In press at Canadian Journal of Fisheries and Aquatic
-Science.
+Conservation Biology. \url{https://doi.org/10.1111/cobi.14184}.
 \item[2023]

### 2024-02-20 cv/cv-body.tex
Add paper
space
Update publications

@@
 \tightlist
+\item[2024]
+Thorson, J.T., \textbf{S.C. Anderson}, P. Goddard, C.N. Rooper.
+tinyVAST: R package with an expressive interface to specify lagged and
+simultaneous effects in multivariate spatio-temporal models. arXiv.
+\url{https://doi.org/10.48550/arXiv.2401.10193}.
 \item[2022]
-\textbf{Anderson, S.C.}, E.J. Ward, P.A. English, L.A.K. Barnett.
-sdmTMB: an R package for fast, flexible, and user-friendly generalized
-linear mixed effects models with spatial and spatiotemporal random
-fields. bioRxiv 2022.03.24.485545.
+\textbf{Anderson, S.C.}, E.J. Ward, P.A. English, L.A.K. Barnett, J.T.
+Thorson. sdmTMB: an R package for fast, flexible, and user-friendly
+generalized linear mixed effects models with spatial and spatiotemporal
+random fields. bioRxiv 2022.03.24.485545.
 \url{https://doi.org/10.1101/2022.03.24.485545}.
@@
 \tightlist
+\item[2024]
+\textbf{Anderson, S.C.}, P.A. English, K.S.P. Gale, D.R. Haggarty, C.K.
+Robb, E.M. Rubidge, P.L. Thompson. Impacts on population indices if
+scientific surveys are excluded from marine protected areas. ICES
+Journal of Marine Science. fsae009.
+\url{https://doi.org/10.1093/icesjms/fsae009}.
+\item[2024]
+Hughes, B.B., Beheshti, K.M., Tinker, M.T., Angelini, C., Endris, C.,
+Murai, L., \textbf{Anderson, S.C.}, Espinosa, S., Staedler, M.,
+Tomoleoni, J.A., Sanchez, M., and Silliman, B.R. Top-predator recovery
+abates geomorphic decline of a coastal ecosystem. Nature 626(7997):
+111--118. \url{https://doi.org/10.1038/s41586-023-06959-9}.
+\item[2024]
+English, P.A., C.M. Picco, J.C. Edwards, D.R. Haggarty, R.E. Forrest,
+\textbf{S.C. Anderson}. Spatial restrictions hinder avoidance of choke
+species in an Indigenous rights-based fishery. People and Nature. 6:
+75--90. \url{https://doi.org/10.1002/pan3.10554}.
 \item[2023]
@@
 in species distribution models demonstrate the need for robust model
-evaluation. In press at Can. J. Fish. Aquat. Sci.
+evaluation. Can. J. Fish. Aquat. Sci. 80(12): 1869-1889.
 \url{https://doi.org/10.1139/cjfas-2022-0279}.
-\item[2023]
+\item[2024]
 Baker, K.D., \textbf{S.C. Anderson}, D.R.J. Mullowney, W. Walkusz, and
@@
 Exploring the limits of spatiotemporal and design-based index
-standardization under reduced survey coverage. In press at ICES Journal
-of Marine Science. \url{https://doi.org/10.1093/icesjms/fsad155}.
+standardization under reduced survey coverage. ICES Journal of Marine
+Science. 80(9): 2368--2379.
+\url{https://doi.org/10.1093/icesjms/fsad155}.
 \item[2023]
 K.L. Blake, D. Veríssimo, \textbf{S.C. Anderson}, A. Gleave. Do species
-receive more attention when named after celebrities? In press at
-Conservation Biology. \url{https://doi.org/10.1111/cobi.14184}.
+receive more attention when named after celebrities? Conservation
+Biology. e14184. \url{https://doi.org/10.1111/cobi.14184}.
 \item[2023]
@@
 bottom-associated marine fish, Atlantic cod (\emph{Gadus morhua}). ICES
-Journal of Marine Science. In press.
+Journal of Marine Science. 80(5): 1539--1550.
 \url{https://doi.org/10.1093/icesjms/fsad084}.
@@
 fisheries: Surprises and lessons from key case studies. Fish and
-Fisheries. In press. \url{https://doi.org/10.1111/faf.12753}.
+Fisheries. 24: 652--674. \url{https://doi.org/10.1111/faf.12753}.
 \item[2023]

### 2024-09-20 cv/cv-body.tex
Add paper
cv

@@
 \tightlist
-\item[2021--23]
+\item[2021--24]
 Research Scientist, Stock Assessment and Research Division, Pacific
 Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, Canada
-\item[2019--23]
+\item[2019--24]
 Associate Editor, the Canadian Journal of Fisheries and Aquatic Sciences
@@
 \item[2024]
+Ward, E.J., \textbf{S.C. Anderson}, L.A.K. Barnett, P.A. English, H.M.
+Berger, C.J.C. Commander, T.E Essington, C.J Harvey, M.E. Hunsicker, M.
+Jacox, K.F Johnson, S. Large, O.R. Liu, K. Richerson, J.F. Samhouri,
+S.A. Siedlecki, A.O. Shelton, K.A. Somers, and J.T. Watson. Win, lose,
+or draw: evaluating dynamic thermal niches of northeast Pacific
+groundfish. EarthArXiv Preprint. \url{https://doi.org/10.31223/X5J113}.
+\item[2024]
+Indivero, J., \textbf{S.C. Anderson}, L.A.K. Barnett, T.E. Essington,
+and E.J. Ward. Estimating physiological mechanisms from monitoring data
+reveals challenges and opportunities for forecasting distribution
+shifts. Authorea Preprint.
+\url{https://doi.org/10.22541/au.172446072.20202938/v1}.
+\item[2024]
 Thorson, J.T., \textbf{S.C. Anderson}, P. Goddard, C.N. Rooper.
@@
 \url{https://doi.org/10.48550/arXiv.2401.10193}.
-\item[2022]
-\textbf{Anderson, S.C.}, E.J. Ward, P.A. English, L.A.K. Barnett, J.T.
-Thorson. sdmTMB: an R package for fast, flexible, and user-friendly
-generalized linear mixed effects models with spatial and spatiotemporal
-random fields. bioRxiv 2022.03.24.485545.
-\url{https://doi.org/10.1101/2022.03.24.485545}.
 \item[2021]
@@
 \tightlist
+\item[2024]
+\textbf{Anderson, S.C.}, E.J. Ward, P.A. English, L.A.K. Barnett, and
+J.T. Thorson. sdmTMB: an R package for fast, flexible, and user-friendly
+generalized linear mixed effects models with spatial and spatiotemporal
+random fields. In press at Journal of Statistical Software. Preprint:
+bioRxiv 2022.03.24.485545.
+\url{https://doi.org/10.1101/2022.03.24.485545}.
+\item[2024]
+Freshwater, C., \textbf{S.C. Anderson}, and J. King. Model-based indices
+of juvenile Pacific salmon abundance highlight species-specific seasonal
+distributions and impacts of changes to survey design. Fisheries
+Research. 277: 107063.
+\url{https://doi.org/10.1016/j.fishres.2024.107063}.
+\item[2024]
+Bass, A.L., Anderson, S.C., Bateman, A.W., Connors, B.M., Peña, M.A.,
+Li, S., Kaukinen, K.H., Patterson, D.A., Hinch, S. G., and Miller, K.M.
+2024. Intrinsic and extrinsic factors associated with the
+spatio-temporal distribution of infectious agents in early marine
+Chinook and coho salmon. Marine Ecology Progress Series. 736: 107--127.
+\url{https://doi.org/10.3354/meps14581}.
+\item[2024]
+Freshwater, C., \textbf{S.C. Anderson}, D.D. Huff, J.M. Smith, D.
+Jackson, B. Hendriks, S.G. Hinch, S. Johnston, A.W. Trites, and J. King.
+Chinook salmon depth distributions on the continental shelf are shaped
+by interactions between location, season, and individual condition.
+Movement Ecology. 12(1): 21.
+\url{https://doi.org/10.1186/s40462-024-00464-y}.
 \item[2024]
@@
 \item[2023]
-K.L. Blake, D. Veríssimo, \textbf{S.C. Anderson}, A. Gleave. Do species
-receive more attention when named after celebrities? Conservation
+K.L. Blake, D. Veríssimo, \textbf{S.C. Anderson}, A. Gleave. Impact on
+species' online attention when named after celebrities. Conservation
 Biology. e14184. \url{https://doi.org/10.1111/cobi.14184}.
